### 2016-11-30 project_report/main.tex
Added report tex file.

\documentclass{llncs}
\usepackage[utf8]{inputenc}
\usepackage{todonotes}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{dirtytalk}
\usepackage{color}

\begin{document}
\frontmatter

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%  HEADER  %%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
\title{Malware Detection Via Machine Learning}
\subtitle{Some Clever Subtitle?}
\titlerunning{Malware Detection Via ML}
\author{João F. Godinho\inst{1} \and Pedro Adão\inst{2}}
\institute{Instituto Superior Técnico\\
\email{enter email}\and
Instituto Superior Técnico\\
\email{enter email}}
\maketitle
\todo[inline]{Validate the title, subtitle, authors, email}

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% ABSTRACT %%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
The abstract should summarize the contents of the paper
using at least 70 and at most 150 words.
\end{abstract}

\clearpage
\todo[inline]{Table of contents?}

\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%  INTRO   %%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec_introduction}
\todo[inline]{Motivation}
\begin{itemize}
	\item Define Malware, problem with that, subjectivity
	\item Quantify malware in recent years (quote vendors/get stats from header analysis)
	\item What is being done about malware?
	\item How is malware identified?
	\begin{itemize}
		\item Types of strategies
		\item Is it effective? efficient?
	\end{itemize}
	\item Being able to identify malware as early as possible would help?
	\item Introduce services like malwr.com/virustotal.com
	\item Introduce Machine Learning capabilites
	\item Automation of detection/classification
\end{itemize}

Malware, short for malicious software, is any type of software developed with the intent of performing some unwanted or unexpected behavior on a computer system, meaning (that) it acts against the user expectancy, knowingly or unknowingly. \todo{quote this, Malware, Rootkits \& Botnets A Beginner's Guide pp.10}The term malware was first introduced in 1990 by Yisrael Radai. Before then most malicious software were known as computer viruses given their ability to self-replicate, similar to their biological counterparts.

\todo[inline]{Maybe use the following as in intro on how to detect malware, followed by the usage of ML for such purpose}
Nowadays, malware comprehends a great variety of types, such as spywares, adwares, rootkits, ransomwares and, like previously mentioned, viruses, to name but a few. Each type will behave differently, both from an internal and external view.

From an external view, adware and ransomware will provide a noisy output to the user, given their function to deliver unwanted advertisements (for adware) and request data ransom (for ransomware). Spyware and rootkit types, on the other hand, will provide little to no user interaction, given their function to spy on user activity (for spyware) and conceal processes, files or system data (for rootkit).

From an internal view, adware and ransomware will behave similarly, since both need interfaces that facilitate user interaction (e.g. create and show dialogs), in contrast to spyware and and rootkit, which thrive on concealing actions (e.g. run hidden processes).

\hfill \break
Nowadays, computer viruses are seen as a type of malware. Other subsets of malware include \textit{spyware}, which monitor on user activity without their knowledge; \textit{adware}, which deliver automatic and mainly unwanted and unexpected advertisements; \textit{rootkit}, which are designed to be invisible and are used to conceal running processes, files or system data; \textit{ransomware}, which holds a computer system, or part of it, captive (usually through encryption) demanding a ransom for a user to regain control. These are but a few types of malware (\todo{Quote this, check comment}more can be seen at [])

Nowadays, computer viruses are seen as a type of malware, the term malware comprehends a great variety of types, such as: spyware, adware, rookit, ransomware,

% Quote this for more types of malware
% https://support.symantec.com/en_US/article.HOWTO54185.html

\todo[inline]{Problem}
\begin{itemize}
	\item Remove the human factor from detection/classification to speed up process.
	\item Introduce lack of temporal consistency here?
	\item Talk how related work creates own corpus and low corpus size/variation?
	\item Refer stats from header analysis? Time vendors take to take knowledge of samples/Time until new malware is detected in most vendors
	\item Reinforce time to classify malware?
\end{itemize}

\todo[inline]{Hypothesis}
\begin{itemize}
	\item Would a more automated approach, based on ML, facilitate malware detection/classification?
	\item Dynamic/Static analysis to detect and classify malware
	\item Temporal consistency is a factor that should be taken into account?

\end{itemize}
\todo[inline]{Objectives/Contributions}
\begin{itemize}
	\item Usage of real world analysis
	\item Try to apply ML to PE32 analysis from malwr.com (Talk about this here?)
	\item Contribute with a corpus of analysis from malwr.com
	\item Maintain a temporal consistency

\end{itemize}
\todo[inline]{Document outline}
The following sections are organized as follows. Section \ref{sec_related_work} describes related work and its influence on the current work. Section \ref{sec_solution} describes the planned approach for solving the problem at hand. Section \ref{sec_conclusions} closes the report and references the work done so far.

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% RELATED  %%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec_related_work}

\todo[inline]{Talk about the state of the art}
\todo[inline]{Talk about the related work}
\begin{itemize}
	\item Talk about products that facilitate analysis?
	\item How detailed should we go? Talk about first detection techniques? Only talk about ML techniques?
\end{itemize}

% explicar as tecnicas e porque, como usa-las, vantagens, what's short
%

\todo[inline]{Explain what is malware, get quotes from different articles}
\todo[inline]{Explain different malware types, get quotes}
\todo[inline]{Explain malware detection techniques}

% #1 Semantics-Aware Malware Detection
% #2 Attacking Malicious Code: A report to the Infosec Research Council
% #3 Malware, Rootkits & Botnets A Beginner's Guide pp.10
% #4 Computer viruses: theory and experiments
% #5 Theory of self-reproducing automata
% #6 The Evolution of Viruses and Worms
% #7 The art of computer virus research and defense
% #8 https://www.av-test.org/en/
% #9 http://www.caro.org/articles/naming.html
% #10 http://members.chello.at/erikajo/vnc99b2.txt

Malware, which can be referred to by other names like malicious software or malicious code, is not easily described by a single definition. For example, Christodorescu et al.[quote 1] simply describe malware as \say{a program that has malicious intent}, which in turn begs the question, what is malicious intent? McGraw and Morrisett[quote 2] give a more concrete definition by saying \say{Malicious code is any code added, changed, or removed from a software system in order to
intentionally cause harm or subvert the intended function of the system}. Both definitions give an overall understanding of \textcolor{red}{(on?)} what is malware, but to better understand the definition, one can briefly look at the history of malware.

The term malware was first used by Yisrael Radai in 1990[quote 3], before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983[quote 4]. Before Cohen, John Von Neumann had already done some similar academic work that closely relates to computer viruses, namely \say{Theory of self-reproducing automata}, published in 1699[quote 5]. The term computer virus is one of the many types of malicious software, but given that it predates the term malware, it is not uncommon to see use both terms used as interchangeably.

Even before the formalization of computer virus by Cohen in 1983, one of the first documented computer virus, although experimental, was developed by Bob Thomas at BBN in 1971[quote 6]. The program was named Creeper and consisted of a self-replicating program that would print a message and move between the nodes of ARPANET.

Another malicious software worth mentioning, given it was one of the first known virus that spread \say{in the wild} (i.e. outside the computer system or laboratory in which it was developed), is Elk Cloner. Created by Rich Skrenta in 1982[quote 7], Elk Cloner spread by floppy disk and had a payload that displayed Skrenta's poem after every 50th use of the infected system.

Since then, the amount of malware has grown at an alarming rate. According to AV-TEST[quote 8], an independent security software group, up to 2012 the amount of total malware per year would not surpass 100 million, whereas 2013 almost reached 200 million, followed by more than 100 million increase per year, with almost 600 million total malware before the end of 2016.

Even though the term malware generalizes the malicious intent of a piece of software, one can classify malware further based on its purpose and behavior. Given the high amount of malware, there is also a high variety of malware, based on purpose and behavior.

From the general purpose and behavior of a malware sample, it can be classified as belonging to one or more malware categories (or classes). Examples of categories are virus, worms, trojans and adware, to name a few. Having the ability to classify malware into a more restrict subset is helpful, but there is still a need to distinguish between samples of the same category.

One attempt at creating a naming convention was made in 1991, by the Computer Antivirus Research Organization (CARO). The convention states the name should consist of four parts, separated by a dot. The four constituents are the family name, group name, major variant name and minor variant name, displayed in that order. An optional modifier could also be appended to the name with a colon. The problem with the aforementioned convention is that it only covered computer viruses, which were more prominent at the time.

An attempt at improving CARO's naming convention was made in 1999, by Gerald Scheidl[quote 10]. His proposal added two other constituents, platform and type, as a prefix to CARO's naming.

Later attempts to standardize malware naming were made, none of which resulted in a standard widely adopted. Due to this, antivirus vendors use their own naming scheme, which are variants of the mentioned schemes.

Since the malware naming scheme is not standardized, and given the amount and variety of malware, different families names with the same behavior emerge from different antivirus vendors. The name redundancy is worsen by how different vendors detect and classify malware.

\todo[inline]{start how detection/classification is made}

\todo[inline]{Is this okay? Use own definition of malware? Based on the previous paragraphs}
Given the subjectivity on the definition of malware, this work accepts the term malware as any type of software developed with the intent of performing some unwanted or unexpected behavior on a computer system, acting against the user expectancy, knowingly or unknowingly.

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% SOLUTION %%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
\section{Solution}\label{sec_solution}

\todo[inline]{Start by the approach?}
\todo[inline]{Talk about the architecture}
\todo[inline]{Should evaluation and planning go inside this section?}
\todo[inline]{Talk about the stats of header analysis here?}
\todo[inline]{Talk about what has already been/being done?}
\begin{itemize}
	\item Begin with extraction from malwr.com
	\item Talk about why use of PE32?
	\item Create a corpus from the analysis
	\item How to classify samples as malware? How to define the threshold?
	\item How to do feature selection? Measure multiple features?
	\item Create a baseline to evaluate against, validate other's work? Ignore temporal consistency?
	\item Introduce temporal consistency in samples, how well one can detect/classify with limited samples
	\item Try to classify/detect samples that were taken after the training samples.
\end{itemize}

% tools existentes q vao ser usadas

% \section{•} % Metodologia de Avaliação do Trabalho
% \section{•} % Calendarização do Trabalho

%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%CONCLUSION%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}\label{sec_conclusions}

\begin{thebibliography}{}

\end{thebibliography}

\end{document}

### 2016-12-01 project_report/main.tex
More introduction...

@@
 \section{Introduction}\label{sec_introduction}
+% #1 Semantics-Aware Malware Detection
+% #2 Attacking Malicious Code: A report to the Infosec Research Council
+% #3 Malware, Rootkits & Botnets A Beginner's Guide pp.10
+% #4 Computer viruses: theory and experiments
+% #5 Theory of self-reproducing automata
+% #6 The Evolution of Viruses and Worms
+% #7 The art of computer virus research and defense
+% #8 https://www.av-test.org/en/
+% #9 http://www.caro.org/articles/naming.html
+% #10 http://members.chello.at/erikajo/vnc99b2.txt
+Malware, which can be referred to by other names like malicious software or malicious code, is not easily described by a single definition.
+For example, Christodorescu et al.[quote 1] simply describe malware as \say{a program that has malicious intent}, which in turn begs the question, what is malicious intent?
+McGraw and Morrisett[quote 2] give a more concrete definition by saying \say{Malicious code is any code added, changed, or removed from a software system in order to
+intentionally cause harm or subvert the intended function of the system}.
+Both definitions give an overall understanding of \textcolor{red}{(on?)} what is malware, but to better understand the definition, one can briefly look at the history of malware.
+
+The term malware was first used by Yisrael Radai in 1990[quote 3], before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983[quote 4].
+Before Cohen, John Von Neumann had already done some similar academic work that closely relates to computer viruses, namely \say{Theory of self-reproducing automata}, published in 1699[quote 5].
+The term computer virus is one of the many types of malicious software, but given that it predates the term malware, it is not uncommon to see both terms used as interchangeably.
+
+Even before the formalization of computer virus by Cohen in 1983, one of the first documented computer virus, although experimental, was developed by Bob Thomas at BBN in 1971[quote 6].
+The program was named Creeper and consisted of a self-replicating program that would print a message and move between the nodes of ARPANET.
+
+Another malicious software worth mentioning, given it was one of the first known virus that spread \say{in the wild} (i.e.
+outside the computer system or laboratory in which it was developed), is Elk Cloner.
+Created by Rich Skrenta in 1982[quote 7], Elk Cloner spread by floppy disk and had a payload that displayed Skrenta's poem after every 50th use of the infected system.
+
+Since then, the amount of malware has grown at an alarming rate.
+According to AV-TEST[quote 8], an independent security software group, up to 2012 the amount of total malware per year would not surpass 100 million, whereas 2013 almost reached 200 million, followed by more than 100 million increase per year, with almost 600 million total malware before the end of 2016.
+
+Even though the term malware generalizes the malicious intent of a piece of software, one can classify malware further based on its purpose and behavior.
+Given the high amount of malware, there is also a high variety of malware, based on purpose and behavior.
+
+From the general purpose and behavior of a malware sample, it can be classified as belonging to one or more malware categories (or classes).
+Examples of categories are virus, worms, trojans and adware, to name a few.
+Having the ability to classify malware into a more restrict subset is helpful, but there is still a need to distinguish between samples of the same category, by giving more detailed names to malware.
+
+One attempt at creating a naming convention was made in 1991, by the Computer Antivirus Research Organization (CARO).
+The convention states the name should consist of four parts, separated by a dot.
+The four constituents are the family name, group name, major variant name and minor variant name, displayed in that order.
+An optional modifier could also be appended to the name with a colon.
+The problem with the aforementioned convention is that it only covered computer viruses, which were more prominent at the time.
+
+An attempt at improving CARO's naming convention was made in 1999, by Gerald Scheidl[quote 10]. His proposal added two other constituents, platform and type, as a prefix to CARO's naming.
+
+Later attempts to standardize malware naming were made, none of which resulted in a standard widely adopted.
+Due to this, antivirus vendors use their own naming scheme, which are variants of the mentioned schemes.
+
+Since the malware naming scheme is not standardized, and given the amount and variety of malware, different family names with the same behavior emerge from different antivirus vendors.
+The name redundancy is worsen by how different vendors detect and classify malware.
+
+With the raise of malware amount and diversity, being able to identify and classify malware becomes essential in order to protect systems.
+
+Before the late eighties, given the low amount of malware, the first antivirus programs were designed to find and remove a handful of malware, mainly through signature based methods, which detect malware by looking for specific patterns that appear in known malware.
+
+In the late eighties and nineties the first antivirus were released, along with more sophisticated detection methods that make use of heuristics, which also detect malware by also looking for patterns, but at a higher level of abstraction.
+
+In more recent years, and given the increment of malware amount and variety, signature and heuristic methods fall short when keeping up with new types of malware.
+The mentioned methods work best for samples of known malware or new malware with small variants of known malware.
+This is because these methods look for known patterns.
+If a new type emerges, which pattern's are not known, it is not flagged as malicious.
+
+Given this limitation, more advanced methods like behavioral analysis are applied together with the previous mentioned methods.
+Behavioral analysis takes into account the actions performed by the malware, introducing dynamic information, from which patterns can be extracted and again matched against known patterns.
+
+% VirusScan (from McAfee) and NOD (from ESET) were first released in 1987; AntiVir (from Avira) and Avast Antivirus were released in 1988.
+Norton Anti-Virus (from Symantec) and F-Secure antivirus were released in 1991.
+
+\todo[inline]{talk about machine learning, how it can be applied with behavioral analysis}
+\hfill \break
+
 \todo[inline]{Motivation}
@@
 
-Malware, short for malicious software, is any type of software developed with the intent of performing some unwanted or unexpected behavior on a computer system, meaning (that) it acts against the user expectancy, knowingly or unknowingly.
-\todo{quote this, Malware, Rootkits \& Botnets A Beginner's Guide pp.10}The term malware was first introduced in 1990 by Yisrael Radai.
-Before then most malicious software were known as computer viruses given their ability to self-replicate, similar to their biological counterparts.
-
-\todo[inline]{Maybe use the following as in intro on how to detect malware, followed by the usage of ML for such purpose}
-Nowadays, malware comprehends a great variety of types, such as spywares, adwares, rootkits, ransomwares and, like previously mentioned, viruses, to name but a few.
-Each type will behave differently, both from an internal and external view.
-
-From an external view, adware and ransomware will provide a noisy output to the user, given their function to deliver unwanted advertisements (for adware) and request data ransom (for ransomware).
+%From an external view, adware and ransomware will provide a noisy output to the user, given their function to deliver unwanted advertisements (for adware) and request data ransom (for ransomware).
 Spyware and rootkit types, on the other hand, will provide little to no user interaction, given their function to spy on user activity (for spyware) and conceal processes, files or system data (for rootkit).
-
-From an internal view, adware and ransomware will behave similarly, since both need interfaces that facilitate user interaction (e.g.
+%
+%From an internal view, adware and ransomware will behave similarly, since both need interfaces that facilitate user interaction (e.g.
 create and show dialogs), in contrast to spyware and and rootkit, which thrive on concealing actions (e.g.
 run hidden processes).
-
-\hfill \break
-Nowadays, computer viruses are seen as a type of malware.
-Other subsets of malware include \textit{spyware}, which monitor on user activity without their knowledge; \textit{adware}, which deliver automatic and mainly unwanted and unexpected advertisements; \textit{rootkit}, which are designed to be invisible and are used to conceal running processes, files or system data; \textit{ransomware}, which holds a computer system, or part of it, captive (usually through encryption) demanding a ransom for a user to regain control.
-These are but a few types of malware (\todo{Quote this, check comment}more can be seen at [])
-
-Nowadays, computer viruses are seen as a type of malware, the term malware comprehends a great variety of types, such as: spyware, adware, rookit, ransomware,
 
@@
 
-\todo[inline]{Explain what is malware, get quotes from different articles}
-\todo[inline]{Explain different malware types, get quotes}
-\todo[inline]{Explain malware detection techniques}
+%\todo[inline]{Explain what is malware, get quotes from different articles}
+%\todo[inline]{Explain different malware types, get quotes}
+%\todo[inline]{Explain malware detection techniques}
 
-% #1 Semantics-Aware Malware Detection
-% #2 Attacking Malicious Code: A report to the Infosec Research Council
-% #3 Malware, Rootkits & Botnets A Beginner's Guide pp.10
-% #4 Computer viruses: theory and experiments
-% #5 Theory of self-reproducing automata
-% #6 The Evolution of Viruses and Worms
-% #7 The art of computer virus research and defense
-% #8 https://www.av-test.org/en/
-% #9 http://www.caro.org/articles/naming.html
-% #10 http://members.chello.at/erikajo/vnc99b2.txt
-
-Malware, which can be referred to by other names like malicious software or malicious code, is not easily described by a single definition.
-For example, Christodorescu et al.[quote 1] simply describe malware as \say{a program that has malicious intent}, which in turn begs the question, what is malicious intent?
-McGraw and Morrisett[quote 2] give a more concrete definition by saying \say{Malicious code is any code added, changed, or removed from a software system in order to
-intentionally cause harm or subvert the intended function of the system}.
-Both definitions give an overall understanding of \textcolor{red}{(on?)} what is malware, but to better understand the definition, one can briefly look at the history of malware.
-
-The term malware was first used by Yisrael Radai in 1990[quote 3], before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983[quote 4].
-Before Cohen, John Von Neumann had already done some similar academic work that closely relates to computer viruses, namely \say{Theory of self-reproducing automata}, published in 1699[quote 5].
-The term computer virus is one of the many types of malicious software, but given that it predates the term malware, it is not uncommon to see use both terms used as interchangeably.
-
-Even before the formalization of computer virus by Cohen in 1983, one of the first documented computer virus, although experimental, was developed by Bob Thomas at BBN in 1971[quote 6].
-The program was named Creeper and consisted of a self-replicating program that would print a message and move between the nodes of ARPANET.
-
-Another malicious software worth mentioning, given it was one of the first known virus that spread \say{in the wild} (i.e.
-outside the computer system or laboratory in which it was developed), is Elk Cloner.
-Created by Rich Skrenta in 1982[quote 7], Elk Cloner spread by floppy disk and had a payload that displayed Skrenta's poem after every 50th use of the infected system.
-
-Since then, the amount of malware has grown at an alarming rate.
-According to AV-TEST[quote 8], an independent security software group, up to 2012 the amount of total malware per year would not surpass 100 million, whereas 2013 almost reached 200 million, followed by more than 100 million increase per year, with almost 600 million total malware before the end of 2016.
-
-Even though the term malware generalizes the malicious intent of a piece of software, one can classify malware further based on its purpose and behavior.
-Given the high amount of malware, there is also a high variety of malware, based on purpose and behavior.
-
-From the general purpose and behavior of a malware sample, it can be classified as belonging to one or more malware categories (or classes).
-Examples of categories are virus, worms, trojans and adware, to name a few.
-Having the ability to classify malware into a more restrict subset is helpful, but there is still a need to distinguish between samples of the same category.
-
-One attempt at creating a naming convention was made in 1991, by the Computer Antivirus Research Organization (CARO).
-The convention states the name should consist of four parts, separated by a dot.
-The four constituents are the family name, group name, major variant name and minor variant name, displayed in that order.
-An optional modifier could also be appended to the name with a colon.
-The problem with the aforementioned convention is that it only covered computer viruses, which were more prominent at the time.
-
-An attempt at improving CARO's naming convention was made in 1999, by Gerald Scheidl[quote 10]. His proposal added two other constituents, platform and type, as a prefix to CARO's naming.
-
-Later attempts to standardize malware naming were made, none of which resulted in a standard widely adopted.
-Due to this, antivirus vendors use their own naming scheme, which are variants of the mentioned schemes.
-
-Since the malware naming scheme is not standardized, and given the amount and variety of malware, different families names with the same behavior emerge from different antivirus vendors.
-The name redundancy is worsen by how different vendors detect and classify malware.
-
-\todo[inline]{start how detection/classification is made}
-
-\todo[inline]{Is this okay? Use own definition of malware? Based on the previous paragraphs}
-Given the subjectivity on the definition of malware, this work accepts the term malware as any type of software developed with the intent of performing some unwanted or unexpected behavior on a computer system, acting against the user expectancy, knowingly or unknowingly.
+%\todo[inline]{Is this okay? Use own definition of malware? Based on the previous paragraphs}
+%Given the subjectivity on the definition of malware, this work accepts the term malware as any type of software developed with the intent of performing some unwanted or unexpected behavior on a computer system, acting against the user expectancy, knowingly or unknowingly.
 

### 2016-12-02 project_report/main.tex
More intro, until malwr description.

@@
 
+Having to run samples in order to extract dynamic information poses the problem of infecting the system running the samples.
+One way to prevent this is to use a sandbox environment, which is a security mechanism to isolate running programs.
+Using a sandbox environment closely relates to computer virtualization.
+
+To provide users the possibility to check files without running the risk of being infected, online services allow users to submit files which are then run in a sandbox environment.
+
+An example of such service is malwr.
+Malwr takes user submissions and runs them in a sandbox environment, specifically Cuckoo Sandbox, providing static and dynamic information about the sample, as well as checking if the file is known in VirusTotal, a similar service that also checks if the submission is known in different antivirus solutions.
+Malwr was first launched in April 2013 and contains nearly 700 000 analyses.
+
 % VirusScan (from McAfee) and NOD (from ESET) were first released in 1987; AntiVir (from Avira) and Avast Antivirus were released in 1988.
@@
 
-\todo[inline]{talk about machine learning, how it can be applied with behavioral analysis}
+\todo[inline]{brief look on what malwr does and provides}
+\todo[inline]{re-focus on high amount of malware, difficulty of having to run all samples and check}
+\todo[inline]{propose machine learning as a solution to deal with the high amounts of data}
 \hfill \break

### 2016-12-16 project_report/main.tex
Removed some phrases.
Related work with background and part of malware detection.

@@
 \usepackage{color}
+\usepackage{glossaries}
+
+% Reset footnotes every page
+%\usepackage{perpage}
+%\MakePerPage{footnote}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%% ACRONYMS %%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%
+\newacronym{caro}{CARO}{Computer Antivirus Research Organization}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%% ABSTRACT %%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}\label{sec_introduction}
+\section{Introduction}\label{sec:introduction}
 % #1 Semantics-Aware Malware Detection
@@
 
+\todo[inline]{Clean previous paragraph. Remove bellow}
+
 The term malware was first used by Yisrael Radai in 1990[quote 3], before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983[quote 4].
@@
 
+\todo[inline]{Up to here}
+
 Since then, the amount of malware has grown at an alarming rate.
 According to AV-TEST[quote 8], an independent security software group, up to 2012 the amount of total malware per year would not surpass 100 million, whereas 2013 almost reached 200 million, followed by more than 100 million increase per year, with almost 600 million total malware before the end of 2016.
+
+\todo[inline]{Clean previous paragraph. Remove bellow}
 
@@
 Malwr takes user submissions and runs them in a sandbox environment, specifically Cuckoo Sandbox, providing static and dynamic information about the sample, as well as checking if the file is known in VirusTotal, a similar service that also checks if the submission is known in different antivirus solutions.
-Malwr was first launched in April 2013 and contains nearly 700 000 analyses.
 
@@
 
-\todo[inline]{brief look on what malwr does and provides}
 \todo[inline]{re-focus on high amount of malware, difficulty of having to run all samples and check}
@@
 The following sections are organized as follows.
-Section \ref{sec_related_work} describes related work and its influence on the current work.
-Section \ref{sec_solution} describes the planned approach for solving the problem at hand.
-Section \ref{sec_conclusions} closes the report and references the work done so far.
+Section \ref{sec:related_work} describes related work and its influence on the current work.
+Section \ref{sec:solution} describes the planned approach for solving the problem at hand.
+Section \ref{sec:conclusion} closes the report and references the work done so far.
 
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec_related_work}
+\section{Related Work}\label{sec:related_work}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Background
+\subsection{Background}\label{subsec:background}
+The term malware was first used by Yisrael Radai in 1990[quote 3], before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983[quote 4].
+Before Cohen, John Von Neumann had already done some similar academic work that closely relates to computer viruses, namely \say{Theory of self-reproducing automata}, published in 1699[quote 5].
+The term computer virus is one of the many types of malicious software, but given that it predates the term malware, it is not uncommon to see both terms used as interchangeably.
+
+Even before the formalization of computer virus by Cohen in 1983, one of the first documented computer virus, although experimental, was developed by Bob Thomas at BBN in 1971[quote 6].
+The program was named Creeper and consisted of a self-replicating program that would print a message and move between the nodes of ARPANET.
+
+Another malicious software worth mentioning, given it was one of the first known virus that spread \say{in the wild} (i.e.
+outside the computer system or laboratory in which it was developed), is Elk Cloner.
+Created by Rich Skrenta in 1982[quote 7], Elk Cloner spread by floppy disk and had a payload that displayed Skrenta's poem after every 50th use of the infected system.
+
+Over the years, malware has grown in size and diversity, which introduces the need to be able to distinguish malware, not only to facilitate its removal, but also to provide safer mechanisms of defense.
+
+Based on the malware general purpose and behavior, a high level classification of the malware can be made by splitting malware into one or more categories (or classes).
+Examples of such classes are viruses, worms, trojans and adware, to name a few.
+This high level classification is helpful, but lacks detail, like target platform, origin or version.
+In order solve this issue, attempts have been made to create a naming convention for malware.
+
+One of the first attempts was made in 1991, by the \gls{caro}\footnote{http://www.caro.org/articles/naming.html}, which states the name should consist of four  obligatory parts, separated by a dot, and an optional fifth part.
+The four main constituents are the family name, group name, major variant and minor variant name.
+The optional fifth part is a modifier and could be appended to the name with a colon.
+\todo[inline]{Describe the parts?}
+
+Although the aforementioned naming convention is well defined, it is aimed at naming viruses, lacking differentiation for other classes as well as target platform.
+In order to solve this, an improvement was proposed in 1999 by Gerald Scheidl\footnote{http://members.chello.at/erikajo/vnc99b2.txt} that suggested the addition of a prefix that consisted of the platform and type.
+
+A more recent attempt at creating a naming convention was made during the Virus Bulletin Conference in 2001\footnote{https://www.virusbulletin.com/uploads/pdf/magazine/2002/200201.pdf}, with the idea that malware naming should consist on three levels of information: malware name, classification and full text description.
+\todo[inline]{Again, describe parts?}
+
+None of the referenced naming conventions is widely adopted, with antivirus vendors having their own naming conventions, that although similar, vary in some way or another.
+For example, Trend Micro\footnote{http://docs.trendmicro.com/all/ent/smex/v11.0/en-us/smex\_11.0\_olh/secrisk\_malware\_name.html} and Symantec\footnote{https://www.symantec.com/security\_response/virusnaming.jsp} use a prefix (which identifies the malware class), threat name (i.e.
+family) and suffix (which indicates a variant of the threat).
+Whereas Microsoft\footnote{https://www.microsoft.com/security/portal/mmpc/shared/malwarenaming.aspx} tries to follow the \gls{caro}'s convention, by using a type, platform, family, variant and information.
+Avira\footnote{https://www.avira.com/en/support-malware-naming-conventions} on the other hand, takes a more simplistic approach by using a prefix (which identifies the malware type) and a threat name.
+
+Although the naming and its constituents are similar between vendors, no convention is set, hence the threat names given to malware varies, increasing the redundancy for the same malware.
+According to an article published by Symantec in 2002\footnote{https://www.symantec.com/connect/articles/virus-any-other-name-virus-naming-practices}, based on a list of 300,000 unique names, the name redundancy is quantified as an average of four different names for the same malware.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Malware detection
+\subsection{Malware Detection}\label{subsec:malware_detection}
+In the early days of malware, due to low amount and variety, antivirus could target malware specifically (e.g.
+Reaper was created to remove Creeper), but with an estimate of more than 600 million total malware by the end of 2016, targeting specific malware becomes increasingly more difficult, hence being able to identify and classify malware efficiently is crucial to protect systems.
+
+What follows is a quick rundown on different detection techniques developed over the years[quote 7].
+
+\subsubsection{First-Generation Scanners.} These methods of detection work at a fairly simple level and use static methods of detection. The applied principles include:
+\begin{itemize}
+	\item String scanning, where a sequence of bytes that is common in a malicious program but not in a regular program is matched against the suspicious file.
+	\item Wildcards, similar to string scanning, but instead of a fixed sequence of bytes, regular expressions can be used in the sequence of bytes, this allows variation and byte skipping when matching.
+	\item Mismatches, also similar to string scanning, in which a sequence of bytes is taken as reference, but a number of bytes $N$ is given allowing up to $N$ bytes to vary in any position of the string (e.g.
+the string \say{\texttt{01 02}} with $N=1$ will match \say{\texttt{XX 02}} or \say{\texttt{01 XX}}).
+\end{itemize}
+
+The described principles can be applied with slight variations that improve the performance of the scanner, such as matching on the beginning and end of the file (top-and-tail scanning) or matching at a certain offset (bookmarks).
+
+They all require a knowledge base with known patterns that must be kept up-to-date in order to detect newer malware.
+
+\subsubsection{Second-Generation Scanners.} These methods improve on the previous generation by doing a static analysis at a higher level of abstraction.
+Where first-generation scanners would match sequence of bytes, second-generation scanners also look at the representation of the bytes.
+The applied principles include:
+\begin{itemize}
+	\item Smart scanning, where the Assembly representation of the bytes is taken into account and instructions that have no effect (e.g.
+\texttt{NOP}) can be skipped.
+The resulting signature is smaller and more easily detects mutations.
+The same principle could be applied to textual malware (e.g.
+scripts), where extra white spaces have no effect.
+	\item Nearly exact identification, where instead of relying on a single string for detection, double-string detection can be used.
+An alternative to double-string detection could also be based on a checksum range selected from the malware body, giving a higher accuracy by taking into account a larger portion of the malicious file.
+	\item Exact identification, which can exactly identify malware variants, is usually used in combination with first-generation methods and instead of using a single range like nearly exact, uses multiple ranges to calculate checksums.
+To be accurate, this method must only take into account constant bits of the malicious file, and given that more ranges are used it is slower then the previous ones.
+\end{itemize}
+
+Like in the previous generation, a knowledge base must be used and kept up-to-date to identify newer malware.
+
+\subsubsection{Algorithmic Scanning Methods.} These methods are not so focused on direct matching of signatures, but on the routines implemented in the scanner that allow the detection of malware.
+Early implementations of algorithmic scanning were hard-coded in the antivirus, but with different types and families emerging, the hard-coded algorithms could easily become obsolete as such.
+To overcome this, vendors would introduce more specific routines for malicious programs when updating the antivirus.
+
+Being able to update the antivirus algorithms frequently enables the vendors to have a faster response to new malware by developing detection routines for new malware and pushing them to end users.
+
+\subsubsection{Code Emulation.} This detection technique uses dynamic analysis to detect malware by simulating code execution.
+A virtual machine is used to simulate the system's components, thus the malicious files are encapsulated and no code is executed by the real processor.
+
+Code emulation can be faster than using algorithmic scanning methods when dealing with encrypted/obfuscated malware, as the code can be emulated and the decrypted/deobfuscated code is much easier obtained.
+
+\subsubsection{Heuristic Analysis.} Weights different "features" to assert if a suspicious file is malware or not, which allows to detect never seen malware types.
+
+\subsubsection{Integrity Checking.}
+
+\subsubsection{Sand-Boxing.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Machine Learning
+\subsection{Machine Learning}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Machine Learning in malware detection
+\subsection{Machine Learning in Malware Detection}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\clearpage
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Solution}\label{sec_solution}
+\section{Solution}\label{sec:solution}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}\label{sec_conclusions}
+\section{Conclusion}\label{sec:conclusion}
 

### 2016-12-22 project_report/main.tex
More related and solution.

@@
 
-\subsubsection{Heuristic Analysis.} Weights different "features" to assert if a suspicious file is malware or not, which allows to detect never seen malware types.
+\subsubsection{Heuristic Analysis.} This detection technique takes into account several characteristics of malicious software, both static and dynamic.
+Examples of such characteristics are code execution starting in uncommon sections, suspicious code redirection or incorrect size of code in header.
 
-\subsubsection{Integrity Checking.}
+When checking a suspicious file, the presence or absence of multiple characteristics is weighted to assert if the file is deemed malicious or not.
+The upside of this technique is that it allows for never seen malware to be correctly flagged, but brings the downside of a higher amount of false positives.
 
-\subsubsection{Sand-Boxing.}
+\subsubsection{Sand-Boxing.} This detection technique builds on the already mentioned code emulation, but with a higher level of complexity.
+With sand-boxing the whole system is emulated (i.e.
+virtualized) and suspicious or malicious programs have access to a copy of the real resources, effectively limiting the effects of the program and consequently protecting the system.
+
+This technique allows to extract dynamic information without compromising the real system, which can then be used to both detect if a suspicious program is malware, by applying the aforementioned techniques, and to help analyze new types of malware.
+
+The downside of using sand-boxing comes in the form of compatibility issues and in the possibility of malware breaking out of the sand-box due to vulnerabilities in the virtualization software.\\
+
+Detecting malware is not as simple as applying one of the mentioned techniques, different methods work best in different scenarios, hence antivirus need to balance the usage of these techniques to keep systems protected.
+
+\todo[inline]{Talk about cuckoo here?}
 
@@
 
+\todo[inline]{Paragraph introducing ML?}
+\todo[inline]{quick rundown on supervised/unsupervised/semi-supervised}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% #11 Reviewer Integration and Performance Measurement for Malware Detection
 
-\clearpage
-
-\todo[inline]{Talk about the state of the art}
-\todo[inline]{Talk about the related work}
-\begin{itemize}
-	\item Talk about products that facilitate analysis?
-	\item How detailed should we go? Talk about first detection techniques? Only talk about ML techniques?
-\end{itemize}
+Miller et al provide an interesting work on combining machine learning techniques with manual review of samples[quote 11].
+They extract both static and dynamic features from binaries and apply logistic regression to achieve a 72\% detection rate with a 0.5\% false positive rate.
+They also note that performance measurement can easily be inflated when ignoring temporal consistency in both binary samples and labels.
+Such can happen when using methods like \textit{cross-validation} for performance measurement, which implies that all types of malware are both seen in evaluation and in training, consequently the notion of unseen malware is not present.
 
@@
 %
-
-%\todo[inline]{Explain what is malware, get quotes from different articles}
-%\todo[inline]{Explain different malware types, get quotes}
-%\todo[inline]{Explain malware detection techniques}
-
-%\todo[inline]{Is this okay? Use own definition of malware? Based on the previous paragraphs}
-%Given the subjectivity on the definition of malware, this work accepts the term malware as any type of software developed with the intent of performing some unwanted or unexpected behavior on a computer system, acting against the user expectancy, knowingly or unknowingly.
 
@@
 \section{Solution}\label{sec:solution}
+
+With the rate at which malware is discovered, having the ability to automatically detect both known and new samples of malware facilitates the protection of systems, while reducing human interaction.
+As such, the current section proposes a solution that takes advantage of machine learning capabilities to detect malware, based on static and dynamic information that can be extract from malware samples.
+
+The solution is divided into four sequential stages, where each stage directly depends on the previous one.
+The four parts are: collecting; parsing; training and validation.
+More details about each stage is described in the corresponding subsection.
+
+With an implemented solution, it is expected to have a comparison on how different models predict a sample class (i.e.
+malware, not malware, what kind of malware), as well as benchmarks on the accuracy of the models.
+
+\subsection{Collecting}
+The first stage of the proposed solution, which consists on obtaining a corpus from which information about the malware can be extracted.
+For the purpose of this work, the obtained dataset consists on reports of static and dynamic information about binaries.
+Working with available reports instead of generating from samples saves the time overhead from both sample obtainment and report generation, allowing for more time better spent in the remaining stages, which are more related to the presented work.
+\todo[inline]{Talk about using malwr reports}
+
+\subsection{Parsing}
+The second state of the proposed solution, that consists on parsing the previously obtained data.
+The reports must be parsed not only to filter relevant information, but also to guarantee data canonicalization.
+Having a well structured dataset eases feature extraction and posterior application of machine learning techniques, hence this stage is of extreme importance for the upcoming stage.
+\todo[inline]{Talk about what can be extracted from this stage (metawise), like percentages, dups, etc}
+
+\subsection{Training}
+The third stage of the proposed solution, in which machine learning techniques are applied to obtain a model that can predict if a given sample is malicious and if so, what kind of malware is it.
+Before applying any machine learning techniques, a set of features must be selected from the dataset.
+Since the model will be trained based on the selected features, it is of uttermost importance that feature selection is done correctly, otherwise poorly chosen features will not be able to provide a good segregation between samples and consequently the final model will not as accurate.
+
+\todo[inline]{talk about unsupervised methods to facilitate feature selection}
+\todo[inline]{talk about thresholds for supervised methods to classify samples}
+
+\subsection{Validation}
+The fourth and last stage of the proposed solution, that consists on measuring how good the previously obtained model is.
+The validation of the model is done by taking a set of samples which were not used for training and predict to what class the sample belongs (i.e.
+malware, not malware, malware type).
+With the model prediction, and since the validation set is labeled, the true positive, true negative, false positive and false negative can be taken into account to assert how accurate the model is with regards to the correct labels.
+
+\todo[inline]{Talk about using cross-validation as a baseline}
+\todo[inline]{move to temporal consistency and how it performs vs cross}
+
+\subsection{Tools (or some more relevant name?)}
+Being able to easily modify and rerun each stage is one of the main concerns, as more time can be spent trying different techniques without spending much time on implementation.
+As such, Python 3 was chosen as the base language for development.
+
+There are a large number of libraries in Python that provide easy access to machine learning capabilities, information retrieval and plotting and statistics.
+The ones that are of main interest for this work are: SciKit-Learn, Scrapy, Numpy, Pandas and MatPlotLib.
+
+\todo[inline]{Talk here about INOV's machine?}
+
+\subsection{Developed work}
+The stages on which most time should be spent are training and validation.
+As such, some work as already been done in the previous stages, specifically in the collecting and parsing stage, so that more time can be dedicated to training and validation.
+
+With regards to the collecting stage, ...
+
+Regarding the parsing stage, ...
+
+\clearpage
 

### 2016-12-26 project_report/main.tex
Machine learning section.

@@
 \newacronym{caro}{CARO}{Computer Antivirus Research Organization}
+\newacronym{ai}{AI}{Artificial Intelligence}
+\newacronym{ml}{ML}{Machine Learning}
 
@@
 
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\todo[inline]{Paragraph introducing ML?}
-\todo[inline]{quick rundown on supervised/unsupervised/semi-supervised}
+Even before the modern concept of computer proposed by Alan Turing, mankind has always searched for ways to either automate or facilitate more complex tasks.
+With the technological advances that provide higher computation power, computer science fields like \gls{ai} have greatly improved in recent years.
+One of the \gls{ai} fields that as benefit from the technological advances is \gls{ml}.
+
+\gls{ml} is a subset of \gls{ai} in which algorithms are applied to learn and make predictions on data, or more formally defined by Thomas Mitchell\cite{mitchell:ml}:\\
+\say{\emph{A computer program is said to learn from experience E with respect
+to some class of tasks T and performance measure P, if its performance at tasks in
+T, as measured by P, improves with experience E.}}
+
+A noteworthy nuance about the given definition, and the \gls{ml} field in general, is that algorithms are not designed to deal with specific problems.
+Meaning that as long as a learning problem is well defined through $T$, $P$ and $E$, an algorithm based on these three features is able to learn, but not necessarily solve, a problem.
+
+After defining a learning problem, to obtain an useful outcome (i.e.
+model) one or more learning algorithms are applied.
+Learning algorithms fall within one of the following classes: supervised learning; unsupervised learning; semi-supervised learning or reinforcement learning.
+Each of the categories is defined on how the program learns the problem.
+A more detailed description, as seen in \cite{norvig:ai} is as follows:
+
+\subsubsection{Supervised Learning.} In this category the program is given a corpus (i.e.
+experience) with input-output pairs and it must learn how to map a never seen input into the correct output.
+Within supervised learning, a learning problem can be further classified into a classification problem, if the output is set of finite values (e.g.
+is it going to rain tomorrow?), or into a regression problem, if the output is continuous (e.g.
+how likely is it to rain tomorrow?).
+
+\subsubsection{Unsupervised Learning.} In this category the program is given a corpus and it must learn patterns from the input, hence no correct or wrong output is expected, only how different input relates.
+The most common task in unsupervised learning is clustering, where input is grouped based on similar characteristics.
+
+\subsubsection{Semi-supervised Learning.} This category is a combination of the previous ones and exists because sometimes the learning problem is based on a dataset that combines both labeled and unlabeled inputs or on a dataset that contains noisy data.
+The combination of supervised and unsupervised learning provides a better overall result in such scenarios.
+
+\subsubsection{Reinforcement Learning.} In this category the program is given a problem through a set of states and a set of actions it can perform.
+At any given state the program performs an action which results in a reward or punishment.
+The cumulative reinforcement at each state teaches the program the best and worse actions for each state.\\
+
+Another important part of \gls{ml} is the performance measure $P$, which dictates how well the program learned the problem.
+Performance measuring is related to the task $T$, meaning the measurement must clearly reflect on the task (e.g.
+if the task is playing checkers, performance should be percent of games won against opponents).
+
+For the purpose of this work, the \gls{ml} categories that fit the problem are supervised and unsupervised learning, as the problem can be seen as a classification problem (i.e.
+to which class a malware belongs) and a clustering problem (i.e.
+how similar is this malware type to known types).
+
+\clearpage
 
@@
 
+\bibitem{mitchell:ml}
+Mitchell, T.:
+Machine Learning
+
+\bibitem{norvig:ai}
+Norvig, P., Russell, S.:
+Artificial Intelligence: A Modern Approach
+
 \end{thebibliography}

### 2017-02-01 project_report/main.tex
Saving changes to doc

@@
 
-\gls{ml} is a subset of \gls{ai} in which algorithms are applied to learn and make predictions on data, or more formally defined by Thomas Mitchell\cite{mitchell:ml}:\\
+\gls{ml} is a subset of \gls{ai} in which algorithms are applied to learn and make predictions on data, or more formally defined by Tom Mitchell\cite{mitchell:ml}:\\
 \say{\emph{A computer program is said to learn from experience E with respect
@@
 
-% #11 Reviewer Integration and Performance Measurement for Malware Detection
+Even though the current work is not aimed at dealing with malicious binaries directly, but with a dataset derived from malware, prudence and methodological rigor must be taken into consideration.
+Rossow et al \cite{rossow:practices} present some guidelines for dealing with malware experiments, including malware datasets.
+Their work surveys 36 academic publications from 2006 to 2011 that rely on malware execution.
+From the guidelines they describe, failures are found regarding problematic assumptions on the use of execution-driven datasets, absence of security precautions taken during experiments and insufficient description of the experimental setup.
+The provided guidelines are split into groups:
+\begin{itemize}
+	\item Correct Datasets
+	\item Transparency
+	\item Realism
+	\item Safety guidelines are not relevant for the current work, since they are aimed at running malware.
+\end{itemize}
 
-Miller et al provide an interesting work on combining machine learning techniques with manual review of samples[quote 11].
+The provided guidelines are split into four groups: correct datasets; transparency; realism and safety.
+
+The mentioned groups are further detailed as follows:
+
+\paragraph{Correct Datasets.} This group include 6 guidelines which are: check if goodware samples should be removed from datasets
+
+Miller et al provide an interesting work on combining machine learning techniques with manual review of samples\cite{miller:rev_int}.
 They extract both static and dynamic features from binaries and apply logistic regression to achieve a 72\% detection rate with a 0.5\% false positive rate.
@@
 Mitchell, T.:
-Machine Learning
+\say{Machine Learning}.
+McGraw Hill, 1997
 
 \bibitem{norvig:ai}
-Norvig, P., Russell, S.:
-Artificial Intelligence: A Modern Approach
+Norvig, P. and Russell, S.:
+\say{Artificial Intelligence: A Modern Approach 3rd Edition}.
+Prentice Hall, 2010
+
+\bibitem{rossow:practices}
+Rossow C., Dietrich C., Grier C., Kreibich C., Paxson V., Pohlmann N., Bos H. and Steen M.:
+\say{Prudent Pratices for Designing Malware Experiments: Status Quo and Outlook} in IEEE Symposium on Security and Privacy, 2012
+
+\bibitem{miller:rev_int}
+Miller, B., Kantchelian, A., Tschantz, M., Afroz S., Bachwani R., Faizullabhoy R., Huang L., Shankar V., Wu T., Yiu G., Joseph A. and Tygar J.:
+\say{Reviewer Integration and Performance Measurement for Malware Detection} in DIMVA, 2016
 

### 2017-06-26 project_report/main.tex
Work at project submission

--- 2016-12-26 project_report/main.tex
@@
 \usepackage{glossaries}
+\usepackage{graphicx}
+\usepackage{float}
+\usepackage{booktabs}
+\usepackage[caption = false]{subfig}
+\graphicspath{ {images/} }
+\usepackage{rotating}
+\usepackage{pdflscape}
 
@@
 \title{Malware Detection Via Machine Learning}
-\subtitle{Some Clever Subtitle?}
+\subtitle{}
 \titlerunning{Malware Detection Via ML}
-\author{João F. Godinho\inst{1} \and Pedro Adão\inst{2}}
+\author{João F. Godinho}
 \institute{Instituto Superior Técnico\\
-\email{enter email}\and
-Instituto Superior Técnico\\
-\email{enter email}}
+\email{joao.f.godinho@tecnico.ulisboa.pt}}
+\let\oldaddcontentsline\addcontentsline
+\def\addcontentsline#1#2#3{}
 \maketitle
-\todo[inline]{Validate the title, subtitle, authors, email}
+\def\addcontentsline#1#2#3{\oldaddcontentsline{#1}{#2}{#3}}
+%\todo[inline]{Validate the title, subtitle}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
+\newacronym{hids}{HIDS}{Host-based Intrusion Detection Systems}
+\newacronym{ml}{ML}{machine learning}
 \newacronym{caro}{CARO}{Computer Antivirus Research Organization}
 \newacronym{ai}{AI}{Artificial Intelligence}
-\newacronym{ml}{ML}{Machine Learning}
+\newacronym{api}{API}{Application Programming Interface}
+\newacronym{bios}{BIOS}{Basic Input/Output System}
+\newacronym{dll}{DLL}{Dynamic-Link Library}
+\newacronym{cc}{CC}{Creative Common's}
+\newacronym{ann}{ANN}{Artificial Neural Networks}
+\newacronym{tpr}{TPR}{True Positive Rate}
+\newacronym{fpr}{FPR}{False Positive Rate}
+\newacronym{svm}{SVM}{Support Vector Machine}
+\newacronym{mist}{MIST}{Malware Instruction Set}
+\newacronym{lr}{LR}{Logistic Regression}
+\newacronym{pe}{PE}{Portable Executable}
 
@@
 \begin{abstract}
-The abstract should summarize the contents of the paper
-using at least 70 and at most 150 words.
+	Malware, which focuses on harming or subverting a system, has grown exponentially in recent years.
+The regular signature and heuristic methods applied by anti-virus vendors can fall short on keeping up with the rise of new types of malware.
+Other works have shown that machine learning has potential for dealing with the task of malware detection.
+With this in mind, the current report elaborates on the problem of malware, how anti-virus detect it, and proposes the development of a machine learning system to deal with the task of malware detection.
+To design the system to be, the current work will make use of a dataset with 388,513 samples for training and evaluation, focusing on how laboratory and real-world conditions can affect the final result.
 \end{abstract}
 
-\clearpage
-\todo[inline]{Table of contents?}
+\setcounter{tocdepth}{2}
+\tableofcontents
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
+
+% Temporary quotes, move to bib
+% 1 - Attacking Malicious Code: A report to the Infosec Research Council
+% 2 - AV-TEST Security Report 2015/2016 - https://www.av-test.org/fileadmin/pdf/security_report/AV-TEST_Security_Report_2015-2016.pdf
+% 3 - McAfee Labs Threats Report April 2017 - https://www.mcafee.com/us/resources/reports/rp-quarterly-threats-mar-2017.pdf
+% 4 - Symantec Internet Security Threat Report - https://www.symantec.com/content/dam/symantec/docs/reports/istr-22-2017-en.pdf
+% 5 - Scanners of The Year 2000: Heuristics - http://vxheaven.org/lib/adg00.html
+% 6 - Thomas Mitchell - Machine Learning
+% 7 - Some Studies in Machine Learning Using the Game of Checkers - Arthur L. Samuel - http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.368.2254&rep=rep1&type=pdf
+% 8 - Prudent Practices for Designing Malware Experiments: Status Quo and Outlook
+
 \section{Introduction}\label{sec:introduction}
+Malware, which can be referred to by other names like malicious software or malicious code, can be described as \say{any code added, changed, or removed from a software system in order to
+intentionally cause harm or subvert the intended function of the system}\cite{mcgraw:mal_code}.
+Malware can be a piece of code attached to a legitimate program, an independent program, or a combination of both.
+The reasoning that leads to malware creation varies greatly, some are created to demonstrate some vulnerability or concept and do not cause direct harm to systems, while others are used to steal information or render systems useless.
+The wide variety and quantity of now-a-day systems and the amount of sensitive information they store provide numerous opportunities to illegally make a profit out of subverting legitimate systems.
+
+The amount of malware has shown an exponential increase since the early 2000s, from 1.7 million in 2005 to over 600 million by the end of 2016\cite{av-test:report,mcafee:report}.
+According to a 2007 study by McAfee\footnote{McAfee Study Finds 4 Percent of Search Results Malicious, by Frederick Lane, June 4, 2007 [http:\/\/www.newsfactor.com\/story.xhtml?story\_id=010000CEUEQO]} about 4\% of the query results by major search engines lead to potentially dangerous websites.
+A more recent study by Norton in 2010\footnote{Symantec Press Release, July 29, 2010 [https://www.symantec.com/en/aa/about/newsroom/press-releases/2010/symantec\_0729\_02]} showed that at least 10\% of top trending search terms returned malicious results.
+From the contact with malicious software, 1.1 billion identities (i.e.
+user's personal information) have been exposed in 2016 alone\cite{symantec:report}.
+
+Given the wide range and quantity of systems being operated by naive users, in the sense that they lack the knowledge to distinguish malicious and legitimate software, the first line of defense against threats are \gls{hids}, commonly known as anti-virus solutions.
+Their main function is to prevent infections by malicious code, which is achieved by detecting a program as malicious and disallow it from running (i.e.
+quarantining the malicious file).
+
+To detect malicious software, anti-virus solutions depend on two main methods.
+\textit{Signature-based} methods check for the presence of unique byte sequences from known malware.
+For this method to be effective, a database of all malware signatures must be available, which leads to an obvious shortcoming: only known malware is detected\cite{gryaznov:scanners}.
+New and unknown malware is not detected until vendors analyze it and update their clients.
+To improve the effectiveness of anti-virus and overcome the problem in aforementioned methods, \textit{heuristic-based} methods are employed.
+These apply a set of rules to decide whether a program is likely to be malware or not.
+Like before, a database of rules must be available, but unlike signature methods, which provide exact matching on malware, heuristic methods are capable, to some degree, of generalizing malware features, hence detecting new and unknown malware.
+Although more sophisticated, heuristic-based methods need a database of rules that must be defined \textit{a priori}, just like signature methods.
+
+As mentioned before, signature methods only detect known malware, which translates to a very low false positive rate (i.e.
+goodware wrongly classified as malware), while having a high detection rate (i.e.
+correctly classified malware from the total malware), assuming the signature database is on par with malware samples, which at the rate new malware emerges becomes a hard task.
+For heuristic methods new and unknown malware can be detected due to similarities to known malware, translating to a higher detection rate with less information, but with the downside of raising the false positive rate, as benign programs (goodware) can be flagged malicious due to characteristics that fit into malware rules\cite{gryaznov:scanners}.
+To improve the effectiveness of anti-virus, both methods must complement each other.
+
+Having the need to either create signatures for new malware or define a set of rules that describe malware characteristics becomes a cumbersome task for the increasing malware numbers.
+To try and improve the detection task of anti-virus software, the current work proposes a \gls{ml} approach to malware detection as an evolution to heuristic-based methods.
+
+The machine learning field studies the development of computer programs that automatically improve with experience.
+A program is given a task and an experience, if its performance at the task improves with the experience given, then the program is learning the task\cite{mitchell:ml}.
+In the proposed work, the task at hand is to distinguish malware from goodware (i.e.
+classify samples) and the experience is a wide variety of samples, both goodware and malware.
+The final result is a program that performs well for the given task (i.e.
+successfully learns to distinguish goodware from malware).
+
+The \gls{ml} field has been around for multiple decades, with some of the earlier works dating back to 1959 by Arthur L.
+Samuel\cite{samuel:studies}, but only in recent years has the field grown in popularity.
+Its recent growth can be accredited to two main factors: availability of large data sets and computation power.
+Larger data sets provide more experience for a program to learn from, but also require more processing power.
+
+The growing number of malware samples provide a good opportunity to make use of \gls{ml} for malware detection.
+This approach is not new and has provided high detection rates with low false positive rates\cite{miller:rev_int,nissim:al_pdf,perdisci:behavior,rieck:dynamic,schultz:data_mining}.
+However, many of the resulting research is done under laboratory conditions that do not represent real world scenarios.
+Some of those conditions have been studied in \cite{rossow:practices} and include the correctness of datasets, transparency and realism.
+Alongside the application of \gls{ml} to detect malware, the proposed work will study how laboratory conditions differ from real world one's and how effectiveness is affected.
+
+To provide a solution that makes use of \gls{ml} to detect malware and its performance under laboratory \textit{vs.} real world conditions, the proposed work will follow an approach that is divided into three stages:
+
+\begin{itemize}
+	\item \textit{Data collection}: experience from which a program can learn must be available, as such the first step must ensure the collection of samples that represent both malware and goodware examples.
+The current work will take advantage of the Malwr\cite{tool:malwr} service, which provides both static (e.g.
+library imports) and dynamic (e.g.
+library call order during runtime) details about a sample.
+The total number of available samples surpasses 388 thousand, dated from April 2013 to October 2016.
+	\item \textit{Data labeling}: before training a classifier to detect malware, samples must be labeled as either benign or malign.
+The Malwr\cite{tool:malwr} service provides details on how multiple anti-virus solutions classify a sample, hence the approach is to use said classification to determine the legitimacy of a sample.
+This step is not linear, as different anti-virus classify samples differently, introducing edge cases (e.g.
+50\% classify malware, the other 50\% classify goodware).
+To overcome this, the current work must develop a metric to deal with such cases.
+	\item \textit{Classifier implementation}: at this stage the collected data is ready to be used as experience for the program.
+Using static and dynamic details together with the labeling of the samples, a classifier can be trained to distinguish between malware and goodware.
+Alongside the classifier creation, this phase must also measure the effectiveness of the solution.
+For this work's scope, the effectiveness is to be measured under laboratory and real world conditions.
+\end{itemize}
+
+In sum, the work to be done as to verify the application of \gls{ml} methods for the task of detecting malware is: collection of samples, labeling of collected samples and finally usage of the resulting dataset to achieve a malware classifier.
+
+The following sections provide deeper analysis on the subject of using \gls{ml} to classify malware, as well as the choices to be made and the reasoning behind them.
+Section \ref{sec:related_work} describes related work that both influenced and motivated the current work.
+Section \ref{sec:solution} details the intended approach, the tools to be used, how the work is to be presented and the already accomplished work.
+Section \ref{sec:conclusion} summarizes and closes the report.
+
+%\todo[inline]{falar dos antivirus para combater isto, falar nas tecnicas basicas: signature based, heuristic based.
+tirar partido de ML para estender metodos heuristicos e melhorar a classificaçao de malware.}
+%
+%
+%\todo[inline]{one liner do que é malware (quote), quantificar (quote), necessidade de detectar malware mais rapidamente (automatizar/generalizar).
+%
+%Avanços na detecçao do malware, pouca coisa. Aplicaçao de ML neste tipo de problema.
+%
+%Contributos do trabalho para este problema, desenvolver um classificador que funcione no mundo real e ajude a resolver o problema.
+%
+%1. Dados recolhidos, o que posso extrair dos dados\\
+%2. Definir o que é malware (thresholds, consenso)\\
+%3. Comparaçao de metodologias (estatatica/dinamica)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%% RELATED  %%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related Work}\label{sec:related_work}
+
 % #1 Semantics-Aware Malware Detection
@@
 % #10 http://members.chello.at/erikajo/vnc99b2.txt
-Malware, which can be referred to by other names like malicious software or malicious code, is not easily described by a single definition.
-For example, Christodorescu et al.[quote 1] simply describe malware as \say{a program that has malicious intent}, which in turn begs the question, what is malicious intent?
-McGraw and Morrisett[quote 2] give a more concrete definition by saying \say{Malicious code is any code added, changed, or removed from a software system in order to
-intentionally cause harm or subvert the intended function of the system}.
-Both definitions give an overall understanding of \textcolor{red}{(on?)} what is malware, but to better understand the definition, one can briefly look at the history of malware.
+% #11 Learning to Detect and Classify Malicious Executables in the Wild
+% #12 The Evolution of Viruses and Worms
+% #13 http://cert.europa.eu/static/SecurityAdvisories/2017/CERT-EU-SA2017-012.pdf
+% #14 https://www.virusbulletin.com/uploads/pdf/magazine/2002/200201.pdf
 
-\todo[inline]{Clean previous paragraph. Remove bellow}
-
-The term malware was first used by Yisrael Radai in 1990[quote 3], before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983[quote 4].
-Before Cohen, John Von Neumann had already done some similar academic work that closely relates to computer viruses, namely \say{Theory of self-reproducing automata}, published in 1699[quote 5].
-The term computer virus is one of the many types of malicious software, but given that it predates the term malware, it is not uncommon to see both terms used as interchangeably.
-
-Even before the formalization of computer virus by Cohen in 1983, one of the first documented computer virus, although experimental, was developed by Bob Thomas at BBN in 1971[quote 6].
-The program was named Creeper and consisted of a self-replicating program that would print a message and move between the nodes of ARPANET.
-
-Another malicious software worth mentioning, given it was one of the first known virus that spread \say{in the wild} (i.e.
-outside the computer system or laboratory in which it was developed), is Elk Cloner.
-Created by Rich Skrenta in 1982[quote 7], Elk Cloner spread by floppy disk and had a payload that displayed Skrenta's poem after every 50th use of the infected system.
-
-\todo[inline]{Up to here}
-
-Since then, the amount of malware has grown at an alarming rate.
-According to AV-TEST[quote 8], an independent security software group, up to 2012 the amount of total malware per year would not surpass 100 million, whereas 2013 almost reached 200 million, followed by more than 100 million increase per year, with almost 600 million total malware before the end of 2016.
-
-\todo[inline]{Clean previous paragraph. Remove bellow}
-
-Even though the term malware generalizes the malicious intent of a piece of software, one can classify malware further based on its purpose and behavior.
-Given the high amount of malware, there is also a high variety of malware, based on purpose and behavior.
-
-From the general purpose and behavior of a malware sample, it can be classified as belonging to one or more malware categories (or classes).
-Examples of categories are virus, worms, trojans and adware, to name a few.
-Having the ability to classify malware into a more restrict subset is helpful, but there is still a need to distinguish between samples of the same category, by giving more detailed names to malware.
-
-One attempt at creating a naming convention was made in 1991, by the Computer Antivirus Research Organization (CARO).
-The convention states the name should consist of four parts, separated by a dot.
-The four constituents are the family name, group name, major variant name and minor variant name, displayed in that order.
-An optional modifier could also be appended to the name with a colon.
-The problem with the aforementioned convention is that it only covered computer viruses, which were more prominent at the time.
-
-An attempt at improving CARO's naming convention was made in 1999, by Gerald Scheidl[quote 10]. His proposal added two other constituents, platform and type, as a prefix to CARO's naming.
-
-Later attempts to standardize malware naming were made, none of which resulted in a standard widely adopted.
-Due to this, antivirus vendors use their own naming scheme, which are variants of the mentioned schemes.
-
-Since the malware naming scheme is not standardized, and given the amount and variety of malware, different family names with the same behavior emerge from different antivirus vendors.
-The name redundancy is worsen by how different vendors detect and classify malware.
-
-With the raise of malware amount and diversity, being able to identify and classify malware becomes essential in order to protect systems.
-
-Before the late eighties, given the low amount of malware, the first antivirus programs were designed to find and remove a handful of malware, mainly through signature based methods, which detect malware by looking for specific patterns that appear in known malware.
-
-In the late eighties and nineties the first antivirus were released, along with more sophisticated detection methods that make use of heuristics, which also detect malware by also looking for patterns, but at a higher level of abstraction.
-
-In more recent years, and given the increment of malware amount and variety, signature and heuristic methods fall short when keeping up with new types of malware.
-The mentioned methods work best for samples of known malware or new malware with small variants of known malware.
-This is because these methods look for known patterns.
-If a new type emerges, which pattern's are not known, it is not flagged as malicious.
-
-Given this limitation, more advanced methods like behavioral analysis are applied together with the previous mentioned methods.
-Behavioral analysis takes into account the actions performed by the malware, introducing dynamic information, from which patterns can be extracted and again matched against known patterns.
-
-Having to run samples in order to extract dynamic information poses the problem of infecting the system running the samples.
-One way to prevent this is to use a sandbox environment, which is a security mechanism to isolate running programs.
-Using a sandbox environment closely relates to computer virtualization.
-
-To provide users the possibility to check files without running the risk of being infected, online services allow users to submit files which are then run in a sandbox environment.
-
-An example of such service is malwr.
-Malwr takes user submissions and runs them in a sandbox environment, specifically Cuckoo Sandbox, providing static and dynamic information about the sample, as well as checking if the file is known in VirusTotal, a similar service that also checks if the submission is known in different antivirus solutions.
-
-% VirusScan (from McAfee) and NOD (from ESET) were first released in 1987; AntiVir (from Avira) and Avast Antivirus were released in 1988.
-Norton Anti-Virus (from Symantec) and F-Secure antivirus were released in 1991.
-
-\todo[inline]{re-focus on high amount of malware, difficulty of having to run all samples and check}
-\todo[inline]{propose machine learning as a solution to deal with the high amounts of data}
-\hfill \break
-
-\todo[inline]{Motivation}
-\begin{itemize}
-	\item Define Malware, problem with that, subjectivity
-	\item Quantify malware in recent years (quote vendors/get stats from header analysis)
-	\item What is being done about malware?
-	\item How is malware identified?
-	\begin{itemize}
-		\item Types of strategies
-		\item Is it effective? efficient?
-	\end{itemize}
-	\item Being able to identify malware as early as possible would help?
-	\item Introduce services like malwr.com/virustotal.com
-	\item Introduce Machine Learning capabilites
-	\item Automation of detection/classification
-\end{itemize}
-
-%From an external view, adware and ransomware will provide a noisy output to the user, given their function to deliver unwanted advertisements (for adware) and request data ransom (for ransomware).
-Spyware and rootkit types, on the other hand, will provide little to no user interaction, given their function to spy on user activity (for spyware) and conceal processes, files or system data (for rootkit).
-%
-%From an internal view, adware and ransomware will behave similarly, since both need interfaces that facilitate user interaction (e.g.
-create and show dialogs), in contrast to spyware and and rootkit, which thrive on concealing actions (e.g.
-run hidden processes).
-
-% Quote this for more types of malware
-% https://support.symantec.com/en_US/article.HOWTO54185.html
-
-\todo[inline]{Problem}
-\begin{itemize}
-	\item Remove the human factor from detection/classification to speed up process.
-	\item Introduce lack of temporal consistency here?
-	\item Talk how related work creates own corpus and low corpus size/variation?
-	\item Refer stats from header analysis? Time vendors take to take knowledge of samples/Time until new malware is detected in most vendors
-	\item Reinforce time to classify malware?
-\end{itemize}
-
-\todo[inline]{Hypothesis}
-\begin{itemize}
-	\item Would a more automated approach, based on ML, facilitate malware detection/classification?
-	\item Dynamic/Static analysis to detect and classify malware
-	\item Temporal consistency is a factor that should be taken into account?
-
-\end{itemize}
-\todo[inline]{Objectives/Contributions}
-\begin{itemize}
-	\item Usage of real world analysis
-	\item Try to apply ML to PE32 analysis from malwr.com (Talk about this here?)
-	\item Contribute with a corpus of analysis from malwr.com
-	\item Maintain a temporal consistency
-
-\end{itemize}
-\todo[inline]{Document outline}
-The following sections are organized as follows.
-Section \ref{sec:related_work} describes related work and its influence on the current work.
-Section \ref{sec:solution} describes the planned approach for solving the problem at hand.
-Section \ref{sec:conclusion} closes the report and references the work done so far.
-
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%% RELATED  %%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec:related_work}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Background}\label{subsec:background}
-The term malware was first used by Yisrael Radai in 1990[quote 3], before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983[quote 4].
-Before Cohen, John Von Neumann had already done some similar academic work that closely relates to computer viruses, namely \say{Theory of self-reproducing automata}, published in 1699[quote 5].
-The term computer virus is one of the many types of malicious software, but given that it predates the term malware, it is not uncommon to see both terms used as interchangeably.
+Malware, which can be referred to by other names like malicious software or malicious code, can be described as \say{a program that has malicious intent}\cite{christodorescu:semantics}, or more detailed as any code added, changed, or removed from a software system in order to intentionally cause harm or subvert the intended function of the system\cite{mcgraw:mal_code}.
 
-Even before the formalization of computer virus by Cohen in 1983, one of the first documented computer virus, although experimental, was developed by Bob Thomas at BBN in 1971[quote 6].
-The program was named Creeper and consisted of a self-replicating program that would print a message and move between the nodes of ARPANET.
+The term malware was first used by Yisrael Radai in 1990\cite{elisan:malware}, before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983\cite{cohen:virus}.
+Given term computer virus predates the term malware, it is not uncommon to see both terms used interchangeably.
+For the purpose of this work the term malware is adopted, for reasons explained in the following paragraphs.
 
-Another malicious software worth mentioning, given it was one of the first known virus that spread \say{in the wild} (i.e.
-outside the computer system or laboratory in which it was developed), is Elk Cloner.
-Created by Rich Skrenta in 1982[quote 7], Elk Cloner spread by floppy disk and had a payload that displayed Skrenta's poem after every 50th use of the infected system.
+To the exponential growth of malware\cite{av-test:report}, the ability to distinguish between malware samples is crucial.
+In the biological world, different types of infections reckon distinctive disinfections, the same applies to the digital world.
+The infection method, behavior and subsequently purpose of malware samples vary, hence being able to make a distinction facilitates prevention and disinfection by anti-virus solutions.
 
-Over the years, malware has grown in size and diversity, which introduces the need to be able to distinguish malware, not only to facilitate its removal, but also to provide safer mechanisms of defense.
+There exists no ground truth when it comes to distinguishing malware, leading to subtle differences on the classification and naming of the same malware instance by different parties.
+To facilitate the reader's understanding of the topic, the current work will focus on defining malware types based on their propagation and purpose.
 
-Based on the malware general purpose and behavior, a high level classification of the malware can be made by splitting malware into one or more categories (or classes).
-Examples of such classes are viruses, worms, trojans and adware, to name a few.
-This high level classification is helpful, but lacks detail, like target platform, origin or version.
-In order solve this issue, attempts have been made to create a naming convention for malware.
+With respect to the propagation method, malware can be split into three classes\cite{kolter:learning}.
 
-One of the first attempts was made in 1991, by the \gls{caro}\footnote{http://www.caro.org/articles/naming.html}, which states the name should consist of four  obligatory parts, separated by a dot, and an optional fifth part.
-The four main constituents are the family name, group name, major variant and minor variant name.
-The optional fifth part is a modifier and could be appended to the name with a colon.
-\todo[inline]{Describe the parts?}
+\paragraph{Virus.}\footnote{This work uses virus as a subset of malicious code, hence the choice of using the term malware over virus.}This type of malware inherits its name from resembling to biological viruses.
+A virus is usually composed by two main subroutines.
+The first subroutine is responsible for infecting other programs by attaching the virus code, while the second subroutine is the actual malware payload that contains the malware purpose (i.e.
+virus payload)\cite{chen:evolution}.
+Viruses are attached to programs and propagate when an infected program is run, hence depending on other programs (as hosts) and user interaction.
+For viruses to infect new systems, an infected file must be carried between them by a user (e.g.
+USB pen, email attachment), reinforcing the user's role in this propagation method.
 
-Although the aforementioned naming convention is well defined, it is aimed at naming viruses, lacking differentiation for other classes as well as target platform.
-In order to solve this, an improvement was proposed in 1999 by Gerald Scheidl\footnote{http://members.chello.at/erikajo/vnc99b2.txt} that suggested the addition of a prefix that consisted of the platform and type.
+\paragraph{Worm.} The increasing number of network connected devices facilitates the application of worms to propagate malware.
+A worm shares the self-replicating ability of viruses, but discards user interaction and the need for a host.
+This is possible by making a worm an independent program that exploits networks to find vulnerable systems to infect with a copy of themselves\cite{chen:evolution}.
+Similarly to viruses, worms can contain additional payload to perform malicious actions on infected systems (e.g.
+Blaster worm exploited a vulnerability in RPC for propagation while its payload flooded a Microsoft domain\footnote{W32/Blaster worm, CERT, August 11, 2003 [http://www.cert.org/historical/advisories/CA-2003-20.cfm]}).
 
-A more recent attempt at creating a naming convention was made during the Virus Bulletin Conference in 2001\footnote{https://www.virusbulletin.com/uploads/pdf/magazine/2002/200201.pdf}, with the idea that malware naming should consist on three levels of information: malware name, classification and full text description.
-\todo[inline]{Again, describe parts?}
+\paragraph{Trojan.} While viruses and worms focus on self-propagation, trojans focus on deceiving users into executing them, disregarding propagation.
+Trojans try to appeal users with some useful functionality as to allure into running the program\cite{szor:art}.
+By hoaxing users into running them, trojans bypass the need for a host (in the virus case) or an exploit (in the worm case) to perform malicious actions.\\
 
-None of the referenced naming conventions is widely adopted, with antivirus vendors having their own naming conventions, that although similar, vary in some way or another.
-For example, Trend Micro\footnote{http://docs.trendmicro.com/all/ent/smex/v11.0/en-us/smex\_11.0\_olh/secrisk\_malware\_name.html} and Symantec\footnote{https://www.symantec.com/security\_response/virusnaming.jsp} use a prefix (which identifies the malware class), threat name (i.e.
-family) and suffix (which indicates a variant of the threat).
-Whereas Microsoft\footnote{https://www.microsoft.com/security/portal/mmpc/shared/malwarenaming.aspx} tries to follow the \gls{caro}'s convention, by using a type, platform, family, variant and information.
-Avira\footnote{https://www.avira.com/en/support-malware-naming-conventions} on the other hand, takes a more simplistic approach by using a prefix (which identifies the malware type) and a threat name.
+These three classes are not exclusive and can be used in conjunction to facilitate propagation (e.g.
+a virus that relies on a trojan to start propagating).
+Given the inexistent standards for malware naming, some literature\cite{szor:art} define worms as a subset of viruses, as some worms start by attaching to files.
+With the general infection methods layered out, the following will succinctly describe the general naming when referring to the malware purpose (i.e.
+its payload).
 
-Although the naming and its constituents are similar between vendors, no convention is set, hence the threat names given to malware varies, increasing the redundancy for the same malware.
-According to an article published by Symantec in 2002\footnote{https://www.symantec.com/connect/articles/virus-any-other-name-virus-naming-practices}, based on a list of 300,000 unique names, the name redundancy is quantified as an average of four different names for the same malware.
+Given the high numbers of malware multiple purposes imply multiple names, as such the following lists some of the most commonly heard names and definitions as given by McAfee\cite{mcafee:glossary}:
+
+\begin{itemize}
+	\item \textit{Adware}: Program that automatically displays advertisement to the user. Most do not cause direct harm.
+	\item \textit{Backdoor}: Installs or takes advantage of an unknown entry point to a user's system, allowing remote control of a system.
+	\item \textit{Dropper}: Program that facilitates the introduction of other malware instances into a system.
+	\item \textit{Keylogger}: Program that monitors a user's input (e.g. keyboard strokes) to steal private information.
+	\item \textit{Password Stealer (PWS)}: Program that specifically targets a user's personal information, like usernames and passwords.
+	\item \textit{Ransomware}: Program that restricts access to a system/files (usually by encryption) and demands a ransom to allow a user his access.
+	\item \textit{Rootkit}: Program designed to hide the existence of other (malicious) programs, hindering detection of malicious applications.
+	\item \textit{Scareware}: Program that fakes the existence of malware and tricks the user into paying for disinfection or into downloading real malware.
+	\item \textit{Spyware}: Program that spies on a user's activity through multiple vectors (e.g. taking prints of a user's desktop/webcam).
+\end{itemize}
+
+The ability to name a malware given its propagation method and purpose facilitates detection and disinfection, but given the high quantities of malware, being able to particularly discriminate an instance is helpful.
+To that purpose, attempts have been made to create naming conventions for malware.
+
+One of the first attempts was made in 1991, by the \gls{caro}\cite{caro:naming}.
+This attempt was aimed at viruses only and focused on discriminating and relating virus samples.
+The convention defines an hierarchical naming with three levels.
+The base of the hierarchy is the family name, where structurally similar virus should be grouped.
+To further distinguish instances, a group name and major variant name is applied.
+To differentiate versions of the same instance and provide additional information about a sample, minor variant name and modifier fields are to be used.
+In sum, a virus name should look as following:
+
+\begin{center}\texttt{family.group.major.minor:modifier}\end{center}
+
+To enable differentiation from other types of malware, an improvement to \gls{caro}'s naming was proposed in 1999 by Geral Scheidl\cite{scheidl:naming}.
+Scheidl's proposal included a prefix that consisted of the platform targeted by the malware and a type (i.e.
+previously defined classes and names), rendering a more in-depth name:
+
+\begin{center}\texttt{platform.type/family.group.major.minor:modifier}\end{center}
+
+In the Virus Bulletin magazine of January 2002\cite{virus_bulletin} the naming problem is again addressed, referencing that only a small number of vendors apply \gls{caro}'s names and that the lack of cooperation between anti-virus vendors lead to multiple aliases for malware names.
+Again a naming convention is proposed with focus on detail by using three levels of information: malware name, classification and full text description.
+
+Given that no naming convention is widely adopted, anti-virus vendors make use of their own naming that although similar, vary slightly.
+This variation raises the redundancy for instances of the same malware as shown by an article published by Symantec in 2002\footnote{A Virus by Any Other Name: Virus Naming Pratices, Costin Riau, June 2, 2002 [https://www.symantec.com/connect/articles/virus-any-other-name-virus-naming-practices]}, stating that the same malware instance has an average of four different names.
+
+The naming problem, although not directly related to the current work, must be taken into consideration as to not influence the classification results.
 
@@
 %%% Malware detection
-\subsection{Malware Detection}\label{subsec:malware_detection}
-In the early days of malware, due to low amount and variety, antivirus could target malware specifically (e.g.
-Reaper was created to remove Creeper), but with an estimate of more than 600 million total malware by the end of 2016, targeting specific malware becomes increasingly more difficult, hence being able to identify and classify malware efficiently is crucial to protect systems.
+\subsection{Malware Detection}\label{subsec:mal_detec}
 
-What follows is a quick rundown on different detection techniques developed over the years[quote 7].
+In section \ref{sec:introduction} a brief description of malware detection methods was given.
+This section will provide a deeper description of detection methods, from the simplest techniques to more complex ones.
+The following is based on the work by Peter Szor, \textit{The Art of Computer Virus Research and Defense}\cite{szor:art}.
 
-\subsubsection{First-Generation Scanners.} These methods of detection work at a fairly simple level and use static methods of detection. The applied principles include:
-\begin{itemize}
-	\item String scanning, where a sequence of bytes that is common in a malicious program but not in a regular program is matched against the suspicious file.
-	\item Wildcards, similar to string scanning, but instead of a fixed sequence of bytes, regular expressions can be used in the sequence of bytes, this allows variation and byte skipping when matching.
-	\item Mismatches, also similar to string scanning, in which a sequence of bytes is taken as reference, but a number of bytes $N$ is given allowing up to $N$ bytes to vary in any position of the string (e.g.
-the string \say{\texttt{01 02}} with $N=1$ will match \say{\texttt{XX 02}} or \say{\texttt{01 XX}}).
-\end{itemize}
+To detect malware instances, anti-virus applications are used.
+These can also be called scanners, as early detection methods simply scanned for byte sequences to detect malware.
+Before going into details on the different detection techniques, it is worth mentioning that scanners can be used as \textit{on-demand} or \textit{on-access} scanners.
+On-demand anti-virus are executed only at the user's request.
+On-access anti-virus are memory-resident, meaning they load as an application and intercept actions related to file and disk access.
+This makes it so that when files are opened, created or closed, the anti-virus scans the changed object.
 
-The described principles can be applied with slight variations that improve the performance of the scanner, such as matching on the beginning and end of the file (top-and-tail scanning) or matching at a certain offset (bookmarks).
+\subsubsection{First-Generation Scanners.} These type of scanners work at a fairly simple level and focus searching for known byte sequences (i.e.
+known to be malware sequences).
+The applied principles include:
 
-They all require a knowledge base with known patterns that must be kept up-to-date in order to detect newer malware.
+\paragraph{String Scanning.} A sequence of bytes that is common in a malicious program but not in a regular program is matched against a suspicious file.
 
-\subsubsection{Second-Generation Scanners.} These methods improve on the previous generation by doing a static analysis at a higher level of abstraction.
-Where first-generation scanners would match sequence of bytes, second-generation scanners also look at the representation of the bytes.
+\paragraph{Wildcards.} This method is similar to string scanning, but instead of searching for a fixed sequence of bytes, a more flexible matching is made, allowing to skip bytes or byte ranges.
+The method can also include the use of regular expressions for matching.
+
+\paragraph{Mismatches.} Also similar to string scanning, yet there can be $N$ number of bytes in the string that can take any value and in any position (e.g.
+the string \say{\texttt{01 02}} with mistmatch of 1 matches \say{\texttt{01 AA}} and \say{\texttt{AA 02}}).
+
+\paragraph{Generic Detection.} This method typically applies both wildcards and mismatches to create generic strings that can scan for several variants of a family of malware.
+
+\paragraph{Hashing.} This term is not a method by itself, but techniques that speed up searching algorithms, enhancing previous methods.
+
+\paragraph{Bookmarks.} This method is both used to provide more accurate detections and disinfections. Bookmarks provide specific locations for where strings are to be searched (i.e. offsets).
+
+\paragraph{Top-and-Tail Scanning.} Like hashing, this principle is used to speed up detection by scanning the start and end of a file, instead of the entire file.
+The technique is mainly effective when dealing with viruses that prefix or append data to files.
+
+\paragraph{Entry-Point and Fixed-Point Scanning.} This technique also improves the speed of anti-virus scanners.
+By taking advantage of the structure of binary objects this method can start scanning at the entry-point of an object, or alternatively use the entry-point plus some fixed size as a fixed-point from where to start scanning.
+
+\paragraph{Hyperfast Disk Access.} This last technique also provides a speed improvement.
+It does so by bypassing operating system level \gls{api} to read the disk directly with the \gls{bios}.
+It is worth noting that this technique is hard to implement nowadays due to the high variety of file systems and disk controllers.\\
+
+The described methods, excluding the ones that focus on speed improvement, can be assigned to the signature-based type of methods, as they depend on searching some previously defined byte sequence to be effective.
+Their effectiveness depends on an updated knowledge base with the strings to be matched.
+
+\subsubsection{Second-Generation Scanners.} These methods improve on the previous generation by doing an analysis at a higher level of abstraction.
+Where first-generation scanners would match sequences of bytes, second-generation scanners also look at the bytes representation.
 The applied principles include:
-\begin{itemize}
-	\item Smart scanning, where the Assembly representation of the bytes is taken into account and instructions that have no effect (e.g.
+
+\paragraph{Smart Scanning.} This method takes the assembly representation of the bytes into account and instructions that have no effect (e.g.
 \texttt{NOP}) can be skipped.
 The resulting signature is smaller and more easily detects mutations.
-The same principle could be applied to textual malware (e.g.
+The same principle can be applied to textual malware (e.g.
 scripts), where extra white spaces have no effect.
-	\item Nearly exact identification, where instead of relying on a single string for detection, double-string detection can be used.
-An alternative to double-string detection could also be based on a checksum range selected from the malware body, giving a higher accuracy by taking into account a larger portion of the malicious file.
-	\item Exact identification, which can exactly identify malware variants, is usually used in combination with first-generation methods and instead of using a single range like nearly exact, uses multiple ranges to calculate checksums.
-To be accurate, this method must only take into account constant bits of the malicious file, and given that more ranges are used it is slower then the previous ones.
-\end{itemize}
 
-Like in the previous generation, a knowledge base must be used and kept up-to-date to identify newer malware.
+\paragraph{Skeleton Detection.} This method is specifically useful in detecting textual malware (e.g.
+macros) families.
+By parsing the macro statements line by line, the scanner can drop nonessential statements.
+The result can be seen as a skeleton of the full macro that contains only the relevant code that appears in malware.
+
+\paragraph{Nearly Exact Identification.} This method provides a more accurate way to detect malware.
+First-generation scanners base detection mainly on a byte sequence, whereas nearly exact identification makes use of double-string detection.
+This is specifically useful to identify same family malware.
+Another nearly exact identification method applies a checksum to a range from the malware body, the resulting checksum can then be used to detect and disinfection files.
+
+\paragraph{Exact Identification.} This method enhances the previous method by adding the ability to differentiate malware variants.
+It does so by using as many checksum ranges as needed to identify variants of the same malware, improving the accuracy and facilitating disinfection.\\
+
+Like in the previous generation, these methods fit into signature-based, hence still being highly dependent on an updated knowledge base.
 
 \subsubsection{Algorithmic Scanning Methods.} These methods are not so focused on direct matching of signatures, but on the routines implemented in the scanner that allow the detection of malware.
-Early implementations of algorithmic scanning were hard-coded in the antivirus, but with different types and families emerging, the hard-coded algorithms could easily become obsolete as such.
+Early implementations of algorithmic scanning were hard-coded in the antivirus, but with different types and families emerging, the hard-coded algorithms could easily become obsolete.
 To overcome this, vendors would introduce more specific routines for malicious programs when updating the antivirus.
+Examples of these methods are filtering, where signatures are only applied if an object meets a particular filter; static decryptor, which detects encrypted malware; and X-ray, which tries to decrypt malware that use simple encryption methods (e.g.
+\say{\texttt{XOR}} with constant key).
 
-Being able to update the antivirus algorithms frequently enables the vendors to have a faster response to new malware by developing detection routines for new malware and pushing them to end users.
+Algorithmic scanning methods are more related to heuristic-based methods, since the focus is developing rules to facilitate malware detection and removal.
+These methods also allow for vendors to push new detection algorithms to clients without having to update the entire anti-virus.
 
-\subsubsection{Code Emulation.} This detection technique uses dynamic analysis to detect malware by simulating code execution.
+\subsubsection{Code Emulation.} This detection technique uses dynamic analysis (i.e.
+analysis during runtime) to detect malware by simulating code execution.
 A virtual machine is used to simulate the system's components, thus the malicious files are encapsulated and no code is executed by the real processor.
-
-Code emulation can be faster than using algorithmic scanning methods when dealing with encrypted/obfuscated malware, as the code can be emulated and the decrypted/deobfuscated code is much easier obtained.
+Code emulation can be faster than using algorithmic scanning methods when dealing with encrypted/obfuscated malware, as the code can be emulated and the decrypted/deobfuscated code is much easier to obtain.
+The downside to emulation comes from a time perspective, as more resources are needed to emulate code.
 
@@
 Examples of such characteristics are code execution starting in uncommon sections, suspicious code redirection or incorrect size of code in header.
-
-When checking a suspicious file, the presence or absence of multiple characteristics is weighted to assert if the file is deemed malicious or not.
+Using the previously described methods, heuristic analysis can take into account different characteristics that are present or absent to assert if a file is deemed malicious or not.
 The upside of this technique is that it allows for never seen malware to be correctly flagged, but brings the downside of a higher amount of false positives.
@@
 
-This technique allows to extract dynamic information without compromising the real system, which can then be used to both detect if a suspicious program is malware, by applying the aforementioned techniques, and to help analyze new types of malware.
-
-The downside of using sand-boxing comes in the form of compatibility issues and in the possibility of malware breaking out of the sand-box due to vulnerabilities in the virtualization software.\\
+This technique allows to extract dynamic information without compromising the real system, which can then be both used to detect if a suspicious program is malware, by applying the aforementioned techniques, and to help analyze new types of malware.
+This technique demands more resources then previous ones and may not be ideal to use inside an anti-virus application.
+Other downsides of sand-boxing come as compatibility issues, the possibility of malware detecting virtualization, which can hinder analysis, or even if the virtualization software is vulnerable and malicious code can escape the sandbox into the host machine.\\
 
@@
 
-\todo[inline]{Talk about cuckoo here?}
+\subsection{Cuckoo Sandbox and Malwr}\label{subsec:cuckoo}
 
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Machine Learning
-\subsection{Machine Learning}
+Building on the last described method of sand-boxing, this section will describe a specific implementation of sand-boxing, namely Cuckoo.
+Cuckoo Sandbox is a malware analysis system created by Claudio Guarnieri\cite{tool:cuckoo}.
+
+When analyzing a malware sample, an analyst will have to create his own virtual environment, transfer the sample into that environment and manually interact with the sample to see how it behaves and what it does.
+Cuckoo facilitates analysis by automating this process.
+The Cuckoo system will take the sample to analyze and run it inside a predefined virtual environment (e.g.
+Windows XP).
+The system is then capable of simulating user interaction, take screenshots of the environment, extract static and dynamic information about a sample, monitor network traffic and perform memory analysis.
+The time spent doing the analysis can be configured by the user.
+Upon finishing the analysis the system provides reports about the sample.
+Examples of what is contained in these reports include: what functions were imported from which \gls{dll}, what processes were created and the \gls{api} calls order and which files were created/modified.
+
+The creator of Cuckoo and one of its developers, Alessandro Tanasi, also created a free malware analysis service, Malwr\cite{tool:malwr}.
+This service implements the Cuckoo system and allows users to upload their own files, which are then run with Cuckoo.
+The service is presented as a web application that takes in samples and returns a report with Cuckoo's results.
+An example of the behavioral analysis (i.e.
+created process and \gls{api} calls) can be seen in figure \ref{fig:malwr_sample}.
+
+\begin{figure}[h]
+	\includegraphics[width=\textwidth]{malwr_sample}
+	\caption{Example of a behavioral report from Malwr}
+	\label{fig:malwr_sample}
+\end{figure}
+
+For the purpose of detecting malware via \gls{ml}, the reports provided in Malwr supply various information that can be used to fulfill the current work.
+The whole website content is provided under \gls{cc} CC-BY-NC-SA 3.0\cite{cc-by-nc-sa} license, allowing the use of the data for this work's purpose, as long as credits are maintained.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%% Machine Learning
+\subsection{Machine Learning}\label{subsec:ml}
 
@@
 With the technological advances that provide higher computation power, computer science fields like \gls{ai} have greatly improved in recent years.
-One of the \gls{ai} fields that as benefit from the technological advances is \gls{ml}.
+One of the \gls{ai} fields that has benefit from the technological advances is \gls{ml}.
+The current subsection will provide a more detailed description of \gls{ml} and its use for the current work.
 
-\gls{ml} is a subset of \gls{ai} in which algorithms are applied to learn and make predictions on data, or more formally defined by Thomas Mitchell\cite{mitchell:ml}:\\
+\gls{ml} is a subset of \gls{ai} in which algorithms are applied to learn and make predictions on data, or more formally defined by Tom Mitchell\cite{mitchell:ml}:\\
 \say{\emph{A computer program is said to learn from experience E with respect
@@
 A noteworthy nuance about the given definition, and the \gls{ml} field in general, is that algorithms are not designed to deal with specific problems.
-Meaning that as long as a learning problem is well defined through $T$, $P$ and $E$, an algorithm based on these three features is able to learn, but not necessarily solve, a problem.
+Meaning that as long as a learning problem is well defined through $T$, $P$ and $E$, an algorithm based on these three features is able to learn, but not necessarily solve a problem.
 
-After defining a learning problem, to obtain an useful outcome (i.e.
+After defining a learning problem, to obtain a useful outcome (i.e.
 model) one or more learning algorithms are applied.
@@
 experience) with input-output pairs and it must learn how to map a never seen input into the correct output.
-Within supervised learning, a learning problem can be further classified into a classification problem, if the output is set of finite values (e.g.
+Within supervised learning, a learning problem can be further classified into a classification problem, if the output is a set of finite values (e.g.
 is it going to rain tomorrow?), or into a regression problem, if the output is continuous (e.g.
@@
 
-Another important part of \gls{ml} is the performance measure $P$, which dictates how well the program learned the problem.
-Performance measuring is related to the task $T$, meaning the measurement must clearly reflect on the task (e.g.
-if the task is playing checkers, performance should be percent of games won against opponents).
+The proposed work will fit into the supervised learning category, with the purpose of creating a classifier to assert whether a given sample is malware or not.
+This problem can be seen as whether a pure classification problem, with a binary output, or as a regression problem, with the likelihood of a sample being malware.
 
-For the purpose of this work, the \gls{ml} categories that fit the problem are supervised and unsupervised learning, as the problem can be seen as a classification problem (i.e.
-to which class a malware belongs) and a clustering problem (i.e.
-how similar is this malware type to known types).
+When designing a learning system, multiple choices must be taken into consideration. A brief description of what should be taken into account is given as seen in\cite{mitchell:ml}.
 
-\clearpage
+The first choice is the \textit{training experience}, as it can have a significant impact on the success or failure of the learner.
+The feedback provided by the experience can be direct or indirect, affecting the end result of the system.
+Another important aspect is how the experience represents the distribution over which the performance is measured, meaning how well the experience represents the environment where the system is to be used.
+
+The next choice is the \textit{target function}, representing what is to be learned.
+The function takes as input the set of features that represent the problem and (ideally) outputs a solution to the task.
+
+With a well defined target function, what remains is \textit{choosing a representation for the target function}, where a learning model is applied to learn the given task.
+The choice is of great importance, given it will influence how well the system performs under evaluation.
+
+Given the nature of the purposed work as a classification problem, many models can be applied as a representation for the target function.
+The following describe some models that should be taken into consideration during implementation.
+
+\paragraph{Decision Tree Learning.}\cite{mitchell:ml} These methods are used to approximate discrete-valued target functions, where the learned function is represented as a decision tree.
+The decision tree is built based on how much information each attribute gives (i.e.
+information gain).
+These are best suited for problems where instances are represented by attribute-value pairs of discrete values and problems where data may contain errors.
+The shortcomings with decision tree methods include how deep to grow the decision tree, how to handle continuous attributes and handling attributes with different weights.
+
+\paragraph{Artificial Neural Networks.}\cite{mitchell:ml} These methods can be used for real-valued, discrete-valued and vector-valued target functions.
+\gls{ann} are inspired in part by neurons in the biological world, hence being represented as directed graphs.
+\gls{ann}'s can be applied to the same type of problems as decision trees, but these can also handle instances represented by many attribute-value pairs.
+\gls{ann}'s have the downsides of being time consuming to train and to easily overfit the training data (i.e.
+training data is learned well, but fail to generalize to new data).
+
+\paragraph{Bayesian Learning.}\cite{mitchell:ml} These methods use a probabilistic approach to inference.
+The reasoning is that the optimal decision can be related to some probabilistic distribution of the data.
+Bayesian learning proves to be effective for text classification tasks and also provides useful insight on other learning algorithms that do not deal with probabilities.
+The shortcomings can be related to the need of requiring initial knowledge, as well as computational cost to achieve the best hypothesis, which grows with the number of hypothesis.\\
+
+Even though the field of \gls{ml} is broad, the main focus is always the ability to generalize a given task.
+For every learning task multiple models can be applied, which complicates the task of choosing a model.
+Being able to try different models how they compare to each other benefits the end result, which is to be taken into consideration for the current work.
 
@@
 %%% Machine Learning in malware detection
-\subsection{Machine Learning in Malware Detection}
+\subsection{Machine Learning in Malware Detection}\label{subsec:ml_md}
 
-% #11 Reviewer Integration and Performance Measurement for Malware Detection
+The previous subsections gave a general understanding on how anti-virus software detects malware and how the \gls{ml} field can be applied to learn tasks without having to define specific algorithms.
+This section will describe research that aggregates both malware detection and machine learning.
+Authors make use of specific \gls{ml} classifiers like Naive Bayes (NB), \gls{lr} and \gls{svm} to design systems that learn to detect malware.
 
-Miller et al provide an interesting work on combining machine learning techniques with manual review of samples[quote 11].
-They extract both static and dynamic features from binaries and apply logistic regression to achieve a 72\% detection rate with a 0.5\% false positive rate.
-They also note that performance measurement can easily be inflated when ignoring temporal consistency in both binary samples and labels.
-Such can happen when using methods like \textit{cross-validation} for performance measurement, which implies that all types of malware are both seen in evaluation and in training, consequently the notion of unseen malware is not present.
+The main focus is to note the choices made by the authors relative to the datasets, classifiers and their evaluation, all of which are to be taken into consideration for the proposed work.
 
-% explicar as tecnicas e porque, como usa-las, vantagens, what's short
-%
+Before analyzing the work done with \gls{ml} in malware detection, it is worth noting how such work is to be conducted.
+Rossow et al\cite{rossow:practices} present some guidelines for designing malware experiments.
+Their work surveys 36 academic publications from 2006 to 2011 that rely on malware execution.
+The authors note frequent shortcomings in the surveyed publications regarding assumptions on the use of execution-driven datasets, absence of security precautions taken during experiments and insufficient description of the experimental setup.
+The provided guidelines are split into four groups: correct datasets, transparency, realism and safety.
+
+Regarding correct datasets, the publication guidelines include whether goodware samples are to be present in datasets, how the dataset is balanced over malware families and if training and evaluation datasets should have different families.
+The authors note that at least nine distinct papers suffer from significant problems relating to the correctness criteria.
+
+With regards to transparency, the given guidelines include the interpretation of false positives, false negatives and true positives, family names of employed malware samples and how they were named.
+In the surveyed papers, it is noted that the majority do not interpret the numeric results and a fifth do not name the malware families contained in the datasets.
+
+Regarding realism, the paper introduces guidelines which include relevance and variety of malware families and real-world evaluations.
+The authors note that only a minority of papers include real-world evaluations and very few offer significant sample sizes, making it difficult to judge practical use of the methodology.
+
+With regards to safety, the given guidelines focus on containment policies when analyzing malware samples. The authors note that most papers did not deploy or adequately describe containment.
+
+As mentioned in subsection \ref{subsec:cuckoo}, the current work is to make use of the reports in Malwr, hence some the guidelines provided in the aforementioned research do not apply, specifically the safety guidelines.
+As for the other guidelines, it is important to take them into consideration when designing the proposed learning system.
+
+On the same note of guidelines, Shabtai et al\cite{shabtai:survey} provide a survey directed at the application of machine learning classifiers to detect malware from static features.
+Their work concerns the design and evaluation of such systems.
+The process is divided into two phases: training and testing.
+
+The training phase includes the creation of a vector to represent each file and posterior processing of such vectors to generate a classifier, while the testing phase measures the performance of the classifier by employing classical measures like \gls{tpr}, which measures the sensitivity (i.e.
+detection rate) of the classifier, and \gls{fpr}, which measures fall-out.
+
+The authors outline that the work in designing learning systems include how the samples are to be represented (e.g.
+byte n-grams, portable executable features, strings), methods to select features (e.g.
+gain ratio, document frequency) and finally the classification algorithm (e.g.
+\gls{ann}, \gls{svm}).
+
+It is noted by the authors that the use of multiple classifiers with different weights (ensembles) is beneficial to the classification task, as different classifiers work better on different features.
+
+Some problems addressed that are of importance for the proposed work include the imbalance problem and chronological evaluation.
+Having a balanced dataset of malware and goodware does not accurately represent the real world.
+For chronological evaluation the question if whether the usage of old samples for training can aid or hinder the classification of new samples.
+
+When it comes to malware detection from static features, Schultz et al\cite{schultz:data_mining} present a data-mining framework to detect new, unseen malicious executables.
+Their work uses a dataset of 4,266 programs, with 3,625 malicious binaries and 1,001 clean programs, labeled by a commercial anti-virus.
+The features include \gls{dll}'s used by the binary, list of \gls{dll} function calls, number of different function calls within each \gls{dll}, strings from the binary and byte sequences.
+
+They apply three different models (RIPPER, Naive Bayes and Multi-Naive Bayes) to different features, achieving a detection rate between 52\% and 98\% and a false positive rate between 5\% and 9\%.
+It is worth noting how the application of different models to different features provide very distinct detection rates.
+Also worth noting is the size and imbalance of the dataset, where only 23\% of samples are benign in a total universe of 4,266 samples.
+
+The authors describe the malware samples were obtained from various FTP sites and that 5\% were trojans, while the remaining 95\% consisted of viruses, with no remarks are made regarding the involved families.
+The majority of benign samples were gathered from a fresh installation of Windows 98, which may not represent the average benign sample.
+
+Another application of \gls{ml} to malware detection from static features is done by Nissim et al\cite{nissim:al_pdf}.
+The authors apply active learning to the detection of malicious PDF files based on static features.
+They use a dataset of 6,774 PDF files, including 1629 malicious and 5,145 benign files, where the benign files were mark as such by Kaspersky anti-virus software.
+
+Their framework works by filtering out all known benign and malicious PDF's and passing the remaining files to a \gls{svm} model, which linearly separates files based on an initial training set of malicious and benign files.
+The ones deemed informative are flagged for manual inspecting and used in future iterations of the \gls{svm} model training.
+The core idea in this work is to focus on detecting new and unseen instances of malware files with the help of human interaction.
+
+The authors only apply an \gls{svm} model and justify its use due to previous work that provided good results in malware detection, the fact that it is hard for attackers to understand how the model separates files, that it is efficient when combined with active learning and the ability to handle large number of features.
+
+They evaluate their framework by creating 10 subsets of 620 files, with the remaining 574 files used as initial training for the \gls{svm} model.
+For each subset the \gls{svm} is trained and choses the most informative files, these are manually inspected, labeled and added to the training set.
+By incrementally using a bigger training set with files manually labeled, the authors show a growing detection rate from 92\% in the first subset to nearly 97\% with the last subset.
+This is accompanied by a false positive rate from 0.6\% to under 0.2\%.
+These results show that their framework is capable of detecting new and unseen files without a very large training set, but with the help of manual inspection.
+
+Their approach on creating subsets for evaluation and using more benign files in the sets can more closely relate to a real world scenario, as new malicious content is created every day, but it is still less then the amount of benign content.
+
+The authors fail to describe why the chosen features work, this might be related to the use of an \gls{svm}, which makes it difficult interpret the parameters of the model.
+
+With regards to malware detection from dynamic features, Rieck et al\cite{rieck:dynamic} propose a framework for automatic analysis of malware behavior using machine learning.
+Their approach applies both unsupervised (clustering) and supervised (classification) learning to identify novel malware classes and assign unknown to the discovered classes.
+Worth noting is that the mentioned work does not deal with goodware samples, focusing only on identifying different malware families.
+
+Their framework starts by executing and monitoring the malware binaries in a sandbox environment, producing sequential reports that contain the operations and actions performed (system calls and their arguments).
+The reports are embedded in a high-dimensional vector space and \gls{ml} techniques for clustering and classification are applied.
+
+To optimize the processing of reports, the authors propose a special representation of behavior, namely \gls{mist}.
+This representation encodes the monitored system call and its arguments using short numeric identifiers.
+Different levels of specificity can be obtained since the arguments are arranged in blocks.
+For variable-length arguments an index number is used, which can be mapped to the original content.
+Figure \ref{fig:mist} shows the schematic overview of \gls{mist} instructions.
+
+\begin{figure}[h]
+	\includegraphics[width=\textwidth]{mist}
+	\caption{\gls{mist} representation as seen in \cite{rieck:dynamic}}
+	\label{fig:mist}
+\end{figure}
+
+An example of a MIST instruction could be as following:
+
+\begin{center}\texttt{03 05 | 0100000001 | 00006ce5 | 000066fc}\end{center}
+
+This instruction contains four levels.
+The first level contains the category and operation, while the remaining levels each contain an argument block.
+By variating the level, different granularities can be achieved.
+
+The authors create a vector space for each report that represents the behavior of the sample.
+The vector is conceived by creating \textit{Q-grams}.
+These are the result of applying a sliding window of size $q$ to the MIST report and constructing a binary vector where every position is whether the report contains the sequence of size $q$ or not.
+The size of the vector varies with the number of possible instructions and the size $q$ of the window.
+
+The authors implement their own clustering and classification algorithms for the task at hand.
+The evaluation data consists on two sets, a reference set of 3,133 malware reports with 24 different malware families, to be used to evaluate and calibrate the framework, and an application set of 33,698 unknown malware reports taken over a course of seven days.
+
+On the reference set authors apply F-measure for evaluation, which uses precision and recall as parameters, the maximization of F-measure reflects high precison and recall.
+The results yield an F-measure over 0.936, which surpasses previous research for both clustering and classification.
+
+On the application set authors apply their methodology and achieve 434 clusters (i.e.
+malware families).
+To the ten largest clusters authors take the most frequent anti-virus label for each cluster using Kaspersky anti-virus.
+They note that one anti-virus label is prevalent for each of the ten clusters, assuming that their framework successfully discovers and classifies unknown malware families.
+
+By applying their framework to two different sets, where one was created from real-world samples, the authors show how their framework is applicable to a real-world scenario.
+
+Miller et al\cite{miller:rev_int} provide a system which relates to \cite{nissim:al_pdf}, as it applies expert reviewers as a limited labeling resource.
+Their work uses both static and dynamic features to detect malware and uses manual label together with \gls{ml} to improve the overall result.
+The authors evaluate their design on an enormous dataset of 1.1 million binaries spanning 2.5 years, taken from submissions made at VirusTotal, a service similar to Malwr.
+Alongside the enormous dataset, the authors also make sure that training data predates evaluation data, contributing to an evaluation with high realism.
+
+With regards to the design of the system, it consists of a detection and a training pipeline.
+The detection pipeline takes a binary, extracts the features and applies the current model to classify the binary as malicious or benign.
+The training pipeline takes binaries labeled by the manual reviewer, VirusTotal and from the current model, extracts the features and trains the new model.
+
+The system uses multiple static and dynamic features from binaries, including: binary metadata, digital signing, static imports, dynamic imports, file operations, mutex operations, network operations, processes and Windows \gls{api} calls.
+The authors chose to use \gls{lr} as the classifier, basing their choice on the fact that it predicts a real-valued quantity, which enables the use of a threshold and therefore enables the creation of a tradeoff between true and false positive rates, by adjusting the threshold.
+\gls{lr} also assigns a weight to each feature, which allows to comprehend which features are more discriminating.
+
+On evaluation, authors show that when ignoring temporal consistency, results inflate up to 20\% at a 0.5\% false positive rate.
+When considering temporal consistency, results range from 65\% to 75\% true positive rate with a false positive rate range from 0.1\% to 1\%.
+
+Another remark made in this article is that vendors prefer false negatives over false positives (i.e.
+wrongly classifying malware as goodware over wrongly classifying goodware as malware).
+When looking at duplicated samples, 29.6\% of those samples increased in the number of detections as malware and only 0.25\% decreased the number of detections.
+
+Perdisci et al\cite{perdisci:behavior} use an unsupervised learning approach to cluster malware using malicious network traces.
+Although their work does not directly relate to the proposed work, it also deals with the malware naming problem.
+To validate the obtained clusters, the cohesion and separation among clusters is measured in terms of the labels given by anti-virus vendors.
+
+To derive the cohesion and separation, a graph for each cluster is constructed.
+Each node is a malware label given by some vendor and an edge connects two nodes if the labels appear together in any malware sample.
+Weights are assigned as $1 - \dfrac{m}{n}$, where $m$ is the number of times the different labels appear together and $n$ is the number of samples in the cluster.
+From the graphs the authors can then measure the cohesion and separation of clusters.
+
+The authors use the aforementioned method to evaluate their results, but the approach can be useful to label and balance malware datasets that contain various anti-virus labels.\\
+
+Overall the presented research provides useful insight for the current work.
+Decisions like the handling of the dataset, how and which features are to be extracted from samples, which and why \gls{ml} classifiers should be used and how and why the evaluation should be made.
 
@@
 
-With the rate at which malware is discovered, having the ability to automatically detect both known and new samples of malware facilitates the protection of systems, while reducing human interaction.
-As such, the current section proposes a solution that takes advantage of machine learning capabilities to detect malware, based on static and dynamic information that can be extract from malware samples.
+With the knowledge given from previous sections, the current section will describe in detail the work be done as to achieve the objective of detecting malware samples by applying \gls{ml} classifiers.
 
-The solution is divided into four sequential stages, where each stage directly depends on the previous one.
-The four parts are: collecting; parsing; training and validation.
-More details about each stage is described in the corresponding subsection.
+As mentioned in section \ref{sec:introduction}, the approach to be taken to achieve the desired system is divided into three states: data collection, data labeling and classifier implementation.
+The following will describe each stage with more detail, together with work that as already been done.
 
-With an implemented solution, it is expected to have a comparison on how different models predict a sample class (i.e.
-malware, not malware, what kind of malware), as well as benchmarks on the accuracy of the models.
+\subsection{Data Collection}\label{subsec:data_collection}
 
-\subsection{Collecting}
-The first stage of the proposed solution, which consists on obtaining a corpus from which information about the malware can be extracted.
-For the purpose of this work, the obtained dataset consists on reports of static and dynamic information about binaries.
-Working with available reports instead of generating from samples saves the time overhead from both sample obtainment and report generation, allowing for more time better spent in the remaining stages, which are more related to the presented work.
-\todo[inline]{Talk about using malwr reports}
+The core objective of this stage is to obtain a corpus from which information about malware can be extracted.
+As mentioned in subsection \ref{subsec:cuckoo}, reports from Malwr are to be used, as they provide both static and dynamic information.
+The choice of using the already available reports instead of generating reports from collected samples is two-fold, it saves the overhead time of doing analysis and removes any bias regarding the sample choice, as real-world submissions are being used.
+The downside of using these reports is assuming the analysis was correctly done, which may not hold and can introduce noise to the extracted information.
 
-\subsection{Parsing}
-The second state of the proposed solution, that consists on parsing the previously obtained data.
-The reports must be parsed not only to filter relevant information, but also to guarantee data canonicalization.
-Having a well structured dataset eases feature extraction and posterior application of machine learning techniques, hence this stage is of extreme importance for the upcoming stage.
-\todo[inline]{Talk about what can be extracted from this stage (metawise), like percentages, dups, etc}
+The data provided by each report can be either static or dynamic, as mentioned in subsection \ref{subsec:cuckoo}.
+Malwr also provides additional information regarding static analysis, as it checks if the sample was submitted to VirusTotal, if so it lists the classification given by multiple anti-virus vendors.
+This information can then be used to assert if a sample is malicious or not.
 
-\subsection{Training}
-The third stage of the proposed solution, in which machine learning techniques are applied to obtain a model that can predict if a given sample is malicious and if so, what kind of malware is it.
-Before applying any machine learning techniques, a set of features must be selected from the dataset.
-Since the model will be trained based on the selected features, it is of uttermost importance that feature selection is done correctly, otherwise poorly chosen features will not be able to provide a good segregation between samples and consequently the final model will not as accurate.
+Although Malwr enables the user to submit various type of objects, this work will only deal with \gls{pe} files (e.g.
+32bit executables, \gls{dll}'s), as the Cuckoo implementation used by Malwr only provides static analysis for such files.
 
-\todo[inline]{talk about unsupervised methods to facilitate feature selection}
-\todo[inline]{talk about thresholds for supervised methods to classify samples}
+To check if the \gls{pe} provided a good amount of samples, an initial scrape of the list of analysis\footnote{Malwr Analysis List [https://malwr.com/analysis/]} was made using Scrapy\cite{tool:scrapy} and the present file types were analyzed.
+At the time of scrape there was a total of 642,698 submissions, where 388,702 (60.48\%) were \gls{pe} files, dated between April 16th, 2013 and October 10th, 2016.
+Figure \ref{fig:dist_submissions} and \ref{fig:dist_submissions_types} show the submissions by month, the percentage of duplicates per month and how the sample types are distributed, worth noting is that in August 2014 no samples were collected, possibly due to service downtime.
 
-\subsection{Validation}
-The fourth and last stage of the proposed solution, that consists on measuring how good the previously obtained model is.
-The validation of the model is done by taking a set of samples which were not used for training and predict to what class the sample belongs (i.e.
-malware, not malware, malware type).
-With the model prediction, and since the validation set is labeled, the true positive, true negative, false positive and false negative can be taken into account to assert how accurate the model is with regards to the correct labels.
+Knowing that \gls{pe} still account for a good part of all samples, a scrape of those reports was made.
+The reports were saved as HTML in the gzip format to save disk space.
+A total of 388,513 reports were obtained (different from the previous amount as some errors could have occurred during scraping) occupying a total of 26GB.
 
-\todo[inline]{Talk about using cross-validation as a baseline}
-\todo[inline]{move to temporal consistency and how it performs vs cross}
+Having the reports locally, the next step is to parse the HTML format to extract the information to be used.
+Related work\cite{miller:rev_int,nissim:al_pdf,rieck:dynamic,schultz:data_mining,shabtai:survey} provide good examples of static and dynamic information to extract.
+So far only the static imports made by each executable were extracted (i.e.
+\gls{dll} functions), leaving much information to be extracted still.
+To speed parsing, Celery\cite{tool:celery} was used to distribute the work.
+Since it has proven to be useful in extracting static imports, it should also be applied for the remaining parsing.
 
-\subsection{Tools (or some more relevant name?)}
-Being able to easily modify and rerun each stage is one of the main concerns, as more time can be spent trying different techniques without spending much time on implementation.
-As such, Python 3 was chosen as the base language for development.
+\begin{figure}[!h]
+	\subfloat[Total samples per month]{
+		\includegraphics[width=\textwidth]{samples_count}
+	}\\
+	\subfloat[Percentage of duplicated (same MD5) samples per month]{
+		\includegraphics[width=\textwidth]{samples_dups}
+	}
+	\caption{Statistics for the scraped submissions. Count and percentage of duplicates}
+	\label{fig:dist_submissions}
+\end{figure}
 
-There are a large number of libraries in Python that provide easy access to machine learning capabilities, information retrieval and plotting and statistics.
-The ones that are of main interest for this work are: SciKit-Learn, Scrapy, Numpy, Pandas and MatPlotLib.
+\begin{figure}[!h]
+	\includegraphics[width=\textwidth]{samples_types}
+	\caption{Statistics for the scraped submissions. Distribution of file types}
+	\label{fig:dist_submissions_types}
+\end{figure}
 
-\todo[inline]{Talk here about INOV's machine?}
+\subsection{Data Labeling}\label{subsec:data_labeling}
 
-\subsection{Developed work}
-The stages on which most time should be spent are training and validation.
-As such, some work as already been done in the previous stages, specifically in the collecting and parsing stage, so that more time can be dedicated to training and validation.
+While the previous section dealt with extracting information about an executable, the current section deals with labeling each executable as malicious or benign.
+This task is not trivial, as each sample is classified with different labels by different vendors.
+Not only that, but the number of anti-virus classifications varies among samples, as only vendors that have seen the sample classify it.
+Metrics have to be defined as to split the samples into malware or goodware.
 
-With regards to the collecting stage, ...
+To help labeling the dataset, this work proposes to filter samples by only keeping samples classified by a subset of vendors.
+The chosen subset is created by ordering the vendors by the number of classified samples (from the total number of samples) and choosing the top 20.
+This guarantees that every vendor has a classification for the sample at a reduced number of total samples.
 
-Regarding the parsing stage, ...
+To label a sample as benign, the current work defines that if a sample is considered as clean by all vendors, it is labeled as goodware.
+When dealing with duplicated samples, the most recent report is taken into account.
+Although this method may seen obvious in the first place, it can fail if at the time the report was made, vendors weren't able to detect the threat and labeled the sample benign.
+As authors in \cite{miller:rev_int} have shown, this problem occurred in almost 30\% of their duplicated samples.
 
-\clearpage
+To label a sample as malicious, the number of detections has to be taken into account.
+The imposed question is at which threshold should a sample be considered malware?
+On top of that, the type of malware should also be taken into consideration to provide a uniform (i.e balanced) dataset, as mentioned in \cite{rossow:practices}.
+A possible method to help balance the malware variety is to adopt some of the work in \cite{perdisci:behavior} regarding the naming problem: creating a graph where the nodes are all existing labels and edges connect each node based on how many times labels appear together in a sample.
 
-\todo[inline]{Start by the approach?}
-\todo[inline]{Talk about the architecture}
-\todo[inline]{Should evaluation and planning go inside this section?}
-\todo[inline]{Talk about the stats of header analysis here?}
-\todo[inline]{Talk about what has already been/being done?}
-\begin{itemize}
-	\item Begin with extraction from malwr.com
-	\item Talk about why use of PE32?
-	\item Create a corpus from the analysis
-	\item How to classify samples as malware? How to define the threshold?
-	\item How to do feature selection? Measure multiple features?
-	\item Create a baseline to evaluate against, validate other's work? Ignore temporal consistency?
-	\item Introduce temporal consistency in samples, how well one can detect/classify with limited samples
-	\item Try to classify/detect samples that were taken after the training samples.
-\end{itemize}
+Regarding data labeling some work was done to better understand the problem.
+For the top 20 anti-virus vendors, a filter was applied as to limit samples where all vendors classified as either goodware or malware, resulting in a total of 269,563 samples.
+A histogram of how many vendors classified a sample as malware was created, resulting in figure \ref{fig:dist_mal_class}.
+This shows how the threshold to label a sample as malware can be a difficult choice, since vendors often disagree on what is malware.
 
-% tools existentes q vao ser usadas
+\begin{figure}[]
+	\includegraphics[width=\textwidth]{samples_top20_class}
+	\caption{Distribution of how many vendors classify a sample as malware.}
+	\label{fig:dist_mal_class}
+\end{figure}
 
-% \section{•} % Metodologia de Avaliação do Trabalho
-% \section{•} % Calendarização do Trabalho
+Further work was done to understand the naming variation.
+Again for the top 20 anti-virus vendors a filter was applied as to limit samples where all vendors classified as malware, resulting in a total of 10,868 (as can be seen in the last column of figure \ref{fig:dist_mal_class}).
+The unique number of names was counted, together with the percentage of repeating names.
+The results can be seen in table \ref{tab:statistics}, showing a varying number of unique names per vendor, with an average of 2,403 unique malware names per vendor.
+For the ten most repeating names, it can be seen that the variation is smaller and that on average the top ten most repeating names account for 28.7\% of the samples.
+
+\subsection{Classifier Implementation}\label{subsec:class_implementation}
+
+This final stage takes the work done in the previous stages to generate a model that is able to detect malware based on static and dynamic features.
+This stage is to take into consideration the guidelines given in \cite{rossow:practices,shabtai:survey} and the classifiers in \cite{miller:rev_int,nissim:al_pdf,rieck:dynamic,schultz:data_mining}.
+Ideally different classifiers should be tested, as some perform better over others for some set of features.
+The possibility of using multiple classifiers for different features can also be applied to improve the overall result.
+
+\paragraph{Evaluation.} The final result is to be measured using the common evaluation metrics that others\cite{miller:rev_int,nissim:al_pdf,schultz:data_mining} have applied, specifically the \gls{tpr} and \gls{fpr}.
+The \gls{tpr} measures the detection rate of the resulting model, while the \gls{fpr} measures the fall-out.
+
+The described metrics are to be applied in two different conditions: laboratory and real-world.
+For laboratory conditions, the dataset is to be randomly split, using one part for training and other for evaluation.
+For real-world conditions, the dataset is to be split by date, using older samples for training and more recent samples for evaluation.
+By forcing this temporal consistency, the current work will also try to validate the assumption in \cite{miller:rev_int}.
+
+\subsection{Experimental Work}\label{subsec:exp_work}
+
+As a motivation for the work still to be done, a small experiment was conducted to check how easily the creation of a classifier to detect malware is.
+A dataset was created using the static imports as features.
+As for the sample labeling, the classification as given by Microsoft's Windows Defender was used.
+
+Using scikit-learn\cite{tool:sklearn} each sample was converted to a binary vector, where each position corresponded to the presence (or absence) of a static import.
+This feature representation lead to vectors with nearly 30,000 positions (i.e.
+there were almost 30,000 different static imports).
+A logistic regression classifier was trained with 75\% of the samples, using the remaining 25\% for evaluation.
+To check how the dataset balancing affected the classifier, as mentioned in \cite{rossow:practices}, three variants of the dataset were used.
+
+The first dataset was the default and contained more goodware than malware (154,556 goodware, 76,495 malware), this yielded a detection rate 57.2\% with a \gls{fpr} of 28.4\%.
+The second dataset used the same amount of goodware and malware (76,495 samples each), resulting in a detection rate of 69.2\% and a \gls{fpr} of 23.7\%.
+The last dataset contained more malware than goodware (76,495 malware, 38,247 goodware) and gave a detection rate of 92\% with a \gls{fpr} of 22.4\%.
+Table \ref{tab:experimental} summarizes the results and notes the results from \cite{miller:rev_int} as a comparison.
+
+\begin{table}[h!]
+	\centering
+	\def\arraystretch{1.5}
+	\begin{tabular}{lcc}
+		\toprule
+		& Detection Rate        & False Positive Rate        \\
+		\midrule
+		First Dataset  & 0.572     & 0.284      \\ \hline
+		Second Dataset & 0.692     & 0.237      \\ \hline
+		Third Dataset  & 0.92      & 0.224      \\ \hline
+		Miller et al   & [0.65, 0.75] & [0.001, 0.01] \\
+		\bottomrule
+	\end{tabular}
+	\caption{Summary of experimental results compared to Miller et al\cite{miller:rev_int}}
+	\label{tab:experimental}
+\end{table}
+
+This quick test gave some interesting results, although they do not compare to related work, it showed that with little to no work on the dataset and classifier a simple model was achieved.
+It also provided some experience with scikit-learn, which is to be used in the work to come.\\
+
+The work that will result in the proposed system is to be developed between June 2017 and March 2018.
+
+The planned schedule is to use the month of June to finish the data collection phase, together with the labeling phase during the months of June and July.
+The final phase of implementation is planned to take six months, from August to January, as many variables are present and more experiments should provide better results.
+Regarding the writing of the thesis document, the months of February and March are to be used.
+A more detailed planning can be seen in figure \ref{fig:schedule}.
+
+\begin{figure}[h]
+	\includegraphics[width=\textwidth]{gantt}
+	\caption{Planned Schedule}
+	\label{fig:schedule}
+\end{figure}
 
@@
 
-\begin{thebibliography}{}
+With the growing amount of malware instances, being able to effectively detect malware as it appears is a must, otherwise cases like WannaCry malware\footnote{Wanna Cry Profile [https://www.fireeye.com/blog/threat-research/2017/05/wannacry-malware-profile.html]} will continue to have real and serious effects\footnote{Huffing Post Article [http://www.huffingtonpost.com/entry/britain-cyber-attack-national-health-service\_us\_5915e62ae4b00f308cf4fc86]}.
 
-\bibitem{mitchell:ml}
-Mitchell, T.:
-Machine Learning
+When new types of malware arise, anti-virus may be unable to detect them with the usual signature and heuristic methods.
+The application of \gls{ml} to the detection task as shown to be possible and effective\cite{kolter:learning,miller:rev_int,nissim:al_pdf,perdisci:behavior,rieck:dynamic,schultz:data_mining}.
+With that in mind, the current work will develop a \gls{ml} system capable of detecting malicious files, as well as test how different types of evaluations can impact the effectiveness of the solution.
 
-\bibitem{norvig:ai}
-Norvig, P., Russell, S.:
-Artificial Intelligence: A Modern Approach
+The proposed solution, as described in section \ref{sec:solution}, is to be split in three steps: data collection, data labeling and classifier implementation.
+Some work as been done in each step as to get a feel on what to expect and the problems that might surge along the way.
 
-\end{thebibliography}
+The next step is to fully terminate each of the sequential stages, with special focus on classifier implementation as many variations can be tested, providing a better research and final result.
+
+Hopefully the current work enlightened the reader on the problem, how the current technologies deal with it and how the proposed work could improve the detection task.
+
+\begin{landscape}
+	\begin{table}
+		\centering
+		\def\arraystretch{1.1}
+		\subfloat{
+			\begin{tabular}{rcccccccccc}
+				\toprule
+				{} &  thehacker &  bitdefender &  cat-quickheal &  fortinet &   virobot &    ikarus &     gdata &    f-prot &  k7antivirus &    clamav \\
+				\midrule
+				Unique names   &   3090 &     3223 &       2286 &  2144 &  3617 &  1139 &  3227 &  2515 &     1869 &  3195 \\
+				\midrule
+				1st most used  &      0.068 &        0.066 &          0.063 &     0.057 &     0.063 &     0.075 &     0.066 &     0.052 &        0.042 &     0.026 \\
+				2nd most used  &      0.047 &        0.047 &          0.052 &     0.050 &     0.028 &     0.040 &     0.047 &     0.035 &        0.035 &     0.022 \\
+				3rd most used  &      0.039 &        0.029 &          0.040 &     0.032 &     0.027 &     0.028 &     0.029 &     0.032 &        0.030 &     0.022 \\
+				4th most used  &      0.030 &        0.025 &          0.030 &     0.027 &     0.025 &     0.027 &     0.025 &     0.030 &        0.029 &     0.017 \\
+				5th most used  &      0.025 &        0.020 &          0.028 &     0.027 &     0.022 &     0.023 &     0.020 &     0.029 &        0.023 &     0.017 \\
+				6th most used  &      0.021 &        0.019 &          0.024 &     0.025 &     0.015 &     0.021 &     0.019 &     0.028 &        0.023 &     0.017 \\
+				7th most used  &      0.019 &        0.017 &          0.023 &     0.019 &     0.013 &     0.017 &     0.017 &     0.026 &        0.023 &     0.017 \\
+				8th most used  &      0.018 &        0.017 &          0.021 &     0.018 &     0.013 &     0.016 &     0.017 &     0.019 &        0.021 &     0.015 \\
+				9th most used  &      0.015 &        0.017 &          0.019 &     0.017 &     0.012 &     0.015 &     0.017 &     0.017 &        0.020 &     0.014 \\
+				10th most used &      0.015 &        0.014 &          0.014 &     0.015 &     0.012 &     0.015 &     0.014 &     0.015 &        0.018 &     0.013 \\
+				\midrule
+				Total          &      0.297 &        0.272 &          0.315 &     0.287 &     0.231 &     0.277 &     0.272 &     0.283 &        0.263 &     0.179 \\
+				\bottomrule
+			\end{tabular}
+		}\\
+		\subfloat{
+			\begin{tabular}{rcccccccccc}
+				\toprule
+				{} &    mcafee &  microsoft &  nprotect &    comodo &     avast &       avg &     vba32 &  kaspersky &  mcafee-gw-edition &  symantec \\
+				\midrule
+				Unique names   &  2560 &   1644 &  4332 &  1959 &  1570 &  2644 &  1366 &   2572 &           2436 &   667 \\
+				\midrule
+				1st most used  &     0.069 &      0.064 &     0.060 &     0.068 &     0.068 &     0.044 &     0.069 &      0.064 &              0.051 &     0.071 \\
+				2nd most used  &     0.037 &      0.040 &     0.030 &     0.061 &     0.062 &     0.032 &     0.030 &      0.052 &              0.036 &     0.060 \\
+				3rd most used  &     0.030 &      0.030 &     0.025 &     0.050 &     0.058 &     0.031 &     0.025 &      0.046 &              0.024 &     0.060 \\
+				4th most used  &     0.030 &      0.028 &     0.020 &     0.036 &     0.043 &     0.025 &     0.025 &      0.041 &              0.024 &     0.053 \\
+				5th most used  &     0.028 &      0.027 &     0.018 &     0.031 &     0.042 &     0.023 &     0.023 &      0.030 &              0.019 &     0.033 \\
+				6th most used  &     0.028 &      0.025 &     0.016 &     0.028 &     0.030 &     0.017 &     0.022 &      0.028 &              0.018 &     0.032 \\
+				7th most used  &     0.024 &      0.025 &     0.016 &     0.027 &     0.024 &     0.017 &     0.017 &      0.025 &              0.017 &     0.032 \\
+				8th most used  &     0.019 &      0.025 &     0.014 &     0.021 &     0.019 &     0.017 &     0.016 &      0.021 &              0.016 &     0.030 \\
+				9th most used  &     0.017 &      0.020 &     0.014 &     0.019 &     0.019 &     0.015 &     0.016 &      0.020 &              0.013 &     0.028 \\
+				10th most used &     0.017 &      0.019 &     0.013 &     0.015 &     0.017 &     0.015 &     0.013 &      0.019 &              0.012 &     0.026 \\
+				\midrule
+				Total          &     0.300 &      0.304 &     0.228 &     0.356 &     0.383 &     0.237 &     0.258 &      0.345 &              0.230 &     0.425 \\
+				\bottomrule
+			\end{tabular}
+		}
+	\caption{Malware name statistics for samples where all top 20 vendors classify as malware (10,868 samples). Tables show the unique different names together with repeating names percentage.}
+	\label{tab:statistics}	
+	\end{table}
+\end{landscape}
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%REFERENCES%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+\nocite{*}
+\bibliography{reportbib}{}
+\bibliographystyle{plain}
 

### 2017-06-26 lib/plotting.py

@@ -4,7 +4,7 @@ import matplotlib.ticker as mticker
 
 def plotByMonth(df, name, title):
     ticklabels = [item.strftime('%b %Y') for item in df.index]
-    ax = df.plot(kind='bar', figsize=(20, 10), alpha=0.6)
+    ax = df.plot(kind='bar', figsize=(15, 5), alpha=0.6)
     ax.xaxis.set_major_formatter(mticker.FixedFormatter(ticklabels))
     plt.title(title)
     plt.ylabel('#Analyses')
@@ -14,7 +14,7 @@ def plotByMonth(df, name, title):
 
 def plotByMonthPercentage(df, name, title):
     ticklabels = [item.strftime('%b %Y') for item in df.index]
-    ax = df.plot(kind='bar', stacked=True, figsize=(20, 10), alpha=0.6)
+    ax = df.plot(kind='bar', stacked=True, figsize=(15, 5), alpha=0.6)
     ax.xaxis.set_major_formatter(mticker.FixedFormatter(ticklabels))
     ax.set_ylim([0, 1])
     plt.title(title)

### 2018-04-26 thesis_report/Thesis_Background.tex
Thesis report update.

--- 2017-06-26 project_report/main.tex
@@
-\documentclass{llncs}
-\usepackage[utf8]{inputenc}
-\usepackage{todonotes}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{dirtytalk}
-\usepackage{color}
-\usepackage{glossaries}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{booktabs}
-\usepackage[caption = false]{subfig}
-\graphicspath{ {images/} }
-\usepackage{rotating}
-\usepackage{pdflscape}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                      %
+%     File: Thesis_Background.tex                                      %
+%     Tex Master: Thesis.tex                                           %
+%                                                                      %
+%     Author: João C. Godinho                                          %
+%     Last modified : Apr 2018                                         %
+%                                                                      %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Reset footnotes every page
-%\usepackage{perpage}
-%\MakePerPage{footnote}
+\chapter{Related Work}
+\label{chapter:related_work}
 
-\begin{document}
-\frontmatter
+The following chapter will go over prior work that closely relates to the topic of our research and areas of contribution.
+We approach the definition of malware and how it can hard to detect and label by providing some background on the subject.
+We then proceed to elucidate on different malware detection techniques, from simple methods to more complex approaches, focusing on a specific malware analysis technique that provided the basis for our work.
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%  HEADER  %%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Malware Detection Via Machine Learning}
-\subtitle{}
-\titlerunning{Malware Detection Via ML}
-\author{João F. Godinho}
-\institute{Instituto Superior Técnico\\
-\email{joao.f.godinho@tecnico.ulisboa.pt}}
-\let\oldaddcontentsline\addcontentsline
-\def\addcontentsline#1#2#3{}
-\maketitle
-\def\addcontentsline#1#2#3{\oldaddcontentsline{#1}{#2}{#3}}
-%\todo[inline]{Validate the title, subtitle}
+We advance to the general topic of \gls{ml}, after which we describe related work that studies the use of \gls{ml} in malware detection.
+Lastly, we reference prior work that not only tackles our same goal, but also takes into account the impact of laboratory \textit{vs.} real world scenarios, as well as the notion of temporal based detection, or concept drift.
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%% ACRONYMS %%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\newacronym{hids}{HIDS}{Host-based Intrusion Detection Systems}
-\newacronym{ml}{ML}{machine learning}
-\newacronym{caro}{CARO}{Computer Antivirus Research Organization}
-\newacronym{ai}{AI}{Artificial Intelligence}
-\newacronym{api}{API}{Application Programming Interface}
-\newacronym{bios}{BIOS}{Basic Input/Output System}
-\newacronym{dll}{DLL}{Dynamic-Link Library}
-\newacronym{cc}{CC}{Creative Common's}
-\newacronym{ann}{ANN}{Artificial Neural Networks}
-\newacronym{tpr}{TPR}{True Positive Rate}
-\newacronym{fpr}{FPR}{False Positive Rate}
-\newacronym{svm}{SVM}{Support Vector Machine}
-\newacronym{mist}{MIST}{Malware Instruction Set}
-\newacronym{lr}{LR}{Logistic Regression}
-\newacronym{pe}{PE}{Portable Executable}
+\newpage
+\todo[inline]{CSF Related work}
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%% ABSTRACT %%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-	Malware, which focuses on harming or subverting a system, has grown exponentially in recent years.
-The regular signature and heuristic methods applied by anti-virus vendors can fall short on keeping up with the rise of new types of malware.
-Other works have shown that machine learning has potential for dealing with the task of malware detection.
-With this in mind, the current report elaborates on the problem of malware, how anti-virus detect it, and proposes the development of a machine learning system to deal with the task of malware detection.
-To design the system to be, the current work will make use of a dataset with 388,513 samples for training and evaluation, focusing on how laboratory and real-world conditions can affect the final result.
-\end{abstract}
+We use this section to present prior work that closely relates to the topic of our research and our areas of contribution.
+Specifically, detecting malicious software by training supervised models on static information, and validation methodologies that resemble real-world conditions.
 
-\setcounter{tocdepth}{2}
-\tableofcontents
+Shabtai et al.\cite{shabtai:survey} provide a survey directed
+at the application of \gls{ml} classifiers to detect malware from static
+features.
+Their work concerns the design and evaluation of such systems.
+Our contributions are inspired by the concern of how to correctly evaluate \gls{ml} classifiers, regarding size, reliability of labeling metrics and chronological evaluation.
 
-\mainmatter
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%  INTRO   %%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
+In the topic of methodologies that resemble real-world conditions, Srndi\'c et al.\cite{vsrndic2013detection}
+train and validate their malicious PDF detector under laboratory and real-world conditions.
+Laboratory conditions consist on applying regular cross-validation, whereas real-world conditions validate a newer dataset with a model created from outdated data (i.e.\ older then the validation), and also validate the model when the validation set spans one week and the training is gathered in the previous 4 weeks.
+They show that laboratory conditions inflate the results, when compared to real-world conditions.
 
-% Temporary quotes, move to bib
-% 1 - Attacking Malicious Code: A report to the Infosec Research Council
-% 2 - AV-TEST Security Report 2015/2016 - https://www.av-test.org/fileadmin/pdf/security_report/AV-TEST_Security_Report_2015-2016.pdf
-% 3 - McAfee Labs Threats Report April 2017 - https://www.mcafee.com/us/resources/reports/rp-quarterly-threats-mar-2017.pdf
-% 4 - Symantec Internet Security Threat Report - https://www.symantec.com/content/dam/symantec/docs/reports/istr-22-2017-en.pdf
-% 5 - Scanners of The Year 2000: Heuristics - http://vxheaven.org/lib/adg00.html
-% 6 - Thomas Mitchell - Machine Learning
-% 7 - Some Studies in Machine Learning Using the Game of Checkers - Arthur L. Samuel - http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.368.2254&rep=rep1&type=pdf
-% 8 - Prudent Practices for Designing Malware Experiments: Status Quo and Outlook
+Miller et al.\cite{miller:rev_int} also introduce sample temporal consistency.
+They show the impact of performance measurement technique on a dataset containing 1.1 million samples, when using cross-validation and temporally consistent samples.
+As noted by others, regular cross-validation showed inflated results when compared to temporally consistent samples.
 
-\section{Introduction}\label{sec:introduction}
-Malware, which can be referred to by other names like malicious software or malicious code, can be described as \say{any code added, changed, or removed from a software system in order to
-intentionally cause harm or subvert the intended function of the system}\cite{mcgraw:mal_code}.
-Malware can be a piece of code attached to a legitimate program, an independent program, or a combination of both.
-The reasoning that leads to malware creation varies greatly, some are created to demonstrate some vulnerability or concept and do not cause direct harm to systems, while others are used to steal information or render systems useless.
-The wide variety and quantity of now-a-day systems and the amount of sensitive information they store provide numerous opportunities to illegally make a profit out of subverting legitimate systems.
+Our work enhances these methodologies by analyzing the performance variation when the distance between the training and validation set increases and decreases, as well as analysis on how reducing the size of the training without compromising the results.
 
-The amount of malware has shown an exponential increase since the early 2000s, from 1.7 million in 2005 to over 600 million by the end of 2016\cite{av-test:report,mcafee:report}.
-According to a 2007 study by McAfee\footnote{McAfee Study Finds 4 Percent of Search Results Malicious, by Frederick Lane, June 4, 2007 [http:\/\/www.newsfactor.com\/story.xhtml?story\_id=010000CEUEQO]} about 4\% of the query results by major search engines lead to potentially dangerous websites.
-A more recent study by Norton in 2010\footnote{Symantec Press Release, July 29, 2010 [https://www.symantec.com/en/aa/about/newsroom/press-releases/2010/symantec\_0729\_02]} showed that at least 10\% of top trending search terms returned malicious results.
-From the contact with malicious software, 1.1 billion identities (i.e.
-user's personal information) have been exposed in 2016 alone\cite{symantec:report}.
+Kolter et al.\cite{kolter:learning} learn to detect malicious executables in a dataset with under 4,000 samples, obtained from reliable sources, and evaluate their model under standard cross-validation and by gathering newer malware samples to validate for new and unseen samples.
+They show how the model provides optimal results under cross-validation, but for the unseen samples lower scores are obtained.
+Our work considers these results to compare how reliability affects performance.
 
-Given the wide range and quantity of systems being operated by naive users, in the sense that they lack the knowledge to distinguish malicious and legitimate software, the first line of defense against threats are \gls{hids}, commonly known as anti-virus solutions.
-Their main function is to prevent infections by malicious code, which is achieved by detecting a program as malicious and disallow it from running (i.e.
-quarantining the malicious file).
+Sebastián, M.
+et al.\cite{sebastian2016avclass} develop \textit{AVClass}, a tool that given a set of antivirus vendors, outputs the most likely family name.
+They test their tool under 10 datasets, totaling 8.9 M samples, with results showing an F1 measure up to 93.9 on labeled datasets.
+The tool takes as input the labels as seen in VirusTotal, tokenizes the labels, replaces known aliases and general names (e.g.\ win32, trojan, generic), ending with possible family names.
+These remaining names are counted and the most frequent one is given as family name.
+Our work takes advantage of this tool, not to label malware families, but minimal modifications, to label more general malware classes (e.g.\ trojan, virus).
 
-To detect malicious software, anti-virus solutions depend on two main methods.
-\textit{Signature-based} methods check for the presence of unique byte sequences from known malware.
-For this method to be effective, a database of all malware signatures must be available, which leads to an obvious shortcoming: only known malware is detected\cite{gryaznov:scanners}.
-New and unknown malware is not detected until vendors analyze it and update their clients.
-To improve the effectiveness of anti-virus and overcome the problem in aforementioned methods, \textit{heuristic-based} methods are employed.
-These apply a set of rules to decide whether a program is likely to be malware or not.
-Like before, a database of rules must be available, but unlike signature methods, which provide exact matching on malware, heuristic methods are capable, to some degree, of generalizing malware features, hence detecting new and unknown malware.
-Although more sophisticated, heuristic-based methods need a database of rules that must be defined \textit{a priori}, just like signature methods.
+Deo, A.
+et al.~\cite{deo2016prescience} focus over the problem of \gls{ml} models becoming antiquated over time, given how malware evolves.
+This problem is seen as concept drift, where the performance of a model over time diminishes, as the statistical properties of malware (i.e.\ features) change over time.
+They study how probabilistic predictors can help minimize the aforementioned problem, by indicating when retraining of a given model is necessary.
 
-As mentioned before, signature methods only detect known malware, which translates to a very low false positive rate (i.e.
-goodware wrongly classified as malware), while having a high detection rate (i.e.
-correctly classified malware from the total malware), assuming the signature database is on par with malware samples, which at the rate new malware emerges becomes a hard task.
-For heuristic methods new and unknown malware can be detected due to similarities to known malware, translating to a higher detection rate with less information, but with the downside of raising the false positive rate, as benign programs (goodware) can be flagged malicious due to characteristics that fit into malware rules\cite{gryaznov:scanners}.
-To improve the effectiveness of anti-virus, both methods must complement each other.
+Jordaney, R.
+et al.~\cite{jordaney2017transcend} also study the problem of concept drift in malware classification models, focusing on providing metrics, based on statistical comparison of samples, to detect when should a model be retrained.
 
-Having the need to either create signatures for new malware or define a set of rules that describe malware characteristics becomes a cumbersome task for the increasing malware numbers.
-To try and improve the detection task of anti-virus software, the current work proposes a \gls{ml} approach to malware detection as an evolution to heuristic-based methods.
+Both ~\cite{deo2016prescience} and ~\cite{jordaney2017transcend} are on the subject of our work, regarding the problem of concept drift, but differ on the study-case.
+In our work we acknowledge the concept but focus on its relative effects and how one can balance the size of needed training data vs.
+validation data, when maintaining a temporal consistent dataset.
+Whereas their work provide indicators for when should a model be retrained when the problem of concept drift becomes significant.
 
-The machine learning field studies the development of computer programs that automatically improve with experience.
-A program is given a task and an experience, if its performance at the task improves with the experience given, then the program is learning the task\cite{mitchell:ml}.
-In the proposed work, the task at hand is to distinguish malware from goodware (i.e.
-classify samples) and the experience is a wide variety of samples, both goodware and malware.
-The final result is a program that performs well for the given task (i.e.
-successfully learns to distinguish goodware from malware).
+\newpage
+\todo[inline]{Project Report related work}
 
-The \gls{ml} field has been around for multiple decades, with some of the earlier works dating back to 1959 by Arthur L.
-Samuel\cite{samuel:studies}, but only in recent years has the field grown in popularity.
-Its recent growth can be accredited to two main factors: availability of large data sets and computation power.
-Larger data sets provide more experience for a program to learn from, but also require more processing power.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Background}
+\label{section:background}
 
-The growing number of malware samples provide a good opportunity to make use of \gls{ml} for malware detection.
-This approach is not new and has provided high detection rates with low false positive rates\cite{miller:rev_int,nissim:al_pdf,perdisci:behavior,rieck:dynamic,schultz:data_mining}.
-However, many of the resulting research is done under laboratory conditions that do not represent real world scenarios.
-Some of those conditions have been studied in \cite{rossow:practices} and include the correctness of datasets, transparency and realism.
-Alongside the application of \gls{ml} to detect malware, the proposed work will study how laboratory conditions differ from real world one's and how effectiveness is affected.
-
-To provide a solution that makes use of \gls{ml} to detect malware and its performance under laboratory \textit{vs.} real world conditions, the proposed work will follow an approach that is divided into three stages:
-
-\begin{itemize}
-	\item \textit{Data collection}: experience from which a program can learn must be available, as such the first step must ensure the collection of samples that represent both malware and goodware examples.
-The current work will take advantage of the Malwr\cite{tool:malwr} service, which provides both static (e.g.
-library imports) and dynamic (e.g.
-library call order during runtime) details about a sample.
-The total number of available samples surpasses 388 thousand, dated from April 2013 to October 2016.
-	\item \textit{Data labeling}: before training a classifier to detect malware, samples must be labeled as either benign or malign.
-The Malwr\cite{tool:malwr} service provides details on how multiple anti-virus solutions classify a sample, hence the approach is to use said classification to determine the legitimacy of a sample.
-This step is not linear, as different anti-virus classify samples differently, introducing edge cases (e.g.
-50\% classify malware, the other 50\% classify goodware).
-To overcome this, the current work must develop a metric to deal with such cases.
-	\item \textit{Classifier implementation}: at this stage the collected data is ready to be used as experience for the program.
-Using static and dynamic details together with the labeling of the samples, a classifier can be trained to distinguish between malware and goodware.
-Alongside the classifier creation, this phase must also measure the effectiveness of the solution.
-For this work's scope, the effectiveness is to be measured under laboratory and real world conditions.
-\end{itemize}
-
-In sum, the work to be done as to verify the application of \gls{ml} methods for the task of detecting malware is: collection of samples, labeling of collected samples and finally usage of the resulting dataset to achieve a malware classifier.
-
-The following sections provide deeper analysis on the subject of using \gls{ml} to classify malware, as well as the choices to be made and the reasoning behind them.
-Section \ref{sec:related_work} describes related work that both influenced and motivated the current work.
-Section \ref{sec:solution} details the intended approach, the tools to be used, how the work is to be presented and the already accomplished work.
-Section \ref{sec:conclusion} summarizes and closes the report.
-
-%\todo[inline]{falar dos antivirus para combater isto, falar nas tecnicas basicas: signature based, heuristic based.
-tirar partido de ML para estender metodos heuristicos e melhorar a classificaçao de malware.}
-%
-%
-%\todo[inline]{one liner do que é malware (quote), quantificar (quote), necessidade de detectar malware mais rapidamente (automatizar/generalizar).
-%
-%Avanços na detecçao do malware, pouca coisa. Aplicaçao de ML neste tipo de problema.
-%
-%Contributos do trabalho para este problema, desenvolver um classificador que funcione no mundo real e ajude a resolver o problema.
-%
-%1. Dados recolhidos, o que posso extrair dos dados\\
-%2. Definir o que é malware (thresholds, consenso)\\
-%3. Comparaçao de metodologias (estatatica/dinamica)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%% RELATED  %%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec:related_work}
-
-% #1 Semantics-Aware Malware Detection
-% #2 Attacking Malicious Code: A report to the Infosec Research Council
-% #3 Malware, Rootkits & Botnets A Beginner's Guide pp.10
-% #4 Computer viruses: theory and experiments
-% #5 Theory of self-reproducing automata
-% #6 The Evolution of Viruses and Worms
-% #7 The art of computer virus research and defense
-% #8 https://www.av-test.org/en/
-% #9 http://www.caro.org/articles/naming.html
-% #10 http://members.chello.at/erikajo/vnc99b2.txt
-% #11 Learning to Detect and Classify Malicious Executables in the Wild
-% #12 The Evolution of Viruses and Worms
-% #13 http://cert.europa.eu/static/SecurityAdvisories/2017/CERT-EU-SA2017-012.pdf
-% #14 https://www.virusbulletin.com/uploads/pdf/magazine/2002/200201.pdf
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Background
-\subsection{Background}\label{subsec:background}
 Malware, which can be referred to by other names like malicious software or malicious code, can be described as \say{a program that has malicious intent}\cite{christodorescu:semantics}, or more detailed as any code added, changed, or removed from a software system in order to intentionally cause harm or subvert the intended function of the system\cite{mcgraw:mal_code}.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Malware detection
-\subsection{Malware Detection}\label{subsec:mal_detec}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Malware Detection}
+\label{section:mal_detec}
 
@@
 
-\subsection{Cuckoo Sandbox and Malwr}\label{subsec:cuckoo}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Cuckoo Sandbox and Malwr}
+\label{section:cuckoo}
 
@@
 
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{malwr_sample}
-	\caption{Example of a behavioral report from Malwr}
-	\label{fig:malwr_sample}
-\end{figure}
+%\begin{figure}[h]
+%	\includegraphics[width=\textwidth]{malwr_sample}
+%	\caption{Example of a behavioral report from Malwr}
+%	\label{fig:malwr_sample}
+%\end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%% Machine Learning
-\subsection{Machine Learning}\label{subsec:ml}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Machine Learning}
+\label{section:ml}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Machine Learning in malware detection
-\subsection{Machine Learning in Malware Detection}\label{subsec:ml_md}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Machine Learning in Malware Detection}
+\label{section:ml_md}
 
@@
 Before analyzing the work done with \gls{ml} in malware detection, it is worth noting how such work is to be conducted.
-Rossow et al\cite{rossow:practices} present some guidelines for designing malware experiments.
+Rossow et al\ \cite{rossow:practices} present some guidelines for designing malware experiments.
 Their work surveys 36 academic publications from 2006 to 2011 that rely on malware execution.
@@
 
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{mist}
-	\caption{\gls{mist} representation as seen in \cite{rieck:dynamic}}
-	\label{fig:mist}
-\end{figure}
+%\begin{figure}[h]
+%	\includegraphics[width=\textwidth]{mist}
+%	\caption{\gls{mist} representation as seen in \cite{rieck:dynamic}}
+%	\label{fig:mist}
+%\end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%% SOLUTION %%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Solution}\label{sec:solution}
-
-With the knowledge given from previous sections, the current section will describe in detail the work be done as to achieve the objective of detecting malware samples by applying \gls{ml} classifiers.
-
-As mentioned in section \ref{sec:introduction}, the approach to be taken to achieve the desired system is divided into three states: data collection, data labeling and classifier implementation.
-The following will describe each stage with more detail, together with work that as already been done.
-
-\subsection{Data Collection}\label{subsec:data_collection}
-
-The core objective of this stage is to obtain a corpus from which information about malware can be extracted.
-As mentioned in subsection \ref{subsec:cuckoo}, reports from Malwr are to be used, as they provide both static and dynamic information.
-The choice of using the already available reports instead of generating reports from collected samples is two-fold, it saves the overhead time of doing analysis and removes any bias regarding the sample choice, as real-world submissions are being used.
-The downside of using these reports is assuming the analysis was correctly done, which may not hold and can introduce noise to the extracted information.
-
-The data provided by each report can be either static or dynamic, as mentioned in subsection \ref{subsec:cuckoo}.
-Malwr also provides additional information regarding static analysis, as it checks if the sample was submitted to VirusTotal, if so it lists the classification given by multiple anti-virus vendors.
-This information can then be used to assert if a sample is malicious or not.
-
-Although Malwr enables the user to submit various type of objects, this work will only deal with \gls{pe} files (e.g.
-32bit executables, \gls{dll}'s), as the Cuckoo implementation used by Malwr only provides static analysis for such files.
-
-To check if the \gls{pe} provided a good amount of samples, an initial scrape of the list of analysis\footnote{Malwr Analysis List [https://malwr.com/analysis/]} was made using Scrapy\cite{tool:scrapy} and the present file types were analyzed.
-At the time of scrape there was a total of 642,698 submissions, where 388,702 (60.48\%) were \gls{pe} files, dated between April 16th, 2013 and October 10th, 2016.
-Figure \ref{fig:dist_submissions} and \ref{fig:dist_submissions_types} show the submissions by month, the percentage of duplicates per month and how the sample types are distributed, worth noting is that in August 2014 no samples were collected, possibly due to service downtime.
-
-Knowing that \gls{pe} still account for a good part of all samples, a scrape of those reports was made.
-The reports were saved as HTML in the gzip format to save disk space.
-A total of 388,513 reports were obtained (different from the previous amount as some errors could have occurred during scraping) occupying a total of 26GB.
-
-Having the reports locally, the next step is to parse the HTML format to extract the information to be used.
-Related work\cite{miller:rev_int,nissim:al_pdf,rieck:dynamic,schultz:data_mining,shabtai:survey} provide good examples of static and dynamic information to extract.
-So far only the static imports made by each executable were extracted (i.e.
-\gls{dll} functions), leaving much information to be extracted still.
-To speed parsing, Celery\cite{tool:celery} was used to distribute the work.
-Since it has proven to be useful in extracting static imports, it should also be applied for the remaining parsing.
-
-\begin{figure}[!h]
-	\subfloat[Total samples per month]{
-		\includegraphics[width=\textwidth]{samples_count}
-	}\\
-	\subfloat[Percentage of duplicated (same MD5) samples per month]{
-		\includegraphics[width=\textwidth]{samples_dups}
-	}
-	\caption{Statistics for the scraped submissions. Count and percentage of duplicates}
-	\label{fig:dist_submissions}
-\end{figure}
-
-\begin{figure}[!h]
-	\includegraphics[width=\textwidth]{samples_types}
-	\caption{Statistics for the scraped submissions. Distribution of file types}
-	\label{fig:dist_submissions_types}
-\end{figure}
-
-\subsection{Data Labeling}\label{subsec:data_labeling}
-
-While the previous section dealt with extracting information about an executable, the current section deals with labeling each executable as malicious or benign.
-This task is not trivial, as each sample is classified with different labels by different vendors.
-Not only that, but the number of anti-virus classifications varies among samples, as only vendors that have seen the sample classify it.
-Metrics have to be defined as to split the samples into malware or goodware.
-
-To help labeling the dataset, this work proposes to filter samples by only keeping samples classified by a subset of vendors.
-The chosen subset is created by ordering the vendors by the number of classified samples (from the total number of samples) and choosing the top 20.
-This guarantees that every vendor has a classification for the sample at a reduced number of total samples.
-
-To label a sample as benign, the current work defines that if a sample is considered as clean by all vendors, it is labeled as goodware.
-When dealing with duplicated samples, the most recent report is taken into account.
-Although this method may seen obvious in the first place, it can fail if at the time the report was made, vendors weren't able to detect the threat and labeled the sample benign.
-As authors in \cite{miller:rev_int} have shown, this problem occurred in almost 30\% of their duplicated samples.
-
-To label a sample as malicious, the number of detections has to be taken into account.
-The imposed question is at which threshold should a sample be considered malware?
-On top of that, the type of malware should also be taken into consideration to provide a uniform (i.e balanced) dataset, as mentioned in \cite{rossow:practices}.
-A possible method to help balance the malware variety is to adopt some of the work in \cite{perdisci:behavior} regarding the naming problem: creating a graph where the nodes are all existing labels and edges connect each node based on how many times labels appear together in a sample.
-
-Regarding data labeling some work was done to better understand the problem.
-For the top 20 anti-virus vendors, a filter was applied as to limit samples where all vendors classified as either goodware or malware, resulting in a total of 269,563 samples.
-A histogram of how many vendors classified a sample as malware was created, resulting in figure \ref{fig:dist_mal_class}.
-This shows how the threshold to label a sample as malware can be a difficult choice, since vendors often disagree on what is malware.
-
-\begin{figure}[]
-	\includegraphics[width=\textwidth]{samples_top20_class}
-	\caption{Distribution of how many vendors classify a sample as malware.}
-	\label{fig:dist_mal_class}
-\end{figure}
-
-Further work was done to understand the naming variation.
-Again for the top 20 anti-virus vendors a filter was applied as to limit samples where all vendors classified as malware, resulting in a total of 10,868 (as can be seen in the last column of figure \ref{fig:dist_mal_class}).
-The unique number of names was counted, together with the percentage of repeating names.
-The results can be seen in table \ref{tab:statistics}, showing a varying number of unique names per vendor, with an average of 2,403 unique malware names per vendor.
-For the ten most repeating names, it can be seen that the variation is smaller and that on average the top ten most repeating names account for 28.7\% of the samples.
-
-\subsection{Classifier Implementation}\label{subsec:class_implementation}
-
-This final stage takes the work done in the previous stages to generate a model that is able to detect malware based on static and dynamic features.
-This stage is to take into consideration the guidelines given in \cite{rossow:practices,shabtai:survey} and the classifiers in \cite{miller:rev_int,nissim:al_pdf,rieck:dynamic,schultz:data_mining}.
-Ideally different classifiers should be tested, as some perform better over others for some set of features.
-The possibility of using multiple classifiers for different features can also be applied to improve the overall result.
-
-\paragraph{Evaluation.} The final result is to be measured using the common evaluation metrics that others\cite{miller:rev_int,nissim:al_pdf,schultz:data_mining} have applied, specifically the \gls{tpr} and \gls{fpr}.
-The \gls{tpr} measures the detection rate of the resulting model, while the \gls{fpr} measures the fall-out.
-
-The described metrics are to be applied in two different conditions: laboratory and real-world.
-For laboratory conditions, the dataset is to be randomly split, using one part for training and other for evaluation.
-For real-world conditions, the dataset is to be split by date, using older samples for training and more recent samples for evaluation.
-By forcing this temporal consistency, the current work will also try to validate the assumption in \cite{miller:rev_int}.
-
-\subsection{Experimental Work}\label{subsec:exp_work}
-
-As a motivation for the work still to be done, a small experiment was conducted to check how easily the creation of a classifier to detect malware is.
-A dataset was created using the static imports as features.
-As for the sample labeling, the classification as given by Microsoft's Windows Defender was used.
-
-Using scikit-learn\cite{tool:sklearn} each sample was converted to a binary vector, where each position corresponded to the presence (or absence) of a static import.
-This feature representation lead to vectors with nearly 30,000 positions (i.e.
-there were almost 30,000 different static imports).
-A logistic regression classifier was trained with 75\% of the samples, using the remaining 25\% for evaluation.
-To check how the dataset balancing affected the classifier, as mentioned in \cite{rossow:practices}, three variants of the dataset were used.
-
-The first dataset was the default and contained more goodware than malware (154,556 goodware, 76,495 malware), this yielded a detection rate 57.2\% with a \gls{fpr} of 28.4\%.
-The second dataset used the same amount of goodware and malware (76,495 samples each), resulting in a detection rate of 69.2\% and a \gls{fpr} of 23.7\%.
-The last dataset contained more malware than goodware (76,495 malware, 38,247 goodware) and gave a detection rate of 92\% with a \gls{fpr} of 22.4\%.
-Table \ref{tab:experimental} summarizes the results and notes the results from \cite{miller:rev_int} as a comparison.
-
-\begin{table}[h!]
-	\centering
-	\def\arraystretch{1.5}
-	\begin{tabular}{lcc}
-		\toprule
-		& Detection Rate        & False Positive Rate        \\
-		\midrule
-		First Dataset  & 0.572     & 0.284      \\ \hline
-		Second Dataset & 0.692     & 0.237      \\ \hline
-		Third Dataset  & 0.92      & 0.224      \\ \hline
-		Miller et al   & [0.65, 0.75] & [0.001, 0.01] \\
-		\bottomrule
-	\end{tabular}
-	\caption{Summary of experimental results compared to Miller et al\cite{miller:rev_int}}
-	\label{tab:experimental}
-\end{table}
-
-This quick test gave some interesting results, although they do not compare to related work, it showed that with little to no work on the dataset and classifier a simple model was achieved.
-It also provided some experience with scikit-learn, which is to be used in the work to come.\\
-
-The work that will result in the proposed system is to be developed between June 2017 and March 2018.
-
-The planned schedule is to use the month of June to finish the data collection phase, together with the labeling phase during the months of June and July.
-The final phase of implementation is planned to take six months, from August to January, as many variables are present and more experiments should provide better results.
-Regarding the writing of the thesis document, the months of February and March are to be used.
-A more detailed planning can be seen in figure \ref{fig:schedule}.
-
-\begin{figure}[h]
-	\includegraphics[width=\textwidth]{gantt}
-	\caption{Planned Schedule}
-	\label{fig:schedule}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%CONCLUSION%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}\label{sec:conclusion}
-
-With the growing amount of malware instances, being able to effectively detect malware as it appears is a must, otherwise cases like WannaCry malware\footnote{Wanna Cry Profile [https://www.fireeye.com/blog/threat-research/2017/05/wannacry-malware-profile.html]} will continue to have real and serious effects\footnote{Huffing Post Article [http://www.huffingtonpost.com/entry/britain-cyber-attack-national-health-service\_us\_5915e62ae4b00f308cf4fc86]}.
-
-When new types of malware arise, anti-virus may be unable to detect them with the usual signature and heuristic methods.
-The application of \gls{ml} to the detection task as shown to be possible and effective\cite{kolter:learning,miller:rev_int,nissim:al_pdf,perdisci:behavior,rieck:dynamic,schultz:data_mining}.
-With that in mind, the current work will develop a \gls{ml} system capable of detecting malicious files, as well as test how different types of evaluations can impact the effectiveness of the solution.
-
-The proposed solution, as described in section \ref{sec:solution}, is to be split in three steps: data collection, data labeling and classifier implementation.
-Some work as been done in each step as to get a feel on what to expect and the problems that might surge along the way.
-
-The next step is to fully terminate each of the sequential stages, with special focus on classifier implementation as many variations can be tested, providing a better research and final result.
-
-Hopefully the current work enlightened the reader on the problem, how the current technologies deal with it and how the proposed work could improve the detection task.
-
-\begin{landscape}
-	\begin{table}
-		\centering
-		\def\arraystretch{1.1}
-		\subfloat{
-			\begin{tabular}{rcccccccccc}
-				\toprule
-				{} &  thehacker &  bitdefender &  cat-quickheal &  fortinet &   virobot &    ikarus &     gdata &    f-prot &  k7antivirus &    clamav \\
-				\midrule
-				Unique names   &   3090 &     3223 &       2286 &  2144 &  3617 &  1139 &  3227 &  2515 &     1869 &  3195 \\
-				\midrule
-				1st most used  &      0.068 &        0.066 &          0.063 &     0.057 &     0.063 &     0.075 &     0.066 &     0.052 &        0.042 &     0.026 \\
-				2nd most used  &      0.047 &        0.047 &          0.052 &     0.050 &     0.028 &     0.040 &     0.047 &     0.035 &        0.035 &     0.022 \\
-				3rd most used  &      0.039 &        0.029 &          0.040 &     0.032 &     0.027 &     0.028 &     0.029 &     0.032 &        0.030 &     0.022 \\
-				4th most used  &      0.030 &        0.025 &          0.030 &     0.027 &     0.025 &     0.027 &     0.025 &     0.030 &        0.029 &     0.017 \\
-				5th most used  &      0.025 &        0.020 &          0.028 &     0.027 &     0.022 &     0.023 &     0.020 &     0.029 &        0.023 &     0.017 \\
-				6th most used  &      0.021 &        0.019 &          0.024 &     0.025 &     0.015 &     0.021 &     0.019 &     0.028 &        0.023 &     0.017 \\
-				7th most used  &      0.019 &        0.017 &          0.023 &     0.019 &     0.013 &     0.017 &     0.017 &     0.026 &        0.023 &     0.017 \\
-				8th most used  &      0.018 &        0.017 &          0.021 &     0.018 &     0.013 &     0.016 &     0.017 &     0.019 &        0.021 &     0.015 \\
-				9th most used  &      0.015 &        0.017 &          0.019 &     0.017 &     0.012 &     0.015 &     0.017 &     0.017 &        0.020 &     0.014 \\
-				10th most used &      0.015 &        0.014 &          0.014 &     0.015 &     0.012 &     0.015 &     0.014 &     0.015 &        0.018 &     0.013 \\
-				\midrule
-				Total          &      0.297 &        0.272 &          0.315 &     0.287 &     0.231 &     0.277 &     0.272 &     0.283 &        0.263 &     0.179 \\
-				\bottomrule
-			\end{tabular}
-		}\\
-		\subfloat{
-			\begin{tabular}{rcccccccccc}
-				\toprule
-				{} &    mcafee &  microsoft &  nprotect &    comodo &     avast &       avg &     vba32 &  kaspersky &  mcafee-gw-edition &  symantec \\
-				\midrule
-				Unique names   &  2560 &   1644 &  4332 &  1959 &  1570 &  2644 &  1366 &   2572 &           2436 &   667 \\
-				\midrule
-				1st most used  &     0.069 &      0.064 &     0.060 &     0.068 &     0.068 &     0.044 &     0.069 &      0.064 &              0.051 &     0.071 \\
-				2nd most used  &     0.037 &      0.040 &     0.030 &     0.061 &     0.062 &     0.032 &     0.030 &      0.052 &              0.036 &     0.060 \\
-				3rd most used  &     0.030 &      0.030 &     0.025 &     0.050 &     0.058 &     0.031 &     0.025 &      0.046 &              0.024 &     0.060 \\
-				4th most used  &     0.030 &      0.028 &     0.020 &     0.036 &     0.043 &     0.025 &     0.025 &      0.041 &              0.024 &     0.053 \\
-				5th most used  &     0.028 &      0.027 &     0.018 &     0.031 &     0.042 &     0.023 &     0.023 &      0.030 &              0.019 &     0.033 \\
-				6th most used  &     0.028 &      0.025 &     0.016 &     0.028 &     0.030 &     0.017 &     0.022 &      0.028 &              0.018 &     0.032 \\
-				7th most used  &     0.024 &      0.025 &     0.016 &     0.027 &     0.024 &     0.017 &     0.017 &      0.025 &              0.017 &     0.032 \\
-				8th most used  &     0.019 &      0.025 &     0.014 &     0.021 &     0.019 &     0.017 &     0.016 &      0.021 &              0.016 &     0.030 \\
-				9th most used  &     0.017 &      0.020 &     0.014 &     0.019 &     0.019 &     0.015 &     0.016 &      0.020 &              0.013 &     0.028 \\
-				10th most used &     0.017 &      0.019 &     0.013 &     0.015 &     0.017 &     0.015 &     0.013 &      0.019 &              0.012 &     0.026 \\
-				\midrule
-				Total          &     0.300 &      0.304 &     0.228 &     0.356 &     0.383 &     0.237 &     0.258 &      0.345 &              0.230 &     0.425 \\
-				\bottomrule
-			\end{tabular}
-		}
-	\caption{Malware name statistics for samples where all top 20 vendors classify as malware (10,868 samples). Tables show the unique different names together with repeating names percentage.}
-	\label{tab:statistics}	
-	\end{table}
-\end{landscape}
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%REFERENCES%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-\nocite{*}
-\bibliography{reportbib}{}
-\bibliographystyle{plain}
-
-\end{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Theoretical Overview}
+%\label{section:overview}
+%
+%Some overview of the underlying theory about the topic...
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Theoretical Model 1}
+%\label{section:theory1}
+%
+%The research should be supported with a comprehensive list of references.
+%These should appear whenever necessary, in the limit, from the first to the last chapter.
+%
+%A reference can be cited in any of the following ways:
+%%
+%\begin{itemize}
+%  \item Citation mode \#1 - \quad \cite{jameson:adjointns}
+%  \item Citation mode \#2 - \quad \citet{jameson:adjointns}
+%  \item Citation mode \#3 - \quad \citep{jameson:adjointns}
+%  \item Citation mode \#4 - \quad \citet*{jameson:adjointns}
+%  \item Citation mode \#5 - \quad \citep*{jameson:adjointns}
+%  \item Citation mode \#6 - \quad \citealt{jameson:adjointns}
+%  \item Citation mode \#7 - \quad \citealp{jameson:adjointns}
+%  \item Citation mode \#8 - \quad \citeauthor{jameson:adjointns}
+%  \item Citation mode \#9 - \quad \citeyear{jameson:adjointns}
+%  \item Citation mode \#10 - \quad \citeyearpar{jameson:adjointns}
+%\end{itemize}
+%%
+%Several citations can be made simultaneously as \citep{nocedal:opt,marta:ijcfd}. \\
+%
+%This is often the default bibliography style adopted (numbers following the citation order), according to the options:\\
+%{\tt \textbackslash usepackage\{natbib\}} in file {\tt Thesis\_Preamble.tex},\\
+%{\tt \textbackslash bibliographystyle\{abbrvnat\}} in file {\tt Thesis.tex}.\\
+%
+%Notice however that this style can be changed from numerical citation order to authors' last name with the options: \\
+%{\tt \textbackslash usepackage[numbers]\{natbib\}} in file {\tt Thesis\_Preamble.tex},\\
+%{\tt \textbackslash bibliographystyle\{abbrvunsrtnat\}} in file {\tt Thesis.tex}. \\
+%
+%Multiple citations are compressed when using the {\tt sort\&compress} option when loading the {\tt natbib} package as {\tt \textbackslash usepackage[numbers,sort\&compress]\{natbib\}} in file {\tt Thesis\_Preamble.tex}, resulting in citations like \citep{marta:ijcfd1,marta:ijcfd2,marta:ijcfd3,marta:ijcfd4}.
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Theoretical Model 2}
+%\label{section:theory2}
+%
+%Other models...

### 2018-04-26 thesis_report/Thesis_Data_Collection.tex
Thesis report update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Data_Collection.tex                                 %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified :  Apr 2018                                        %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Data Collection}
\label{chapter:data_collection}

In the current chapter we detail our initial approach to construct a machine learning model that detects malware based on static and dynamic features.
This first approach was to create a dataset of malware and goodware to feed into a ML model.
We start by describing how the training data was collected, followed by how it was analyzed.
We then proceed to label the samples with help of the aforementioned analysis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Sources}
\label{section:data_sources}

The purpose of this step is to obtain a corpus from which information
about malware and goodware can be extracted.
To minimize any possible bias towards data collection, and to save the overhead time of collecting and analyzing samples, we chose to take advantage of the publicly available reports from Malwr \cite{tool:malwr} service.
As previously mentioned in Section \ref{section:cuckoo}, Malwr is a free online service that does static and dynamic analysis on submitted files using Cuckoo \cite{tool:cuckoo}, which are then accessible in an HTML report.

To enrich our knowledge about samples' \textit{ground truth} we resorted to two other online repositories: \gls{nsrl} \cite{tool:nsrl} and VirusShare.com \cite{tool:virusshare}, these provide metadata (\eg\ MD5 hash) regarding known goodware and malware samples, respectively.
As \gls{nsrl} contains a collection of digital signatures of known, traceable software applications, if a sample is present in this collection, we are more confident it is indeed goodware.
On the other hand, VirusShare.com is a repository of malware samples, hence a sample present in this repository gives us a higher confidence it is indeed malware.

Our data extraction methodology for Malwr and VirusShare.com was based on applying a common data extraction technique, where online data is saved locally or in a central database for further analysis. This technique, called \textit{scraping}, was done using Scrapy \cite{tool:scrapy}, a Python scraping library.

As for the \gls{nsrl} repository, information was provided in textual format, which led us to use Pandas \cite{tool:pandas}, a Python data analysis library, to extract and analyze the data.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Analysis}
\label{section:data_analysis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{High Level Analysis}
\label{subsection:high_level_analysis}

Our first approach was to scrape the list of analysis in Malwr\footnote{Malwr Analysis list [https://malwr.com/analysis/]}.
It is worth noting this scraping did not provide us the full analysis report, only information regarding the type of submission (\eg\ binary, text, archive), its submission date and the sample MD5 hash.

After our initial scraping to obtain a high level view of the available reports, we observed that Malwr had a total of 642,698 submissions, dated between April 16th, 2013 and October 10th, 2016. As can be seen in Figure \ref{fig:samples_count}, there was an increase in sample submission from 2013 up to all time highs in May, June and July of 2014, followed by possibly a downtime period in August 2014.
The service went back up in September 2014, after which it stabilized around 20,000 submissions per month.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.95\textwidth]{Figures/samples_count.png}
	\caption[Number of monthly submissions.]{Number of monthly submissions.}
	\label{fig:samples_count}
\end{figure}

Following this initial observation, we measured the percentage of duplicated submissions per month, which vary between 10\% and 20\%, and the percentage of different file types, as shown in Figure \ref{fig:monthly_distributions}.
The file types distribution is interesting, as it shows that executables (\ie\ compiled binary files) took the great majority of submissions up to the end of 2014, after which documents (\eg\ Word, Excel) and text files (\eg\ HTML, JavaScript) started to increase.
As a side observation, this increase might indicate how users are more suspicious of these type of files, given that it is not uncommon to see malware being distributed in documents (under email attachments) and in websites (as JavaScript scripts) in more recent years.

\begin{figure}[!htb]
	\centering
	\begin{subfigmatrix}{2}
		\subfigure[Percentage of duplicated submissions per month (based on MD5).]{\includegraphics[width=0.95\linewidth]{Figures/samples_dups.png}}
		\subfigure[Distribution of file types per month.]{\includegraphics[width=0.95\linewidth]{Figures/samples_types.png}}
	\end{subfigmatrix}
	\caption[Monthly distributions of submissions.]{Monthly distributions of submissions.}
	\label{fig:monthly_distributions}
\end{figure}

Having a better overall understanding on the available reports, and taking into account the limitations of static analysis in \textcolor{red}{Cuckoo} as mentioned in Section \ref{section:cuckoo}, led us to our first data selection criteria: collect reports from \gls{pe} (\eg\ 32bit executables, DLL's) submissions only.
This selection reduces the number of available reports to 388,702 (60.48\% of the total) but enables us to work with more static information.

We went on to extract these reports, again using Scrapy, as HTML files and kept them in a gzip compressed format to save disk space.
A set $\mathcal{R}$ of raw reports, totaling 388,507 submissions was obtained, (different from the
previous amount as some errors could have occurred during scraping) occupying a total of 26GB.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{VirusTotal Classifications}
\label{subsection:virustotal_classifications}

As previously mentioned in Section \ref{section:cuckoo}, Malwr aggregates the scan report as provided by VirusTotal (if the sample is already known by VirusTotal), which entails multiple antivirus classifications.
Our next analysis takes advantage of this information to better understand how the samples are labeled and how reliable using the VirusTotal report can be to label our dataset.

After counting the number of different antivirus vendors present in at least one submission, we decided to narrow down to the amount of unique vendors by requiring their presence in at least 95\% of the classified reports.
Doing so guarantees that vendors did not show occasionally and have a strong presence in all classified samples.

By applying this criteria, we narrowed from 97 unique vendors to a smaller set $\VV$ of 38.
This led us to a smaller set $\mathcal{C}\subseteq\mathcal{R}$ of classified reports, which totaled 284,880 submissions. With this new set $\mathcal{C}$ we went on to analyze how reliable the vendors classification is.

Reiterating what was already said, defining and classifying malware is not a trivial task.
The \textit{ground truth} of a sample is hard to obtain, as both antivirus vendors and users disagree on what is malware.
To come to terms with this obstacle, we make use of the classifications given by our set of vendors $\mathcal{V}$ to define a simple malware/goodware criteria: what is the minimum/maximum threshold of positive malware classifications as to label a sample as malware/goodware.

Although this criteria bases itself on a vote policy, adding robustness to the criteria, it also inherits some of the vendors problems: false positives and false negatives. With this in mind, we now aim to understand vendors' accuracy and how they change their classifications.

To study how the classifications vary, we took advantage of the duplicated classified submissions in $\mathcal{C}$. There are 27,798 samples submitted more than once to Malwr, for a total of of 74,916
duplicated submissions $\mathcal{C}_{dups}$ averaging 2.7 submissions per
duplicate.

We now show how we studied the differences between the number of positive classifications on the first and last submission, inspired by Miller's et al. \cite{miller:rev_int} work.

For each duplicated sample in $\mathcal{C}_{dups}$ we counted the number
of positive classifications on the first and last submissions, $\mathcal{P}_f$ and $\mathcal{P}_l$ respectively.
If $\mathcal{P}_f > \mathcal{P}_l$ then we are looking
at a possible false positive (FP), as the number of vendors
classifying the sample as malware decreased.
Conversely, if
$\mathcal{P}_l > \mathcal{P}_f$ we are looking at a potential false negative (FN).
For the case $\mathcal{P}_f = \mathcal{P}_l$ we conclude that vendors are confident regarding their classification for the sample. It is worth noting that there may be cases where $\mathcal{P}_f = \mathcal{P}_l$, but with different vendors (i.e. an equal number of vendors simply switch classification), but given we are looking for a criteria based on a minimum threshold, these cases are not relevant for our analysis.

Figure \ref{fig:dups_frequency} shows the frequency of $\mathcal{P}_l-\mathcal{P}_f$ for each duplicated
sample.
We first note that 44.32\% of duplicated samples change in classification, among which 38.72\% increase its classification (green bars), whereas only 5.61\% decrease (red bars).
There is a clear discrepancy between positive and negative changes, suggesting that vendors favor false negatives (\ie\ increase in positive classifications) over false positives (\ie\ decrease in positive classifications), as also noted by Miller et al. \cite{miller:rev_int}.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.95\textwidth]{Figures/dups_frequency.png}
	\caption[Distribution of duplicated samples.]{Distribution of duplicated samples in terms of the changes in the number of
positive classifications between last and first submissions, $\mathcal{P}_l-\mathcal{P}_f$.}
	\label{fig:dups_frequency}
\end{figure}

Another interesting analysis on our samples is understanding the vendors \emph{\gls{dr}} (or True Positive Rate), and \emph{\gls{fpr}}.
Although these formulas are trivially defined respectively as
\begin{eqnarray}
\dr&=&\frac\tp{\#malware}=\frac\tp{\tp+\fn} \\ 
\fpr&=&\frac\fp{\#goodware}=\frac\fp{\tn+\fp}
\end{eqnarray}
we lack ground truth for what is $\#malware$ and $\#goodware$.
To solve this, we must use the available information and propose relative metrics to determine what is positive ($\#malware$) and negative ($\#goodware$).

Our first metric falls back on duplicated submissions again, to define an accuracy metric, $\Madups$. As we have previously shown, 44.32\% of duplicated samples change in classification, which can be translated into vendors acknowledging their own errors.

With that in mind, for each vendor $v\in\VV$, we define a duplicated sample for $v$ (according to $\Madups$) as:
\begin{itemize}
	\item $\tp_v$, \emph{true positive for $v$}, if $v$ classified it positively in both the first and last submissions;
	\item $\tn_v$, \emph{true negative for $v$}, if $v$ classified it negatively in both the first and last submissions;
	\item $\fp_v$, \emph{false positive for $v$}, if $v$ classified it positively in the first submission and negatively in the last submission;
	\item $\fn_v$, \emph{false negative for $v$}, if $v$ classified it negatively in the first submission and positively in the last submission.
\end{itemize}

Figure~\ref{fig:dr_fpr_own} plots each vendors' $\dr_v$ vs.\ $\fpr_v$, under the previously defined metric $\Madups$.
Our first observation is that vendors do acknowledge their own classification errors, as we see a varying $\dr_v$ from 56.82\% to 85.29\%, with a $\fpr_v$ rate ranging from 0.03\% to 6.91\%.
If this was not the case (i.e. vendors would not change their classifications), one would have that $\Pl-\Pf = 0$ for every duplicate and consequently $\dr_v = 1$ and $\fpr_v = 0$, which in turn would mean that all clean samples remain clean, hence $\fn_v=0$, and all malicious samples remain malicious, hence $\fp=0$.
As previously noted, it is apparent that vendors favor \fn\ over \fp,\ as the range for $\fpr_v$ (6.88\%) is much smaller than for $\dr_v$ (28.47\%).

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.95\textwidth]{Figures/dr_fpr_own.png}
	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Madups$.}
	\label{fig:dr_fpr_own}
\end{figure}

The previous analysis gave us a better understanding on vendors accuracy, but is highly dependent on vendors acknowledging their own errors.
To overcome this dependency, we take a second approach regarding vendors' accuracy, which takes into account our observations from Figure \ref{fig:dups_frequency} and our dataset $\CC$ (instead of $\mathcal{C}_{dups}$), to define another metric $\Mac$.

This new metric $\Mac$ implements a simple minimum/maximum threshold of positive classifications to label a sample as malware/goodware.
Intuitively a sample is classified as goodware according to this metric (\ie\  negative), if every vendor $v\in\mathcal{V}$ classifies it as clean.
To understand if our intuition is sound, we plot Figure~\ref{fig:distribution_clean}, a subset of Figure~\ref{fig:dups_frequency}, showing the frequency $\Pl-\Pf$ for samples that were classified as clean in their first submission (\ie\ samples with $\Pf=0$).
These account for 4,902 samples, 3,741 (76.32\%) of which do not increase in classification.

To arrive at a positive (\ie\ malware) sample definition, we relate Figure \ref{fig:distribution_clean} with Figure \ref{fig:dups_frequency}.
Specifically we want to find a minimum threshold of positive classifications to define a sample as malware.
We chose five as the threshold, observing that percentage of samples that decrease in 5 or more positive classifications is 0.46\%, meaning it is an upper bound for samples that decrease from 5 or more to zero classifications.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.95\textwidth]{Figures/distribution_clean.png}
	\caption[Distribution of duplicated samples with $\Pf=0$.]{Distribution of samples that started as goodware and changed in the number of positive classifications between the last and first submission, \ie $\Pl-\Pf$ for samples with $\Pf=0$}
	\label{fig:distribution_clean}
\end{figure}

With these minimum/maximum thresholds for malware/goodware we can now define our new metric $\Mac$ as:

\begin{itemize}
	\item $\tp_v$, \emph{true positive for $v$}, if $v$ and at least 5 other vendors in $\mathcal{V}$ classify it positively;
	\item $\tn_v$, \emph{true negative for $v$}, if $v$ and all other vendors in $\mathcal{V}$ classify it negatively;
	\item $\fp_v$, \emph{false positive for $v$}, if $v$ is the only vendor in $\mathcal{V}$ classifying it positively;
	\item $\fn_v$, \emph{false negative for $v$}, if $v$ classifies it negatively and at least 5 other vendors in $\mathcal{V}$ classify it positively.
\end{itemize}

Having the new metric defined, we again plot each vendors' $\dr_v$ vs.\ $\fpr_v$ according to~$\Mac$, presented in Figure \ref{fig:dr_fpr_d}.
Using this metric we note that vendors' detection rate is more scattered than under $\Madups$, ranging from 20.17\% to 83.50\%, whereas false positive rate is similar, ranging from 0.01\% to 5.77\%.
The scattering along the \tpr\ axis is not surprising, as now we are measuring each vendor $v$ against all others.
What is surprising is the amount of vendors below 50\% detection rate under this metric.
Possible reasoning behind this is either vendors disagree on what is malware (\ie\ \textit{ground truth} problem), or some vendors simply are not good at detecting malware.
As for the \fpr\ axis, it is still interesting to note how the scale remains relatively small, again reflecting the false negative over false positive preference.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.95\textwidth]{Figures/dr_fpr_d.png}
	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Mac$.}
	\label{fig:dr_fpr_d}
\end{figure}

After studying vendors' accuracy under our defined metrics, and given the impossibility of finding a source that is able to unanimously label each sample in our dataset, we want to find the vendors that perform the best under both metrics $\Madups$ and $\Mac$.

Our approach to this problem is to filter the top 20 vendors $\VVdups$ and $\VVc$, according to each metric and define $\VVstar=\VVdups\cap\VVc$.
Given we have to maximize two variables, \tpr\ and \fpr, we decided to take advantage of the linear equation in the form $mx+b=y$ to choose the top vendors.
This form allows us to choose an $m$ and $b$ such that there are 20 vendors above the line, the top vendors $\VVstar$.
By tweaking the variable $m$ one can change the line's steepness, reflecting in a preference between \tpr\ and \fpr\:

\begin{itemize}
	\item \tpr\ preference: $m < 1$, less steepness therefore higher \fpr\ and \tpr\ values;
	\item \fpr\ preference: $m > 1$, more steepness, lower \fpr\ and \tpr\ values.
\end{itemize}

Since we have no \emph{a priori} preference between $\tpr$ nor $\fpr$, we search for the maximum $b$ with $m=1$ such that there are exactly 20 vendors above $x+b=y$ in each graphic (Figures~\ref{fig:dr_fpr_own} and~\ref{fig:dr_fpr_d}).

Figure \ref{fig:dr_fpr_top} shows the $\dr_v$ vs. $\fpr_v$ for the resulting 11 vendors under each metric, $\Madups$ in green and $\Mac$ in red, with $\dr_v$ varying from 63.49\% to 83.50\% and $\fpr_v$ from 0.01\% to 3.67\%.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{Figures/dr_fpr_top.png}
	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Madups$ (green) and $\Mac$ (red).}
	\label{fig:dr_fpr_top}
\end{figure}

With a better understanding on the vendors' accuracy, followed by metrics that allowed us to narrow down the amount of vendors to those that perform best under our metrics, we can now proceed to label our samples as either goodware or malware.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data labeling}
\label{section:data_labeling}

Given we are interested not only in providing a model to detect malware, but also in understanding the impacts of \textit{ground truth} in a detection model, we now go into detail on how we made use of the previously defined top vendors $\VVstar$ in conjunction with the information provided by \gls{nsrl} and VirusShare.com to label the available samples.
To do so, we take the previously defined scenarios in Figure \ref{fig:dataset_sizes} to define three labeling metrics.

The first and most real metric we define is $\Mrealv$ that labels a sample $s\in\CC$ as:

\begin{itemize}
	\item $s\in\mal_\real$ if at least 5 vendors in $\VVstar$ classify $s$ positively;
	\item $s\in\good_\real$ if all vendors in $\VVstar$ classify $s$ negatively.
\end{itemize}

Since the labeling information is solely provided by $\VVstar$'s vendors, this metric's ground truth is highly dependent on their performance, which means labeling errors may be present (as we have discussed in \ref{subsection:virustotal_classifications}).
To minimize this dependency, samples between one and four positive classifications are discarded.

\medskip

Our second metric $\Mloosev$, restricts the previous metric $\Mrealv$ to achieve a better ground truth and takes into account the information from \gls{nsrl}, defined by the set $\CNSRL=\NSRL\ \cap\ \CC$, and VirusShare.com, defined by the set $\CVS=\VS\ \cap\ \CC$.

This metric labels a sample $s\in\mathcal{C}$ as:
\begin{itemize}
	\item $s\in\mal_\loose$ if $s\in\mal_\real$ \textbf{and} it belongs to $\CVS$ and does not belong to $\CNSRL$;
	\item $s\in\good_\loose$ if $s\in\good_\real$ \textbf{and} it belongs to $\CNSRL$ and does not belong to $\CVS$.
\end{itemize}
that is
\begin{eqnarray*}
	\mal_\loose&=&\left(\mal_\real\cap\CVS\right)\setminus\CNSRL\\
	\good_\loose&=&\left(\good_\real\cap\CNSRL\right)\setminus\CVS
\end{eqnarray*}

By taking into account the presence in $\NSRL$, that reinforces cleanliness, and VirusShare.com, that reinforces maliciousness, this metric is more reliable, ground truth wise, at the expense of a smaller number of samples.
% (since $\DNSRL \subset\DD$ and $\DVS \subset \DD$).

\medskip

Our third and final metric $\Mstrictv$, is the strictest one, labeling a sample $s\in\CC$ as:
\begin{itemize}
	\item $s\in\mal_\strict$ if all $v\in\VVstar$ classify it positively \textbf{and} $s\in\CVS\setminus\CNSRL$;
	\item $s\in\good_\strict$ if $s\in\good_\loose$.
\end{itemize}

Obviously this is the most reliable metric, in the sense that it is closely related to the samples' ground truth, leaving little room for disagreement. However, this is achieved again at the cost of a smaller number of samples.

Having these three metrics for labeling, we can now proceed to create the datasets for each of these metrics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dataset creation}
\label{section:dataset_creation}

Taking the previously defined metrics, the task of creating labeled datasets based on them is trivial.
We apply each of the previously defined metrics, $\Mstrictv$, $\Mloosev$ and $\Mrealv$ to our classified dataset $\CC$ to obtained three new datasets, $\CC_{strict}\subseteq\CC_{loose}\subseteq\CC_{real}\subseteq\CC$.
Table \ref{tab:dataset_sizes} provides information regarding the size and number of malware and goodware in each of our datasets.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Malware			& 0 & 0 & 0\\
		Goodware		& 0 & 0 & 0\\
		\midrule
		Total			& 0 & 0 & 0\\
		\bottomrule
	\end{tabular}
	\caption{Sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes}
\end{table}

\medskip

In this chapter we analyzed our available data in order to better understand its advantages and disadvantages.
We provided information regarding our data sources and how the extraction process was made, followed how a simple high level analysis provided interesting insights on the available data.

We took some time understanding how vendors classified reports, how they changed their classifications and how it could impact the reliability of our labels - so much so we defined different metrics that provide different levels of \textit{ground truth} reliability.

Out of these metrics we created three datasets, all subsets of $\CC$, in hopes of enlightening the impacts of \textit{ground truth} when it comes to the task of detecting malware. We are now ready to make use of the datasets to build a malware detection model, which will take place in the following chapters.

### 2018-04-26 thesis_report/Thesis_Glossary.tex
Thesis report update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Glossary.tex                                        %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Andre C. Marta                                           %
%     Last modified : 30 Oct 2012                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% The definitions can be placed anywhere in the document body
% and their order is sorted by <symbol> automatically when
% calling makeindex in the makefile
%
% The \glossary command has the following syntax:
%
% \glossary{entry}
%
% The \nomenclature command has the following syntax:
%
% \nomenclature[<prefix>]{<symbol>}{<description>}
%
% where <prefix> is used for fine tuning the sort order,
% <symbol> is the symbol to be described, and <description> is
% the actual description.

% ----------------------------------------------------------------------

% \glossary{name={\textbf{ML}},description={Machine Learning}}
\nomenclature[ML]{a}{dasdad}
\glossary{name={\textbf{MDO}},description={Multi-Disciplinar Optimization is an engineering technique that uses optimization methods to solve design problems incorporating two or more disciplines.}}

\glossary{name={\textbf{CFD}},description={Computational Fluid Dynamics is a branch of fluid mechanics that uses numerical methods and algorithms to solve problems that involve fluid flows.}}

\glossary{name={\textbf{CSM}},description={Computational Structural Mechanics is a branch of structure mechanics that uses numerical methods and algorithms to perform the analysis of structures and its components.}}

### 2018-04-26 thesis_report/Thesis_Implementation.tex
Thesis report update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Implementation.tex                                  %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Andre C. Marta                                           %
%     Last modified :  2 Jul 2015                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Solution}
\label{chapter:implementation}

In the current chapter we detail our approach to construct a machine learning model that detects malware based on static and dynamic features, as well as practicals applications of such model.

We first define how the training data was collected and analyzed, followed by how it was labeled.
Having the data labeled, we detail what features were used for the machine learning model, based on the available data. We then explain the used model that learned to detect malware based on the selected features, its advantages and disadvantages. To measure how well the model fitted the data we describe how the evaluation was made, with special attention on keeping temporal consistent evaluations (i.e.\ training samples predates validation samples). We close the chapter with how we applied our model in a real-world scenario, as to enhance the credibility and usability of the main subject: malware detection via machine learning.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Sources}
\label{section:data_sources}

The purpose of this step is to obtain a corpus from which information
about malware and goodware can be extracted.
To minimize any possible bias towards data collection, and to save the overhead time of collecting and analyzing samples, we chose to take advantage of the publicly available reports from \textcolor{red}{Malwr} service.
As previously mentioned in \textcolor{red}{malwr cuckoo section}, Malwr is a free online service that does static and dynamic analysis on submitted files using \textcolor{red}{Cuckoo}, which are then accessible in an HTML report.

Our first approach was to apply a common data extraction technique, where online data is saved locally or in a central database for further analysis, to the list of analysis in Malwr\footnote{Malwr Analysis list [https://malwr.com/analysis/]}.
This technique, called \textit{scraping}, was done using \textcolor{red}{Scrapy}, a Python scraping library.
It is worth noting this scraping did not provide us the full analysis report, only information regarding the type of submission (e.g. binary, text, archive), its submission date and the sample MD5 hash.
After narrowing down the type of submissions, we applied a second scraping to Malwr, in order to obtain the full HTML reports. In Section \ref{section:data_analysis} we detail on how we narrowed the type of submissions.

To enrich our knowledge about samples' \textit{ground truth} we resorted to two other online repositories: \textcolor{red}{National Software Reference Library (NSRL)} and \textcolor{red}{VirusShare.com}, these provide metadata (e.g. MD5 hash) regarding known goodware and malware samples, respectively.
As \textcolor{red}{NSRL} contains a collection of digital signatures of known, traceable software applications, if a sample is present in this collection, we are more confident it is indeed goodware.
On the other hand, VirusShare.com is a repository of malware samples, hence a sample present in this repository gives us a higher confidence it is indeed malware.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Analysis}
\label{section:data_analysis}

After our initial scraping to obtain a high level view of the available reports, we observed that \textcolor{red}{Malwr} had a total of 642,698 submissions, dated between April 16th, 2013 and October 10th, 2016. As can be seen in Figure \ref{fig:samples_count}, there was an increase in sample submission from 2013 up to all time highs in May, June and July of 2014, followed by possibly a downtime period in August 2014.
The service went back up in September 2014, after which it stabilized around 20,000 submissions per month.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.95\textwidth]{Figures/samples_count.png}
	\caption[Number of monthly submissions.]{Number of monthly submissions.}
	\label{fig:samples_count}
\end{figure}

Followed this initial observation, we measured the percentage of duplicated submissions per month, which vary between 10\% and 20\%, and the percentage of different file types, as shown in Figure \ref{fig:monthly_distributions}.
The file types distribution is interesting, as it shows that executables (i.e.\ compiled binary files) took the great majority of submissions up to the end of 2014, after which documents (e.g.\ Word, Excel) and text files (e.g.\ HTML, JavaScript) started to increase.
As a side observation, this increase might indicate how users are more suspicious of these type of files, given that it is not uncommon to see malware being distributed in documents (under email attachments) and in websites (as JavaScript scripts) in more recent years.

\begin{figure}[!htb]
	\begin{subfigmatrix}{2}
		\subfigure[Percentage of duplicated submissions per month (based on MD5).]{\includegraphics[width=0.95\linewidth]{Figures/samples_dups.png}}
		\subfigure[Distribution of file types per month.]{\includegraphics[width=0.95\linewidth]{Figures/samples_types.png}}
	\end{subfigmatrix}
	\caption[Monthly distributions of submissions.]{Monthly distributions of submissions.}
	\label{fig:monthly_distributions}
\end{figure}

Having a better overall understanding on the available reports, and taking into account the limitations of static analysis in \textcolor{red}{Cuckoo} as mentioned in \textcolor{red}{malwr cuckoo section}, led us to our first data selection criteria: collect reports from \textcolor{red}{PE} (e.g.\ 32bit executables, DLL's) submissions only.
This selection reduces the number of available reports to 388,702 (60.48\% of the total) but enables us to work with more static information.

We went on to extract these reports, again using \textcolor{red}{Scrapy}, as HTML files and kept them in a gzip compressed format to save disk space.
A set $\mathcal{R}$ of raw reports, totaling 388,507 submissions was obtained, (different from the
previous amount as some errors could have occurred during scraping) occupying a total of 26GB.

As previously mentioned in \textcolor{red}{malwr cuckoo section}, Malwr aggregates the scan report as provided by VirusTotal, which entails multiple antivirus classifications.
Our next analysis takes advantage of this information to better understand how the samples are labeled and how reliable using the VirusTotal report can be to label our dataset.

We started by creating a set $\mathcal{C}\subseteq\mathcal{R}$ of 284,880 classified reports.
This set was achieved by selecting vendors that were present in at least 95\% of the classified reports, as to guarantee that vendors did not show occasionally.
The selected vendors $\mathcal{V}$ total to 37, out of 97 that appear at least once in a report.

Our next analysis was to understand how each sample was classified by different vendors.

%\begin{figure}[!htb]
%	\subfloat[Total samples per month]{
%		\includegraphics[width=0.95\textwidth]{samples_dups.png}
%	}\\
%	\subfloat[Percentage of duplicated (same MD5) samples per month]{
%		\includegraphics[width=0.95\textwidth]{samples_types.png}
%	}
%	\caption{Statistics for the scraped submissions. Count and percentage of duplicates}
%	\label{fig:dist_submissions}
%\end{figure}

\newpage

Description of the numerical implementation of the models explained in Chapter~\ref{chapter:background}...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data labeling}
\label{section:data_labeling}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Feature Selection}
\label{section:feature_selection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Selection}
\label{section:model_selection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation}
\label{section:evaluation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Practical Applications}
\label{section:practical_applications}

### 2018-04-26 thesis_report/Thesis_Introduction.tex
Thesis report update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Introduction.tex                                    %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Andre C. Marta                                           %
%     Last modified :  2 Jul 2015                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\label{chapter:introduction}

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{Figures/dataset_sizes.png}
	\caption{Representation of our $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$ scenarios.}
	\label{fig:dataset_sizes}
\end{figure}

Problema de cyber ataques, frequencia, detecçao.

Problema do malware, quantidade, dificuldade em detectar

Aplicaçao de tecnicas de ML para detectar malware, com resultados bons

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}
\label{section:motivation}

Como ja foi referido, problema do malware, quantidade, dificuldade distinguir.

Bastante trabalho ja feito em malware detection com base em ML, mas com pouca enfase na consistencia temporal das amostras, malware evolve ao longo do tempo.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Main Goal}
\label{section:overview}

Pelas razoes ja descritas, o objectivo desta tese é desenvolver um modelo de ML para detectar malware, mas tendo em conta a consistencia temporal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Thesis Outline}
\label{section:outline}

Chapter \ref{chapter:related_work} introduces malware by describing some of its history, followed by different techniques to detect malware. We then describe a practical application of a specific analysis technique, from which we obtained our dataset.
We introduce the notion of machine learning and its applications, followed by related work that specifically uses \gls{ml} to detect malware.

Chapter \ref{chapter:data_collection} describes our initial approach and methodology to help solve the problem of malware detection using \gls{ml}. We detail the data gathering process and how it was analyzed, followed by the labeling process and dataset creation.

Chapter \ref{chapter:static_features} and \ref{chapter:dynamic_features} takes the created datasets to build a malware detection model based on both static and dynamic features, as well as how these perform under different validation methodologies.

Chapter \ref{chapter:practical_applications} describes how our model is practically implemented both as an online malware detection service (identical to Malwr) and how it was inserted into a corporate email server to analyze attachments.

Chapter \ref{chapter:conclusions} provides some final remarks on our work and contributions, together with possible ideas that may improve the presented work.

### 2018-04-26 thesis_report/Thesis_Practical_Applications.tex
Thesis report update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Practical_Applications.tex                          %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified : Apr 2018                                         %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Practical Applications}
\label{chapter:practical_applications}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Malware Detection Service}
\label{section:malware_service}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Email Attachment Scanner}
\label{section:email_scanner}

### 2018-04-26 thesis_report/Thesis_Static_Features.tex
Thesis report update.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Static_Features.tex                                 %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified : Apr 2018                                         %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Static Features}
\label{chapter:static_features}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Selection and Evaluation}
\label{section:model_selection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Feature Selection}
\label{section:feature_selection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Single Layer Results}
\label{section:single_layer_results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi Layer Results}
\label{section:multi_layer_results}

### 2018-05-02 thesis_report/Thesis_Static_Features.tex
More writing.

@@
 
+In the present chapter we go into details about how we built a machine learning model to detect malware based on static features.
+We start by detailing the model selection and its evaluation, followed by the selected features.
+We then provide the results of our model and some improvements that were made to the model.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Model Selection and Evaluation}
+\label{section:model_selection_evaluation}
+
+In this section we go over the classifier used to create the model that separates malware from goodware, as well as our evaluation methodology.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Model Selection}
 \label{section:model_selection}
+
+Our main concerns when choosing a classifier regard the ability to produce a probabilistic output (\ie\ probability of being malware), good scaling for large number of features and samples and ease of use.
+
+Taking into consideration the guidelines given in~\cite{rossow:practices,shabtai:survey} and related work in~\cite{miller:rev_int,nissim:al_pdf,rieck:dynamic,schultz:data_mining}, we decide to use \textit{\acrfull{lr}} as our model.
+This model fits our needs as it gives the probability of a random variable $X$ being 0 or 1, given a set of constraints (\ie~features), scales well with samples and features and it is readily available from several libraries, facilitating implementation~\cite{friedman2001elements}.
+
+\gls{lr} can be defined with the form
+\begin{eqnarray*}
+	\rho(x) = \dfrac{1}{1 + e^{-x}},&x = \beta_0 + \beta_1x_1 + ... + \beta_nx_n
+\end{eqnarray*}
+where $\beta_n$ is the learned weight for feature $x_n$.
+This weight is learned through iteration in order to minimize the error between the predicted values and the actual values.
+In other words, given an \textit{n-th} dimensional set of features, \gls{lr} will try to create an hyperplane that divides samples from two classes.
+
+As \gls{lr} is based on the logistic function (or sigmoid function, visually represented in Figure \ref{fig:sigmoid_function}), each feature $x_n$ can vary from $-\infty$ to $+\infty$ and still the output is contained between 0 and 1, hence providing probabilistic values.
+
+\begin{figure}[!h]
+	\centering
+	\begin{tikzpicture}
+	\begin{axis}
+	\addplot[
+		color=red,
+		domain=-5:5
+	]
+	{1/(1+exp(-x))};
+	\addlegendentry{$1/(1+e^{-x})$}
+	\end{axis}
+	\end{tikzpicture}
+	\caption{Sigmoid function $f(x)=\dfrac{1}{1+e^{-x}}$}
+	\label{fig:sigmoid_function}
+\end{figure}
+
+To implement this model we take advantage of scikit-learn~\cite{tool:sklearn} library, a Python library that provides a multitude of models and tools that facilitate the creation of \gls{ml} models.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evaluation}
+\label{section:evaluation}
+
+With regards to our evaluation methodology, as we have previously mentioned, our purpose is to understand how laboratory conditions compare to real-world conditions.
+We now detail how we achieve and compare these conditions.
+
+Given the purpose of our work, we choose to measure our results by plotting an \gls{auroc} graph, which measures the \tpr~at different \fpr~levels, metrics that are commonly used across similar work~\cite{miller:rev_int,nissim:al_pdf,schultz:data_mining}.
+
+Our first evaluation methodology is to apply a common \gls{ml} validation technique named \textit{k-fold cross-validation}, with $k=10$, depicted in Figure \ref{fig:xval}.
+This method works by splitting the dataset into $k$ equally sized parts (folds) and then trains using $k-1$ folds, using the remaining fold as validation.
+This process is repeated $k$ times, guaranteeing that every fold is used at least once for validation.
+This methodology is useful to prevent over-fitting, as the validation fold is never used together with the training folds.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_xvalidation.png}
+	\caption[Cross-validation evaluation example.]{Cross-validation evaluation example with $k=10$ folds.}
+	\label{fig:xval}
+\end{figure}
+
+This evaluation methodology is our baseline, hence we define it as our \textit{laboratory} conditions, where no time based consistency is taken into account.
+
+\medskip
+
+Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples.
+Since we want to measure the when training samples predate validation, we now define a couple of temporal based validations, based on the concept of cross-validation:
+
+\paragraph{\textit{Past-to-Present}:} This validation, depicted in Figure \ref{fig:past_present}, can be resumed as an iterative methodology where the validation set is a fixed percentage of dataset's most recent samples, whereas the training set is a fixed percentage of the dataset's oldest.
+The training set is split into $k$ temporal consistent folds and at each iteration it is extended with more recent folds and scored against the training set, until all folds are used.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_pastpresent.png}
+	\caption[Past-to-Present evaluation example.]{Past-to-Present evaluation example with 20\% test, 80\% training, $k=10$ folds in training.}
+	\label{fig:past_present}
+\end{figure}
+
+\paragraph{\textit{Present-to-Past}:} This validation, depicted in Figure \ref{fig:present_past}, is the opposite of \textit{Past-to-Present} with regards to the starting point of the training set.
+Again the validation is fixed at the most recent samples, but the training set starts with the temporally closest fold to the validation set.
+At each iteration the training set is extended, this time with older folds, and scored against the validation, until all folds are used.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_presentpast.png}
+	\caption[Present-to-Past evaluation example.]{Present-to-Past evaluation example with 20\% test, 80\% training, $k=10$ folds in training.}
+	\label{fig:present_past}
+\end{figure}
+
+\medskip
+Both \textit{Past-to-Present} and \textit{Present-to-Past} validations require two parameters: the size of the validation set and how the increments to the training set are made.
+For the purpose of our work, we use the 20\% most recent samples for validation, and split the remaining 80\% into 10 folds, hence the validation is done 10 times, with each iteration increasing the training set size by one fold.
+
+These two validation methodologies give us the ability to account for temporal consistency.
+Moreover, they enable us to compare the importance of older \textit{vs.}\ newer samples to classify recent samples.
+
+\paragraph{\textit{Temporal Window}:} This validation, depicted in Figure \ref{fig:sliding_window}, also splits the dataset into $k$ folds, but takes $w$ temporal consistent and contiguous folds (\ie each fold immediately precedes the next one), using the first $w-1$ folds for training (older samples) and the remaining fold (more recent samples) for validation.
+By starting with the $w$ first folds of the dataset and sliding one fold on each iteration, we apply a sliding window of size $w$ over the dataset.
+For this last validation methodology, we again split the dataset into 10 folds.
+The sliding window size $w$, is chosen during the results phase, as its choice depends on previous results.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[\textit{Temporal Window} evaluation example.]{\textit{Temporal Window} evaluation example with $w=3$ window size over $k=10$ folds.}
+	\label{fig:sliding_window}
+\end{figure}
+
+\medskip
+
+As we have previously described in Section \ref{section:dataset_creation}, we possess 3 different datasets, created from 3 different scenarios.
+With that in mind, all our previously described methodologies: cross-validation, past-to-present and present-to-past; are applied to each of the 3 datasets.
+
+Applying the aforementioned methodologies to our 3 datasets enables us to study not only how temporal consistency impacts the results, but also the impact of \textit{ground-truth}.
 
@@
 
+One of the most important stages in Machine Learning is the selection of the features to analyze, and features based on static imports have shown promising results in \gls{ml} applications for malware detection~\cite{miller:rev_int,schultz:data_mining}.
+In this section we describe the adopted static features that were fed into our model.
+
+Although Cuckoo provides enormous amounts of usable information, we chose to start with simple features as to have a basic understanding of how doable our approach is.
+More so, one of our main concerns is how the same feature gives different results under our different scenarios, hence the performance between scenarios and methodologies is more relevant than absolute performance.
+With that in mind, we chose to use the static imports as features.
+
+Static imports are present in the import table, referencing libraries or other executables that are not part of the current binary.
+Their usage is key, as they provide access to all kind of system level operations (\eg\ file manipulation, network access).
+Given their purpose, our expectation is that a sample's maliciousness can be inferred based on the static imports.
+
+Using Celery~\cite{tool:celery}, a distributed task queue for Python, we optimized the parsing of the available HTML reports, extracting samples that contained information regarding static imports into a new set $\DD_{static}$.
+We then joined the samples with static imports $\FF_{static}$ to the labeled samples $\CC_{real}$, obtaining a total of 189,352 labeled samples with static imports $\DD_{static} = \CC_{real} \cap \FF_{static}$.
+
+We then vectorized imports by creating a binary vector where each position corresponds to a specific import.
+If a given import $i$ is present in a sample, its feature vector $x$ will have the value 1 at that position $x_i$.
+Likewise, if a given import $j$ is not present in a sample, its feature vector $x$ will have the value 0 at that position $x_j$.
+
+Due to the amount of samples and variety of imports, each sample got a vector $x$ of 6,647 dimensions (\ie\ there are 6,647 different imports).
+To reduce this number, and to remove any noise due to incorrect parsing of static imports by Cuckoo, we applied a variance threshold.
+
+The variance threshold calculates the variance for each import, removing those that are below a given threshold.
+In our case, since we are working with a binary vector, each import can be represented as Bernoulli random variable, hence their variance is given by $p(1-p)$.
+With that in mind, we removed any import that did not vary in more than 99\% of samples.
+
+The resulting dataset $\FF_{static}$ got reduced to 187,545 samples, each with a 64 dimensional binary vector.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Having met all the necessary conditions to build a \gls{ml} model: model selection and evaluation and feature selection, we are now ready to provide the results of our first experiments.
+
+Using an Ubuntu Virtual Machine with 16 cores and 16GB of RAM, provided by INOV - Inesc Inovação, and taking advantage of scikit-learn~\cite{tool:sklearn} to train the models, Pandas~\cite{tool:pandas} for data analysis and Jupyter Interactive Notebooks~\cite{tool:jupyter} to interact with the Python scripts, we now detail our logistic regression model results.
+
+We define this our single layer model $\LR$ results, as we are directly taking a set of features (static imports in this case) and outputting the probability of a sample being malware.
+
+\medskip
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsection{Laboratory Conditions}
+% \label{section:single_layer_laboratory}
+
+We start with what we define as \textit{laboratory conditions}, ideal conditions for the problem of malware detection.
+These conditions are met when we take the labeled dataset $\CC_{strict}$ and features $\DD_{static}$, providing scenario $\mathcal{S}_{\strict}$.
+
+Using cross-validation, our $\LR$ scored with an \gls{auroc} of \miss{score}, as shown by the red line in Figure \ref{fig:xval_results}.
+We argue that such high values are easily attained from factors like a small and reliable dataset, and the use of cross-validation, which mixes samples and ignores possible dependencies on malware samples.
+
+When we relax our ideal conditions to more \textit{real-world conditions}, by testing the labeled dataset $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{static}$ for scenarios $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, respectively, the score under \gls{auroc} decreases to \miss{score} for $\mathcal{S}_{loose}$ (blue line in Figure \ref{fig:xval_results}) and \miss{score} for $\mathcal{S}_{real}$ (green line in Figure \ref{fig:xval_results}).
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[Single layer results for static features in laboratory conditions.]{Single layer results for static features in laboratory conditions.}
+	\label{fig:xval_results}
+\end{figure}
+
+From $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$, the only change is the amount of malware labeled samples, which significantly increase.
+The difference is interesting, as although the number of malware labeled samples increase significantly, the results are not that affected.
+This suggests that although the reliability for malware decreases, its impact is not as noticeable as expected.
+This might also suggest that vendors do converge on their definition of malware, under our $\Mloosev$ metric.
+If vendors did not converge on what is malware, adding more samples would culminate in worse results, as separation between malware and goodware would become harder.
+
+When looking at the changes from $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$, not only the amount of malware labeled samples increase, but also the number of goodware labeled samples, both by a significant amount.
+The way this impacts the results is pretty significant, as we observe a high decrease in the \gls{auroc} for both models.
+The metric $\Mrealv$ that labels malware and goodware for this scenario $\mathcal{S}_{\real}$ disregards the cross-check from outside repositories, which in turn degrade the reliability significantly, as well as increase the dataset size notably.
+We attribute the results' degradation mainly to the unreliability of goodware labeling, not only because we have previously seen that increase in malware does not significantly impact results (from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$), but also due to the tendency for false negatives in vendors (Figure \ref{fig:dups_frequency}), which in turn lead us to incorrectly label goodware for the samples in $\CC_{real}$.
+
+The results we described show how relaxing \textit{laboratory conditions} to more \textit{real-world conditions} degrade the model's performance.
+We now focus on using our previously defined temporal based methodologies to further converge into a real-world scenario.
+
+\medskip
+
+We start by applying our \textit{Past-to-Present} validation to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
+As previously defined, this validation starts with an older set of training samples and iteratively adds newer samples, validating each iteration on a fixed set of the most recent samples.
+Since our interest is to measure performance variation over time, we plot in Figure~\ref{fig:pastpresent} the \gls{auroc} at every iteration (\ie, fold), for each of our three scenarios.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[Single layer results for static features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation.
+Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
+fold 1)}
+	\label{fig:pastpresent}
+\end{figure}
+
+When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note a decrease from \miss{score} to \miss{score} for $\mathcal{S}_{\strict}$, \miss{score} to \miss{score} for $\mathcal{S}_{\loose}$, and \miss{score} to \miss{score} for $\mathcal{S}_{\real}$.
+This decrease is intuitive to the methodology, as we are forcing temporal consistency between samples.
+
+When looking at results from $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$, we see they closely relate.
+This relation has already been noticed on previous cross-validation results, as their metrics $\Mstrictv$ and $\Mloosev$ are not very different.
+As for $\mathcal{S}_{\real}$, the degradation is higher, as the reliability of the metric $\Mrealv$ goes down.
+
+Our main observation for this validation methodology is that there is a slight tendency for \gls{auroc} to increase, as we move forward in time, close to the validation set.
+
+Both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ behave similarly under our \textit{Past-to-Present} evaluation, decreasing the \gls{auroc} until fold \miss{fold}, after which both increase, stabilizing around \miss{score} \gls{auroc} by the last \miss{two} folds.
+As for $\mathcal{S}_{\real}$, its \gls{auroc} does not vary significantly, although a subtle increase over time is noticeable.
+
+\medskip
+
+Our next result, which uses our \textit{Present-to-Past} validation methodology will further help analyze the aforementioned detail.
+The \textit{Present-to-Past} validation enhances the previous results under real-world conditions.
+This methodology starts by fixing the validation set to the most recent samples, but with the training set starting at the temporally closest samples to validation.
+At each iteration, older samples are added to the training set and validated on the fixed, most recent, samples.
+
+By applying this methodology to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we plot Figure \ref{fig:presentpast}, where the X axis increases as older samples are added to the training set (\ie\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time.
+Similarly to the previous observation, the average \gls{auroc} suffers a decrease when compared to cross-validation.
+For $\mathcal{S}_{\strict}$ we note a change from \miss{score} to \miss{score}, for $\mathcal{S}_{\loose}$ the decrease is from \miss{score} to \miss{score}, and for $\mathcal{S}_{\real}$ \miss{score} to \miss{score}.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[Single layer results for static features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation.
+Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
+	\label{fig:presentpast}
+\end{figure}
+
+The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ display very similar results, with $\mathcal{S}_{\real}$ dropping behind due to its less reliable labeling metric.
+
+With these results, our original observation that samples closer to the validation set benefit the model becomes more convincing.
+In fact, we argue that there should be an ideal number of necessary training folds, temporally consistent with the validation fold (\ie\ any fold from training predates validation), needed to maximize the overall score.
+
+\medskip
+
+Finally, we analyze how does such reduced training set behaves in our scenarios; for this purpose, we define a sliding window that moves forward in time through each scenario for training and validation.
+We propose a reduction on the training size to \miss{$w=3$} folds predating the validation fold.
+We choose \miss{$w = 3$}, since we have seen that the scores either do not improve (for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$) or actually go down (for $\mathcal{S}_{\real}$) with higher folds.
+In summary, we have selected \miss{30\%} of each dataset for training purposes and the next \miss{10\%} for validation \miss{(3 training folds, 1 validation fold)}, and then started moving the window forward in time (1 fold at a time) to obtain the following results (Figure~\ref{fig:slidingwindow}): for  $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we obtain \gls{auroc} values of \miss{score, score and score}, respectively.
+These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[Single layer results for static features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
+	\label{fig:slidingwindow}
+\end{figure}
+
+Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from \miss{0.93 to 0.87 for $\mathcal{S}_{\strict}$, from 0.92 to 0.86 for $\mathcal{S}_{\loose}$ and from 0.79 to 0.76 for $\mathcal{S}_{\real}$.}
+We should highlight that the results that use temporal-consistency should better reflect the reality than standard cross-validation, since we are requiring temporal consistency.
+Another important idea that should be stressed is that for cross validation we used a fairly reasonable amount of data for training purposes, whereas in this last case we used a restricted amount of data.
+This might be a relevant issue in a few year's time.
+The results obtained are summarized in Table \ref{tab:singlelayer_results}
+
+\begin{table}[!htb]
+	\renewcommand{\arraystretch}{1.2} % more space between rows
+	\centering
+	\begin{tabular}{ccccc}
+	\hline AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
+	\hline Cross-Validation & 0.93 & 0.92 & 0.79 & 90 / 10\\
+	\hline Past-to-Present & 0.88 & 0.87 & 0.71 & 10 to 90 / 10\\
+	\hline Present-to-Past & 0.90 & 0.86 & 0.72 & 10 to 90 / 10\\
+	\hline Sliding-Window & 0.87 & 0.86 & 0.76 & 30 / 10\\
+	\hline
+\end{tabular}
+\caption{Single layer results summary.}
+\label{tab:singlelayer_results}
+\end{table}
+
+\medskip
+
+With a better understanding of how the model behaves under different methodologies, we now diverge to how we improved not only the overall results, but also the information provided by the model.
+
+\todo[inline]{Diria talvez começar um novo capitulo, so de improvements, tal como o multilayer e uso de mais features}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-05-04 thesis_report/Thesis_Model_Improvements.tex
More on improvements.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Model_Improvements.tex                              %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified : Apr 2018                                         %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Model Improvements}
\label{chapter:model_improvements}

Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this chapter to present the improvements made in order to obtain a more robust model to detect malware.
We start by describing our first improvement, applying a multi layer model to extract more information regarding a sample.
We then take this enhanced model and increase the number of features to include dynamic content and how it impacted the model's results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi Layer Model}
\label{section:improvements_multi_layer}

On the previous chapter we ended up with a simple \gls{lr} model $\LR$ that given a set of static imports from a sample, would give the probability of it being malware.
Although that was the end purpose of our work, it does not give any deeper understanding of the sample, other than being or not malware.
With that in mind, we propose a new model that not only provides the probability of a sample being malware, but also the probability of being from a specific malware type.

This new model $\mathcal{E}$ comprises a simple ensemble stacking approach, which instead of using a single \gls{lr} classifier, multiple ones are used, layered into two steps.

The first step (layer $\mathcal{E}_{\mathcal{L}_{0}}$) is composed of $n$ \gls{lr} models, where $n$ is the number of possible classes.
Each model is trained to output the likelihood of sample belonging to one of the $n$ classes, in a \textit{one-vs-all} methodology (\ie\ a sample either belongs to $\mathcal{C}_{n}$ or not), having as input the raw features (\eg\ static imports).

The second step (layer $\mathcal{E}_{\mathcal{L}_{1}}$) is identical to $\mathcal{LR}$, but now takes as features the output of each classifier from the previous layer, outputting the likelihood of a sample being malware.

In summary, as depicted in Figure \ref{fig:dia_multilayer}, we define a 2 layer ensemble stacking with $n$ classifiers on the first layer to a single classifier in the second layer.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/dia_multilayer.png}
	\caption{Multi layer model representation.}
	\label{fig:dia_multilayer}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Malware Classes}

With this new model defined, we now present our approach on selecting the $n$ classes of interest.
This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware as belonging to some subclass.

To choose the malware classes for our datasets, we take into account our description in Chapter \ref{chapter:related_work} regarding propagation methods and malware names based on purpose.
With this in mind, we chose 6 malware classes: \textit{virus}, \textit{trojan}, \textit{worm}, \textit{ransom}, \textit{spyware} and \textit{other}.
The first three classes, \textit{virus}, \textit{trojan} and \textit{worm}, were chosen from the propagation methods, whereas \textit{ransom} and \textit{spyware}
where chosen due to their popularity in recent years. The last class, \textit{other}, serves for any malware that does not fit the previous five.

As mentioned in Chapter \ref{chapter:related_work}, there is no agreed upon naming convention for malware, which translates into different names for the same malware sample.
To minimize this problem, we referenced a tool by Sebastián, M. et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, using the names provided by VirusTotal~\cite{tool:virustotal}.

We took advantage of this tool and modified it such that instead of providing a family name, it would provide one (or more) of the 6 previously defined classes.
Specifically, we changed it in a way that given a set of malware names, the output would be a distribution over the 6 malware classes.

To calculate each class weight we apply the following formula
\begin{eqnarray*}
	\mathcal{W}_c = \dfrac{f_c}{\sum\limits_{c}f_c}
\end{eqnarray*}

where $f_c$ is the frequency for the class $c$ and $\sum_{c}f_c$ is the number of times all classes appear.
For example, if a given set of names contain the name \textit{trojan} 3 times and the name \textit{virus} one time, then the weights would be
\begin{eqnarray*}
	\mathcal{W}_{trojan}=\dfrac{3}{4}=0.75,~\mathcal{W}_{virus}=\dfrac{1}{4}=0.25,~ \mathcal{W}_{c}=0, c \in \{worm, spyware, other, ransom\}
\end{eqnarray*}

\medskip

Having these malware classes defined for our multi layer model, we also added the \textit{goodware} class for samples that are not malware.
Doing so gives us 7 possible classes, 6 of which are malware only.
It is worth mentioning that if a sample belongs to the \textit{goodware} class, it cannot belong to any other, likewise if it belongs to any malware class, it cannot belong to the \textit{goodware} class.

In sum, we added a layer of labeling to our dataset, where a sample can either belong to the \textit{goodware} class, or to a set of the other 6 malware classes.
This concludes our first improvement to our model, what follows are our improvements regarding features.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamic Features}
\label{section:improvements_dynamic_features}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Improved Model Results}
\label{section:improvements_results}

### 2018-05-08 thesis_report/Thesis_Data_Collection.tex
More writing.
...

@@
 
+The following provide a detailed analysis of the available data sources, done by using Pandas~\cite{tool:pandas}, a Python Data Analysis Library.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\centering
-	\includegraphics[width=0.95\textwidth]{Figures/samples_count.png}
+	\includegraphics[width=\textwidth]{Figures/samples_count.png}
 	\caption[Number of monthly submissions.]{Number of monthly submissions.}
@@
 	\begin{subfigmatrix}{2}
-		\subfigure[Percentage of duplicated submissions per month (based on MD5).]{\includegraphics[width=0.95\linewidth]{Figures/samples_dups.png}}
-		\subfigure[Distribution of file types per month.]{\includegraphics[width=0.95\linewidth]{Figures/samples_types.png}}
+		\subfigure[Percentage of duplicated submissions per month (based on MD5).]{\includegraphics[width=\linewidth]{Figures/samples_dups.png}}
+		\subfigure[Distribution of file types per month.]{\includegraphics[width=\linewidth]{Figures/samples_types.png}}
 	\end{subfigmatrix}
@@
 	\centering
-	\includegraphics[width=0.95\textwidth]{Figures/dups_frequency.png}
+	\includegraphics[width=\textwidth]{Figures/dups_frequency.png}
 	\caption[Distribution of duplicated samples.]{Distribution of duplicated samples in terms of the changes in the number of
@@
 	\centering
-	\includegraphics[width=0.95\textwidth]{Figures/dr_fpr_own.png}
+	\includegraphics[width=\textwidth]{Figures/dr_fpr_own.png}
 	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Madups$.}
@@
 	\centering
-	\includegraphics[width=0.95\textwidth]{Figures/distribution_clean.png}
+	\includegraphics[width=\textwidth]{Figures/distribution_clean.png}
 	\caption[Distribution of duplicated samples with $\Pf=0$.]{Distribution of samples that started as goodware and changed in the number of positive classifications between the last and first submission, \ie $\Pl-\Pf$ for samples with $\Pf=0$}
@@
 	\centering
-	\includegraphics[width=0.95\textwidth]{Figures/dr_fpr_d.png}
+	\includegraphics[width=\textwidth]{Figures/dr_fpr_d.png}
 	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Mac$.}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{Figures/dr_fpr_top.png}
+	\includegraphics[width=\textwidth]{Figures/dr_fpr_top.png}
 	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Madups$ (green) and $\Mac$ (red).}

### 2018-05-08 thesis_report/Thesis_Model_Improvements.tex
...

@@
 Doing so gives us 7 possible classes, 6 of which are malware only.
-It is worth mentioning that if a sample belongs to the \textit{goodware} class, it cannot belong to any other, likewise if it belongs to any malware class, it cannot belong to the \textit{goodware} class.
+It is worth mentioning that if a sample belongs to the \textit{goodware} class, it cannot belong to any other, likewise, if it belongs to any malware class, it cannot belong to the \textit{goodware} class.
+
+By applying this new labeling, we refine our previous datasets $\CC_{strict}$, $\CC_{loose}$ and $\CC_{real}$ as shown in Table \ref{tab:dataset_sizes_new}.
+
+\begin{table}[!htb]
+	\renewcommand{\arraystretch}{1.2} % more space between rows
+	\centering
+	\begin{tabular}{lccc}
+		\toprule
+		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
+		\midrule
+		Trojan			& 0 & 0 & 0\\
+		Virus			& 0 & 0 & 0\\
+		Worm			& 0 & 0 & 0\\
+		Spyware			& 0 & 0 & 0\\
+		Ransom			& 0 & 0 & 0\\
+		Other			& 0 & 0 & 0\\
+		\midrule
+		Total Malware	& 0 & 0 & 0\\
+		Total Goodware	& 0 & 0 & 0\\
+		\midrule
+		\midrule
+		Total			& 0 & 0 & 0\\
+		\bottomrule
+	\end{tabular}
+	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
+	\label{tab:dataset_sizes_new}
+\end{table}
 
 In sum, we added a layer of labeling to our dataset, where a sample can either belong to the \textit{goodware} class, or to a set of the other 6 malware classes.
-This concludes our first improvement to our model, what follows are our improvements regarding features.
+This concludes the description of our first improvement to the model, what follows are our improvements regarding features.
 
@@
 
+Having described our first model improvement, we now proceed to our second enhancement, which takes advantage of the amount of information available in Cuckoo's~\cite{tool:cuckoo} reports.
+
+In the previous chapter, we used static imports as features for our malware detection model.
+Although the results are reasonable, the information which can be retrieved from static imports alone is limited.
+As an example, if a sample is compressed, encrypted or packed, its behavior cannot be inferred from static imports only.
+To overcome these limitations, we resort to more dynamic information provided by Cuckoo.
+
+Given that Cuckoo runs the provided samples inside a virtual machine, it monitors the sample from a dynamic point of view.
+This monitoring includes information like the sequence of library calls and their categories and network and file activity.
+With this in mind, we now detail what dynamic information we extracted from the reports.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Category Calls}
+\label{section:improvements_categories}
+
+The first type of dynamic information we extracted were the number of dynamic category calls.
+When Cuckoo runs and monitors a sample, it registers some low level library calls, which it then assigns to a fixed number of categories.
+
+There are a total of 14 different categories defined by Cuckoo: \textit{anomaly}, \textit{device}, \textit{filesystem}, \textit{hooking}, \textit{misc}, \textit{network}, \textit{process}, \textit{registry}, \textit{services}, \textit{socket}, \textit{synchronization}, \textit{system}, \textit{threading} and \textit{windows}.
+After using Celery~\cite{tool:celery} to extract the number of each category calls for the samples, we obtained a total of 148,036 samples with information regarding category calls.
+
+Figure \ref{fig:boxplot_category_calls} shows a boxplot for the category calls, where one can see the categories \textit{filesystem}, \textit{registry} and \textit{system} are the ones with higher mean, hence more used than the other categories.
+Although not depicted, some outliers for the number of category calls go up to 1,400,000, which do not provide useful information other than the number of category calls is very high.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=\textwidth]{Figures/boxplot_category_calls.png}
+	\caption{Category calls boxplot.}
+	\label{fig:boxplot_category_calls}
+\end{figure}
+
+With that in mind, we decide to transform the values into a $[0,1]$ range.
+We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
+This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large values fall into the same bin.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Library Calls}
+\label{section:improvements_api_unigrams}
+
+Our second type of dynamic information are the number of library calls, which closely relate to the previous feature.
+While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call, hence being a subset of the previous.
+
+Cuckoo~\cite{tool:cuckoo} registers the number of calls for 163 different functions, ranging from opening and closing files, to opening and closing sockets.
+Again we used Celery~\cite{tool:celery} to extract these numbers, obtaining information from 148,036 samples.
+
+Given we are dealing with a high number of features (163 different library calls), we decided to apply the same variance threshold as in Section \ref{section:feature_selection}, to remove features that do not vary in most samples.
+By choosing a threshold of 80\%, we remove library calls that do not vary in more than 80\% of the samples, effectively reducing the number of library calls to 144.
+With regards to how these features can vary from 0 to $+\infty$, as before, we again apply a quantile transformer with a normal distribution.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Cuckoo Signatures}
+\label{section:improvements_categories}
+
+For our third and last type of dynamic information, we resort to Cuckoo~\cite{tool:cuckoo} custom signatures.
+These signatures are built from certain activities that Cuckoo deems malicious or suspicious.
+For example, if a sample allocates memory and then makes it executable, it might suggest some sort of packing or obfuscation.
+
+Although some signatures are defined by a sequence of library calls, which could also be attained by creating n-grams of our previously defined feature, using the already defined Cuckoo signatures saves training time.
+As an example, creating bi-grams of library calls, given there are 144 unique calls, would yield over 20,000 possible features, hence being unpractical.
+
+To extract these signatures, we use Celery~\cite{tool:celery} and obtain a total of 124,821 samples and 61 different signatures.
+As with our static import features, we use a binary vector for each sample, where each position corresponds to a specific signature.
+Re-iterating on how we represent this, if a given signature $i$ is present in a sample, its feature vector $x$ will have the value 1 at that position $x_i$.
+Likewise, if a given signature $j$ is not present in a sample, its feature vector $x$ will have the value 0 at that position $x_j$.
+
+\medskip
+
+Having these new features, we create a new set with all features (static, category and library calls and signatures) $\FF_{dynamic}$.
+We joined these samples $\FF_{dynamic}$ to the labeled samples $\CC_{real}$, obtaining a total of \miss{size} labeled samples with dynamic imports $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$.
+
+% To reduce the number of features into those that are more representative, we apply, as in other features, a variance threshold, removing signatures that do not vary in more than 80\% of the samples.
+% By doing so, we reduce the total number of signatures to
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{section:improvements_results}
+
+In the previous sections we described all our modifications to the original model $\LR$: a two layer approach that splits malware into subclasses and the addition of three dynamic features.
+We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{section:evaluation}.
+Specifically we test the model using the baseline cross-validation methodology, followed by our three temporally consistent scenarios: \textit{Past-to-Present}, \textit{Present-to-Past} and \textit{Temporal Window}.
+We test each methodology using the three different scenarios: $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
+
+We again use the resources provided by INOV - Inesc Inovação, together with scikit-learn~\cite{tool:sklearn}, Pandas~\cite{tool:pandas} and Jupyter Interactive Notebooks~\cite{tool:jupyter} to train and analyze the data.
+
+\medskip
+
+Starting with \textit{laboratory conditions}, we apply the cross-validation evaluation to model $\mathcal{E}$ with the labeled dataset $\CC_{strict}$ and features $\DD_{dynamic}$, providing scenario $\mathcal{S}_{\strict}$.
+For this scenario, we obtain an \gls{auroc} of \miss{score}, as presented by the red line in Figure \ref{fig:improved_xval}.
+These provide the highest results, provided by factors like a small and reliable dataset, and the use of cross-validation.
+
+Relaxing to more \textit{real-world conditions}, under the form of a less reliable ground truth, we test the datasets $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{dynamic}$.
+As shown in Figure \ref{fig:improved_xval}, the score under \gls{auroc} decreases to \miss{socre} for $\mathcal{S}_{\loose}$ (blue line) and \miss{score} for $\mathcal{S}_{\real}$ (green line).
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption{Improved model results with dynamic features in laboratory conditions.}
+	\label{fig:improved_xval}
+\end{figure}
+
+% Adicionar seccao para resultados do multi layer apenas.

### 2018-05-08 thesis_report/Thesis_Static_Features.tex
More on improvements.
...

@@
 We start by detailing the model selection and its evaluation, followed by the selected features.
-We then provide the results of our model and some improvements that were made to the model.
+We finish off by providing how our model performed under various scenarios.
 
@@
 
-Using Celery~\cite{tool:celery}, a distributed task queue for Python, we optimized the parsing of the available HTML reports, extracting samples that contained information regarding static imports into a new set $\DD_{static}$.
+Using Celery~\cite{tool:celery}, a distributed task queue for Python, we optimized the parsing of the available HTML reports, extracting samples that contained information regarding static imports into a new set $\FF_{static}$.
 We then joined the samples with static imports $\FF_{static}$ to the labeled samples $\CC_{real}$, obtaining a total of 189,352 labeled samples with static imports $\DD_{static} = \CC_{real} \cap \FF_{static}$.
@@
 
-The resulting dataset $\FF_{static}$ got reduced to 187,545 samples, each with a 64 dimensional binary vector.
+The resulting dataset $\DD_{static}$ got reduced to 187,545 samples, each with a 64 dimensional binary vector.
 
@@
 With a better understanding of how the model behaves under different methodologies, we now diverge to how we improved not only the overall results, but also the information provided by the model.
-
-\todo[inline]{Diria talvez começar um novo capitulo, so de improvements, tal como o multilayer e uso de mais features}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Multi Layer Results}
-\label{section:multi_layer_results}

### 2018-05-08 thesis_report/Thesis_Introduction.tex
Intro v1

@@
 
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=\columnwidth]{Figures/dataset_sizes.png}
-	\caption{Representation of our $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$ scenarios.}
-	\label{fig:dataset_sizes}
-\end{figure}
+The number of reported security breaches due to \emph{virus}, \emph{trojans}, \emph{ransomwares}, etc, has been growing considerably in recent years~\cite{av-test:report} with reports of infections due to \emph{malware}  making the headlines, now more than ever.
+Almost every week one such security vulnerability is reported which may be seen as a failure by the security community on the control and detection of malicious content.
 
-Problema de cyber ataques, frequencia, detecçao.
-
-Problema do malware, quantidade, dificuldade em detectar
-
-Aplicaçao de tecnicas de ML para detectar malware, com resultados bons
+Due to the significant number of malware attacks, and taking into consideration the increasing popularity and huge success of \gls{ml} methodologies in classification in different domains~\cite{lee2003learning,joachims2002learning,li2010object,ding2001multi}, it is only natural to see these techniques applied to complement classical methodologies for malicious-content detection~\cite{arp2014drebin,christodorescu:semantics,kolosnjaji2016deep,kolter:learning,miller:rev_int,nissim:al_pdf,perdisci:behavior,rieck:dynamic,santos2013opcode,schultz:data_mining,schwenk2012autonomous,vsrndic2013detection,deo2016prescience,gandotra2014malware,jordaney2017transcend,rossow:practices}, in particular, supervised learning techniques.
 
@@
 
-Como ja foi referido, problema do malware, quantidade, dificuldade distinguir.
+Malware, which can be referred to by other names like malicious software or malicious code, can be described as \say{any code added, changed, or removed from a software system in order to	intentionally cause harm or subvert the intended function of the system}~\cite{mcgraw:mal_code}.
+Malware can be a piece of code attached to a legitimate program, an independent program, or a combination of both.
 
-Bastante trabalho ja feito em malware detection com base em ML, mas com pouca enfase na consistencia temporal das amostras, malware evolve ao longo do tempo.
+The reasoning that leads to malware creation varies greatly, some are created to demonstrate some vulnerability or concept and do not cause direct harm to systems, while others are used to steal information or render systems useless.
+The wide variety and quantity of now-a-day systems and the amount of sensitive information they store provide numerous opportunities to illegally make a profit out of subverting legitimate systems.
+
+The amount of malware has shown an exponential increase since the early 2000s, from 1 million in 2006 to over 500 million by the end of 2016~\cite{av-test:report,mcafee:report}.
+According to a 2007 study by McAfee\footnote{McAfee Study Finds 4 Percent of Search Results Malicious, by Frederick Lane, June 4, 2007 [http:\/\/www.newsfactor.com\/story.xhtml?story\_id=010000CEUEQO]} about 4\% of the query results by major search engines lead to potentially dangerous websites.
+A more recent study by Norton in 2010\footnote{Symantec Press Release, July 29, 2010 [https://www.symantec.com/en/aa/about/newsroom/press-releases/2010/symantec\_0729\_02]} showed that at least 10\% of top trending search terms returned malicious results.
+From the contact with malicious software, 1.1 billion identities (\ie\ user's personal information) have been exposed in 2016 alone~\cite{symantec:report}.
+
+To detect malicious software, anti-virus solutions rely on two main methods: \textit{signature-based}, which uses a database of known malware to detect malware and \textit{heuristic-based}, which make use of malicious patterns and sets of rules to detect both known and new malware~\cite{gryaznov:scanners}.
+
+To help in the task of detecting both new and old malware, the use of \gls{ml} methodologies have shown promising results~\cite{arp2014drebin,christodorescu:semantics,kolosnjaji2016deep,kolter:learning,miller:rev_int,nissim:al_pdf,perdisci:behavior,rieck:dynamic,santos2013opcode,schultz:data_mining,schwenk2012autonomous,vsrndic2013detection} when dealing with this task, but not without its pitfalls~\cite{deo2016prescience,gandotra2014malware,jordaney2017transcend,rossow:practices}.
+
+The very first pitfall is the definition of \textit{malware}, not only the lack of an agreed upon definition brings the problem of having different anti-virus vendors classifying the same program differently, but also some programs fall within a gray area for which no clear classification can be deemed correct.
+An example is what is called \textit{adware}, advertising-supported software, that although not performing directly malicious actions, perform arguably non-requested actions.
+The non-existence of concrete metrics and properties that uniquely distinguish malware from goodware requires an extra effort in the preparation of datasets for evaluation of malware detectors.
+
+Other problematic when using \gls{ml} for malware detection is how to evaluate the trained model.
+While many classification tasks are time independent, one might argue that malware detection is not.
+As such, applying common evaluation strategies that disregard for any temporal consistency affects the model performance~\cite{deo2016prescience,jordaney2017transcend,rossow:practices}.
 
@@
 
-Pelas razoes ja descritas, o objectivo desta tese é desenvolver um modelo de ML para detectar malware, mas tendo em conta a consistencia temporal.
+The \acrfull{ml} field studies the development of computer programs that improve with experience.
+A program is given a task and an experience, if its performance at the task improves with the experience given, then the program is learning the task~\cite{mitchell:ml}.
+
+In this work we took advantage of \gls{ml} capabilities and the growing number of malware samples to investigate how one can develop a model that takes into account the aforementioned pitfalls.
+
+To achieve this, we took an incremental approach to build a malware detection model, enabling us to measure the impacts of \textit{ground-truth} (\ie\ the gray area when it comes to malware), as well as how enforcing temporal consistent learning influences the final results.
+
+To achieve our \textit{ground-truth} impact objective we made a comparative analysis of a supervised learning approach in three different scenarios (depicted in Figure~\ref{fig:scenarios}):
+
+\begin{itemize}
+	\item \emph{strict scenario} where only very well-characterized samples are considered;
+	\item \emph{loose scenario} where a wider set of still well-studied samples is considered;
+	\item \emph{realistic scenario} where we get closer to the reality faced by vendors of malware detection solutions.
+\end{itemize}
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{Figures/dataset_sizes.png}
+	\caption{Representation of our $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$ scenarios.}
+	\label{fig:scenarios}
+\end{figure}
+
+To understand the impacts of temporal consistent learning, we perform a comparative analysis of the three above scenarios under two distinct environments:
+
+\begin{itemize}
+	\item \textit{laboratory conditions} where traditional cross-validation methodologies can be applied;
+	\item \emph{real-world conditions} where time is relevant and we analyze the behavior of the classifier with temporal-based methodologies.
+\end{itemize}
+
+We also provide several improvements to the model.
+One of these come under the form of a multi-layer approach, which instead of classifying a sample as malware or goodware, also provides information regarding the type of malware.
+Our other improvement regards providing more information about a sample to the model, in order to obtain higher scores.
 
@@
 
-Chapter \ref{chapter:static_features} and \ref{chapter:dynamic_features} takes the created datasets to build a malware detection model based on both static and dynamic features, as well as how these perform under different validation methodologies.
+Chapter \ref{chapter:static_features} and \ref{chapter:model_improvements} takes the created datasets to build a malware detection model based on both static and dynamic features, as well as how these perform under different validation methodologies.
 
-Chapter \ref{chapter:practical_applications} describes how our model is practically implemented both as an online malware detection service (identical to Malwr) and how it was inserted into a corporate email server to analyze attachments.
+Chapter \ref{chapter:practical_applications} describes how our model is practically implemented both as an online malware detection service and how it was inserted into a corporate email server to analyze attachments.
 

### 2018-05-08 thesis_report/Thesis_Model_Improvements.tex
more thesis.

@@
 For this scenario, we obtain an \gls{auroc} of \miss{score}, as presented by the red line in Figure \ref{fig:improved_xval}.
-These provide the highest results, provided by factors like a small and reliable dataset, and the use of cross-validation.
+These provide the highest results, provided by factors like a small and reliable dataset, more features and the use of cross-validation.
 
@@
 % Adicionar seccao para resultados do multi layer apenas.
+
+As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are not highly affected, when the change between the scenarios is merely in the number of malware samples.
+
+Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the lowest score when using the most realistic dataset.
+
+\todo[inline]{}
+
+When looking at the changes from $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$, not only the amount of malware labeled samples increase, but also the number of goodware labeled samples, both by a significant amount.
+The way this impacts the results is pretty significant, as we observe a high decrease in the \gls{auroc} for both models.
+The metric $\Mrealv$ that labels malware and goodware for this scenario $\mathcal{S}_{\real}$ disregards the cross-check from outside repositories, which in turn degrade the reliability significantly, as well as increase the dataset size notably.
+We attribute the results' degradation mainly to the unreliability of goodware labeling, not only because we have previously seen that increase in malware does not significantly impact results (from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$), but also due to the tendency for false negatives in vendors (Figure \ref{fig:dups_frequency}), which in turn lead us to incorrectly label goodware for the samples in $\CC_{real}$.
+
+The results we described show how relaxing \textit{laboratory conditions} to more \textit{real-world conditions} degrade the model's performance.
+We now focus on using our previously defined temporal based methodologies to further converge into a real-world scenario.
+
+\medskip
+
+We start by applying our \textit{Past-to-Present} validation to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
+As previously defined, this validation starts with an older set of training samples and iteratively adds newer samples, validating each iteration on a fixed set of the most recent samples.
+Since our interest is to measure performance variation over time, we plot in Figure~\ref{fig:pastpresent} the \gls{auroc} at every iteration (\ie, fold), for each of our three scenarios.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[Single layer results for static features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation.
+Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
+		fold 1)}
+	\label{fig:pastpresent}
+\end{figure}
+
+When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note a decrease from \miss{score} to \miss{score} for $\mathcal{S}_{\strict}$, \miss{score} to \miss{score} for $\mathcal{S}_{\loose}$, and \miss{score} to \miss{score} for $\mathcal{S}_{\real}$.
+This decrease is intuitive to the methodology, as we are forcing temporal consistency between samples.
+
+When looking at results from $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$, we see they closely relate.
+This relation has already been noticed on previous cross-validation results, as their metrics $\Mstrictv$ and $\Mloosev$ are not very different.
+As for $\mathcal{S}_{\real}$, the degradation is higher, as the reliability of the metric $\Mrealv$ goes down.
+
+Our main observation for this validation methodology is that there is a slight tendency for \gls{auroc} to increase, as we move forward in time, close to the validation set.
+
+Both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ behave similarly under our \textit{Past-to-Present} evaluation, decreasing the \gls{auroc} until fold \miss{fold}, after which both increase, stabilizing around \miss{score} \gls{auroc} by the last \miss{two} folds.
+As for $\mathcal{S}_{\real}$, its \gls{auroc} does not vary significantly, although a subtle increase over time is noticeable.
+
+\medskip
+
+Our next result, which uses our \textit{Present-to-Past} validation methodology will further help analyze the aforementioned detail.
+The \textit{Present-to-Past} validation enhances the previous results under real-world conditions.
+This methodology starts by fixing the validation set to the most recent samples, but with the training set starting at the temporally closest samples to validation.
+At each iteration, older samples are added to the training set and validated on the fixed, most recent, samples.
+
+By applying this methodology to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we plot Figure \ref{fig:presentpast}, where the X axis increases as older samples are added to the training set (\ie\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time.
+Similarly to the previous observation, the average \gls{auroc} suffers a decrease when compared to cross-validation.
+For $\mathcal{S}_{\strict}$ we note a change from \miss{score} to \miss{score}, for $\mathcal{S}_{\loose}$ the decrease is from \miss{score} to \miss{score}, and for $\mathcal{S}_{\real}$ \miss{score} to \miss{score}.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[Single layer results for static features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation.
+Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
+	\label{fig:presentpast}
+\end{figure}
+
+The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ display very similar results, with $\mathcal{S}_{\real}$ dropping behind due to its less reliable labeling metric.
+
+With these results, our original observation that samples closer to the validation set benefit the model becomes more convincing.
+In fact, we argue that there should be an ideal number of necessary training folds, temporally consistent with the validation fold (\ie\ any fold from training predates validation), needed to maximize the overall score.
+
+\medskip
+
+Finally, we analyze how does such reduced training set behaves in our scenarios; for this purpose, we define a sliding window that moves forward in time through each scenario for training and validation.
+We propose a reduction on the training size to \miss{$w=3$} folds predating the validation fold.
+We choose \miss{$w = 3$}, since we have seen that the scores either do not improve (for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$) or actually go down (for $\mathcal{S}_{\real}$) with higher folds.
+In summary, we have selected \miss{30\%} of each dataset for training purposes and the next \miss{10\%} for validation \miss{(3 training folds, 1 validation fold)}, and then started moving the window forward in time (1 fold at a time) to obtain the following results (Figure~\ref{fig:slidingwindow}): for  $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we obtain \gls{auroc} values of \miss{score, score and score}, respectively.
+These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\caption[Single layer results for static features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
+	\label{fig:slidingwindow}
+\end{figure}
+
+Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from \miss{0.93 to 0.87 for $\mathcal{S}_{\strict}$, from 0.92 to 0.86 for $\mathcal{S}_{\loose}$ and from 0.79 to 0.76 for $\mathcal{S}_{\real}$.}
+We should highlight that the results that use temporal-consistency should better reflect the reality than standard cross-validation, since we are requiring temporal consistency.
+Another important idea that should be stressed is that for cross validation we used a fairly reasonable amount of data for training purposes, whereas in this last case we used a restricted amount of data.
+This might be a relevant issue in a few year's time.
+The results obtained are summarized in Table \ref{tab:singlelayer_results}
+
+\begin{table}[!htb]
+	\renewcommand{\arraystretch}{1.2} % more space between rows
+	\centering
+	\begin{tabular}{ccccc}
+		\hline AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
+		\hline Cross-Validation & 0.93 & 0.92 & 0.79 & 90 / 10\\
+		\hline Past-to-Present & 0.88 & 0.87 & 0.71 & 10 to 90 / 10\\
+		\hline Present-to-Past & 0.90 & 0.86 & 0.72 & 10 to 90 / 10\\
+		\hline Sliding-Window & 0.87 & 0.86 & 0.76 & 30 / 10\\
+		\hline
+	\end{tabular}
+	\caption{Single layer results summary.}
+	\label{tab:singlelayer_results}
+\end{table}

### 2018-05-09 thesis_report/Thesis_Background.tex
Background v1

@@
 The following chapter will go over prior work that closely relates to the topic of our research and areas of contribution.
-We approach the definition of malware and how it can hard to detect and label by providing some background on the subject.
+We approach the definition of malware and how it can be hard to detect and label by providing some background on the subject.
 We then proceed to elucidate on different malware detection techniques, from simple methods to more complex approaches, focusing on a specific malware analysis technique that provided the basis for our work.
@@
 We advance to the general topic of \gls{ml}, after which we describe related work that studies the use of \gls{ml} in malware detection.
-Lastly, we reference prior work that not only tackles our same goal, but also takes into account the impact of laboratory \textit{vs.} real world scenarios, as well as the notion of temporal based detection, or concept drift.
-
-\newpage
-\todo[inline]{CSF Related work}
-
-We use this section to present prior work that closely relates to the topic of our research and our areas of contribution.
-Specifically, detecting malicious software by training supervised models on static information, and validation methodologies that resemble real-world conditions.
-
-Shabtai et al.\cite{shabtai:survey} provide a survey directed
-at the application of \gls{ml} classifiers to detect malware from static
-features.
-Their work concerns the design and evaluation of such systems.
-Our contributions are inspired by the concern of how to correctly evaluate \gls{ml} classifiers, regarding size, reliability of labeling metrics and chronological evaluation.
-
-In the topic of methodologies that resemble real-world conditions, Srndi\'c et al.\cite{vsrndic2013detection}
-train and validate their malicious PDF detector under laboratory and real-world conditions.
-Laboratory conditions consist on applying regular cross-validation, whereas real-world conditions validate a newer dataset with a model created from outdated data (i.e.\ older then the validation), and also validate the model when the validation set spans one week and the training is gathered in the previous 4 weeks.
-They show that laboratory conditions inflate the results, when compared to real-world conditions.
-
-Miller et al.\cite{miller:rev_int} also introduce sample temporal consistency.
-They show the impact of performance measurement technique on a dataset containing 1.1 million samples, when using cross-validation and temporally consistent samples.
-As noted by others, regular cross-validation showed inflated results when compared to temporally consistent samples.
-
-Our work enhances these methodologies by analyzing the performance variation when the distance between the training and validation set increases and decreases, as well as analysis on how reducing the size of the training without compromising the results.
-
-Kolter et al.\cite{kolter:learning} learn to detect malicious executables in a dataset with under 4,000 samples, obtained from reliable sources, and evaluate their model under standard cross-validation and by gathering newer malware samples to validate for new and unseen samples.
-They show how the model provides optimal results under cross-validation, but for the unseen samples lower scores are obtained.
-Our work considers these results to compare how reliability affects performance.
-
-Sebastián, M.
-et al.\cite{sebastian2016avclass} develop \textit{AVClass}, a tool that given a set of antivirus vendors, outputs the most likely family name.
-They test their tool under 10 datasets, totaling 8.9 M samples, with results showing an F1 measure up to 93.9 on labeled datasets.
-The tool takes as input the labels as seen in VirusTotal, tokenizes the labels, replaces known aliases and general names (e.g.\ win32, trojan, generic), ending with possible family names.
-These remaining names are counted and the most frequent one is given as family name.
-Our work takes advantage of this tool, not to label malware families, but minimal modifications, to label more general malware classes (e.g.\ trojan, virus).
-
-Deo, A.
-et al.~\cite{deo2016prescience} focus over the problem of \gls{ml} models becoming antiquated over time, given how malware evolves.
-This problem is seen as concept drift, where the performance of a model over time diminishes, as the statistical properties of malware (i.e.\ features) change over time.
-They study how probabilistic predictors can help minimize the aforementioned problem, by indicating when retraining of a given model is necessary.
-
-Jordaney, R.
-et al.~\cite{jordaney2017transcend} also study the problem of concept drift in malware classification models, focusing on providing metrics, based on statistical comparison of samples, to detect when should a model be retrained.
-
-Both ~\cite{deo2016prescience} and ~\cite{jordaney2017transcend} are on the subject of our work, regarding the problem of concept drift, but differ on the study-case.
-In our work we acknowledge the concept but focus on its relative effects and how one can balance the size of needed training data vs.
-validation data, when maintaining a temporal consistent dataset.
-Whereas their work provide indicators for when should a model be retrained when the problem of concept drift becomes significant.
-
-\newpage
-\todo[inline]{Project Report related work}
+Lastly, we reference prior work that not only tackles our same goal, but also takes into account the impact of \textit{laboratory} \textit{vs.} \textit{real world} scenarios, as well as the notion of temporal based detection, or concept drift.
 
@@
 
-Malware, which can be referred to by other names like malicious software or malicious code, can be described as \say{a program that has malicious intent}\cite{christodorescu:semantics}, or more detailed as any code added, changed, or removed from a software system in order to intentionally cause harm or subvert the intended function of the system\cite{mcgraw:mal_code}.
+The term malware was first used by Yisrael Radai in 1990~\cite{elisan:malware}, before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983~\cite{cohen:virus}.
+Given term computer virus predates the term malware, it is not uncommon to see both terms used interchangeably.
+% For the purpose of this work the term malware is adopted, for reasons explained in the following paragraphs.
 
-The term malware was first used by Yisrael Radai in 1990\cite{elisan:malware}, before then malicious software was referred to as computer viruses, a notion which was first formalized by Cohen in 1983\cite{cohen:virus}.
-Given term computer virus predates the term malware, it is not uncommon to see both terms used interchangeably.
-For the purpose of this work the term malware is adopted, for reasons explained in the following paragraphs.
-
-To the exponential growth of malware\cite{av-test:report}, the ability to distinguish between malware samples is crucial.
+To the exponential growth of malware~\cite{av-test:report}, the ability to distinguish between malware samples is crucial.
 In the biological world, different types of infections reckon distinctive disinfections, the same applies to the digital world.
@@
 
-With respect to the propagation method, malware can be split into three classes\cite{kolter:learning}.
+With respect to the propagation method, malware can be split into three classes~\cite{kolter:learning}:
 
-\paragraph{Virus.}\footnote{This work uses virus as a subset of malicious code, hence the choice of using the term malware over virus.}This type of malware inherits its name from resembling to biological viruses.
+\paragraph{Virus.}%\footnote{This work uses virus as a subset of malicious code, hence the choice of using the term malware over virus.}
+This type of malware inherits its name from resembling to biological viruses.
 A virus is usually composed by two main subroutines.
-The first subroutine is responsible for infecting other programs by attaching the virus code, while the second subroutine is the actual malware payload that contains the malware purpose (i.e.
-virus payload)\cite{chen:evolution}.
-Viruses are attached to programs and propagate when an infected program is run, hence depending on other programs (as hosts) and user interaction.
-For viruses to infect new systems, an infected file must be carried between them by a user (e.g.
-USB pen, email attachment), reinforcing the user's role in this propagation method.
+The first subroutine is responsible for infecting other programs by attaching the virus code, while the second subroutine is the actual malware payload that contains the malware purpose (\ie\ virus payload)~\cite{chen:evolution}.
+Viruses are attached to programs and propagate when an infected program is run, either consciously (\eg\ clicking on executables) or unconsciously (\eg\ auto-run features), hence depending on other programs (as hosts) and user interaction.
+For viruses to infect new systems, an infected file must be carried between them by a user (\eg\ USB pen, email attachment), reinforcing the user's role in this propagation method.
 
@@
 A worm shares the self-replicating ability of viruses, but discards user interaction and the need for a host.
-This is possible by making a worm an independent program that exploits networks to find vulnerable systems to infect with a copy of themselves\cite{chen:evolution}.
-Similarly to viruses, worms can contain additional payload to perform malicious actions on infected systems (e.g.
-Blaster worm exploited a vulnerability in RPC for propagation while its payload flooded a Microsoft domain\footnote{W32/Blaster worm, CERT, August 11, 2003 [http://www.cert.org/historical/advisories/CA-2003-20.cfm]}).
+This is possible by making a worm an independent program that exploits networks to find vulnerable systems to infect with a copy of themselves~\cite{chen:evolution}.
+Similarly to viruses, worms can contain additional payload to perform malicious actions on infected systems (\eg\ Blaster worm exploited a vulnerability in RPC for propagation while its payload flooded a Microsoft domain\footnote{W32/Blaster worm, CERT, August 11, 2003 [http://www.cert.org/historical/advisories/CA-2003-20.cfm]}).
 
 \paragraph{Trojan.} While viruses and worms focus on self-propagation, trojans focus on deceiving users into executing them, disregarding propagation.
-Trojans try to appeal users with some useful functionality as to allure into running the program\cite{szor:art}.
-By hoaxing users into running them, trojans bypass the need for a host (in the virus case) or an exploit (in the worm case) to perform malicious actions.\\
+Trojans try to appeal users with some useful functionality as to allure into running the program~\cite{szor:art}.
+By hoaxing users into running them, trojans bypass the need for a host (in the virus case) or an exploit (in the worm case) to perform malicious actions.
 
-These three classes are not exclusive and can be used in conjunction to facilitate propagation (e.g.
-a virus that relies on a trojan to start propagating).
-Given the inexistent standards for malware naming, some literature\cite{szor:art} define worms as a subset of viruses, as some worms start by attaching to files.
-With the general infection methods layered out, the following will succinctly describe the general naming when referring to the malware purpose (i.e.
-its payload).
+\medskip
 
-Given the high numbers of malware multiple purposes imply multiple names, as such the following lists some of the most commonly heard names and definitions as given by McAfee\cite{mcafee:glossary}:
+These three classes are not exclusive and can be used in conjunction to facilitate propagation (\eg\ a virus that relies on a trojan to start propagating).
+Given the inexistent standards for malware naming, some literature~\cite{szor:art} define worms as a subset of viruses, as some worms start by attaching to files.
+With the general infection methods layered out, the following will succinctly describe the general naming when referring to the malware purpose (\ie\ its payload).
+
+Given the high numbers of malware multiple purposes imply multiple names, as such the following lists some of the most commonly heard names and definitions as given by McAfee~\cite{mcafee:glossary}:
 
@@
 	\item \textit{Backdoor}: Installs or takes advantage of an unknown entry point to a user's system, allowing remote control of a system.
+	\item \textit{Bot}: Short for \say{robot}; A program that receives commands from a cybercriminal and executes them blindly, due to this characteristic, compromised computers are called zombies.
 	\item \textit{Dropper}: Program that facilitates the introduction of other malware instances into a system.
-	\item \textit{Keylogger}: Program that monitors a user's input (e.g. keyboard strokes) to steal private information.
+	\item \textit{Keylogger}: Program that monitors a user's input (\eg\ keyboard strokes) to steal private information.
 	\item \textit{Password Stealer (PWS)}: Program that specifically targets a user's personal information, like usernames and passwords.
 	\item \textit{Ransomware}: Program that restricts access to a system/files (usually by encryption) and demands a ransom to allow a user his access.
+	\item \textit{Remote Administration Tool - RAT}: Program designed to give an administrator remote control of a system.
+Similar to \textit{Backdoor} and \textit{Bot}, but usually has higher privileges on infected systems.
 	\item \textit{Rootkit}: Program designed to hide the existence of other (malicious) programs, hindering detection of malicious applications.
 	\item \textit{Scareware}: Program that fakes the existence of malware and tricks the user into paying for disinfection or into downloading real malware.
-	\item \textit{Spyware}: Program that spies on a user's activity through multiple vectors (e.g. taking prints of a user's desktop/webcam).
+	\item \textit{Spyware}: Program that spies on a user's activity through multiple vectors (\eg\ taking prints of a user's desktop/webcam).
 \end{itemize}
@@
 
-One of the first attempts was made in 1991, by the \gls{caro}\cite{caro:naming}.
+One of the first attempts was made in 1991, by the \gls{caro}~\cite{caro:naming}.
 This attempt was aimed at viruses only and focused on discriminating and relating virus samples.
@@
 
-To enable differentiation from other types of malware, an improvement to \gls{caro}'s naming was proposed in 1999 by Geral Scheidl\cite{scheidl:naming}.
-Scheidl's proposal included a prefix that consisted of the platform targeted by the malware and a type (i.e.
-previously defined classes and names), rendering a more in-depth name:
+To enable differentiation from other types of malware, an improvement to \gls{caro}'s naming was proposed in 1999 by Geral Scheidl~\cite{scheidl:naming}.
+Scheidl's proposal included a prefix that consisted of the platform targeted by the malware and a type (\ie\ previously defined classes and names), rendering a more in-depth name:
 
@@
 
-In the Virus Bulletin magazine of January 2002\cite{virus_bulletin} the naming problem is again addressed, referencing that only a small number of vendors apply \gls{caro}'s names and that the lack of cooperation between anti-virus vendors lead to multiple aliases for malware names.
+In the Virus Bulletin magazine of January 2002~\cite{virus_bulletin} the naming problem is again addressed, referencing that only a small number of vendors apply \gls{caro}'s names and that the lack of cooperation between anti-virus vendors lead to multiple aliases for malware names.
 Again a naming convention is proposed with focus on detail by using three levels of information: malware name, classification and full text description.
@@
 
-In section \ref{sec:introduction} a brief description of malware detection methods was given.
 This section will provide a deeper description of detection methods, from the simplest techniques to more complex ones.
-The following is based on the work by Peter Szor, \textit{The Art of Computer Virus Research and Defense}\cite{szor:art}.
+The following is based on the work by Peter Szor, \textit{The Art of Computer Virus Research and Defense}~\cite{szor:art}, which although not recent, provides a solid background on detection methods.
 
@@
 
-\subsubsection{First-Generation Scanners.} These type of scanners work at a fairly simple level and focus searching for known byte sequences (i.e.
-known to be malware sequences).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{First-Generation Scanners}
+\label{subsection:first_gen_scanners}
+
+These type of scanners work at a fairly simple level and focus searching for known byte sequences (\ie\ known to be malware sequences).
 The applied principles include:
@@
 
-\paragraph{Mismatches.} Also similar to string scanning, yet there can be $N$ number of bytes in the string that can take any value and in any position (e.g.
-the string \say{\texttt{01 02}} with mistmatch of 1 matches \say{\texttt{01 AA}} and \say{\texttt{AA 02}}).
+\paragraph{Mismatches.} Also similar to string scanning, yet there can be $N$ number of bytes in the string that can take any value and in any position (\eg\ the string \say{\texttt{01 02}} with mistmatch of 1 matches \say{\texttt{01 AA}} and \say{\texttt{AA 02}}).
 
-\paragraph{Generic Detection.} This method typically applies both wildcards and mismatches to create generic strings that can scan for several variants of a family of malware.
+\paragraph{Generic Detection.} This method typically applies both wildcards and mismatches to create generic strings that can scan for several variants of a family of malware, hence being more robust to subtle changes in malware.
+
+\medskip
+
+These principles can be assigned to the signature-based type of methods, as they depend on searching some previously defined byte sequence to be effective.
+Their effectiveness depends on an updated knowledge base with the strings to be matched.
+
+Alongside these basic principles, the following can be used in conjunction to improve the scanning speed and reliability:
 
 \paragraph{Hashing.} This term is not a method by itself, but techniques that speed up searching algorithms, enhancing previous methods.
+It takes advantage of the hash table data structure to match malware signatures.
 
-\paragraph{Bookmarks.} This method is both used to provide more accurate detections and disinfections. Bookmarks provide specific locations for where strings are to be searched (i.e. offsets).
+\paragraph{Bookmarks.} This method is both used to provide more accurate detections and disinfections. Bookmarks provide specific locations for where strings are to be searched (\ie\ offsets).
 
@@
 It does so by bypassing operating system level \gls{api} to read the disk directly with the \gls{bios}.
-It is worth noting that this technique is hard to implement nowadays due to the high variety of file systems and disk controllers.\\
+In addition to improve the scanners performance, this principle is also useful as an anti-stealth technique, given it works below the filesystem layer.
+It is worth noting that this technique is hard to implement nowadays due to the high variety of file systems and disk controllers.
 
-The described methods, excluding the ones that focus on speed improvement, can be assigned to the signature-based type of methods, as they depend on searching some previously defined byte sequence to be effective.
-Their effectiveness depends on an updated knowledge base with the strings to be matched.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Second-Generation Scanners}
+\label{subsection:second_gen_scanners}
 
-\subsubsection{Second-Generation Scanners.} These methods improve on the previous generation by doing an analysis at a higher level of abstraction.
+These methods improve on the previous generation by doing an analysis at a higher level of abstraction.
 Where first-generation scanners would match sequences of bytes, second-generation scanners also look at the bytes representation.
@@
 
-\paragraph{Smart Scanning.} This method takes the assembly representation of the bytes into account and instructions that have no effect (e.g.
-\texttt{NOP}) can be skipped.
+\paragraph{Smart Scanning.} This method takes the assembly representation of the bytes into account and instructions that have no effect (\eg\ \texttt{NOP}) can be skipped.
 The resulting signature is smaller and more easily detects mutations.
-The same principle can be applied to textual malware (e.g.
-scripts), where extra white spaces have no effect.
+The same principle can be applied to textual malware (\eg\ scripts), where extra white spaces have no effect.
 
-\paragraph{Skeleton Detection.} This method is specifically useful in detecting textual malware (e.g.
-macros) families.
+\paragraph{Skeleton Detection.} This method is specifically useful in detecting textual malware (\eg\ macros) families.
 By parsing the macro statements line by line, the scanner can drop nonessential statements.
@@
 First-generation scanners base detection mainly on a byte sequence, whereas nearly exact identification makes use of double-string detection.
+The principle comes from using two strings to match a given malware, if only one of those strings match, then it is likely to be a variant of a known malware.
 This is specifically useful to identify same family malware.
-Another nearly exact identification method applies a checksum to a range from the malware body, the resulting checksum can then be used to detect and disinfection files.
+Another nearly exact identification method applies a checksum to a range from the malware body, this has the added benefit of better accuracy without overloading the anti-virus database.
 
 \paragraph{Exact Identification.} This method enhances the previous method by adding the ability to differentiate malware variants.
-It does so by using as many checksum ranges as needed to identify variants of the same malware, improving the accuracy and facilitating disinfection.\\
+It does so by using as many checksum ranges as needed to identify variants of the same malware, improving the accuracy and facilitating disinfection.
+
+\medskip
 
@@
 
-\subsubsection{Algorithmic Scanning Methods.} These methods are not so focused on direct matching of signatures, but on the routines implemented in the scanner that allow the detection of malware.
-Early implementations of algorithmic scanning were hard-coded in the antivirus, but with different types and families emerging, the hard-coded algorithms could easily become obsolete.
-To overcome this, vendors would introduce more specific routines for malicious programs when updating the antivirus.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Algorithmic Scanning Methods}
+\label{subsection:algorithmic_scanning}
+
+These methods are not so focused on direct matching of signatures, but on the routines implemented in the scanner that allow the detection of malware.
+Early implementations of algorithmic scanning were hard-coded in the anti-virus, but with different types and families emerging, the hard-coded algorithms could easily become obsolete.
+To overcome this, vendors would introduce more specific routines for malicious programs when updating the anti-virus.
+
 Examples of these methods are filtering, where signatures are only applied if an object meets a particular filter; static decryptor, which detects encrypted malware; and X-ray, which tries to decrypt malware that use simple encryption methods (e.g.
@@
 
-\subsubsection{Code Emulation.} This detection technique uses dynamic analysis (i.e.
-analysis during runtime) to detect malware by simulating code execution.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Code Emulation}
+\label{subsection:code_emulation}
+
+This detection technique uses dynamic analysis (\ie\ analysis during runtime) to detect malware by simulating code execution.
 A virtual machine is used to simulate the system's components, thus the malicious files are encapsulated and no code is executed by the real processor.
+
 Code emulation can be faster than using algorithmic scanning methods when dealing with encrypted/obfuscated malware, as the code can be emulated and the decrypted/deobfuscated code is much easier to obtain.
@@
 
-\subsubsection{Heuristic Analysis.} This detection technique takes into account several characteristics of malicious software, both static and dynamic.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Heuristic Analysis}
+\label{subsection:heuristic_analysis}
+
+This detection technique takes into account several characteristics of malicious software, both static and dynamic.
 Examples of such characteristics are code execution starting in uncommon sections, suspicious code redirection or incorrect size of code in header.
+
 Using the previously described methods, heuristic analysis can take into account different characteristics that are present or absent to assert if a file is deemed malicious or not.
+
 The upside of this technique is that it allows for never seen malware to be correctly flagged, but brings the downside of a higher amount of false positives.
 
-\subsubsection{Sand-Boxing.} This detection technique builds on the already mentioned code emulation, but with a higher level of complexity.
-With sand-boxing the whole system is emulated (i.e.
-virtualized) and suspicious or malicious programs have access to a copy of the real resources, effectively limiting the effects of the program and consequently protecting the system.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Sand-Boxing}
+\label{subsection:sand_boxing}
 
-This technique allows to extract dynamic information without compromising the real system, which can then be both used to detect if a suspicious program is malware, by applying the aforementioned techniques, and to help analyze new types of malware.
+This detection technique builds on the already mentioned code emulation, but with a higher level of complexity.
+With sand-boxing the whole system is emulated (\ie\ virtualized) and suspicious or malicious programs have access to a copy of the real resources, effectively limiting the effects of the program and consequently protecting the system.
+
+This technique allows to extract dynamic information without compromising the real system, which can then be used to detect if a suspicious program is malware, by applying the aforementioned techniques, and to help analyze new types of malware.
 This technique demands more resources then previous ones and may not be ideal to use inside an anti-virus application.
-Other downsides of sand-boxing come as compatibility issues, the possibility of malware detecting virtualization, which can hinder analysis, or even if the virtualization software is vulnerable and malicious code can escape the sandbox into the host machine.\\
 
-Detecting malware is not as simple as applying one of the mentioned techniques, different methods work best in different scenarios, hence antivirus need to balance the usage of these techniques to keep systems protected.
+Other downsides of sand-boxing come as compatibility issues, the possibility of malware detecting virtualization, which can hinder analysis, or even if the virtualization software is vulnerable and malicious code can escape the sandbox into the host machine.
+
+\medskip
+
+Detecting malware is not as simple as applying one of the mentioned techniques, different methods work best in different scenarios, hence anti-virus need to balance the usage of these techniques to keep systems protected.
 
@@
 Building on the last described method of sand-boxing, this section will describe a specific implementation of sand-boxing, namely Cuckoo.
-Cuckoo Sandbox is a malware analysis system created by Claudio Guarnieri\cite{tool:cuckoo}.
+Cuckoo Sandbox is a malware analysis system created by Claudio Guarnieri~\cite{tool:cuckoo}.
 
@@
 Cuckoo facilitates analysis by automating this process.
-The Cuckoo system will take the sample to analyze and run it inside a predefined virtual environment (e.g.
-Windows XP).
+
+The Cuckoo system will take the sample to analyze and run it inside a predefined virtual environment (\eg\ Windows XP).
 The system is then capable of simulating user interaction, take screenshots of the environment, extract static and dynamic information about a sample, monitor network traffic and perform memory analysis.
 The time spent doing the analysis can be configured by the user.
-Upon finishing the analysis the system provides reports about the sample.
-Examples of what is contained in these reports include: what functions were imported from which \gls{dll}, what processes were created and the \gls{api} calls order and which files were created/modified.
+Upon finishing the analysis the system provides reports about the sample. The reports' information can be divided into 5 sections:
+\begin{itemize}
+	\item \textbf{Quick Overview} which includes information like: file type and details; analysis duration; signatures (\ie\ predefined patterns Cuckoo deems suspicious); screenshots during runtime; hosts and domains contacted; files, mutexes and registries used.
+	\item \textbf{Static Analysis} which includes information like: file version information; file sections; resources (\eg\ icons); statically imported libraries and their functions; output of the \texttt{strings}\footnote{Linux binary that outputs the strings of printable characters in files.} command on the file; VirusTotal~\cite{tool:virustotal} classification report.
+	\item \textbf{Behavioral Analysis} which includes information like: spawned processes; \gls{api} call order for each process.
+	\item \textbf{Network Analysis} which includes information like: domains and their resolved IPs; HTTP requests; IRC traffic; SMTP traffic.
+	\item \textbf{Dropped Files} which includes information about files created by the sample: their size, name and hash.
+\end{itemize}
 
-The creator of Cuckoo and one of its developers, Alessandro Tanasi, also created a free malware analysis service, Malwr\cite{tool:malwr}.
+Some of the defined sections are not always available.
+Their availability depends both on a successful analysis by Cuckoo, as well as the type of submitted file.
+For example, only binary executables (\eg\ \gls{dll}, \gls{pe}) provide static analysis information.
+
+The creator of Cuckoo and one of its developers, Alessandro Tanasi, also created a free malware analysis service, Malwr~\cite{tool:malwr}.
 This service implements the Cuckoo system and allows users to upload their own files, which are then run with Cuckoo.
 The service is presented as a web application that takes in samples and returns a report with Cuckoo's results.
-An example of the behavioral analysis (i.e.
-created process and \gls{api} calls) can be seen in figure \ref{fig:malwr_sample}.
+An example of the behavioral analysis (\ie\ created process and \gls{api} calls) can be seen in Figure \ref{fig:malwr_sample}.
 
-%\begin{figure}[h]
-%	\includegraphics[width=\textwidth]{malwr_sample}
-%	\caption{Example of a behavioral report from Malwr}
-%	\label{fig:malwr_sample}
-%\end{figure}
+\begin{figure}[!htb]
+	\includegraphics[width=\columnwidth]{Figures/malwr_sample.png}
+	\caption{Example of a behavioral report from Malwr.}
+	\label{fig:malwr_sample}
+\end{figure}
 
-For the purpose of detecting malware via \gls{ml}, the reports provided in Malwr supply various information that can be used to fulfill the current work.
-The whole website content is provided under \gls{cc} CC-BY-NC-SA 3.0\cite{cc-by-nc-sa} license, allowing the use of the data for this work's purpose, as long as credits are maintained.
+Unfortunately at the time of writing, Malwr is under maintenance hence the provided image will be outdated once the new version is live.
+
+For the purpose of detecting malware via \gls{ml}, the reports provided in Malwr supply various information that can be used for the current work.
+The whole website content is provided under \gls{cc} CC-BY-NC-SA 3.0\cite{cc-by-nc-sa} license, allowing use of the data for this work's purpose, as long as credits are maintained.
 
@@
 With the technological advances that provide higher computation power, computer science fields like \gls{ai} have greatly improved in recent years.
+
 One of the \gls{ai} fields that has benefit from the technological advances is \gls{ml}.
@@
 
-\gls{ml} is a subset of \gls{ai} in which algorithms are applied to learn and make predictions on data, or more formally defined by Tom Mitchell\cite{mitchell:ml}:\\
-\say{\emph{A computer program is said to learn from experience E with respect
-		to some class of tasks T and performance measure P, if its performance at tasks in
-		T, as measured by P, improves with experience E.}}
+\gls{ml} is a subset of \gls{ai} in which algorithms are applied to learn and make predictions on data, or more formally defined by Tom Mitchell~\cite{mitchell:ml}:
+\begin{center}
+\say{\emph{A computer program is said to learn from experience $E$ with respect to some class of tasks $T$ and performance measure $P$, if its performance at tasks in $T$, as measured by $P$, improves with experience $E$.}}
+\end{center}
 
-A noteworthy nuance about the given definition, and the \gls{ml} field in general, is that algorithms are not designed to deal with specific problems.
-Meaning that as long as a learning problem is well defined through $T$, $P$ and $E$, an algorithm based on these three features is able to learn, but not necessarily solve a problem.
+A noteworthy nuance about the given definition, and the \gls{ml} field in general, is that these algorithms are not designed to deal with specific problems.
+Meaning that as long as a learning problem is well defined through $T$, $P$ and $E$, an algorithm based on these three features is able to learn, but not necessarily solve any problem.
+This creates a contrast to \gls{ai} fields, where algorithms are built for specific tasks.
 
-After defining a learning problem, to obtain a useful outcome (i.e.
-model) one or more learning algorithms are applied.
-Learning algorithms fall within one of the following classes: supervised learning; unsupervised learning; semi-supervised learning or reinforcement learning.
-Each of the categories is defined on how the program learns the problem.
-A more detailed description, as seen in \cite{norvig:ai} is as follows:
+After defining a learning problem, to obtain a useful outcome (\ie\ model) one or more learning algorithms are applied.
+Learning algorithms fall within one of the following classes: \textit{supervised learning}; \textit{unsupervised learning}; \textit{semi-supervised learning} or \textit{reinforcement learning}.
 
-\subsubsection{Supervised Learning.} In this category the program is given a corpus (i.e.
-experience) with input-output pairs and it must learn how to map a never seen input into the correct output.
-Within supervised learning, a learning problem can be further classified into a classification problem, if the output is a set of finite values (e.g.
-is it going to rain tomorrow?), or into a regression problem, if the output is continuous (e.g.
-how likely is it to rain tomorrow?).
+Each of the categories is defined on how the problem is learned . A more detailed description, as seen in \cite{norvig:ai} is as follows:
 
-\subsubsection{Unsupervised Learning.} In this category the program is given a corpus and it must learn patterns from the input, hence no correct or wrong output is expected, only how different input relates.
+\paragraph{Supervised Learning.} In this category the program is given a corpus (\ie\ experience) with input-output pairs and it must learn how to map a never seen input into the correct output.
+Within supervised learning, a learning problem can be further classified into a classification problem, if the output is a set of finite values (\eg\ is it going to rain tomorrow?), or into a regression problem, if the output is continuous (\eg\ how likely is it to rain tomorrow?).
+
+\paragraph{Unsupervised Learning.} In this category the program is given a corpus and it must learn patterns from the input, hence no correct or wrong output is expected, only how different input relates.
 The most common task in unsupervised learning is clustering, where input is grouped based on similar characteristics.
 
-\subsubsection{Semi-supervised Learning.} This category is a combination of the previous ones and exists because sometimes the learning problem is based on a dataset that combines both labeled and unlabeled inputs or on a dataset that contains noisy data.
+\paragraph{Semi-supervised Learning.} This category is a combination of the previous ones and exists because sometimes the learning problem is based on a dataset that combines both labeled and unlabeled inputs or on a dataset that contains noisy data.
 The combination of supervised and unsupervised learning provides a better overall result in such scenarios.
 
-\subsubsection{Reinforcement Learning.} In this category the program is given a problem through a set of states and a set of actions it can perform.
+\paragraph{Reinforcement Learning.} In this category the program is given a problem through a set of states and a set of actions it can perform.
 At any given state the program performs an action which results in a reward or punishment.
-The cumulative reinforcement at each state teaches the program the best and worse actions for each state.\\
+The cumulative reinforcement at each state teaches the program the best and worse actions for each state.
 
-The proposed work will fit into the supervised learning category, with the purpose of creating a classifier to assert whether a given sample is malware or not.
+\medskip
+
+The current work will fit into the supervised learning category, with the purpose of creating a model to assert whether a given sample is malware or not.
 This problem can be seen as whether a pure classification problem, with a binary output, or as a regression problem, with the likelihood of a sample being malware.
 
-When designing a learning system, multiple choices must be taken into consideration. A brief description of what should be taken into account is given as seen in\cite{mitchell:ml}.
+\medskip
 
-The first choice is the \textit{training experience}, as it can have a significant impact on the success or failure of the learner.
+When designing a learning system, multiple choices must be taken into consideration.
+A brief description of what should be taken into account is given as seen in~\cite{mitchell:ml}:
+
+\begin{itemize}
+	\item \textbf{Training experience} is the first choice, as it can have a significant impact on the success or failure of the learner.
 	The feedback provided by the experience can be direct or indirect, affecting the end result of the system.
 	Another important aspect is how the experience represents the distribution over which the performance is measured, meaning how well the experience represents the environment where the system is to be used.
-
-The next choice is the \textit{target function}, representing what is to be learned.
+	Being what is going to be fed to the learner (\ie\ features), it must be well thought out and understood.
+	\item \textbf{Target function} is the next choice, representing what is to be learned.
 	The function takes as input the set of features that represent the problem and (ideally) outputs a solution to the task.
-
-With a well defined target function, what remains is \textit{choosing a representation for the target function}, where a learning model is applied to learn the given task.
+	\item \textbf{Target function representation} is the remaining choice, this is where a learning model is applied to learn the given task.
 	The choice is of great importance, given it will influence how well the system performs under evaluation.
+\end{itemize}
 
 Given the nature of the purposed work as a classification problem, many models can be applied as a representation for the target function.
-The following describe some models that should be taken into consideration during implementation.
+The following describe some models taken into consideration during implementation.
 
-\paragraph{Decision Tree Learning.}\cite{mitchell:ml} These methods are used to approximate discrete-valued target functions, where the learned function is represented as a decision tree.
-The decision tree is built based on how much information each attribute gives (i.e.
-information gain).
+\paragraph{Decision Tree Learning.~\cite{mitchell:ml}} These methods are used to approximate discrete-valued target functions, where the learned function is represented as a decision tree.
+The decision tree is built based on how much information each attribute gives (\ie\ information gain).
 These are best suited for problems where instances are represented by attribute-value pairs of discrete values and problems where data may contain errors.
@@
 
-\paragraph{Artificial Neural Networks.}\cite{mitchell:ml} These methods can be used for real-valued, discrete-valued and vector-valued target functions.
-\gls{ann} are inspired in part by neurons in the biological world, hence being represented as directed graphs.
+\paragraph{\gls{ann}.~\cite{mitchell:ml}} These methods can be used for real-valued, discrete-valued and vector-valued target functions.
+\gls{ann}'s are inspired in part by neurons in the biological world, hence being represented as directed graphs.
 \gls{ann}'s can be applied to the same type of problems as decision trees, but these can also handle instances represented by many attribute-value pairs.
-\gls{ann}'s have the downsides of being time consuming to train and to easily overfit the training data (i.e.
-training data is learned well, but fail to generalize to new data).
+\gls{ann}'s have the downsides of being time consuming to train and to easily overfit the training data (\ie\ training data is learned well, but fail to generalize to new data).
 
-\paragraph{Bayesian Learning.}\cite{mitchell:ml} These methods use a probabilistic approach to inference.
+\paragraph{Bayesian Learning.~\cite{mitchell:ml}} These methods use a probabilistic approach to inference.
 The reasoning is that the optimal decision can be related to some probabilistic distribution of the data.
 Bayesian learning proves to be effective for text classification tasks and also provides useful insight on other learning algorithms that do not deal with probabilities.
-The shortcomings can be related to the need of requiring initial knowledge, as well as computational cost to achieve the best hypothesis, which grows with the number of hypothesis.\\
+The shortcomings can be related to the need of requiring initial knowledge, as well as computational cost to achieve the best hypothesis, which grows with the number of hypothesis.
+
+\medskip
 
@@
 For every learning task multiple models can be applied, which complicates the task of choosing a model.
-Being able to try different models how they compare to each other benefits the end result, which is to be taken into consideration for the current work.
+
+Although being able to try different models and how they compare to each other benefits the end result, this work focused solely on a single model, reasoned in Section \ref{section:model_selection_evaluation}.
+Doing so enabled to us to spend more time understanding the impacts of \textit{ground-truth} and temporal consistency.
 
@@
 
-The previous subsections gave a general understanding on how anti-virus software detects malware and how the \gls{ml} field can be applied to learn tasks without having to define specific algorithms.
+The previous Sections gave a general understanding on how anti-virus software detects malware and how the \gls{ml} field can be applied to learn tasks without having to define specific algorithms.
 This section will describe research that aggregates both malware detection and machine learning.
@@
 
-The main focus is to note the choices made by the authors relative to the datasets, classifiers and their evaluation, all of which are to be taken into consideration for the proposed work.
+The main focus is to note the choices made by the authors relative to the datasets, classifiers and their evaluation, all of which were taken into consideration in the current work.
 
 Before analyzing the work done with \gls{ml} in malware detection, it is worth noting how such work is to be conducted.
-Rossow et al\ \cite{rossow:practices} present some guidelines for designing malware experiments.
+
+Rossow et al.~\cite{rossow:practices} present some guidelines for designing malware experiments.
 Their work surveys 36 academic publications from 2006 to 2011 that rely on malware execution.
@@
 
-With regards to safety, the given guidelines focus on containment policies when analyzing malware samples. The authors note that most papers did not deploy or adequately describe containment.
+With regards to safety, the given guidelines focus on containment policies when analyzing malware samples.
+The authors note that most papers did not deploy or adequately describe containment.
 
-As mentioned in subsection \ref{subsec:cuckoo}, the current work is to make use of the reports in Malwr, hence some the guidelines provided in the aforementioned research do not apply, specifically the safety guidelines.
-As for the other guidelines, it is important to take them into consideration when designing the proposed learning system.
+As mentioned in Section \ref{section:cuckoo}, the current work makes use of the reports in Malwr, hence some the guidelines provided in the aforementioned research do not apply, specifically the safety guidelines.
+As for the other guidelines, they were taken into consideration when designing the presented learning system.
 
-On the same note of guidelines, Shabtai et al\cite{shabtai:survey} provide a survey directed at the application of machine learning classifiers to detect malware from static features.
+\medskip
+
+On the same note of guidelines, Shabtai et al.~\cite{shabtai:survey} provide a survey directed at the application of machine learning classifiers to detect malware from static features.
 Their work concerns the design and evaluation of such systems.
@@
 
-The training phase includes the creation of a vector to represent each file and posterior processing of such vectors to generate a classifier, while the testing phase measures the performance of the classifier by employing classical measures like \gls{tpr}, which measures the sensitivity (i.e.
-detection rate) of the classifier, and \gls{fpr}, which measures fall-out.
+The training phase includes the creation of a feature vector to represent each file and posterior processing of such vectors to generate a classifier.
+The testing phase measures the performance of the classifier by employing classical measures like \gls{tpr} (\ie\ detection rate), and \gls{fpr}.
 
-The authors outline that the work in designing learning systems include how the samples are to be represented (e.g.
-byte n-grams, portable executable features, strings), methods to select features (e.g.
-gain ratio, document frequency) and finally the classification algorithm (e.g.
-\gls{ann}, \gls{svm}).
+The authors outline that the work in designing learning systems include how the samples are to be represented (\eg\ byte n-grams, portable executable features, strings), methods to select features (\eg\ gain ratio, document frequency) and finally the classification algorithm (\eg\ \gls{ann}, \gls{svm}).
 
@@
 
-Some problems addressed that are of importance for the proposed work include the imbalance problem and chronological evaluation.
-Having a balanced dataset of malware and goodware does not accurately represent the real world.
+Some problems addressed that are of importance for the presented work include the imbalance problem and chronological evaluation.
+Having a balanced dataset of malware and goodware might not accurately represent the real world.
 For chronological evaluation the question if whether the usage of old samples for training can aid or hinder the classification of new samples.
 
-When it comes to malware detection from static features, Schultz et al\cite{schultz:data_mining} present a data-mining framework to detect new, unseen malicious executables.
+\medskip
+
+When it comes to malware detection from static features, Schultz et al.~\cite{schultz:data_mining} present a data-mining framework to detect new, unseen malicious executables.
 Their work uses a dataset of 4,266 programs, with 3,625 malicious binaries and 1,001 clean programs, labeled by a commercial anti-virus.
@@
 
-Another application of \gls{ml} to malware detection from static features is done by Nissim et al\cite{nissim:al_pdf}.
-The authors apply active learning to the detection of malicious PDF files based on static features.
-They use a dataset of 6,774 PDF files, including 1629 malicious and 5,145 benign files, where the benign files were mark as such by Kaspersky anti-virus software.
+\medskip
+
+Another application of \gls{ml} to malware detection from static features is done by Nissim et al.~\cite{nissim:al_pdf}.
+The authors apply active learning (semi-supervised learning that takes user input) to the detection of malicious PDF files based on static features.
+They use a dataset of 6,774 PDF files, including 1,629 malicious and 5,145 benign files, where the benign files were mark as such by Kaspersky anti-virus software.
 
@@
 
-With regards to malware detection from dynamic features, Rieck et al\cite{rieck:dynamic} propose a framework for automatic analysis of malware behavior using machine learning.
+\medskip
+
+With regards to malware detection from dynamic features, Rieck et al.~\cite{rieck:dynamic} propose a framework for automatic analysis of malware behavior using machine learning.
 Their approach applies both unsupervised (clustering) and supervised (classification) learning to identify novel malware classes and assign unknown to the discovered classes.
@@
 Different levels of specificity can be obtained since the arguments are arranged in blocks.
-For variable-length arguments an index number is used, which can be mapped to the original content.
-Figure \ref{fig:mist} shows the schematic overview of \gls{mist} instructions.
+For variable-length arguments an index number is used, which can be mapped to the original content. Figure \ref{fig:mist} shows the schematic overview of \gls{mist} instructions.
 
-%\begin{figure}[h]
-%	\includegraphics[width=\textwidth]{mist}
-%	\caption{\gls{mist} representation as seen in \cite{rieck:dynamic}}
-%	\label{fig:mist}
-%\end{figure}
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.7\columnwidth]{Figures/mist.png}
+	\caption{\gls{mist} representation as seen in \cite{rieck:dynamic}}
+	\label{fig:mist}
+\end{figure}
 
@@
 
-On the reference set authors apply F-measure for evaluation, which uses precision and recall as parameters, the maximization of F-measure reflects high precison and recall.
+On the reference set authors apply F-measure for evaluation, which uses precision and recall as parameters, the maximization of F-measure reflects high precision and recall.
 The results yield an F-measure over 0.936, which surpasses previous research for both clustering and classification.
 
-On the application set authors apply their methodology and achieve 434 clusters (i.e.
-malware families).
+On the application dataset authors apply their methodology and achieve 434 clusters (\ie\ malware families).
 To the ten largest clusters authors take the most frequent anti-virus label for each cluster using Kaspersky anti-virus.
@@
 
-Miller et al\cite{miller:rev_int} provide a system which relates to \cite{nissim:al_pdf}, as it applies expert reviewers as a limited labeling resource.
+\medskip
+
+Miller et al.~\cite{miller:rev_int} provide a system which relates to \cite{nissim:al_pdf}, as it applies expert reviewers as a limited labeling resource.
 Their work uses both static and dynamic features to detect malware and uses manual label together with \gls{ml} to improve the overall result.
@@
 The system uses multiple static and dynamic features from binaries, including: binary metadata, digital signing, static imports, dynamic imports, file operations, mutex operations, network operations, processes and Windows \gls{api} calls.
-The authors chose to use \gls{lr} as the classifier, basing their choice on the fact that it predicts a real-valued quantity, which enables the use of a threshold and therefore enables the creation of a tradeoff between true and false positive rates, by adjusting the threshold.
+The authors chose to use \acrfull{lr} as the classifier, basing their choice on the fact that it predicts a real-valued quantity, which enables the use of a threshold and therefore allows the creation of a tradeoff between true and false positive rates, by adjusting the threshold.
 \gls{lr} also assigns a weight to each feature, which allows to comprehend which features are more discriminating.
@@
 
-Another remark made in this article is that vendors prefer false negatives over false positives (i.e.
-wrongly classifying malware as goodware over wrongly classifying goodware as malware).
+Another remark made in this article is that vendors prefer false negatives over false positives (\ie\ wrongly classifying malware as goodware over wrongly classifying goodware as malware).
 When looking at duplicated samples, 29.6\% of those samples increased in the number of detections as malware and only 0.25\% decreased the number of detections.
 
-Perdisci et al\cite{perdisci:behavior} use an unsupervised learning approach to cluster malware using malicious network traces.
-Although their work does not directly relate to the proposed work, it also deals with the malware naming problem.
+\medskip
+
+On the topic of methodologies that resemble real-world conditions, Srndi\'c et al.~\cite{vsrndic2013detection} train and validate their malicious PDF detector under laboratory and real-world conditions.
+Laboratory conditions consist on applying regular cross-validation, whereas real-world conditions validate a newer dataset with a model created from outdated data (\ie\ older then the validation), and also validate the model when the validation set spans one week and the training is gathered in the previous 4 weeks.
+They show that laboratory conditions inflate the results, when compared to real-world conditions.
+
+Our work enhances these methodologies by analyzing the performance variation when the distance between the training and validation set increases and decreases, as well as analysis on how reducing the size of the training without compromising the results.
+
+\medskip
+
+Kolter et al.~\cite{kolter:learning} learn to detect malicious executables in a dataset with under 4,000 samples, obtained from reliable sources, and evaluate their model under standard cross-validation and by gathering newer malware samples to validate for new and unseen samples.
+They show how the model provides optimal results under cross-validation, but for the unseen samples lower scores are obtained.
+Our work considers these results to compare how reliability affects performance.
+
+Deo, A. et al.~\cite{deo2016prescience} focus over the problem of \gls{ml} models becoming antiquated over time, given how malware evolves.
+This problem is seen as concept drift, where the performance of a model over time diminishes, as the statistical properties of malware (\ie\ features) change over time.
+They study how probabilistic predictors can help minimize the aforementioned problem, by indicating when retraining of a given model is necessary.
+
+Jordaney, R.
+et al.~\cite{jordaney2017transcend} also study the problem of concept drift in malware classification models, focusing on providing metrics, based on statistical comparison of samples, to detect when should a model be retrained.
+
+Both \cite{deo2016prescience} and \cite{jordaney2017transcend} are on the subject of our work, regarding the problem of concept drift, but differ on the study-case.
+In our work we acknowledge the concept but focus on its relative effects and how one can balance the size of needed training data \textit{vs.}\ validation data, when maintaining a temporal consistent dataset.
+Whereas their work provide indicators for when should a model be retrained when the problem of concept drift becomes significant.
+
+\medskip
+
+Perdisci et al.~\cite{perdisci:behavior} use an unsupervised learning approach to cluster malware using malicious network traces.
+Although their work does not directly relate to the current work, it also deals with the malware naming problem.
 To validate the obtained clusters, the cohesion and separation among clusters is measured in terms of the labels given by anti-virus vendors.
@@
 Weights are assigned as $1 - \dfrac{m}{n}$, where $m$ is the number of times the different labels appear together and $n$ is the number of samples in the cluster.
-From the graphs the authors can then measure the cohesion and separation of clusters.
+From the graphs authors can then measure the cohesion and separation of clusters.
 
-The authors use the aforementioned method to evaluate their results, but the approach can be useful to label and balance malware datasets that contain various anti-virus labels.\\
+The authors use the aforementioned method to evaluate their results, but the approach can be useful to label and balance malware datasets that contain various anti-virus labels.
+
+\medskip
+
+Regarding the malware naming problem, Sebastián, M. et al.~\cite{sebastian2016avclass} develop \textit{AVClass}, a tool that given a set of antivirus vendors, outputs the most likely family name.
+They test their tool under 10 datasets, totaling 8.9 million samples, with results showing an F1 measure up to 93.9\% on labeled datasets.
+The tool takes as input the labels as seen in VirusTotal, tokenizes the labels, replaces known aliases and general names (\eg\ win32, trojan, generic), ending with possible family names.
+These remaining names are counted and the most frequent one is given as family name, Figure \ref{fig:avclass} exemplifies the stated process.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=\columnwidth]{Figures/avclass.png}
+	\caption{AVClass running example.}
+	\label{fig:avclass}
+\end{figure}
+
+Our work takes advantage of this tool, not to label malware families, but with minimal modifications, to label more general malware classes (\eg\ trojan, virus).
+
+\medskip
 
@@
 Decisions like the handling of the dataset, how and which features are to be extracted from samples, which and why \gls{ml} classifiers should be used and how and why the evaluation should be made.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Theoretical Overview}
-%\label{section:overview}
-%
-%Some overview of the underlying theory about the topic...
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Theoretical Model 1}
-%\label{section:theory1}
-%
-%The research should be supported with a comprehensive list of references.
-%These should appear whenever necessary, in the limit, from the first to the last chapter.
-%
-%A reference can be cited in any of the following ways:
-%%
-%\begin{itemize}
-%  \item Citation mode \#1 - \quad \cite{jameson:adjointns}
-%  \item Citation mode \#2 - \quad \citet{jameson:adjointns}
-%  \item Citation mode \#3 - \quad \citep{jameson:adjointns}
-%  \item Citation mode \#4 - \quad \citet*{jameson:adjointns}
-%  \item Citation mode \#5 - \quad \citep*{jameson:adjointns}
-%  \item Citation mode \#6 - \quad \citealt{jameson:adjointns}
-%  \item Citation mode \#7 - \quad \citealp{jameson:adjointns}
-%  \item Citation mode \#8 - \quad \citeauthor{jameson:adjointns}
-%  \item Citation mode \#9 - \quad \citeyear{jameson:adjointns}
-%  \item Citation mode \#10 - \quad \citeyearpar{jameson:adjointns}
-%\end{itemize}
-%%
-%Several citations can be made simultaneously as \citep{nocedal:opt,marta:ijcfd}. \\
-%
-%This is often the default bibliography style adopted (numbers following the citation order), according to the options:\\
-%{\tt \textbackslash usepackage\{natbib\}} in file {\tt Thesis\_Preamble.tex},\\
-%{\tt \textbackslash bibliographystyle\{abbrvnat\}} in file {\tt Thesis.tex}.\\
-%
-%Notice however that this style can be changed from numerical citation order to authors' last name with the options: \\
-%{\tt \textbackslash usepackage[numbers]\{natbib\}} in file {\tt Thesis\_Preamble.tex},\\
-%{\tt \textbackslash bibliographystyle\{abbrvunsrtnat\}} in file {\tt Thesis.tex}. \\
-%
-%Multiple citations are compressed when using the {\tt sort\&compress} option when loading the {\tt natbib} package as {\tt \textbackslash usepackage[numbers,sort\&compress]\{natbib\}} in file {\tt Thesis\_Preamble.tex}, resulting in citations like \citep{marta:ijcfd1,marta:ijcfd2,marta:ijcfd3,marta:ijcfd4}.
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Theoretical Model 2}
-%\label{section:theory2}
-%
-%Other models...

### 2018-05-09 thesis_report/Thesis_Static_Features.tex
Static Features v1

@@
 Using Celery~\cite{tool:celery}, a distributed task queue for Python, we optimized the parsing of the available HTML reports, extracting samples that contained information regarding static imports into a new set $\FF_{static}$.
-We then joined the samples with static imports $\FF_{static}$ to the labeled samples $\CC_{real}$, obtaining a total of 189,352 labeled samples with static imports $\DD_{static} = \CC_{real} \cap \FF_{static}$.
+We then joined the samples with static imports $\FF_{static}$ to the labeled samples $\CC_{real}$, obtaining a total of 155,057 labeled samples with static imports $\DD_{static} = \CC_{real} \cap \FF_{static}$.
 
@@
 
-Due to the amount of samples and variety of imports, each sample got a vector $x$ of 6,647 dimensions (\ie\ there are 6,647 different imports).
+Due to the amount of samples and variety of imports, each sample got a vector $x$ of 7,280 dimensions (\ie\ there are 7,280 different imports).
 To reduce this number, and to remove any noise due to incorrect parsing of static imports by Cuckoo, we applied a variance threshold.
@@
 
-The resulting dataset $\DD_{static}$ got reduced to 187,545 samples, each with a 64 dimensional binary vector.
+The resulting dataset $\DD_{static}$ got reduced to 153,374 samples, each with a 64 dimensional binary vector.
 
@@
 
-Using cross-validation, our $\LR$ scored with an \gls{auroc} of \miss{score}, as shown by the red line in Figure \ref{fig:xval_results}.
+Using cross-validation, our $\LR$ scored with an \gls{auroc} of 91\%, as shown by the red curve in Figure \ref{fig:xval_results}.
 We argue that such high values are easily attained from factors like a small and reliable dataset, and the use of cross-validation, which mixes samples and ignores possible dependencies on malware samples.
 
-When we relax our ideal conditions to more \textit{real-world conditions}, by testing the labeled dataset $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{static}$ for scenarios $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, respectively, the score under \gls{auroc} decreases to \miss{score} for $\mathcal{S}_{loose}$ (blue line in Figure \ref{fig:xval_results}) and \miss{score} for $\mathcal{S}_{real}$ (green line in Figure \ref{fig:xval_results}).
+When we relax our ideal conditions to more \textit{real-world conditions}, by testing the labeled dataset $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{static}$ for scenarios $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, respectively, the score under \gls{auroc} decreases to 90\% for $\mathcal{S}_{loose}$ (blue curve in Figure \ref{fig:xval_results}) and 75\% for $\mathcal{S}_{real}$ (green curve in Figure \ref{fig:xval_results}).
 
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\includegraphics[width=0.8\textwidth]{Figures/xval_results.png}
 	\caption[Single layer results for static features in laboratory conditions.]{Single layer results for static features in laboratory conditions.}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\includegraphics[width=0.8\textwidth]{Figures/pastpresent.png}
 	\caption[Single layer results for static features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation.
@@
 
-When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note a decrease from \miss{score} to \miss{score} for $\mathcal{S}_{\strict}$, \miss{score} to \miss{score} for $\mathcal{S}_{\loose}$, and \miss{score} to \miss{score} for $\mathcal{S}_{\real}$.
-This decrease is intuitive to the methodology, as we are forcing temporal consistency between samples.
+When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 0.75\% to 0.68\%.
 
-When looking at results from $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$, we see they closely relate.
-This relation has already been noticed on previous cross-validation results, as their metrics $\Mstrictv$ and $\Mloosev$ are not very different.
-As for $\mathcal{S}_{\real}$, the degradation is higher, as the reliability of the metric $\Mrealv$ goes down.
+For both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ we note only a slight increase as new folds are added.
+They still relate, as we have previously noted for cross-validation, arguably given their metrics $\Mstrictv$ and $\Mloosev$ are not very different.
+The small variation to the cross-validation methodology can be justified by using small dataset size for both cases.
+
+As for $\mathcal{S}_{\real}$ we note higher variation and lower overall score, as the reliability of the metric $\Mrealv$ goes down.
+This is expected, not only because we are enforcing temporal consistency between samples, but as new folds are added, the training gets bigger, while the test remains the same.
 
 Our main observation for this validation methodology is that there is a slight tendency for \gls{auroc} to increase, as we move forward in time, close to the validation set.
-
-Both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ behave similarly under our \textit{Past-to-Present} evaluation, decreasing the \gls{auroc} until fold \miss{fold}, after which both increase, stabilizing around \miss{score} \gls{auroc} by the last \miss{two} folds.
-As for $\mathcal{S}_{\real}$, its \gls{auroc} does not vary significantly, although a subtle increase over time is noticeable.
 
@@
 By applying this methodology to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we plot Figure \ref{fig:presentpast}, where the X axis increases as older samples are added to the training set (\ie\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time.
-Similarly to the previous observation, the average \gls{auroc} suffers a decrease when compared to cross-validation.
-For $\mathcal{S}_{\strict}$ we note a change from \miss{score} to \miss{score}, for $\mathcal{S}_{\loose}$ the decrease is from \miss{score} to \miss{score}, and for $\mathcal{S}_{\real}$ \miss{score} to \miss{score}.
+Similarly to the previous observation, the average \gls{auroc} remains identical when compared to cross-validation for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$.
+As for $\mathcal{S}_{\real}$, we note a decrease in score from 75\% to 69\%.
 
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\includegraphics[width=0.8\textwidth]{Figures/presentpast.png}
 	\caption[Single layer results for static features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation.
@@
 The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ display very similar results, with $\mathcal{S}_{\real}$ dropping behind due to its less reliable labeling metric.
+It is noticeable that using the entire dataset does not bring much improvement to the final results. In fact, for $\mathcal{S}_{\real}$ the score even drops after fold $\#2$.
 
@@
 Finally, we analyze how does such reduced training set behaves in our scenarios; for this purpose, we define a sliding window that moves forward in time through each scenario for training and validation.
-We propose a reduction on the training size to \miss{$w=3$} folds predating the validation fold.
-We choose \miss{$w = 3$}, since we have seen that the scores either do not improve (for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$) or actually go down (for $\mathcal{S}_{\real}$) with higher folds.
-In summary, we have selected \miss{30\%} of each dataset for training purposes and the next \miss{10\%} for validation \miss{(3 training folds, 1 validation fold)}, and then started moving the window forward in time (1 fold at a time) to obtain the following results (Figure~\ref{fig:slidingwindow}): for  $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we obtain \gls{auroc} values of \miss{score, score and score}, respectively.
-These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score.
+We propose a reduction on the training size to $w=3$ folds predating the validation fold.
+We choose $w=3$, since we have seen that the scores either do not improve (for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$) or actually go down (for $\mathcal{S}_{\real}$) with higher folds.
+In summary, we have selected 30\% of each dataset for training purposes and the next 10\% for validation (3 training folds, 1 validation fold), and then started moving the window forward in time (1 fold at a time) to obtain the following results (Figure~\ref{fig:slidingwindow}): for  $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we obtain \gls{auroc} values of 89\%, 88\% and 73\%, respectively.
+These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score (2\% for each scenario).
 
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
+	\includegraphics[width=0.8\textwidth]{Figures/slidingwindow.png}
 	\caption[Single layer results for static features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
@@
 
-Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from \miss{0.93 to 0.87 for $\mathcal{S}_{\strict}$, from 0.92 to 0.86 for $\mathcal{S}_{\loose}$ and from 0.79 to 0.76 for $\mathcal{S}_{\real}$.}
-We should highlight that the results that use temporal-consistency should better reflect the reality than standard cross-validation, since we are requiring temporal consistency.
+Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from 91\% to 89\% for $\mathcal{S}_{\strict}$, from 90\% to 88\% for $\mathcal{S}_{\loose}$ and from 75\% to 73\% for $\mathcal{S}_{\real}$.
+We should highlight that the results that use temporal consistency should better reflect the reality than standard cross-validation, since we are requiring temporally ordered samples.
 Another important idea that should be stressed is that for cross validation we used a fairly reasonable amount of data for training purposes, whereas in this last case we used a restricted amount of data.
 This might be a relevant issue in a few year's time.
-The results obtained are summarized in Table \ref{tab:singlelayer_results}
+The results obtained are summarized in Table \ref{tab:singlelayer_results}.
 
@@
 	\hline AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
-	\hline Cross-Validation & 0.93 & 0.92 & 0.79 & 90 / 10\\
-	\hline Past-to-Present & 0.88 & 0.87 & 0.71 & 10 to 90 / 10\\
-	\hline Present-to-Past & 0.90 & 0.86 & 0.72 & 10 to 90 / 10\\
-	\hline Sliding-Window & 0.87 & 0.86 & 0.76 & 30 / 10\\
+	\hline Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
+	\hline Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
+	\hline Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
+	\hline Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
 	\hline

### 2018-05-09 thesis_report/Thesis_Data_Collection.tex
minor fixes.

@@
 The file types distribution is interesting, as it shows that executables (\ie\ compiled binary files) took the great majority of submissions up to the end of 2014, after which documents (\eg\ Word, Excel) and text files (\eg\ HTML, JavaScript) started to increase.
+
 As a side observation, this increase might indicate how users are more suspicious of these type of files, given that it is not uncommon to see malware being distributed in documents (under email attachments) and in websites (as JavaScript scripts) in more recent years.
@@
 
-Having a better overall understanding on the available reports, and taking into account the limitations of static analysis in \textcolor{red}{Cuckoo} as mentioned in Section \ref{section:cuckoo}, led us to our first data selection criteria: collect reports from \gls{pe} (\eg\ 32bit executables, DLL's) submissions only.
+Having a better overall understanding on the available reports, and taking into account the limitations of static analysis in Cuckoo~\cite{tool:cuckoo} as mentioned in Section \ref{section:cuckoo}, led us to our first data selection criteria: collect reports from \gls{pe} (\eg\ 32bit executables, DLL's) submissions only.
 This selection reduces the number of available reports to 388,702 (60.48\% of the total) but enables us to work with more static information.
@@
 
-As previously mentioned in Section \ref{section:cuckoo}, Malwr aggregates the scan report as provided by VirusTotal (if the sample is already known by VirusTotal), which entails multiple antivirus classifications.
+As previously mentioned in Section \ref{section:cuckoo}, Malwr aggregates the scan report as provided by VirusTotal (if the sample is already known by VirusTotal) in its static analysis, which entails multiple anti-virus classifications.
 Our next analysis takes advantage of this information to better understand how the samples are labeled and how reliable using the VirusTotal report can be to label our dataset.
 
-After counting the number of different antivirus vendors present in at least one submission, we decided to narrow down to the amount of unique vendors by requiring their presence in at least 95\% of the classified reports.
+After counting the number of different anti-virus vendors present in at least one submission, we decided to narrow down to the amount of unique vendors by requiring their presence in at least 95\% of the classified reports.
 Doing so guarantees that vendors did not show occasionally and have a strong presence in all classified samples.
@@
 Reiterating what was already said, defining and classifying malware is not a trivial task.
-The \textit{ground truth} of a sample is hard to obtain, as both antivirus vendors and users disagree on what is malware.
+The \textit{ground truth} of a sample is hard to obtain, as both anti-virus vendors and users disagree on what is malware.
 To come to terms with this obstacle, we make use of the classifications given by our set of vendors $\mathcal{V}$ to define a simple malware/goodware criteria: what is the minimum/maximum threshold of positive malware classifications as to label a sample as malware/goodware.
@@
 	\includegraphics[width=\textwidth]{Figures/distribution_clean.png}
-	\caption[Distribution of duplicated samples with $\Pf=0$.]{Distribution of samples that started as goodware and changed in the number of positive classifications between the last and first submission, \ie $\Pl-\Pf$ for samples with $\Pf=0$}
+	\caption[Distribution of duplicated samples with $\Pf=0$.]{Distribution of samples that started as goodware and changed in the number of positive classifications between the last and first submission, \ie\ $\Pl-\Pf$ for samples with $\Pf=0$}
 	\label{fig:distribution_clean}
@@
 Given we are interested not only in providing a model to detect malware, but also in understanding the impacts of \textit{ground truth} in a detection model, we now go into detail on how we made use of the previously defined top vendors $\VVstar$ in conjunction with the information provided by \gls{nsrl} and VirusShare.com to label the available samples.
-To do so, we take the previously defined scenarios in Figure \ref{fig:dataset_sizes} to define three labeling metrics.
+To do so, we take the previously defined scenarios in Figure \ref{fig:scenarios} to define three labeling metrics.
 
@@
 		\midrule
-		Malware			& 0 & 0 & 0\\
-		Goodware		& 0 & 0 & 0\\
+		Malware			& 98,582 & 45,306 & 24,658\\
+		Goodware		& 56,475 & 1,989 & 1,989\\
 		\midrule
-		Total			& 0 & 0 & 0\\
+		Total			& 155,057 & 47,295 & 26,647\\
 		\bottomrule

### 2018-05-10 thesis_report/Thesis_Model_Improvements.tex
minor fixes.
Model Improvements v1

--- 2018-05-08 thesis_report/Thesis_Model_Improvements.tex
@@
 
-By applying this new labeling, we refine our previous datasets $\CC_{strict}$, $\CC_{loose}$ and $\CC_{real}$ as shown in Table \ref{tab:dataset_sizes_new}.
+The following Table \ref{tab:dataset_sizes_new} discriminates the amount of each malware class.
+Note that the classes are not exclusive, as such the represented numbers show how many samples, out of each dataset $\CC_{strict}$, $\CC_{loose}$ and $\CC_{real}$, are classified by at least one vendor as belonging to any of the defined classes.
 
@@
 		\midrule
-		Trojan			& 0 & 0 & 0\\
-		Virus			& 0 & 0 & 0\\
-		Worm			& 0 & 0 & 0\\
-		Spyware			& 0 & 0 & 0\\
-		Ransom			& 0 & 0 & 0\\
-		Other			& 0 & 0 & 0\\
-		\midrule
-		Total Malware	& 0 & 0 & 0\\
-		Total Goodware	& 0 & 0 & 0\\
-		\midrule
-		\midrule
-		Total			& 0 & 0 & 0\\
+		Trojan			& 97,054 & 44,329 & 24,176\\
+		Other			& 49,443 & 24,126 & 14,750\\
+		Worm			& 24,554 & 14,837 & 9,381\\
+		Virus			& 21,055 & 12,531 & 6,899\\
+		Spyware			& 20,724 & 10,172 & 5,955\\
+		Ransom			& 7,761 & 1,924 & 1,160\\
 		\bottomrule
 	\end{tabular}
-	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
+	\caption{Samples belonging to each of the 6 malware classes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
 	\label{tab:dataset_sizes_new}
 \end{table}
+
+One can see that the \textit{trojan} class is present in almost all samples, whereas \textit{ransom} is the least common class.
+More detailed statistics regarding each dataset is presented in Appendix \ref{chapter:appendixClasses}, containing information like the average weight for each class and the minimum and maximum values.
+
+\medskip
 
@@
 
-With that in mind, we decide to transform the values into a $[0,1]$ range.
+With that in mind, we decide to transform the values to follow a normal distribution.
 We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
@@
 
-For our third and last type of dynamic information, we resort to Cuckoo~\cite{tool:cuckoo} custom signatures.
+For our third and last type of dynamic information, we resort to Cuckoo's~\cite{tool:cuckoo} custom signatures.
 These signatures are built from certain activities that Cuckoo deems malicious or suspicious.
@@
 
-Although some signatures are defined by a sequence of library calls, which could also be attained by creating n-grams of our previously defined feature, using the already defined Cuckoo signatures saves training time.
-As an example, creating bi-grams of library calls, given there are 144 unique calls, would yield over 20,000 possible features, hence being unpractical.
+Although some signatures are defined by a sequence of library calls, which could also be attained by creating \textit{n-grams} of our previously defined feature, using the already defined Cuckoo signatures saves training time.
+As an example, creating bi-grams of library calls, given there are 144 unique calls, would yield over 20,000 possible features, which becomes highly unpractical when considering a higher degree of n-grams.
 
@@
 
-Having these new features, we create a new set with all features (static, category and library calls and signatures) $\FF_{dynamic}$.
-We joined these samples $\FF_{dynamic}$ to the labeled samples $\CC_{real}$, obtaining a total of \miss{size} labeled samples with dynamic imports $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$.
+We joined these features $\FF_{dynamic}$ to the labeled samples $\CC_{real}$, obtaining a total of 122,633 labeled samples with the new features $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$.
 
-% To reduce the number of features into those that are more representative, we apply, as in other features, a variance threshold, removing signatures that do not vary in more than 80\% of the samples.
-% By doing so, we reduce the total number of signatures to
+Given there is a lower amount of available samples, we provide in Table \ref{tab:dataset_sizes_improved} the new sizes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$, which take into account the new features $\FF_{dynamic}$.
+
+\begin{table}[!htb]
+	\renewcommand{\arraystretch}{1.2} % more space between rows
+	\centering
+	\begin{tabular}{lccc}
+		\toprule
+		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
+		\midrule
+		Malware			& 94,248 & 42,911 & 23,437\\
+		Goodware		& 28,385 & 1,741 & 1,741\\
+		\midrule
+		Total			& 122,633 & 44,652 & 25,178\\
+		\bottomrule
+	\end{tabular}
+	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
+	\label{tab:dataset_sizes_improved}
+\end{table}
 
@@
 We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{section:evaluation}.
+
 Specifically we test the model using the baseline cross-validation methodology, followed by our three temporally consistent scenarios: \textit{Past-to-Present}, \textit{Present-to-Past} and \textit{Temporal Window}.
@@
 Starting with \textit{laboratory conditions}, we apply the cross-validation evaluation to model $\mathcal{E}$ with the labeled dataset $\CC_{strict}$ and features $\DD_{dynamic}$, providing scenario $\mathcal{S}_{\strict}$.
-For this scenario, we obtain an \gls{auroc} of \miss{score}, as presented by the red line in Figure \ref{fig:improved_xval}.
-These provide the highest results, provided by factors like a small and reliable dataset, and the use of cross-validation.
+For this scenario, we obtain an \gls{auroc} of 98\%, as presented by the red curve in Figure \ref{fig:improved_xval}.
+These provide the highest results, provided by factors like a small and reliable dataset, more features and the use of cross-validation.
 
 Relaxing to more \textit{real-world conditions}, under the form of a less reliable ground truth, we test the datasets $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{dynamic}$.
-As shown in Figure \ref{fig:improved_xval}, the score under \gls{auroc} decreases to \miss{socre} for $\mathcal{S}_{\loose}$ (blue line) and \miss{score} for $\mathcal{S}_{\real}$ (green line).
+As shown in Figure \ref{fig:improved_xval}, the score under \gls{auroc} is 98\% for $\mathcal{S}_{\loose}$ (blue curve) and 95\% for $\mathcal{S}_{\real}$ (green curve).
 
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
-	\caption{Improved model results with dynamic features in laboratory conditions.}
+	\includegraphics[width=0.8\textwidth]{Figures/xval_improved.png}
+	\caption{Multi layer results for dynamic features in laboratory conditions.}
 	\label{fig:improved_xval}
@@
 % Adicionar seccao para resultados do multi layer apenas.
+
+As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are affected at all, when the change between the scenarios is merely in the number of malware samples.
+Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the most realistic dataset.
+
+The comparison between scenarios does not yield any new information from what was seen in Section \ref{section:single_layer_results}.
+What is more interesting is that the absolute values are boosted in all scenarios, which show indeed that using the multi-layer approach with dynamic features improve the model's results.
+
+Having applied the same cross-validation to our modified model, obtaining interesting results, we now go over to test how our temporal based methodologies are affected.
+
+\medskip
+
+We start again with \textit{Past-to-Present} validation to each scenario $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
+Figure \ref{fig:pastpresent_modified} shows the \gls{auroc} at every iteration (\ie\ fold) for each scenario: 96\% for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$, and 92\% $\mathcal{S}_{\real}$.
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/pastpresent_improved.png}
+	\caption[Multi layer results for dynamic features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation.
+Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
+		fold 1)}
+	\label{fig:pastpresent_modified}
+\end{figure}
+
+When comparing the average \gls{auroc} between cross-validation and \textit{Past-to-Present} validation, we note that both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ decrease 2\%, while $\mathcal{S}_{\real}$ decreases 3\%.
+This decrease is not a surprise, given the temporal consistency enforcement between samples.
+
+The results are consistent with was previously seen in Section \ref{section:single_layer_results}, with the added factor that the absolute values are higher, and the tendency to increase is more present as we move forward in time, close to the validation set.
+
+\medskip
+
+Following the previous evaluation order, we now present the results using our \textit{Present-to-Past} validation methodology.
+In Figure \ref{fig:presentpast_improved} we present the \gls{auroc} for each iteration, where higher folds represent older samples.
+Here we see values of 97\% for $\mathcal{S}_{\strict}$, 98\% for $\mathcal{S}_{\loose}$ and 96\% $\mathcal{S}_{\real}$.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/presentpast_improved.png}
+	\caption[Multi layer results for dynamic features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation.
+Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
+	\label{fig:presentpast_improved}
+\end{figure}
+
+When comparing to cross-validation, we note $\mathcal{S}_{\strict}$ is affected as expected, whereas $\mathcal{S}_{\loose}$ is not.
+The fact that the datasets vary in size and so the amount of goodware and malware used for testing in each may vary, can justify how the $\mathcal{S}_{\loose}$ is not affected, whereas $\mathcal{S}_{\strict}$ is.
+
+$\mathcal{S}_{\real}$ seems to have a higher value than in cross-validation, but it is rounded, which slightly inflates the value.
+In practice both cross-validation and \textit{Present-to-Past} are identical for $\mathcal{S}_{\real}$.
+
+In these results, the noticeable increase in the first 3 folds (0, 1 and 2) for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ goes even more in favor with our argument that samples closer to the validation set benefit the model.
+More so as the \gls{auroc} stabilizes from those folds on.
+This effect is not as accentuated for $\mathcal{S}_{\real}$, although using more and older folds do not provide significantly better results.
+
+\medskip
+
+Finally, we retest how a reduced training set behaves in our scenarios by using our \textit{Temporal Window} methodology.
+As previously mentioned, the first 3 folds seem to provide enough information to obtain good results, hence we apply the same sliding window size of $w=3$ as in Section \ref{section:single_layer_results}.
+In summary, we selected 30\% of each dataset for training and 10\% for validation.
+Starting at the oldest fold, we apply this window and slide it by one fold at each iteration.
+Figure \ref{fig:slidingwindow_modified} shows how all our scenarios $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$ score the same \gls{auroc} of 94\%, although with different curves.
+
+For $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ the score is equally and negatively affected by 4\%.
+The jagged curve on both scenarios indicate that slight changes on the \gls{fpr} threshold have significant impact on the \gls{tpr}, this may be caused by the smaller dataset size, which in turn creates uneven folds for malware and goodware.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/slidingwindow_improved.png}
+	\caption[Multi layer results for dynamic features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
+	\label{fig:slidingwindow_modified}
+\end{figure}
+
+For $\mathcal{S}_{\real}$, the fact that it only loses 1\% when compared to cross-validation shows that indeed our argument about reducing the training set size is sustained.
+Its curve is much smoother when compared to the previous scenarios, as the dataset is bigger and more even.
+
+\medskip
+
+As a closing remark, we reaffirm on how it is possible to yield reasonable results when considering temporal consistent samples, specifically in this improved model, which not only uses more features, but has the ability to provide more information regarding a malware sample, as it can discriminate malware classes due to its multi layer design.
+The improved layer results are summarized in Table \ref{tab:multilayer_results}.
+
+\begin{table}[!htb]
+	\renewcommand{\arraystretch}{1.2} % more space between rows
+	\centering
+	\begin{tabular}{ccccc}
+		\toprule
+		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
+		\midrule
+		Cross-Validation & 0.98 & 0.98 & 0.95 & 90 / 10\\
+		Past-to-Present & 0.96 & 0.96 & 0.92 & 10 to 90 / 10\\
+		Present-to-Past & 0.97 & 0.98 & 0.96 & 10 to 90 / 10\\
+		Sliding-Window & 0.94 & 0.94 & 0.94 & 30 / 10\\
+		\bottomrule
+	\end{tabular}
+	\caption{Multi layer results summary.}
+	\label{tab:multilayer_results}
+\end{table}
+
+Having studied how our modifications improved the proposed model, the following chapter goes on to provide practical applications for our work.

### 2018-05-10 thesis_report/Thesis_Practical_Applications.tex
Pratical Applications v1

@@
 
+We take this chapter to briefly describe how we took the previously defined model to the real world, providing practical implementations for it.
+The following work is still on-going as part of a scholarship provided by INOV - Inesc Inovação, as such the end products are still being developed and test.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Our first practical application for the obtained model comes as a malware detection service.
+The idea is to take advantage of Cuckoo~\cite{tool:cuckoo} to implement a web service identical to Malwr~\cite{tool:malwr}, but also adding the malware detection by applying our model.
+
+To implement this service, we developed a simple web application that receives files from users and then sends them to a Cuckoo instance running on a server.
+We created a custom signature for Cuckoo that takes the analysis results and converts it into features suitable for our model, which is then used to classify the sample.
+The output of the model is then appended as a signature the final report.
+After analysis our web application extracts basic information regarding the file, as well as the output of the classifier.
+
+Figure \ref{fig:malware_detetor} shows the service being tested with a \texttt{petya}\footnote{Symantec - Petya Ransomware, October 27, 2017 [https://www.symantec.com/blogs/threat-intelligence/petya-ransomware-wiper]} sample, using early versions of our $\LR$ model.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/malware_detector.png}
+	\caption{Malware Detector in early stages.}
+	\label{fig:malware_detetor}
+\end{figure}
+
+We are currently testing and optimizing the service, as it is not currently using our latest and best model $\mathcal{E}$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{section:email_scanner}
+
+Our second practical application uses the previous service, but in an automated way, to scan email attachments arriving and departing Inesc's mail server.
+The idea is to insert the malware detector in the mail server pipeline as a scanner.
+
+Inesc's mail server uses MailScanner~\cite{tool:mailscanner} to filter and scan emails.
+In Figure \ref{fig:mailscanner} we provide a diagram on the MailScanner pipeline.
+
+Our implementation works by creating a custom \textit{Virus Scanner}, which takes email attachments and sends them to the previous described malware detection service.
+Upon receiving the results, our custom scanner modifies the mail object to include the classification results.
+
+Further down in the \textit{Message Processing} step, we added a custom step to modify an unused \texttt{Header} in the email to include the classification results.
+
+Currently we are in the process of moving the developed code into the mail server to start obtaining real world results.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/mailscanner.png}
+	\caption{MailScanner process diagram.}
+	\label{fig:mailscanner}
+\end{figure}

### 2018-05-10 thesis_report/Thesis_Conclusions.tex
Conclusions v1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Conclusions.tex                                     %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified : May 2018                                         %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Conclusions}
\label{chapter:conclusions}

In this work we analyzed how \gls{ml} techniques fit into the scope of malware detection and how could the chosen dataset influence the results of the classifier.
We collected, analyzed, trained and validated multiple datasets to better understand how \textit{laboratory conditions} vary from \textit{real-world conditions}.
Doing so provided us with \gls{auroc} results varying from as high as 98\% to as low as 67\%.

Given the non-existence of a common agreement on how to label samples in a real world dataset, we have proposed three different metrics $\Mrealv$, $\Mloosev$ and $\Mstrictv$ for labeling these samples, and presented three different scenarios $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, ranging from a more simulated scenario, to more realistic ones.

We have analyzed the different scenarios mainly on two kind of conditions: the \textit{laboratory conditions} where the standard cross-validation methodology was applied discarding the importance of \emph{time} in malware detection, and temporal consistent techniques where we have trained and validated the model in a temporal consistent manner.
We have shown that for a modest compromise in accuracy temporal consistent methodologies are adequate to classify malware samples.

We have also concluded that we can reduce the size of the training dataset to avoid the need of training with all ever seen samples, and argue on how much it can be reduced without compromising optimal results.

Having solid knowledge on the effects of temporal consistency in the task of malware detection, we improved our base model for better results.
This was done by using a multi layer approach to provide the ability to extract information regarding malware classes, and by adding more features to the model.
These improvements boosted our results between 7\% and 20\%, while still guaranteeing temporal consistency.

To show our confidence in the obtained results, we provided two practical applications of our developed model, still under development, but with hopes of being implemented in a production environment soon.

We believe that the pertinent question of how much should we seek for great results on \gls{ml} techniques applied to malware detection is worth to be further discussed, bearing in mind that it leads to classifiers that would not perform better over realistic conditions.

The task we set ourselves to achieve was not without its difficulties, but all in all, we believe our work shows that the path to malware detection via machine learning is feasible, not only theoretically, as related work as shown, but also with practical implications.

% ----------------------------------------------------------------------
\section{Achievements}
\label{section:achievements}

In summary, in this work we achieved the following results:

\begin{itemize}
	\item We provided means to achieve metrics as to facilitate the malware labeling task.
	\item We developed methodologies that test real-world scenarios which take into account time consistency.
	\item We provided a simple model to detect malware and how it behaves under different methodologies.
	\item We improved the previous model to boost the results in all of our tested methodologies.
	\item We started developing practical applications of our model, to be inserted in real world scenarios.
\end{itemize}

% ----------------------------------------------------------------------
\section{Future Work}
\label{section:future}

As future work, the developed model could be improved by adding more features, and by doing better research regarding the impacts of each feature.
Other models could be tested, as well as used in conjunction with our logistic regression, to improve the overall results.
Diverging from a supervised learning methodology to include other types of learning would also bring benefits and possibly achieve a fully automated pipeline to detect malware, with minimal user interaction.

### 2018-08-12 extended_abstract/rebuttal/Rebuttal.md
latest

We would like to thank the reviewers for their constructive reports on our work and will briefly comment below on the raised issues. We will proceed through the remarks presented in each review, merging some of them whenever that makes sense.

- On the public availability of datasets (R1): the data we used for our analysis was collected from 3 different online sources: Malwr.com, NSRL, and VirusShare.com. Although these can be freely used for research, it is not clear from their terms of usage if we could make them available online. We believe however that it would be possible to make them privately available for analysis of the reviewers.

- On the advantages of the stacking methodology (R1): Although not presented in the paper, the stacking methodology presented a slight advantage in terms of accuracy when compared to the linear regression counterpart (~2%). As a side effect of this, we also tackle another important issue in malware classification that is "naming" and division in classes. Even if the end-user of an IDS/IPS is mainly concerned if a given sample is malware or goodware, other users, like researchers, benefit from a more detailed naming, as it helps to analyse and comprehend threats. For this reason, the stacking methodology does give advantages, even under marginal performance advantages.

- Keeping a small set of recent samples vs constant retraining of the model (R1+R3): In our work we were more interested in analysing how much data would be needed to keep a high degree of accuracy. It would not be totally unexpected that to achieve that, the size of the training dataset would constantly increase making this analysis potentially infeasible in a few years time. What our results show is that, for this particular dataset, the compromise that has to be made in terms of accuracy in order to keep a relatively small training dataset is not significant.
The proposed approach by the reviewer would be an interesting one and in fact iteration 2 of Past-to-Present (Fig8) would be the last iteration of that process. One can then see in Figure 14, green line, that although we are just looking at the results of one iteration, the accuracy of this iteration is less than the sliding-window approach.
So, although this comment is valid, and given that the cost of training using the sliding-window is <1min, we believe that the advantages of the sliding-window approach compensate the "low-cost" of the retraining.

- Chosing the "reliable" vendors (R1): Our goal was to select the top-vendors according to each metric (Fig 3 and Fig5). Using the form $mx + b = y$ enables us to balance the weight of FPR over DR. By increasing the slope (i.e. $m>1$) we prioratize low FPR over DR, whereas by decreasing the slope (i.e. $m<1$) we prioratize higher DR over FPR. By letting $m=1$ we did not prioritize any of the solutions.

- As for the organizational aspect mentioned by R1 and R3, we appreciate these comments that will be incorporate in a final version of this paper. Similarly for the images. While we tried to incorporate justifications within the text, some work can be done to improve their readability.

- On the sufficiency of static imports (R2+R3): due to the nature of malware, and the usage of run-time obfuscation techniques, it is conceivable that dynamic analysis significantly outperforms static analysis. In our work, we were interested in analysing how much one could get just by simple analysis of static imports. Our claim is that the simplicity and speed of this analysis is a factor that should be taken into consideration, as the accuracy results show. Also, for low latency applications, such as inline with other email scanning tools, time is a constraining factor, thus static analysis presents itself as a viable option. Of course, combining this with dynamic analysis should only provide better results, although compromising (time) performance.

- The differences Pf-Pl (R2): if Pf-Pl>0, it means that the number of malware classifications reduced, which means that, as a community, we are in the presence of a False Positive. Some vendors classified a sample as malware and then changed their classification (or there was some changes in both directions with a "net-value" in that direction). Symmetrically for Pf-Pl<0. Of course when Pf=Pl, the "positive-vendors in P_first" may be completey disjoint of the "positive-vendors in P_last" but, as a community, the sample did not change its valuation. We have thus an analysis on how the community changed their classification (Figure 2). This curve is not symmetric, which underpins the idea that they do not change the classification in a symmetric fashion. Nevertheless, this fact would not change the statistical meaning of the assessment.

- Why 5 vendors (R2+R3): In Figure 2 one can see that the trend to change valuations from goodware to malware (green bars) is much bigger than the opposite (red bars). Moreover, if we consider the integral up to -5, we get that that value is <0.05%. 
This tells us that less than 5 vendors change their classifications from malware to goodware in 99.5% of the times. 
The reasoning behind this choice for cut-off was that if 5 (or more) vendors classify a sample at the beginning as malware, then in 99.5% of the cases it will still be classified as malware in the final evaluation.

- Distinction between upper case and lower case imports (R2); In fact we tried to normalize the imported library names and realized that we obtained worse results when ignoring case sensitivity. We cannot assert the reason behind it, but it could be possible that goodware development usually resorts to a lot of stacked frameworks and libraries, whereas malware is usually done in a more ad-hoc way, hence it may use less libraries (or less common ones).

- (R2) Due to the categories not being well defined and changing over time, regular machine learning validations should be used with care, as we have shown, obtaining good values is not hard, when no temporal order is imposed. This does not invalidate the use of machine learning as a whole, if proper precautions are taken such as imposing temporal consistency, since we are off-loading manual analysis work from malware analysis.

- Comment on AUROC (R2): This work aims at providing a comparative analysis on which approach fits better malware detection, rather than providing the best model or architecture for malware detection. In this sense, the results should be taken relatively (to each other) and not absolutely: we aim at comparing one approach with another using the same (certainly not the best) model throughout the paper, so that we can draw some comparative conclusions. For this reasons, we do not intend to emphasize the impresiveness of an AUROC above 0.9, but rather its meaning when compared with the remaining approaches. Ideally one could change the model under our defined scenarios and obtain the same relative values between scenarios.

- On the choice of NSRL and VirusShare and identification of samples (R3): we used these two sources as the former is a clear source of goodware, and the latter a clear source of malware. This allowed us to clearly classify some of our samples as good/malware.
Also, in both these repositories, and on malwr.com, samples are identified by the MD5 hash of their payload. This allowed us to not only validate which samples from malwr.com are good/malware but also which samples correspond to repeated submissions.

- 5.56% decrease corresponds to the integral of the red bars in Figure 2. Justified in an earlier comment.

- The analysis of vendors' FP, FN, TP, TN rates stands on their classifications compared against our definition of malware (5+ positive) and goodware (0 positive). We are not sure on what does R3 means with "FT_v". If it means that (exactly) 2 vendors classify a sample positively, that sample would not fit our definitions of malware/goodware and would not be contained in our training example, due to the uncertainty underlying (classical) malware detection.

- The purpose of the table on page 5 was to show the size of the datasets.

- Discussion of proposed evaluation metrics (R3): Cross-validation is a standard ML model but not quite applicable to the context of malware detection due to the temporal order underlying a (real) malware detection scenario where samples are observed in a given order. For this reason, we have proposed the temporal-based models. In the temporal-based scenario, we have analysed the impact of both reducing and sliding the training set, and drew some conclusions on how we could minimize the training data without compromising the optimal values.

- time spent for the training and for the validation: answered above. <1min for sliding window.

### 2018-08-12 extended_abstract/rebuttal/rebuttal.txt
latest

Strategy
1. Address all issues of R1. He seems to be on our side and easily convincible
2. Try to rebut R2. Present clear arguments why he is not correct
3. R3 is somehow indiferent. Does not seem to be convincible to change his opinion either way.

=========================

We would like to thank the reviewers for their constructive reports on our work and will briefly comment below on the most important raised issues.

====> Should we address this?
R1: The definition of ground-truth datasets is interesting, and I'd like to see these datasets available online.
<====

====> Anything smart to rebut this claim? Would this be better in dynamic analysis?
More precisely, the stacking methodology proposed for train a model is not convincingly demonstrated to be superior to its simpler linear regression counterpart, which hardly justifies its use. 
<====

- Keeping a small set of recent samples vs constant retraining of the model. While the comment from the Reviewer~1 is valid, in our work we were more interested in analysing how much data would be needed to keep a high degree of accuracy. It would not be totally unexpected that to achieve that, the size of the training dataset would constantly increase making this analysis potentially infeasible in few years. What our results show is that, for this particular dataset, the compromise that has to be made in terms of accuracy in order to keep a relatively small training dataset is not significant.....
And yes, indeed we keep the most recent data as malware tends to explore new features not necessarily detectable by earlier samples
====> CAN WE RUN AN EXPERIMENT WITH THE PROPOSED SOLUTION BY THE REVIEWER AND REPORT ON THAT? CAN WE REPORT ON THE TIME TAKEN FOR TRAINING?
<====

====> Anything smart to say here?
(2) The discussion of how the set of "reliable" vendors is chosen (Section III.B) is not clear to me. What's the meaning of "we search for the maximum b such that there are exactly 20 vendors above x+b in each graph."? What's "x"? Why is this the rule you use?
<====

====> Probably not much to say here
(1) Key terms and definitions are not properly introduced throughout the paper,
<====
As for the organizational aspect mentioned by Reviewer~1, we appreciate these comments amd will be incorporate them in a final version of this paper. Similarly for the images. While we tried to incorporate justifications within the text, some work can be done to improve their readability.

- On the sufficiency of static imports (R2): due to the nature of malware, and the usage of run-time obfuscation techniques, it is conceivable that dynamic analysis significantly outperforms static analysis. In our work, we were interested in analysing how much one could get just by simple analysis of static imports. Our claim is that the simplicity and speed of this analysis is a factor that should be taken into consideration, as the accuracy results show. Of course, combining this with dynamic analysis can only provide better results.
====> any claim about the time it takes static vs dynamix?
<====

- The differences Pf-Pl (R2): if Pf-Pl>0, it means that the number of malware classifications reduced, which means that, as a community, we are in presence of a False Positive. Some vendors classified a sample as malware and then changed their classification. Symmetrically for Pf-Pl<0. Of course when Pf=Pl, the "positive-vendors in P_first" may be completey disjoint of the "positive-vendors in P_last" but, as a community, the sample did not change its valuation.
====> Is this ok? do we know how much changed like he says? did this ever happened? what was the maximum of chnages from + to - and - to +? This would destroy this argument.

- Why 5 vendors: In Figure 2 one can see that the trend to change valuations from goodware to malware (green bars) is much bigger than the opposite (red bars). Moreover, if we consider the integral up to -5, we get that that value is <0.05%. This tells us that less than 5 vendors change ther classifications from malware to goodware in 99.5% of the times. This means that if 5 vendors classify a sample at the beginning as malware, then in 99.5% of the cases it will still be classified as malware in the final evaluation. This is why we took this cut-off of 5.

====> Anything smart?
Why is the case (upper or lower) of imported library names is taken as a
possible feature? It seems that this could only lead to FNs or FPs,
since it actually has no significance.
<====

====> say something about uniform naming
In the E model, we try to classify not only if a sample is malware,
but whether it is a Trojan, virus, spyware, ransomware or "other".
These categories are not well-defined; they are subjective, they
overlap, their meaning has changed with time, and "other" is
particularly meaningless. It doesn't seem very worthwhile to try to
apply machine learning to such loosely defined terms.
<====

====> comment on AUROC
While AUROC values above 0.9 look impressive, I can't
understand whether they are significant or not. There is no discussion
about what this means. Indeed, the discussion and conclusions are more
like a recap of what has been said before. There seems to be no
surprising or informative or useful conclusions.
<====

====>
- I can't understand the meaning of "the main focus of our work
is to compare results relatively and not absolutely, ..."
<====

- On the choice of NSRL and VirusShare and identification of samples (R3): we used these two sources as the former is a clear source of goodware, and the latter a clear source of malware. This allowed us to cleasly classify some of our samples as good/malware.
Also, in both these repositories, and on malwr.com, samples are identified by the MD5 hash of their payload. This allowed us to not only validate which samples from malwr.com are good/malware but also which samples have repeated submissions.

- 5.56% decrease corresponds to the integral of the red bars in Figure 2. Justified in an earlier comment.

====> Pg4. do not understand
With the previous definitions, we can define a vendors’ v classification (according to MDacc ) as: ... -> I am not 100% convinced of this definition. Why is this the best you could give? 
E.g.,  FT_v is 2 instead of 1 (out of many vendors) why it is still not a false positive? 
This definition is arbitrary and I would expect a better explanation
<====

- The purpose of the table on page 5 was to show the size of the datasets.

====>
Section V I would like to have a better and more convincing explanation of why you chose the proposed evaluation metrics (cross-validation, etc.). Comment more also on the results.
<====

====>
Section B could you please comment on the time spent for the training and for the validation in your experiments?
<====

### 2018-08-12 extended_abstract/abstract.tex
latest

%!TEX root = main.tex

The use of supervised learning techniques for malware detection has been used increasingly to aid classical classification methods. 
In this work we aim at developing a malware detection model, analyzing the impact of the reliability of the training dataset on the final result of the classifier, and metrics to define the \emph{ground-truth}.
For this, we propose three datasets' scenarios whose content range from unambiguous malware and goodware samples to more ambiguous and real ones.
We analyze each scenario in laboratory conditions, where standard cross-validation methodologies are applied, discarding the importance of \emph{time} in malware detection, and also in real-world conditions, where temporal-based dependencies are proposed and applied.
Furthermore, we modify our original model to both enrich the extractable information, by implementing a multi layer model, and to improve the final results, by using dynamic information about the samples.
We then use our temporal-based methodologies to reduce the size of the training dataset without compromising optimal results, concluding that there exists an ideal number of necessary training folds, temporally consistent with the validation fold, that maximizes the overall score.

### 2018-08-12 extended_abstract/appendix.tex
latest

%!TEX root = main.tex
\color{red}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_strict_1}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{strict}$ under \textit{Past-to-Present} .}
	\label{fig:roc_tc_strict_1}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_loose_1}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{loose}$ under \textit{Past-to-Present} .}
	\label{fig:roc_tc_loose_1}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_real_1}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{real}$ under \textit{Past-to-Present} .}
	\label{fig:roc_tc_real_1}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_strict_2}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{strict}$ under \textit{Present-to-Past} .}
	\label{fig:roc_tc_strict_2}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_loose_2}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{loose}$ under \textit{Present-to-Past} .}
	\label{fig:roc_tc_loose_2}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_real_2}
	\caption{\gls{roc} and \gls{auroc} for $\mathcal{S}_{real}$ under \textit{Present-to-Past} .}
	\label{fig:roc_tc_real_2}
\end{figure}

### 2018-08-12 extended_abstract/conclusion.tex
latest

%!TEX root = main.tex

\section{Conclusion}\label{sec:conclusion}

In this paper we analyzed how \gls{ml} techniques fit into the scope of malware detection and how could the chosen dataset influence the results of the classifier.

Given the non-existence of a common agreement on how to label samples in a real world dataset, we have proposed three different metrics for labeling these samples, and presented three different scenarios, ranging from a more simulated scenario, where better results are achieved, to more realistic ones, where the \gls{auroc} results can go down by 23\%.
We have analyzed the different scenarios mainly on two kind of conditions: the laboratory conditions where the standard cross-validation methodology was applied discarding the importance of \emph{time} in malware detection, and temporal-consistent techniques where we have trained and validated the model in a temporal-consistent manner.
We have shown that for a modest compromise in accuracy temporal-consistent methodologies are adequate to classify malware samples.

We have also concluded that we can reduce the size of the training dataset to avoid the need of training with all ever seen samples, and argue on how much it can be reduced without compromising optimal results.

Having a sound understanding of the effects of different methodologies, we improved our model to yield higher results.

We believe that the pertinent question of how much should we seek for great results on \gls{ml} techniques applied to malware detection is worth to be further discussed, bearing in mind that it leads to classifiers that would not perform better over realistic conditions. 
As future work we aim at optimizing our logistic regression model, at increasing and optimizing the features, and finally, at developing a supervised learning methodology to classify malware samples according to the main malware families.

### 2018-08-12 extended_abstract/datacollection-old.tex
latest

%!TEX root = main.tex
\color{red}

\section{Data Collection and Labeling-OLD}\label{sec:data_collection}

This section aims at providing a brief overview of the available \textit{corpus} and how it can be used to define data labeling metrics. On top of that, we propose a vendor selection to improve the reliability of data labeling.

\subsection{Data Collection}

The \textit{corpus} used to train and evaluate the current work was obtained from Malwr\cite{tool:malwr}. This service provides free malware analysis by using Cuckoo Sandbox\cite{tool:cuckoo} to analyze files. Alongside the analysis results, Malwr also aggregates antivirus' signatures given by VirusTotal\cite{tool:virustotal} at the time of the analysis.

In general, malware detection is not a deterministic task, several nuances make vendors disagree on what is malware. For this reason, we propose to enrich the available information with an additional couple of repositories, namely, VirusShare\cite{tool:virusshare} for malware, and \gls{nsrl}\cite{tool:nsrl} for goodware, enabling us to cross-check VirusTotal's classifications. We observe that for VirusShare there are 105,251 common samples, and for \gls{nsrl} 4,756. On top of that, they share 426 samples. The cross-checking with these repositories will be a key element to increase the reliability of the ground truth, at a later stage.

We now proceed with an initial filtering of the dataset. For the purpose of this work we are interested in reports from samples that provided static information to be used as features, specifically static imports (e.g.\ \gls{dll}), and that are known in VirusTotal (i.e.\ classified), to provide a way to label samples.

We start by discarding vendors that have not seen at least 90\% of \textit{corpus} obtained from Malwr, let $\mathcal{V}_{>0.9}$ denote the remaining vendors. From now on, we only consider the reports classified by some vendor in $\mathcal{V}_{>0.9}$. We refer to the set of classified samples as $\mathcal{D}_{class}$, which contains 292,127 reports.

Our model's features will be obtained from the reports' static imports, so let us denote by $\mathcal{D}_{static}$, $\mathcal{D}_{static} \subseteq \mathcal{D}_{class}$, the classified samples that contain static imports. This dataset contains 287,117 reports.

\subsection{Data Labeling}

We now turn our focus into labeling the reports as goodware or malware. To do so, we use $\mathcal{D}_{class}$ together with VirusShare and \gls{nsrl} information to derive three different metrics to label the reports as benign or malicious, over a set of vendors $\mathcal{V}$.

The first and tightest metric we define, $\mathcal{M}_{strict}^\mathcal{V}$, labels a sample from $\mathcal{D}_{class}$ as:

\begin{itemize}
	\item \textbf{Malware} if all vendors in $\mathcal{V}$ classify it positively, it belongs to VirusShare repository, and does not belong to \gls{nsrl} repository;
	\item \textbf{Goodware} if all vendors in $\mathcal{V}$ classify it negatively, it belongs to \gls{nsrl} repository, and does not belong to VirusShare repository.
\end{itemize}

Obviously this is the most reliable metric, in the sense that it is closely related to the samples' ground truth, leaving little room for disagreement. However, it is very likely that it
will not be applicable to majority of the samples.

To overcome the strictness of the previous metric, we now define a couple of new metrics that will balance the vendors' agreement on the classification. Indeed, given the reports are classified by multiple antivirus solutions, it is not surprising to see disagreement on whether a sample is malware or not. Due to this, and following a methodology identical to \cite{miller:rev_int}, we analyze duplicated submissions and establish a minimum number of vendors needed to label a sample as malware.

Let us denote by $\mathcal{D}_{dups}$, $\mathcal{D}_{dups} \subseteq \mathcal{D}_{class}$, the samples submitted more than once. These duplicated samples constitute 26.86\% of $\mathcal{D}_{class}$, with a total of 28,996 samples. Figure \ref{fig:distribution_changes} shows the distribution of samples that increase or decrease in positive classifications between the last and the first submissions.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dist_changes}
	\caption{Distribution of samples that increase or decrease in the number of positive classifications between last and first submissions.}
	%\label{fig:distribution_changes}
\end{figure}

We first note that 40.18\% of the duplicated samples change in classification. From those, 34.55\% increase in classification (i.e.\ more vendors classifying as malware), whereas only 5.63\% decrease in classification (i.e.\ less vendors classifying as malware). Such discrepancy between positive and negative changes suggest a preference for false negatives over false positives, as also noted in \cite{miller:rev_int}.

Following the approach in \cite{miller:rev_int} and \cite{vsrndic2013detection}, in the next scenarios we will require at least 5 positive classifications to label a sample as malware. This threshold is supported by the results presented in Figure \ref{fig:distribution_changes}, where we can see that samples that decrease in 5 or more positive represent 0.45\% of samples.

With regards to goodware labeling, we require that a sample must have zero positive classifications. We further analyze this requirement, as there is a high possibility that a sample labeled as goodware increases to some positive classification, given that 34.55\% of samples from $\mathcal{D}_{dups}$ increase in classification.

We are interested in knowing the possibility of a sample that is labeled as goodware, to become malware according to the previous malware labeling requirement. For this purpose, Figure \ref{fig:distribution_changes_zero} presents the same analysis as Figure \ref{fig:distribution_changes}, but restricted to samples that start with zero positive classifications.

% We first identify that 17.77\% of samples increased from zero to some positive classification. From those, 7.51\% increase from zero to 5 or more (our malware threshold), which indicates that there is some chance of wrongly labeling a malware sample as goodware.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dist_changes_zero}
	\caption{Distribution of samples that start with zero classifications and increase in the number of positive classifications to the last submission.}
	\label{fig:distribution_changes_zero}
\end{figure}

We verify that 17.77\% of samples increased from zero to some positive classification. From those, 7.51\% increase from zero to 5 or more (our malware threshold), which indicates that it is possible to wrongly label a malware sample as goodware, by solely using this criterion.

Based on the previous analyses, we define an intermediate metric, characterized by not being as strict as $\mathcal{M}_{strict}^\mathcal{V}$, but still trying to overcome the aforementioned vendors' classification inconsistency. Consider this new metric, $\mathcal{M}_{loose}^\mathcal{V}$, which labels a sample from $\mathcal{D}_{class}$ as:

\begin{itemize}
	\item \textbf{Malware} if at least 5 vendors in $\mathcal{V}$ classify it positively, it belongs to VirusShare repository, and does not belong to \gls{nsrl} repository;
	\item \textbf{Goodware} if all vendors in $\mathcal{V}$ classify it negatively, it belongs to \gls{nsrl} repository, and does not belong to VirusShare repository.
\end{itemize}

Although more prone to incorrect labeling, when compared to $\mathcal{M}_{strict}^\mathcal{V}$, metric $\mathcal{M}_{loose}^\mathcal{V}$ encompasses a larger variety of samples, being much closer to reality.

As seen in Figure \ref{fig:distribution_changes}, it is widely accepted that malware detection is imperfect and prone to error. We can argue that although strict metrics provide more reliable labeling, and consequently provide better models and results, they lack the ability to cope with real-world conditions.

For this reason, we propose a third metric, $\mathcal{M}_{real}^\mathcal{V}$, that solely relies on vendors' classification, hence including their imperfection and uncertainty. It labels a sample from $\mathcal{D}_{class}$ as:

\begin{itemize}
	\item \textbf{Malware} if at least 5 vendors in $\mathcal{V}$ classify it positively;
	\item \textbf{Goodware} if all vendors in $\mathcal{V}$ classify it negatively.
\end{itemize}

This metric is closer to antivirus vendors' reality, and incorporates the underlying classification uncertainty.

%We first note that out of 28,996 samples, 11,651 (40.18\%) change in classification; furthermore, we note that 34.55\% of samples increase in classification (i.e.\ more vendors classifying as malware), while only 5.63\% decrease in classifications (i.e.\ less vendors classifying as malware). This tendency suggests that vendors favor false negatives over false positives, as also noted in \textbf{[quote]}.

% follow the same approach and use 5

%Having an understanding on how vendors tend to change their classification, we choose a minimum threshold of 5 positive classifications to label a sample as malware\textbf{cite}. \textbf{Figure X} shows that samples that decrease in 5 or more positive classifications represent 0.45\% of samples, meaning that the probability of a given sample where 5 or more vendors classify it malware actually turns out to be goodware is lower than 0.45\%.

%The actual value is lower because our graph does not take into account the starting number of positive classifications, meaning it also accounts for cases where a sample starts with 20 positive classifications and ends with 5.

%To label a sample as goodware, we choose a metric where there must be zero positive classifications in a sample. Although intuitive, this metric may have drawbacks when considering that 34.55\% of samples increase in classification. To better understand how likely it is for a sample with zero classifications to go above the previously defined malware threshold (5 or more), we reproduced \textbf{Figure X} but only taking into account changes for samples that start with zero classifications. The results can be seen in \textbf{Figure X}.

%We first identify that 17.77\% of samples increased from zero to some positive classification. From those, 7.51\% increase from zero to 5 or more (our malware threshold), which indicates that there is some chance of wrongly labeling a malware sample as goodware.

%\todo[inline]{To have more room to improve the reliability of the dataset; enhance the metric with info from VS and NSRL.}

%To minimize the chance of wrongly labeling goodware, and to achieve a better overall ground truth, we applied the knowledge from VirusShare and \gls{nsrl}.

%The applied methodology is straightforward, instead of relying solely on the reports' classification, we also require samples to be present in VirusShare and \gls{nsrl} to be labeled as malware and goodware, respectively. Doing so results in having a 86,503 malware samples and 4,000 goodware samples. It is worth noting that samples both in Virus Share and \gls{nsrl} dataset were discarded, as their actual label is dubious.

%These results gave us a sound metric to label samples as malware or goodware. The next subsection builds on this labeling process to narrow down the total amount of vendors, by measuring their detection and false positive score, based on our metric.

% \todo[inline]{Summarize the 3 metrics}

\subsection{Top Vendors}

Since our metrics are highly dependent on vendor's classification, in this subsection we aim at narrowing down the total amount of vendors, by assessing their \gls{dr} and \gls{fpr} according to our strictest metric $\mathcal{M}_{strict}^{\mathcal{V}_{>0.9}}$.

For each vendor $v$ in $\mathcal{V}_{>0.9}$, we considered the metric $\mathcal{M}_{strict}^{\mathcal{W}}$, over the set of vendors $\mathcal{W} = \mathcal{V}_{>0.9}\setminus \{v\}$ to label the samples as malware and goodware. We calculate $v$'s \gls{dr} as the ratio of $v$'s positive detections over the total malware. Furthermore, we calculate $v$'s \gls{fpr} as the ratio of $v$'s positive detections over the total goodware.

The \gls{dr} and \gls{fpr} for each vendor are depicted in Figure's \ref{fig:dr_fpr} left and right Y axis, respectively.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dr_fpr}
	\caption{Calculated \gls{dr} and \gls{fpr} for $\mathcal{M}_{strict}^{\mathcal{V}_{>0.9}}$ metric, ordered by \gls{dr}.}
	\label{fig:dr_fpr}
\end{figure}

We proceed with the 20 vendors with highest \gls{dr}. Let $\mathcal{V}^*$, $\mathcal{V}^* \subseteq \mathcal{V}_{>0.9}$, represent the top 20 vendors, and consider this refinement of classified samples, $\mathcal{D}_{class}^* \subseteq \mathcal{D}_{class}$, by filtering according to $\mathcal{V}^*$. Similarly, consider the refined set $\mathcal{D}_{static}^* \subseteq \mathcal{D}_{static}$, the classified samples from $\mathcal{D}_{class}^*$ that contain static imports.

### 2018-08-12 extended_abstract/datacollection.tex
latest

%!TEX root = main.tex
\color{black}

\section{Data Collection, Analysis and Labeling}\label{sec:data}

Malware classification is non-deterministic task, as several nuances make vendors disagree on what they classify as malware. Not only subtleties related to programs such as \emph{adware} or \emph{remote management consoles}, for which one can find arguments to classify them in either class, but also because some vendors are more accurate than others when classifying malware.

In order to be able to study how \emph{temporal consistency} and \emph{ground truth} influences a model trained to detect malicious and non-malicious applications, we gathered a large dataset of publicly available samples that includes both \emph{malware} and \emph{goodware} with no a priori labeling.

In this section we will provide an overview of the collected data, in particular its sources, followed by an analysis on these samples, and a discussion on the cross-checking mechanisms we used to perform the labeling. In the end of this section we will provide 3 metrics for analysis that differ in the confidence one can provide on the samples' labeling.

\subsection{Data Sources}\label{sub_sec:data_collection}

To study how \emph{temporal consistency} and \emph{ground truth} influences a model trained to detect malicious and non-malicious applications, we started by gathering a dataset of malware and goodware.
This dataset was obtained from \emph{Malwr}~\cite{tool:malwr}, an online service that runs static and dynamic analysis on user submitted files using \emph{Cuckoo Sandbox}~\cite{tool:cuckoo}.
Although Malwr accepts any kind of file, we are interested solely in \gls{pe} files, not only because Windows' systems are relevant targets of malicious applications, but also due to the fact that our initial approach will focus on static information obtained from these \gls{pe} files.

Malwr service provides the analysis of the submitted files as well as the MD5 of such files, but no labeling as whether a sample corresponds to a malicious or non-malicious application.
To perform this labeling we use the anti-virus' signatures given by \emph{VirusTotal}~\cite{tool:virustotal} at the time of analysis (incorporated in Malwr's reports), as a means of labeling the gathered reports.
VirusTotal is an online service where one can submit a suspicious file and in return obtain a list with the analysis performed by a significant number of vendors on whether the sample is clean or not.

However, and as mentioned before, these classifications are not unanimous, and we still need to assign a labeling to each sample.
For this reason, we enrich our knowledge regarding samples' ground truth by aggregating metadata from \emph{\gls{nsrl}}~\cite{tool:nsrl} and \emph{VirusShare.com}~\cite{tool:virusshare}, specifically the samples' MD5. 
NSRL contains a collection of digital signatures of known, traceable software applications, whereas VirusShare.com is a repository of malware samples, so a sample belonging to the \gls{nsrl} set gives us a higher confidence that it is indeed goodware, whereas one belonging to VirusShare.com set gives us a higher confidence that it is malware.

In summary, we collected reports from the \gls{pe} samples available in Malwr, and complemented it with metadata from \gls{nsrl} and VirusShare. The following subsection quantifies our \textit{corpus}, providing a better understanding of the available samples.

\subsection{Data Analysis}\label{sub_sec:data_analysis}
We collected our data from Malwr between April 16th, 2013 and October 10th, 2016.
Our data can be divided in three sets:
\begin{enumerate*}[label={\alph*)},font={\color{red!80!black}\bfseries}]
	\item a set $\RR$ of raw reports, containing 388,702 \gls{pe} samples; 
	\item a set $\CC\subseteq\RR$ of 284,880 classified reports by 38 vendors ($\mathcal{V}$) that is obtained by restricting the original set $\RR$ (that includes 97 vendors) to those reports whose vendors are present in at least 95\% of the classified samples.
\end{enumerate*}
% Our working dataset $\DD=\CC\subseteq\RR$, is of size 280,391 (72.17\% of $\RR$).

With regards to duplicated submissions, there are 27,798 samples submitted more than once, for a total of 74,916 duplicated submissions $\Cdups$ averaging 2.7 submissions per duplicate.
Understanding how vendors change their signatures on samples is crucial, as we use it to label the dataset. With this in mind, and inspired by Miller et al.~\cite{miller:rev_int}, we start our analysis by studying the differences between the number of positive (\ie\ a sample of malware) and negative (\ie\ a sample that is not malware) classifications over the last and first submission of the same sample.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% From Data Analysis 2 notebook
For each duplicated sample in $\Cdups$ we counted the number of positive classifications on the first and last submissions, $\Pf$ and $\Pl$ respectively.
If $\Pf > \Pl$ then we are looking at a possible \emph{false positive (FP)}, as the number of vendors classifying the sample as malware decreased. Conversely,  if $\Pl > \Pf$ we are looking at a potential \emph{false negative (FN)}. For the case $\Pf=\Pl$ we conclude that vendors are confident regarding their classification for the sample.

Figure \ref{fig:distribution_changes} shows the frequency of $\Pl-\Pf$ for each duplicated sample. We first note that 44.32\% of duplicated samples change in classification, among which 38.72\% increase its classification, whereas only 5.61\% decrease.
Such discrepancy between positive and negative changes suggest a preference for false negatives over false positives, as also noted by Miller et al.~\cite{miller:rev_int}.

\begin{figure*}[]
% 	\centering
	\includegraphics[width=\textwidth]{dups_frequency}
	\caption{Distribution of samples in terms of the changes in the number of positive classifications between last and first submissions, ($\Pl-\Pf$).}
	\label{fig:distribution_changes}
\end{figure*}

\medskip

Another interesting analysis on our dataset is understanding the vendors \emph{\gls{dr}} (or True Positive Rate), and \emph{\gls{fpr}}. 
Although these formulas are trivially defined respectively as
\begin{eqnarray}
\dr&=&\frac\tp{\#malware}=\frac\tp{\tp+\fn} \\ 
\fpr&=&\frac\fp{\#goodware}=\frac\fp{\tn+\fp}
\end{eqnarray}
%{\#goodware}, 
we lack ground truth for what is ${\#malware}$ and $\#goodware$.
To solve this, we propose relative metrics to compute what is positive ($\#malware$) and negative ($\#goodware$).

Our first approach is to take advantage of duplicated submissions to define an accuracy metric, $\Madups$.
As we have previously shown, 44.32\% of duplicated samples change in classification, which can be translated into vendors acknowledging their own errors.

With that in mind, for each vendor $v\in\VV$, we define a duplicated sample for $v$ (according to $\Madups$) as:
\begin{itemize}%[\IEEEsetlabelwidth{Z}]
	\item $\tp_v$, \emph{true positive for $v$}, if $v$ classified it positively in both the first and last submissions;
	\item $\tn_v$, \emph{true negative for $v$}, if $v$ classified it negatively in both the first and last submissions;
	\item $\fp_v$, \emph{false positive for $v$}, if $v$ classified it positively in the first submission and negatively in the last submission;
	\item $\fn_v$, \emph{false negative for $v$}, if $v$ classified it negatively in the first submission and positively in the last submission.
\end{itemize}

Figure~\ref{fig:dr_fpr_own} plots each vendors' $\dr_v$ vs.\ $\fpr_v$. We note that vendors do acknowledge their classification errors, as we see a detection rate from 56.82\% to 85.29\%, with a false positive rate ranging from 0.03\% to 6.91\%. 
Had they kept their original classification, one would have that $\Pl-\Pf = 0$ for every duplicate and consequently $\dr_v = 1$ and $\fpr_v = 0$. 
Notice that by keeping the original classification all clean samples remain clean, hence $\fn_v=0$, and all malicious samples remain malicious, hence $\fp=0$.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dr_fpr_own}
	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Madups$.}
	\label{fig:dr_fpr_own}
\end{figure}

\medskip

For our second approach regarding vendors' accuracy, we take into account our observations from Figure~\ref{fig:distribution_changes} and our dataset $\CC$ to define another metric $\Mad$.

Intuitively a sample is classified as goodware according to this metric, \ie\ negative, if every vendor $v\in\mathcal{V}$ classifies it as clean.
To understand if our intuition is sound, we plot Figure~\ref{fig:distribution_clean}, a subset of Figure~\ref{fig:distribution_changes}, showing the frequency $\Pl-\Pf$ for samples that were classified as clean in their first submission, \ie, samples with $\Pf=0$. 
These account for 4,902 samples, 3,741 (76.32\%) of which do not increase in classification.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{distribution_clean}
	\caption{Distribution of samples that started as goodware and changed in the number of positive classifications between the last and first submission, \ie, $\Pl-\Pf$ for samples with $\Pf=0$}
	\label{fig:distribution_clean}
\end{figure}

To arrive at a positive (\ie\ malware) sample definition, we relate Figure \ref{fig:distribution_clean} with Figure \ref{fig:distribution_changes}.
Specifically we want to find a minimum threshold of positive classifications to define a sample as malware.
We chose five as the threshold, observing that percentage of samples that decrease in 5 or more positive classifications is 0.46\%, meaning it is an upper bound for samples that decrease from 5 or more to zero classifications.

With the previous definitions, we can define a vendors' $v$ classification (according to $\Mad$) as:

\begin{itemize}%[\IEEEsetlabelwidth{Z}]
	\item $\tp_v$, \emph{true positive for $v$}, if $v$ and at least 5 other vendors in $\mathcal{V}$ classify it positively;
	\item $\tn_v$, \emph{true negative for $v$}, if $v$ and all other vendors in $\mathcal{V}$ classify it negatively;
	\item $\fp_v$, \emph{false positive for $v$}, if $v$ is the only vendor in $\mathcal{V}$ classifying it positively;
	\item $\fn_v$, \emph{false negative for $v$}, if $v$ classifies it negatively and at least 5 other vendors in $\mathcal{V}$ classify it positively.
\end{itemize}

Figure \ref{fig:dr_fpr_d} plots each vendors' $\dr_v$ vs. $\fpr_v$ according to~$\Mad$.
Using this metric we note that vendors' detection rate is more scattered than under $\Madups$, ranging from 20.17\% to 83.50\%, whereas false positive rate is similar, ranging from 0.01\% to 5.77\%.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dr_fpr_d}
	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Mad$.}
	\label{fig:dr_fpr_d}
\end{figure}

Given the impossibility of finding a source that is able to unanimously label each sample in our dataset, we want to find the vendors that perform the best under both metrics $\Madups$ and $\Mad$.
%We do so by filtering the top 20 vendors $\VVdups$ and $\VVd$, according to each metric and define $\VVstar=\VVdups\cap\VVd$.

Our approach to this problem is to filter the top 20 vendors $\VVdups$ and $\VVd$, according to each metric and define $\VVstar=\VVdups\cap\VVd$.
Given we have to maximize two variables, \tpr\ and \fpr, we decided to take advantage of the linear equation in the form $mx+b=y$ to choose the top vendors.
This form allows us to choose an $m$ and $b$ such that there are 20 vendors above the line, the top vendors $\VVstar$.
By tweaking the variable $m$ one can change the line's steepness, reflecting in a preference between \tpr\ and \fpr\:

\begin{itemize}
	\item \tpr\ preference: $m < 1$, less steepness therefore higher \fpr\ and \tpr\ values;
	\item \fpr\ preference: $m > 1$, more steepness, lower \fpr\ and \tpr\ values.
\end{itemize}

% Since we have no \emph{a priori} preference between $\tpr$ nor $\fpr$, we search for the maximum $b$ with $m=1$ such that there are exactly 20 vendors above $x+b=y$ in each graphic (Figures~\ref{fig:dr_fpr_own} and~\ref{fig:dr_fpr_d}).

Since we have no \emph{a priori} preference between $\tpr$ nor $\fpr$, we search for the maximum $b$ such that there are exactly 20 vendors above $x+b$ in each graphic (Figures~\ref{fig:dr_fpr_own} and~\ref{fig:dr_fpr_d}).

Figure \ref{fig:dr_fpr_top} shows the $\dr_v$ vs. $\fpr_v$ for the resulting 11 vendors under each metric, $\Madups$ in green and $\Mad$ in red, with $\dr_v$ varying from 63.49\% to 83.50\% and $\fpr_v$ from 0.01\% to 3.67\%.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dr_fpr_top}
	\caption{$\tpr_v$ vs. $\fpr_v$ according to $\Madups$ (green) and $\Mad$ (red).}
	\label{fig:dr_fpr_top}
\end{figure}

% To conclude the analysis of our datasets, we discuss the samples obtained from $\NSRL$~($\DNSRL$) and VirusShare.com~($\DVS$).
%We end this subsection with how the sets from \gls{nsrl} $\mathcal{D}_{NSRL}$ and VirusShare $\mathcal{D}_{VS}$ intersect our dataset $\mathcal{D}$. 
% Samples in $\DNSRL \cap \DD$ account for 4,756, while the number is 105,251 for $\DVS \cap \DD$.

% Finally, and due to the ambiguity in the definition of malware, we also have that 426 samples are common to the three sets, \ie, are in $\DNSRL \cap \DVS \cap \DD$.

\subsection{Data Labelling}\label{sub_sec:data_labeling}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% From Data Labeling notebook
Having defined a collegiate set of vendors $\VVstar$ to classify the samples in our dataset, we can now turn our focus into labeling the reports as goodware or malware. 
To do so, we use $\CC$, together with $\NSRL$ and $\VS$, to derive three different metrics to label the reports as benign or malicious, over the set of vendors $\mathcal{V}^*$.

The first and most real metric we define is $\Mrealv$ that labels a sample $s\in\CC$ as:

\begin{itemize}
	\item $s\in\mal_\real$ if at least 5 vendors in $\VVstar$ classify $s$ positively;
	\item $s\in\good_\real$ if all vendors in $\VVstar$ classify $s$ negatively.
\end{itemize}

Since the labeling information is solely provided by $\VVstar$'s vendors, this metric's ground truth is highly dependent on their performance, which means labeling errors may be present (as we have discussed in \ref{sub_sec:data_analysis}). 
Due to this, samples that are classified positively by no more than four vendors, are discarded.

\medskip

Our second metric $\Mloosev$, restricts the previous metric $\Mrealv$ to achieve a better ground truth. We do this by including information from $\NSRL$ and VirusShare.com. 
This metric labels a sample $s\in\mathcal{C}$ as:

\begin{itemize}
	\item $s\in\mal_\loose$ if $s\in\mal_\real$ \textbf{and} it belongs to $\DVS$ and does not belong to $\DNSRL$;
	\item $s\in\good_\loose$ if $s\in\good_\real$ \textbf{and} it belongs to $\DNSRL$ and does not belong to $\DVS$.
\end{itemize}
that is
\begin{eqnarray*}
\mal_\loose&=&\left(\mal_\real\cap\DVS\right)\setminus\DNSRL\\
\good_\loose&=&\left(\good_\real\cap\DNSRL\right)\setminus\DVS
\end{eqnarray*}

By taking into account the presence in $\NSRL$, that reinforces cleanliness, and VirusShare.com, that reinforces maliciousness, this metric is more reliable, ground truth wise, at the expense of a smaller number of samples.

\medskip

Our third and final metric $\Mstrictv$, is the strictest one, labeling a sample $s\in\DD$ as:
\begin{itemize}
	\item $s\in\mal_\strict$ if all $v\in\VVstar$ classify it positively \textbf{and} $s\in\DVS\setminus\DNSRL$;
	\item $s\in\good_\strict$ if $s\in\good_\loose$.
\end{itemize}

Obviously this is the most reliable metric, in the sense that it is closely related to the samples' ground truth, leaving little room for disagreement. However, this is achieved again at the cost of a smaller number of samples.

\medskip

Taking the previously defined metrics, the task of creating labeled datasets based on them is trivial.
We apply each of the previously defined metrics, $\Mstrictv$, $\Mloosev$ and $\Mrealv$ to our classified dataset $\CC$ to obtained three new datasets, $\CC_{strict}\subseteq\CC_{loose}\subseteq\CC_{real}\subseteq\CC$.
Table \ref{tab:dataset_sizes} provides information regarding the size and number of malware and goodware in each of our datasets.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{l|ccc}
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\hline
		Malware			& 98,582 & 45,306 & 24,658\\
		Goodware		& 56,475 & 1,989 & 1,989\\
		\hline
		Total			& 155,057 & 47,295 & 26,647\\
	\end{tabular}
	\smallskip
	\caption{Sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes}
\end{table}

### 2018-08-12 extended_abstract/discussion.tex
latest

%!TEX root = main.tex

\section{Discussion}\label{sec:discussion}

We have proposed different scenarios, based on different labeling metrics, to study laboratory \textit{vs} real-world conditions. Our scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, vary both in reliability and size, going from a more reliable and small dataset to a larger and less reliable one.
This (un)reliability is due to the fact that for real world samples there is usually no agreement among vendors on how to classify a given sample, and for that we had to assign a labeling to such samples according to our proposed metrics.
We have developed several comparative analysis between these three scenarios, to evaluate how much the nature of the dataset can influence the results.
We have split the analysis into two major validation conditions: the cross validation methodology, where the time consistency is discarded; and temporal-based methodologies.
We tested our methodologies by using a simple \gls{lr} model, which was then improved to transmit better information and to use more features.

Following a cross-validation methodology, we have confirmed our intuitions: $\mathcal{S}_{\strict}$ showed up an \gls{auroc} of 0.91, $\mathcal{S}_{\loose}$ have presented 0.9, whereas $\mathcal{S}_{\real}$ decreased to 0.75. 
As we have argued, the results on $\mathcal{S}_{\strict}$ are justified by factors like a small and reliable dataset, and the use of cross-validation, which mixes samples and ignores possible dependencies between them.
This scenario is composed by very well-known and analyzed samples.
Although $\mathcal{S}_{\loose}$ slightly relaxes these requirements, it is still composed by very well-known samples, which partially justifies the comparable \gls{auroc} (0.90).
But this difference is interesting, as although the number of malware labeled samples increased significantly, the results are not that affected.
As we have noticed, this might also suggest that vendors do converge on their definition of malware, under our $\mathcal{M}_{\loose}$ metric. 
The changes observed from $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ are more remarkable, but somehow expected.
The metric that labels malware and goodware for the scenario $\mathcal{S}_{\real}$ disregards the cross-check from outside repositories, which in turn degrades the reliability significantly, while increasing the dataset size notably. 
As we have already noticed, we attribute the result's degradation mainly to the unreliability of goodware labeling, not only because we have previously seen that the increase in malware does not significantly impact the results (from $\SS_\strict$ to $\SS_\loose$), but also due to the tendency for false negatives in vendors (Figure~\ref{fig:distribution_changes}), which in turn lead us to incorrectly label as goodware some of the malicious samples in $\DD$.

When temporal consistency comes into play, the results on different scenarios do not differ much, nevertheless we can observe more pronounce trends.
The great conclusion that we can take stands on the relative position of the training set with respect to the validation set and its size.
Indeed, samples closer to the validation set seem to benefit the model.
We argue that there should be an ideal number of necessary training folds (30\% of the dataset), temporally consistent with the validation fold (10\% of the dataset), needed to maximize the overall score.
This supports our argument that we can reduce the size of the training set, without losing any significant score.

We finished our analysis by validating this temporal-based results.
For this purpose, we have defined a sliding window for each scenario, with the above parameters, that moved forward in time (1 fold at a time) and obtained the \gls{auroc} values of 0.89 for $\mathcal{S}_{\strict}$, 0.88 for $\mathcal{S}_{\loose}$ and 0.73 for $\mathcal{S}_{\real}$.
Comparing these results with the baseline cross-validation, we note a very slight decrease for each scenario.
This decrease, although not significant, was more than expected due to the enforcement of temporal consistency as well as the significant reduction of the size of the training set. 
We should highlight that these results should be much closer to reality than the ones provided by cross validation techniques, since we are requiring temporal consistency and also a reasonable amount of data for training purposes, which might be a relevant issue in a few year's time.
Indeed, aiming at complementing antivirus' vendors techniques with machine learning, we should not expect to gather and use all the samples ever seen for training purposes, and these results may be very useful on the choice of the right training set.

Finally, we describe multiple improvements to our base model $\LR$ in order to improve the overall results.
We started by using a multi layer approach to build a new model $\mathcal{E}$, which enables the extraction of more detailed information regarding a malicious sample, specifically the malware class it belongs.
We also introduced three new dynamic features, to improve the amount of information obtained from the samples.
After applying the same evaluation methodologies to our new model $\mathcal{E}$, we observed an increase in all cases.
We note that the bigger the dataset, the higher the improvement, as $\mathcal{S}_{\strict}$ increased by 0.07 (cross-validation), $\mathcal{S}_{\loose}$ by 0.08 and $\mathcal{S}_{\real}$ (cross-validation) by 0.27 (\textit{present-to-past}).
This comes to show how the model was better able to learn from the new features.

### 2018-08-12 extended_abstract/evaluation-old.tex
latest

%!TEX root = main.tex
\color{red}
\section{Evaluation and Results}\label{sec:eval_results}

In this section we aim at doing a comparative analysis on three different scenarios $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, built on top of the previously defined metrics. This comparison is done using standard cross-validation methodologies and a proposed temporal-based methodology. We further provide an analysis on how to reduce the size of the training set, without compromising the final results.

\subsection{Evaluation}

The three scenarios that we will focus on will rely on metrics $\mathcal{M}_{strict}^\mathcal{V^*}$, $\mathcal{M}_{loose}^\mathcal{V^*}$ and $\mathcal{M}_{real}^\mathcal{V^*}$, over the dataset $\mathcal{D}_{class}^*$:

\begin{itemize}
	\item Strict Scenario $\mathcal{S}_{strict}$, labels as malware samples that: all vendors in $\mathcal{V^*}$ classify positively, belong to VirusShare repository, and do not belong to \gls{nsrl} repository; labels as goodware samples that: all vendors in $\mathcal{V^*}$ classify negatively, belong to \gls{nsrl}, and do not belong to VirusShare repository. This scenario contains 1,361 malware samples and 4,000 goodware samples.
	\item Loose Scenario $\mathcal{S}_{loose}$, labels as malware samples that: at least 5 vendors in $\mathcal{V^*}$ classify positively, belong to VirusShare repository, and do not belong to \gls{nsrl} repository; labels as goodware samples that: all vendors in $\mathcal{V^*}$ classify negatively, belong to \gls{nsrl}, and do not belong to VirusShare repository. This scenario encloses 85,503 malware samples and 4,000 goodware samples.
	\item Real Scenario $\mathcal{S}_{real}$, labels as malware samples that: at least 5 vendors in $\mathcal{V^*}$ classify positively; labels as goodware samples that: all vendors in $\mathcal{V^*}$ classify negatively. This scenario contains 186,444 malware samples and 49,783 goodware samples.
\end{itemize}

\paragraph{Cross validation}
To gain a better insight on how the model generalizes our scenarios, we apply a k-fold cross-validation, with $k=10$. This methodology splits the dataset into $k$ subsets (i.e.\ folds), selects a single fold for validation and uses the remaining $k-1$ folds as training. This process is repeated $k$ times, ensuring every fold is used for validation and training.

Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples. Since we want to measure the score when training samples pre-date validation samples, we now define a couple of temporal consistent validations.

\paragraph{Temporal consistent validation}
The first temporal consistent validation, which we designate as \textit{Past-to-Present} validation, can be resumed as an iterative methodology where the validation set is fixed with the most recent samples, and the training set with the oldest. At each iteration the training set is extended with more recent samples and scored against the validation, until all samples are used.

The seconds temporal consistent validation, which we designate as \textit{Present-to-Past} validation, is the opposite of \textit{Past-to-Present} with regards to the starting position of the training set. Again the validation is fixed the most recent samples, but now the training set starts with the temporally closest samples to the validation set. At each iteration the training set is extended, this time with older samples and scored against the validation, until all samples are used.

\textit{Past-to-Present} and \textit{Present-to-Past} validations both require two parameters, specifically the size of the validation set, and how the increments to the training set are made. For our evaluation, we use the 20\% most recent samples as validation, and split the remaining 80\% into 10 folds, hence the validation is done 10 times, with each iteration increasing the training size by one fold.

These two validation methodologies give us the ability to account for temporal consistency. Moreover, they enable us to compare the importance of older \textit{vs} newer samples to classify recent samples.

We designate the third and last temporal consistent validation as \textit{Temporal Window}. This validation methodology is inspired on regular cross-validation, in the sense it splits the dataset into folds, but changes how the folds are used. Specifically it takes $n$  temporal consistent and followed folds (i.e.\ each fold immediately precedes the next one), using the last fold (more recent samples) as validation and the previous folds as training (older samples). By starting with the $n$ first folds and sliding one fold on each iteration, we apply a sliding window of size $n$ over the dataset.

For this last validation methodology, we again split the dataset into 10 folds. The sliding window size, $n$, is chosen during the results phase, as its choice depends on previous results.

The different scenarios will be compared by plotting their \glsreset{roc}\gls{roc} curve and measuring the \glsreset{auroc}\gls{auroc}. The \gls{roc} plots the True Positive Rate (or \gls{dr}) against the \gls{fpr} at different thresholds, which are the most relevant metrics when dealing malware detection.

We measure the \gls{auroc} during each validation's iteration and use the average measurement to discuss the results.

\subsection{Results}

We implement our experiments in Python, by using Jupyter Interactive Notebooks\cite{tool:jupyter} to facilitate data visualization. We use scikit-learn\cite{tool:sklearn} for \gls{ml}, and Pandas\cite{tool:pandas} for data analysis. Our experiments were conducted on an Ubuntu Virtual Machine with 16 cores and 16GB of RAM, in order to minimize training and validation times.

We now focus on applying the evaluation methodologies to our scenarios. This enables us to compare the different conditions, and consequently results, that affect malware detection.

We start with what we determine as \textit{laboratory conditions}, ideal conditions for the problem of malware detection. These are met when we apply the strict metrics $\mathcal{M}_{strict}^V$, to the dataset $\mathcal{D}_{class}^*$, obtaining scenario $\mathcal{S}_{strict}$.

Under these conditions, the model provides the best results, with an \gls{auroc} of 0.97, as shown in Figure \ref{fig:roc_xval}. We argue that such high values are easily attained from factors like a small and reliable dataset, and the use of cross-validation, which mixes samples and ignores possible dependencies between.

To understand how reliability influences the model's result, we use scenarios $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, which are less reliable and include more samples.

Under these more relaxed, \textit{real-world} conditions, the model's results unsurprisingly decrease to an \gls{auroc} of 0.95 for $\mathcal{S}_{loose}$ and 0.77 for $\mathcal{S}_{real}$, as shown in Figure \ref{fig:roc_xval}.

From $\mathcal{S}_{strict}$ to $\mathcal{S}_{loose}$, the only change is the amount of malware labeled samples, which significantly increase. The difference is interesting, as although the number of malware labeled samples increase significantly, the results are not that affected. This suggests that although the reliability for malware decreases, its impact is not as noticeable as expected. This might also suggest that vendors do converge on their definition of malware, under our $\mathcal{M}_{loose}$ metric.

When looking at the changes from $\mathcal{S}_{loose}$ to $\mathcal{S}_{real}$, not only the amount of malware labeled samples increase, but also the number of goodware labeled samples, both by a significant amount. The way this impacts the results is pretty significant, as we observe a high decrease in the \gls{auroc}, from 0.95 to 0.77. The metric $\mathcal{M}_{real}$ that labels malware and goodware for this scenario $\mathcal{S}_{real}$ disregards the cross-check from outside repositories, which in turn degrade the reliability significantly, as well as increase the dataset size notably. We attribute the result's degradation mainly to the unreliability of goodware labeling, not only because we have previously seen that increase in malware does not significantly impact results, but also due to the tendency for false negatives in vendors, which in turn lead us to incorrectly label goodware under samples in $\mathcal{D}_{class}^*$.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{roc_xval}
	\caption{Cross-Validation \gls{roc} and \gls{auroc} for our three scenarios.}
	\label{fig:roc_xval}
\end{figure}

The results we described show how moving from \textit{laboratory conditions} to more \textit{real-world conditions} degrade the model's performance. We now focus on using our previously defined temporal consistency methodologies to further converge into a real-world scenario.

We start by applying our \textit{Past-to-Present} validation to the three scenarios, $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$. As previously defined, this validation starts with an older set of training samples and iteratively adds newer samples, validating each iteration on a fixed set of the most recent samples. Since our interest is to measure performance variance over time, we plot in Figure \ref{fig:roc_tc_1} the \gls{auroc} at every iteration (i.e.\ fold), for each of out three scenarios.

When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note a decrease from 0.97 to 0.94 for $\mathcal{S}_{strict}$, 0.95 to 0.93 for $\mathcal{S}_{loose}$, and 0.77 to 0.72 for $\mathcal{S}_{real}$. This decrease is intuitive to the methodology, as we are forcing temporal consistency between samples.

When looking at results from $\mathcal{S}_{strict}$ and $\mathcal{S}_{loose}$, we see they closely relate. This relation has already been noticed on the previous cross-validation results, as the their metrics $\mathcal{M}_{strict}$ and $\mathcal{M}_{loose}$ are not very different. As for $\mathcal{S}_{real}$, the degradation is higher, as the reliability of the metric $\mathcal{M}_{real}$ goes down.

Our main observation for this validation methodology is how there is a slight tendency for \gls{auroc} to increase, as we move forward in time, close to the validation set.

When looking at $\mathcal{S}_{strict}$, we note that from the initial fold 0 to fold 3, the \gls{auroc} increases by 0.02, without going significantly further down, from that fold forward. This observation can be seen in Figure \ref{fig:roc_tc_strict_1}. Similarly, looking at $\mathcal{S}_{loose}$, we note it holds the \gls{auroc} up until fold 7, after which it there is a slight increase, although not very noticeable, Figure \ref{fig:roc_tc_loose_1} shows more clearly the change in scores. As for $\mathcal{S}_{real}$, we see a fluctuation down from the initial folds, followed by a steady increase until the last fold. Figure \ref{fig:roc_tc_real_1} shows the actual fluctuation values.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_all_1}
	\caption{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation. Folds order consists with temporal order (i.e.\ fold 0 contains older samples than fold 1)}
	\label{fig:roc_tc_1}
\end{figure}

From these observations, we argue about the possibility that with fixed validation set of the most recent samples, a model benefits by using samples temporally closer to validation. Our next result, which uses our \textit{Present-to-Past} validation methodology will further help analyze the aforementioned detail. The \textit{Present-to-Past} validation enhance the previous results under real-world conditions. This methodology starts by fixing the validation set to the most recent samples, but with the training set starting at the temporally closest samples to validation. At each iteration, older samples are added to the training set and validated on the fixed, most recent, samples.

By applying this methodology to the three scenarios, $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, we plot Figure \ref{fig:roc_tc_2}, where the X axis increases as older samples are added to the training set (e.g.\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time. Similarly to the previous observation, the average \gls{auroc} suffers a decrease when compared to cross-validation. For $\mathcal{S}_{strict}$ we note a change from 0.97 to 0.94, for $\mathcal{S}_{loose}$ the decrease is from 0.95 to 0.94, and for $\mathcal{S}_{real}$ from 0.77 to 0.75.

%From these observations, we argue about the possibility that with fixed validation set of the most recent samples, a model benefits by using samples temporally closer to validation. Our next result, which uses our \textit{Present-to-Past} validation methodology will further help analyze the aforementioned detail.

%The \textit{Present-to-Past} validation enhance the previous results under real-world conditions. This methodology starts by fixing the validation set to the most recent samples, but with the training set starting at the temporally closest samples to validation. At each iteration, older samples are added to the training set and validated on the fixed, most recent samples.

%By applying this methodology to the three scenarios, $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, we plot Figure \ref{fig:roc_tc_2}, where the X axis increases as older samples are added to the training set (e.g.\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time.

%Similarly to the previous observation, the average \gls{auroc} suffers a decrease when compared to cross-validation. For $\mathcal{S}_{strict}$ we note a change from 0.97 to 0.XX, for $\mathcal{S}_{loose}$ the decrease is from 0.95 to 0.XX, and for $\mathcal{S}_{real}$ from 0.77 to 0.XX.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{roc_tc_all_2}
	\caption{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation. Folds order is the inverse of temporal order (i.e.\ fold 0 contains newer samples than fold 1)}
	\label{fig:roc_tc_2}
\end{figure}

The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{strict}$ and $\mathcal{S}_{loose}$ are display very similar results, with $\mathcal{S}_{real}$ dropping behind due to its less reliable labeling metric.

Compared to the \textit{Past-to-Present} validation, the average \gls{auroc} increases slightly. Furthermore, we note that the score maxes out at early folds, specifically between fold 0 to fold 4.

For scenario $\mathcal{S}_{strict}$, there is a slight increase of 0.02 between fold 0 and fold 4, maximum out from there on. For $\mathcal{S}_{loose}$ there is barely any change throughout folds. In scenario $\mathcal{S}_{real}$ the increase in early folds is very small, but what we note is that the older samples weight down the score, with maximum values between fold 1 and fold 4, from fold 5 and onward, the score decreases significantly. The concrete values for each fold for $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$ can be observed in Figure \ref{fig:roc_tc_strict_2}, \ref{fig:roc_tc_loose_2}, \ref{fig:roc_tc_real_2}, respectively.

%With these results, our observation that samples closer to the validation set benefit the model becomes more convincing. In fact, we argue that there should be an ideal number of necessary training folds, temporally consistent with the validation fold (i.e.\ any fold from training predates validation), needed to maximum the overall score.

%Our final results test if indeed we can reduce the training size to some number of folds $n$, which predate the validation fold. We choose $n = 3$, for we have seen the scores either do not improve (for $\mathcal{S}_{strict}$ and $\mathcal{S}_{loose}$) or actually go down (for $\mathcal{S}_{real}$) with higher folds, and apply the final \textit{Temporal Window} validation, which consists in a sliding window of size 4 (3 training folds, 1 validation) to the scenarios $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$.

%As presented in Figure \ref{fig:roc_sliding_window}, for the scenarios $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, we obtain \gls{auroc} values of 0.95, 0.91 and 0.74, respectively.

With these results, our observation that samples closer to the validation set benefit the model becomes more convincing. In fact, we argue that there should be an ideal number of necessary training folds, temporally consistent with the validation fold (i.e.\ any fold from training predates validation), needed to maximize the overall score. For a validation set composed by 10\% of the total dataset, the preceding 30\% should stand for training.

Finally, we analyse how does such reduced training set behaves in our scenarios; for this purpose, we define a sliding window that moves forward in time through each scenario for training and validation. We propose a reduction on the training size to $n=3$ folds predating the validation fold. We chose $n = 3$, since we have seen the scores either do not improve (for $\mathcal{S}_{strict}$ and $\mathcal{S}_{loose}$) or actually go down (for $\mathcal{S}_{real}$) with higher folds. In summary, we have selected 30\% of each dataset (read, scenario) for training purposes and the next 10\% for validation (3 training folds, 1 validation fold), then started moving the window forward in time (1 fold at a time) and obtained the results: as presented in Figure \ref{fig:roc_sliding_window}, for the scenarios $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, we obtain \gls{auroc} values of 0.95, 0.91 and 0.74, respectively. These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{roc_sliding_window}
	\caption{\gls{roc} and \gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
	\label{fig:roc_sliding_window}
\end{figure}

Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from 0.97 to 0.95 for $\mathcal{S}_{strict}$, from 0.95 to 0.91 for $\mathcal{S}_{loose}$ and from 0.77 to 0.74 for $\mathcal{S}_{real}$. An expected result, as we are enforcing temporal validation. We should highlight that these results should be much closer to reality, since we are requiring temporal consistency and also a reasonable amount of data for training purposes (which might be a relevant issue in a few year's time).

%%%%

%Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from 0.97 to 0.95 for $\mathcal{S}_{strict}$, from 0.95 to 0.91 for $\mathcal{S}_{loose}$ and from 0.77 to 0.74 for $\mathcal{S}_{real}$. An expected result, as we are enforcing temporal validation.

%When doing the comparison to the average \gls{auroc} from \textit{Past-to-Present} and \textit{Present-to-Past}, the \textit{Temporal Window} averages either hold similar scores or are actually better, specifically when compared to \textit{Past-to-Present} validation.

%These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score.

%\todo[inline]{tabela de todos os resultados finais}

%\todo[inline]{modelo do strict, validado no real vs modelo real validado no real}

% We start with baseline results, that take all three scenarios, $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$

%\todo[inline]{Cross-validation para os 3 datasets}
%\todo[inline]{Referir como o resultado é optimo para strict}

%ROC/AUROC c/ x-validation, p/ 3 datasets, discussion (0.96; optimo em strict dataset, validar depois com loose e real)

% ROC/AUROC c/ tc validation, p/ 3 datasets, discussion

%discussion entre c/ e s/ TC, comprova perda, mais p/ real, pq é menos reliable

%passado mais recente mais relevante, verificar isso com sliding window
%
%ROC/AUROC c/ tc sliding window, p/ 3 datasets, discussion - possibilidade de descardar dados sem perder resultados.
%
%discussion - resumo dos resultados anteriores
%
%future work - usar nomes do malware, usar mais features, maximizar resultadosb
%
%\todo[inline]{}

%\subsection{Classifier Implementation}\label{sub_sec:classifier_implementation}

%In this subsection we go over the classifier used to create the model that separates malware from goodware. We chose \textit{Logistic Regression} linear classifier for this task, as it deals well with high number of samples and features, but also enables a better understand on what the model is learning, given it assigns weights to each feature.

%\todo[inline]{elaborar isto, ate q ponto?}

%\subsection{Evaluation and Results}\label{sub_sec:results}

%We now put together all the previous subsections to describe evaluation and results. This subsection is hence divided into evaluation, where we describe how our results are measured and why they are measured that way, and results, where we discuss the outcome of our work.

%Our evaluation is based on measuring the \gls{dr} and \gls{fpr}. This metric allows us to directly compare our model to vendors' results, as well as plotting a \gls{roc} curve, which measures the \gls{dr} and \gls{fpr} ate different thresholds. To obtain the overall score we measure the \gls{auroc} curve, that varies between 0 and 1, indicating 0.0 \gls{dr} and 1.0 \gls{fpr} and 1.0 \gls{dr} and 0.0 \gls{fpr}, respectively.

%We evaluated our work by creating different datasets based on the available samples. By using different datasets we compare how the balance of malware/goodware influences the final results. We cross-validated each dataset using k-fold (with $k=10$) and plot its \gls{roc} curve. Furthermore, we apply a validation where we maintain temporal consistency of samples, meaning the training set always predates the validation set. Comparing the curves with and without temporal consistency gives us a better approximation on how models behave under laboratory \textit{vs} real world conditions.

% The used datasets were created by applying the different metrics described in \ref{sub_sec:data_labeling}. In practice we used 3 datasets:

% \subsubsection{\gls{tts}} This dataset is the most fine-grained, and smallest. The malware samples are those where every top vendor classifies as malware, are in the VirusShare set and are not in the \gls{nsrl} set. As for goodware samples, these must all be classified as clean, are in the \gls{nsrl} and are not in the VirusShare set. Numbers wise, this dataset contains 1,361 malware samples and 4,000 goodware samples.

% \subsubsection{\gls{tgs}} This dataset is less strict than the previous, differing only by applying the minimum threshold necessary to label a sample as malware (at least 5 malware classifications). In terms of size, this dataset contains 86,503 malware samples and 4,000 goodware samples.

% \subsubsection{\gls{tbs}} This dataset ignores the VirusShare and \gls{nsrl} sets, meaning a sample is labeled as malware if it has at least 5 top vendors classifying it as malware, and labels a sample as goodware if all top vendors classify it as clean. Since it is less strict dataset, it contains 186,444 malware samples and 49,783 goodware samples.

% \todo[inline]{talvez criar dataset 4 e 5 de fixar malware bom vs fixar goodware bom}

%For temporal consistency validation we sorted the dataset by date and removed the last 20\% of samples for validation. We split the remaining 80\% into 10 temporal consistent folds (i.e.\ samples from fold 0 predate samples from fold 1), we then measure the \gls{roc} curve after applying the algorithms presented in \textbf{Algorithms X}, which provide an incremental validation approach.

%\todo[inline]{algoritmo com old to new, começar do fold 0, ir juntando folds novos}
%\todo[inline]{algoritmo com new to old, começar do fold 9, ir juntando folds velhos}

%This validation provides insight on the dependency between newer and older samples, enabling to understand how recent (and how large) the training set needs to be in order to maximize the validation score.

%Following the previous temporal consistency validation, we apply another approach to temporal consistency. Specifically we split the dataset again into 10 temporal consistent folds, and apply a sliding window over the folds, maintaining the training set older than the validation set.

%\todo[inline]{se calhar separar resultados numa subseccao propria}
%\todo[inline]{onde introduzir arquitectura, sklearn e etc}

%We now present the results of our evaluation, starting by applying the cross-validation for all the datasets, serving as a baseline for \textit{laboratory conditions}.

%For \gls{tts} we obtain an average \gls{auroc} of 0.96, going down to 0.94 for \gls{tgs} and even further down to 0.77 for \gls{tbs}, as seen in \textbf{Figure X}.

%For each dataset we used the exact same methodology to derive a model, varying only the amount of samples. This size increase between datasets, and consequently score decrease, is due to relaxing the labeling methodology, meaning less accurate ground truth which in turn includes samples that are harder for the model to separate.

%Our experiments showed how a reliable model can be obtained when the dataset is built in a way that facilitates separation, even without much work in feature selection and classifier optimization. The limitation is how a good dataset does not represent real world samples, which in turn are much harder to separate, as seen in the results for \gls{tbs}.

%To better approximate \textit{real-world conditions}, our next experiments go over temporal consistency when validating the model. We applied the algorithms described in \textbf{Algorithms X} to the 3 datasets.

%When applying \textbf{Algorithm x.1}, we observed a \gls{auroc} variation from 0.92 to 0.93 for \gls{tts} and \gls{tgs}, where \gls{tbs} varies between 0.69 to 0.74, as seen in \textbf{Figure X.1}.

%Again, the same methodology was applied to all datasets, but by incrementing the training set size with newer samples on every fold, we noticed the tendency is for the \gls{auroc} to increase. We only note that for the first two folds in \gls{tbs} the score is maximum, although it goes down with the third fold and follows the same incrementing pattern as the other sets.

%When applying \textbf{Algorithm x.2}, we observed a \gls{auroc} variation from 0.92 to 0.93 for \glossary{tts} and \gls{tgs}, identical to the previous experiment, but for \gls{tbs} the variation was between 0.72 and 0.76, as observed in \textbf{Figure X.2}.

%With the same methodology for all datasets, but now the training set using older samples as it grows in size, we note that increasing the training with older samples does not improve the score for \gls{tts} and \gls{tgs}, even more, for \gls{tbs} we note the score decreases as we use older samples.

%Comparing results from both temporal consistent validation algorithms, we are led to believe they describe inverse patterns. This fact reinforces the intuition that newer samples are more related to the a close past, rather than a distant one.

%By identifying that with the 3 closest folds to the validation we achieve a maximum or near maximum score, we use this knowledge to apply the last experiment on temporal consistency. We apply the sliding window with 3 folds for training and the forth for validation, incrementing one fold on each iteration.

%By looking at the \gls{auroc}, we note that for \gls{tts} we got an average score of 0.94, for \gls{tgs} the average was 0.88 and for \gls{tbs} 0.74, as depicted in \textbf{Figure X}. For all but \gls{tgs} sets we got scores that are up to par with the scores when using a larger training size. We can argue that using smaller training sets, but close in time to the validation is more beneficial than using a larger dataset that includes samples further away in time.

%To summarize the results, we showed that some conditions to optimize a malware classification model's score exist (i.e.\ laboratory conditions), such as choosing samples that are further apart, by enforcing a stricter labeling method, and using validations techniques that are ideal for independent and identically distributed samples, like k-fold cross-validation.

%When labeling conditions are less strict, the overall scores decrease, as shown by the three different datasets; furthermore, when a time-wise dependency between samples is taken into account, the overall score is negatively impacted, when compared to a normal cross-validation. Nonetheless we showed that it is possible to obtain good results on temporal consistency, without having to take into account the whole past, only the recent past.

### 2018-08-12 extended_abstract/evaluation.tex
latest

%!TEX root = main.tex
\color{black}
\section{Evaluation and Results}\label{sec:eval_results}

In this section we aim at doing a comparative analysis on three different scenarios $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, built on top of the previously defined metrics. This comparison is done using standard cross-validation methodologies and a proposed temporal-based methodology.

We further provide an analysis on how to reduce the size of the training set, without compromising the final results.

\subsection{Evaluation}

With regards to our evaluation methodology, as we have previously mentioned, our purpose is to understand how laboratory conditions compare to real-world conditions.
We now detail how we achieve and compare these conditions.

Given the purpose of our work, we choose to measure our results by plotting an \gls{auroc} graph, which measures the \tpr~at different \fpr~levels, metrics that are commonly used across similar work~\cite{miller:rev_int,nissim:al_pdf,schultz:data_mining}.

The three scenarios that we will focus on will rely on metrics $\Mrealv$, $\Mloosev$ and $\Mstrictv$, over the dataset $\DD$:

\begin{itemize}
	\item Real Scenario $\mathcal{S}_{\real}$, applies the metric $\Mrealv$, containing 98,582 malware samples and 56,475 goodware samples.
	\item Loose Scenario $\mathcal{S}_{\loose}$, applies the metric $\Mloosev$, containing 45,306 malware samples and 1,989 goodware samples.
	\item Strict Scenario $\mathcal{S}_{\strict}$, applies the metric $\Mstrictv$, containing 24,658 malware samples and 1,989 goodware samples.
\end{itemize}

Given these three scenarios, we consider the following evaluation metrics:

\paragraph{Cross-validation (Figure \ref{fig:dia_xvalidation})}
To gain insight on how each model generalizes our scenarios, we apply a k-fold cross-validation, with $k=10$. This methodology splits the dataset into $k$ subsets (\ie, folds), selecting a single fold for validation and the remaining $k-1$ folds for training. This process is repeated $k$ times, ensuring every fold is used for validation and training.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dia_xvalidation}
	\caption{Cross-Validation evaluation example with 10 folds.}
	\label{fig:dia_xvalidation}
\end{figure}

Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples. Since we want to measure the score when training samples pre-date the validation samples, we now define a couple of temporal based validations. These are validated on the best performing model from cross-validation for all 3 scenarios.

\paragraph{Temporal based validation}
The first temporal based validation, which we designate as \textit{Past-to-Present validation}, Figure~\ref{fig:dia_pastpresent}, can be resumed as an iterative methodology where the validation set is fixed with the most recent samples, and the training set with the oldest. At each iteration the training set is extended with more recent samples and scored against the validation, until all samples are used.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dia_pastpresent}
	\caption{Past-to-Present evaluation example with 20/80 test/training, 10 folds in training.}
	\label{fig:dia_pastpresent}
\end{figure}

The second temporal based validation, which we designate as \textit{Present-to-Past validation}, Figure~\ref{fig:dia_presentpast}, is the opposite of \textit{Past-to-Present} with regards to the starting position of the training set. Again the validation is fixed the most recent samples, but now the training set starts with the temporally closest samples to the validation set. At each iteration the training set is extended, this time with older samples and scored against the validation, until all samples are used.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dia_presentpast}
	\caption{Present-to-Past evaluation example with 20/80 test/training, 10 folds in training.}
	\label{fig:dia_presentpast}
\end{figure}

\textit{Past-to-Present} and \textit{Present-to-Past} validations both require two parameters, specifically the size of the validation set, and how the increments to the training set are made. For our evaluation, we use the 20\% most recent samples as validation, and split the remaining 80\% into 10 folds, hence the validation is done 10 times, with each iteration increasing the training size by one fold.

These two validation methodologies give us the ability to account for temporal consistency. Moreover, they enable us to compare the importance of older \textit{vs} newer samples to classify recent samples.

We designate the third and last temporal based validation as \textit{Temporal Window validation}, Figure~\ref{fig:dia_slidingwindow}. This validation methodology is inspired on regular cross-validation, in the sense that it splits the dataset into folds, but changes how the folds are used. Specifically it takes $n$  temporal consistent and contiguous folds, \ie, each fold immediately precedes the next one, and uses the last fold (more recent samples) for validation, and the previous folds for training (older samples). By starting with the $n$ first folds and sliding one fold on each iteration, we apply a sliding window of size $n$ over the dataset.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{dia_slidingwindow}
	\caption{Temporal window evaluation example with window of size 3 over a 10 fold dataset.}
	\label{fig:dia_slidingwindow}
\end{figure}

For this last validation methodology, we again split the dataset into 10 folds. The sliding window size, $n$, is chosen during the results phase, as its choice depends on previous results.

We measure the \gls{auroc} during each validation's iteration and use the average measurement to discuss the results.

\subsection{Results}
\label{section:single_layer_results}

We implement our experiments in Python, by using Jupyter Interactive Notebooks\cite{tool:jupyter} to facilitate data visualization. We use scikit-learn\cite{tool:sklearn} for \gls{ml}, and Pandas\cite{tool:pandas} for data analysis. Our experiments were conducted on an Ubuntu Virtual Machine with 16 cores and 16GB of RAM, in order to minimize training and validation times.

We now focus on applying the evaluation methodologies to our scenarios. This enables us to compare the different conditions, and consequently results, that affect malware detection.

%We compare each model $\mathcal{LR}$ and $\mathcal{E}$ under each scenario $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.

We start with what we determine as \textit{laboratory conditions}, ideal conditions for the problem of malware detection. These are met when we apply the strict metrics $\Mstrictv$, to the dataset $\mathcal{C}$, obtaining scenario $\mathcal{S}_{\strict}$.

Under these conditions, our model provides the best results, with an \gls{auroc} of 0.91, as shown by the red curve in Figure \ref{fig:xval_results}. We argue that such high values are easily attained from factors like a small and reliable dataset, and the use of cross-validation, which mixes samples and ignores possible dependencies on malware samples.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{xval_results}
	\caption{Cross-Validation \gls{roc} and \gls{auroc} for our model under $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.}
	\label{fig:xval_results}
\end{figure}

To understand how reliability influences the models' result, we use scenarios $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, which are less reliable and include more samples.

Under these more relaxed, \textit{real-world} conditions, the model's results hold an \gls{auroc} of 0.90 under $\mathcal{S}_{\loose}$, as shown by the blue curve in Figure \ref{fig:xval_results}, and an \gls{auroc} of 0.75 under $\mathcal{S}_{\real}$, as seen by the green curve in Figure \ref{fig:xval_results}.

From $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$, the only change is the amount of malware labeled samples, which significantly increase.
The difference is interesting, as although the number of malware labeled samples increase significantly, the results are not that affected.
This suggests that although the reliability for malware decreases, its impact is not as noticeable as expected.
This might also suggest that vendors do converge on their definition of malware, under our $\Mloosev$ metric.
If vendors did not converge on what is malware, adding more samples would culminate in worse results, as separation between malware and goodware would become harder.

When looking at the changes from $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$, not only the amount of malware labeled samples increase, but also the number of goodware labeled samples, both by a significant amount. 
The way this impacts the results is pretty significant, as we observe a high decrease in the \gls{auroc}.
The metric $\Mrealv$ that labels malware and goodware for this scenario $\mathcal{S}_{\real}$ disregards the cross-check from outside repositories, which in turn degrade the reliability significantly, as well as increase the dataset size notably.
We attribute the results' degradation mainly to the unreliability of goodware labeling, not only because we have previously seen that increase in malware does not significantly impact results (from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$), but also due to the tendency for false negatives in vendors (Figure \ref{fig:distribution_changes}), which in turn lead us to incorrectly label goodware for the samples in $\mathcal{C}$.

The results we described show how moving from \textit{laboratory conditions} to more \textit{real-world conditions} degrade the model's performance. We now focus on using our previously defined temporal based methodologies to further converge into a real-world scenario.

\medskip

We start by applying our \textit{Past-to-Present} validation to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
As previously defined, this validation starts with an older set of training samples and iteratively adds newer samples, validating each iteration on a fixed set of the most recent samples.
Since our interest is to measure performance variation over time, we plot in Figure~\ref{fig:pastpresent} the \gls{auroc} at every iteration (\ie, fold), for each of our three scenarios.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{pastpresent}
	\caption{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation. Folds order consistent with temporal order (\ie, fold 0 contains older samples than fold 1)}
	\label{fig:pastpresent}
\end{figure}

When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 0.75 to 0.67.
This decrease is intuitive to the methodology, as we are forcing temporal consistency between samples.

For both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ we note only a slight increase as new folds are added.
They still relate, as we have previously noted for cross-validation, arguably given their metrics $\Mstrictv$ and $\Mloosev$ are not very different.
The small variation to the cross-validation methodology can be justified by using small dataset size for both cases.

As for $\mathcal{S}_{\real}$ we note higher variation and lower overall score, as the reliability of the metric $\Mrealv$ goes down.
This is expected, not only because we are enforcing temporal consistency between samples, but as new folds are added, the training gets bigger, while the test remains the same.

Our main observation for this validation methodology is that there is a slight tendency for \gls{auroc} to increase, as we move forward in time, close to the validation set.

From these observations, we argue about the possibility that with fixed validation set of the most recent samples, a model benefits by using samples temporally closer to validation.

\medskip

Our next result, which uses our \textit{Present-to-Past} validation methodology will further help analyze the aforementioned detail.
The \textit{Present-to-Past} validation enhances the previous results under real-world conditions. 
This methodology starts by fixing the validation set to the most recent samples, but with the training set starting at the temporally closest samples to validation.
At each iteration, older samples are added to the training set and validated on the fixed, most recent, samples.

By applying this methodology to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we plot Figure \ref{fig:presentpast}, where the X axis increases as older samples are added to the training set (\eg\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time. Similarly to the previous observation, the average \gls{auroc} suffers a decrease when compared to cross-validation.
For $\mathcal{S}_{\strict}$ we note a change from 0.91 to 0.90, for $\mathcal{S}_{\loose}$ the score is the same, and for $\mathcal{S}_{\real}$ 0.75 to 0.69.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{presentpast}
	\caption{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation. Folds order is the inverse of temporal order (\ie, fold 0 contains newer samples than fold 1)}
	\label{fig:presentpast}
\end{figure}

The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ display very similar results, with $\mathcal{S}_{\real}$ dropping behind due to its less reliable labeling metric.
It is noticeable that using the entire dataset does not bring much improvement to the final results. In fact, for $\mathcal{S}_{\real}$ the score even drops after fold $\#2$.

With these results, our original observation that samples closer to the validation set benefit the model becomes more convincing. In fact, we argue that there should be an ideal number of necessary training folds, temporally consistent with the validation fold (\ie\ any fold from training predates validation), needed to maximize the overall score.

\medskip

Finally, we analyze how does such reduced training set behaves in our scenarios; for this purpose, we define a sliding window that moves forward in time through each scenario for training and validation.
We propose a reduction on the training size to $n=3$ folds predating the validation fold.
We choose $n = 3$, since we have seen that the scores either do not improve (for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$) or actually go down (for $\mathcal{S}_{\real}$) with higher folds.
In summary, we have selected 30\% of each dataset for training purposes and the next 10\% for validation (3 training folds, 1 validation fold), and then started moving the window forward in time (1 fold at a time) to obtain the following results (Figure~\ref{fig:slidingwindow}): for  $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we obtain \gls{auroc} values of 0.89, 0.88 and 0.76, respectively. 
These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score.

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{slidingwindow}
	\caption{\gls{roc} and \gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
	\label{fig:slidingwindow}
\end{figure}

Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from 0.91 to 0.89 for $\mathcal{S}_{\strict}$, from 0.90 to 0.88 for $\mathcal{S}_{\loose}$ and from 0.75 to 0.73 for $\mathcal{S}_{\real}$.
We should highlight that the results that use temporal consistency should better reflect the reality than standard cross-validation, since we are requiring temporally ordered samples.
Another important idea that should be stressed is that for cross validation we used a fairly reasonable amount of data for training purposes, whereas in this last case we used a restricted amount of data. This might be a relevant issue in a few year's time. The results obtained are summarized in Table \ref{tab:singlelayer_results}.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{l|cccc}
		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
		\hline
		Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
		Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
		Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
		Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
		\hline
	\end{tabular}
	\smallskip
	\caption{Single layer results summary.}
	\label{tab:singlelayer_results}
\end{table}

\medskip

With a better understanding of how the model behaves under different methodologies, we now diverge to how we improved not only the overall results, but also the information provided by the model.

### 2018-08-12 extended_abstract/features-old.tex
latest

%!TEX root = main.tex
{\color{red}

\section{Feature and Model Selection}\label{sec:feature_model}

In this subsection we describe our approach to feature selection and linear model choice. Since the main focus of our work is to compare results relatively and not absolutely, we do not emphasize much on this topic.

\subsection{Feature Selection}

Following the work of \cite{miller:rev_int,schultz:data_mining} features based on static imports have shown very promising results in \gls{ml} applications for malware detection.

For the purpose of this work we use static imports as features. Given that Cuckoo sometimes fails in parsing static imports, some reports contain noise which should be removed. To do so we apply a regular expression to filter invalid ASCII and special characters (e.g.\ \texttt{!}, \texttt{\&}, \texttt{*}). Doing so removes imports like \texttt{*invalid*} and \texttt{MSVCRT.dll\textbackslash x90} (where \text{\textbackslash x90} represents a byte with value 90 in hexadecimal).

We select static imports from $\mathcal{D}_{static}^*$, collecting a total number of 9,698 imports, applying the previous criterion we remove 40, obtaining a final number of 9,658 imports.

These discriminate case sensitivity (i.e.\ \texttt{kernel32.dll} is different from \texttt{KERNEL32.dll}), which increases the number of features, but provides more information for the classifier to separate classes. From the 9,658 imports, 8,861 (91.75\%) end with the \texttt{dll} extension, while the other 797 (8.25\%) used different extensions (e.g.\ \texttt{bpl}, \texttt{exe}, \texttt{sys}).

On the topic of features, we close the subsection by describing how imports are converted to a vector of features. We vectorize imports by creating a binary vector where each position corresponds to a specific import. If a given import $i$ is present in a sample, its feature vector $x$ will have the value at that position $x_i$ set to 1. Likewise, if a given import $j$ is not present in a sample, its feature vector $x$ will have the value at that position $x_j$ set to 0.

\subsection{Model Selection}

In this subsection we go over the classifier used to create the model that separates malware from goodware. Our main concerns when choosing a classifier regard the ability to produce a probabilistic output, good scaling for large number of features and samples, and ease of use.

With this in mind, we choose the linear logistic regression model. This model gives the probability of a random variable $X$, being 0 or 1, based on experimental data. For our problem, the random variable $X$ is an unknown sample and the outcome is the probability of being either goodware or malware\cite{friedman2001elements}. By having a probabilistic output we can fine tune the threshold at which a sample is labeled malicious. Due to its popularity, it is also readily available from several libraries.}

### 2018-08-12 extended_abstract/features.tex
latest

--- 2018-05-09 thesis_report/Thesis_Static_Features.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                                                                      %
-%     File: Thesis_Static_Features.tex                                 %
-%     Tex Master: Thesis.tex                                           %
-%                                                                      %
-%     Author: João C. Godinho                                          %
-%     Last modified : Apr 2018                                         %
-%                                                                      %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%!TEX root = main.tex
+\color{black}
+\section{Feature and Model Selection}\label{sec:feature_model}
 
-\chapter{Static Features}
-\label{chapter:static_features}
+In this subsection we describe our approach to feature selection and linear model choice.
 
-In the present chapter we go into details about how we built a machine learning model to detect malware based on static features.
-We start by detailing the model selection and its evaluation, followed by the selected features.
-We finish off by providing how our model performed under various scenarios.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Model Selection and Evaluation}
-\label{section:model_selection_evaluation}
-
-In this section we go over the classifier used to create the model that separates malware from goodware, as well as our evaluation methodology.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Model Selection}
-\label{section:model_selection}
-
-Our main concerns when choosing a classifier regard the ability to produce a probabilistic output (\ie\ probability of being malware), good scaling for large number of features and samples and ease of use.
-
-Taking into consideration the guidelines given in~\cite{rossow:practices,shabtai:survey} and related work in~\cite{miller:rev_int,nissim:al_pdf,rieck:dynamic,schultz:data_mining}, we decide to use \textit{\acrfull{lr}} as our model.
-This model fits our needs as it gives the probability of a random variable $X$ being 0 or 1, given a set of constraints (\ie~features), scales well with samples and features and it is readily available from several libraries, facilitating implementation~\cite{friedman2001elements}.
-
-\gls{lr} can be defined with the form
-\begin{eqnarray*}
-	\rho(x) = \dfrac{1}{1 + e^{-x}},&x = \beta_0 + \beta_1x_1 + ... + \beta_nx_n
-\end{eqnarray*}
-where $\beta_n$ is the learned weight for feature $x_n$.
-This weight is learned through iteration in order to minimize the error between the predicted values and the actual values.
-In other words, given an \textit{n-th} dimensional set of features, \gls{lr} will try to create an hyperplane that divides samples from two classes.
-
-As \gls{lr} is based on the logistic function (or sigmoid function, visually represented in Figure \ref{fig:sigmoid_function}), each feature $x_n$ can vary from $-\infty$ to $+\infty$ and still the output is contained between 0 and 1, hence providing probabilistic values.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}
-	\begin{axis}
-	\addplot[
-		color=red,
-		domain=-5:5
-	]
-	{1/(1+exp(-x))};
-	\addlegendentry{$1/(1+e^{-x})$}
-	\end{axis}
-	\end{tikzpicture}
-	\caption{Sigmoid function $f(x)=\dfrac{1}{1+e^{-x}}$}
-	\label{fig:sigmoid_function}
-\end{figure}
-
-To implement this model we take advantage of scikit-learn~\cite{tool:sklearn} library, a Python library that provides a multitude of models and tools that facilitate the creation of \gls{ml} models.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{section:evaluation}
-
-With regards to our evaluation methodology, as we have previously mentioned, our purpose is to understand how laboratory conditions compare to real-world conditions.
-We now detail how we achieve and compare these conditions.
-
-Given the purpose of our work, we choose to measure our results by plotting an \gls{auroc} graph, which measures the \tpr~at different \fpr~levels, metrics that are commonly used across similar work~\cite{miller:rev_int,nissim:al_pdf,schultz:data_mining}.
-
-Our first evaluation methodology is to apply a common \gls{ml} validation technique named \textit{k-fold cross-validation}, with $k=10$, depicted in Figure \ref{fig:xval}.
-This method works by splitting the dataset into $k$ equally sized parts (folds) and then trains using $k-1$ folds, using the remaining fold as validation.
-This process is repeated $k$ times, guaranteeing that every fold is used at least once for validation.
-This methodology is useful to prevent over-fitting, as the validation fold is never used together with the training folds.
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_xvalidation.png}
-	\caption[Cross-validation evaluation example.]{Cross-validation evaluation example with $k=10$ folds.}
-	\label{fig:xval}
-\end{figure}
-
-This evaluation methodology is our baseline, hence we define it as our \textit{laboratory} conditions, where no time based consistency is taken into account.
-
-\medskip
-
-Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples.
-Since we want to measure the when training samples predate validation, we now define a couple of temporal based validations, based on the concept of cross-validation:
-
-\paragraph{\textit{Past-to-Present}:} This validation, depicted in Figure \ref{fig:past_present}, can be resumed as an iterative methodology where the validation set is a fixed percentage of dataset's most recent samples, whereas the training set is a fixed percentage of the dataset's oldest.
-The training set is split into $k$ temporal consistent folds and at each iteration it is extended with more recent folds and scored against the training set, until all folds are used.
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_pastpresent.png}
-	\caption[Past-to-Present evaluation example.]{Past-to-Present evaluation example with 20\% test, 80\% training, $k=10$ folds in training.}
-	\label{fig:past_present}
-\end{figure}
-
-\paragraph{\textit{Present-to-Past}:} This validation, depicted in Figure \ref{fig:present_past}, is the opposite of \textit{Past-to-Present} with regards to the starting point of the training set.
-Again the validation is fixed at the most recent samples, but the training set starts with the temporally closest fold to the validation set.
-At each iteration the training set is extended, this time with older folds, and scored against the validation, until all folds are used.
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_presentpast.png}
-	\caption[Present-to-Past evaluation example.]{Present-to-Past evaluation example with 20\% test, 80\% training, $k=10$ folds in training.}
-	\label{fig:present_past}
-\end{figure}
-
-\medskip
-Both \textit{Past-to-Present} and \textit{Present-to-Past} validations require two parameters: the size of the validation set and how the increments to the training set are made.
-For the purpose of our work, we use the 20\% most recent samples for validation, and split the remaining 80\% into 10 folds, hence the validation is done 10 times, with each iteration increasing the training set size by one fold.
-
-These two validation methodologies give us the ability to account for temporal consistency.
-Moreover, they enable us to compare the importance of older \textit{vs.}\ newer samples to classify recent samples.
-
-\paragraph{\textit{Temporal Window}:} This validation, depicted in Figure \ref{fig:sliding_window}, also splits the dataset into $k$ folds, but takes $w$ temporal consistent and contiguous folds (\ie each fold immediately precedes the next one), using the first $w-1$ folds for training (older samples) and the remaining fold (more recent samples) for validation.
-By starting with the $w$ first folds of the dataset and sliding one fold on each iteration, we apply a sliding window of size $w$ over the dataset.
-For this last validation methodology, we again split the dataset into 10 folds.
-The sliding window size $w$, is chosen during the results phase, as its choice depends on previous results.
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
-	\caption[\textit{Temporal Window} evaluation example.]{\textit{Temporal Window} evaluation example with $w=3$ window size over $k=10$ folds.}
-	\label{fig:sliding_window}
-\end{figure}
-
-\medskip
-
-As we have previously described in Section \ref{section:dataset_creation}, we possess 3 different datasets, created from 3 different scenarios.
-With that in mind, all our previously described methodologies: cross-validation, past-to-present and present-to-past; are applied to each of the 3 datasets.
-
-Applying the aforementioned methodologies to our 3 datasets enables us to study not only how temporal consistency impacts the results, but also the impact of \textit{ground-truth}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Feature Selection}
+\subsection{Feature Selection}
 \label{section:feature_selection}
@@
 With that in mind, we chose to use the static imports as features.
-
-Static imports are present in the import table, referencing libraries or other executables that are not part of the current binary.
-Their usage is key, as they provide access to all kind of system level operations (\eg\ file manipulation, network access).
-Given their purpose, our expectation is that a sample's maliciousness can be inferred based on the static imports.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Single Layer Results}
-\label{section:single_layer_results}
+\subsection{Model Selection}
 
-Having met all the necessary conditions to build a \gls{ml} model: model selection and evaluation and feature selection, we are now ready to provide the results of our first experiments.
+In this subsection we go over the classifier used to create the model that separates malware from goodware.
+Our main concerns when choosing a classifier regard the ability to produce a probabilistic output, good scaling for large number of features and samples, and ease of use.
 
-Using an Ubuntu Virtual Machine with 16 cores and 16GB of RAM, provided by INOV - Inesc Inovação, and taking advantage of scikit-learn~\cite{tool:sklearn} to train the models, Pandas~\cite{tool:pandas} for data analysis and Jupyter Interactive Notebooks~\cite{tool:jupyter} to interact with the Python scripts, we now detail our logistic regression model results.
+Taking into consideration the guidelines given in~\cite{rossow:practices,shabtai:survey} and related work in~\cite{miller:rev_int,nissim:al_pdf,rieck:dynamic,schultz:data_mining}, we decide to use \textit{\acrfull{lr}} as our model.
+This model fits our needs as it gives the probability of a random variable $X$ being 0 or 1, given a set of constraints (\ie~features), scales well with samples and features and it is readily available from several libraries, facilitating implementation~\cite{friedman2001elements}.
 
-We define this our single layer model $\LR$ results, as we are directly taking a set of features (static imports in this case) and outputting the probability of a sample being malware.
+\gls{lr} can be defined with the form
+\begin{eqnarray*}
+	\rho(x) = \dfrac{1}{1 + e^{-x}},&x = \beta_0 + \beta_1x_1 + ... + \beta_nx_n
+\end{eqnarray*}
+where $\beta_n$ is the learned weight for feature $x_n$.
+This weight is learned through iteration in order to minimize the error between the predicted values and the actual values.
+In other words, given an \textit{n-th} dimensional set of features, \gls{lr} will try to create an hyperplane that divides samples from two classes.
 
-\medskip
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsection{Laboratory Conditions}
-% \label{section:single_layer_laboratory}
+As \gls{lr} is based on the logistic function (or sigmoid function), each feature $x_n$ can vary from $-\infty$ to $+\infty$ and still the output is contained between 0 and 1, hence providing probabilistic values.
 
-We start with what we define as \textit{laboratory conditions}, ideal conditions for the problem of malware detection.
-These conditions are met when we take the labeled dataset $\CC_{strict}$ and features $\DD_{static}$, providing scenario $\mathcal{S}_{\strict}$.
-
-Using cross-validation, our $\LR$ scored with an \gls{auroc} of 91\%, as shown by the red curve in Figure \ref{fig:xval_results}.
-We argue that such high values are easily attained from factors like a small and reliable dataset, and the use of cross-validation, which mixes samples and ignores possible dependencies on malware samples.
-
-When we relax our ideal conditions to more \textit{real-world conditions}, by testing the labeled dataset $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{static}$ for scenarios $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$, respectively, the score under \gls{auroc} decreases to 90\% for $\mathcal{S}_{loose}$ (blue curve in Figure \ref{fig:xval_results}) and 75\% for $\mathcal{S}_{real}$ (green curve in Figure \ref{fig:xval_results}).
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/xval_results.png}
-	\caption[Single layer results for static features in laboratory conditions.]{Single layer results for static features in laboratory conditions.}
-	\label{fig:xval_results}
-\end{figure}
-
-From $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$, the only change is the amount of malware labeled samples, which significantly increase.
-The difference is interesting, as although the number of malware labeled samples increase significantly, the results are not that affected.
-This suggests that although the reliability for malware decreases, its impact is not as noticeable as expected.
-This might also suggest that vendors do converge on their definition of malware, under our $\Mloosev$ metric.
-If vendors did not converge on what is malware, adding more samples would culminate in worse results, as separation between malware and goodware would become harder.
-
-When looking at the changes from $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$, not only the amount of malware labeled samples increase, but also the number of goodware labeled samples, both by a significant amount.
-The way this impacts the results is pretty significant, as we observe a high decrease in the \gls{auroc} for both models.
-The metric $\Mrealv$ that labels malware and goodware for this scenario $\mathcal{S}_{\real}$ disregards the cross-check from outside repositories, which in turn degrade the reliability significantly, as well as increase the dataset size notably.
-We attribute the results' degradation mainly to the unreliability of goodware labeling, not only because we have previously seen that increase in malware does not significantly impact results (from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$), but also due to the tendency for false negatives in vendors (Figure \ref{fig:dups_frequency}), which in turn lead us to incorrectly label goodware for the samples in $\CC_{real}$.
-
-The results we described show how relaxing \textit{laboratory conditions} to more \textit{real-world conditions} degrade the model's performance.
-We now focus on using our previously defined temporal based methodologies to further converge into a real-world scenario.
-
-\medskip
-
-We start by applying our \textit{Past-to-Present} validation to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
-As previously defined, this validation starts with an older set of training samples and iteratively adds newer samples, validating each iteration on a fixed set of the most recent samples.
-Since our interest is to measure performance variation over time, we plot in Figure~\ref{fig:pastpresent} the \gls{auroc} at every iteration (\ie, fold), for each of our three scenarios.
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/pastpresent.png}
-	\caption[Single layer results for static features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation.
-Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
-fold 1)}
-	\label{fig:pastpresent}
-\end{figure}
-
-When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 0.75\% to 0.68\%.
-
-For both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ we note only a slight increase as new folds are added.
-They still relate, as we have previously noted for cross-validation, arguably given their metrics $\Mstrictv$ and $\Mloosev$ are not very different.
-The small variation to the cross-validation methodology can be justified by using small dataset size for both cases.
-
-As for $\mathcal{S}_{\real}$ we note higher variation and lower overall score, as the reliability of the metric $\Mrealv$ goes down.
-This is expected, not only because we are enforcing temporal consistency between samples, but as new folds are added, the training gets bigger, while the test remains the same.
-
-Our main observation for this validation methodology is that there is a slight tendency for \gls{auroc} to increase, as we move forward in time, close to the validation set.
-
-\medskip
-
-Our next result, which uses our \textit{Present-to-Past} validation methodology will further help analyze the aforementioned detail.
-The \textit{Present-to-Past} validation enhances the previous results under real-world conditions.
-This methodology starts by fixing the validation set to the most recent samples, but with the training set starting at the temporally closest samples to validation.
-At each iteration, older samples are added to the training set and validated on the fixed, most recent, samples.
-
-By applying this methodology to the three scenarios, $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we plot Figure \ref{fig:presentpast}, where the X axis increases as older samples are added to the training set (\ie\ fold 0 contains newer samples than fold 1), hence measuring the performance variance over time.
-Similarly to the previous observation, the average \gls{auroc} remains identical when compared to cross-validation for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$.
-As for $\mathcal{S}_{\real}$, we note a decrease in score from 75\% to 69\%.
-
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/presentpast.png}
-	\caption[Single layer results for static features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation.
-Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
-	\label{fig:presentpast}
-\end{figure}
-
-The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ display very similar results, with $\mathcal{S}_{\real}$ dropping behind due to its less reliable labeling metric.
-It is noticeable that using the entire dataset does not bring much improvement to the final results. In fact, for $\mathcal{S}_{\real}$ the score even drops after fold $\#2$.
-
-With these results, our original observation that samples closer to the validation set benefit the model becomes more convincing.
-In fact, we argue that there should be an ideal number of necessary training folds, temporally consistent with the validation fold (\ie\ any fold from training predates validation), needed to maximize the overall score.
-
-\medskip
-
-Finally, we analyze how does such reduced training set behaves in our scenarios; for this purpose, we define a sliding window that moves forward in time through each scenario for training and validation.
-We propose a reduction on the training size to $w=3$ folds predating the validation fold.
-We choose $w=3$, since we have seen that the scores either do not improve (for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$) or actually go down (for $\mathcal{S}_{\real}$) with higher folds.
-In summary, we have selected 30\% of each dataset for training purposes and the next 10\% for validation (3 training folds, 1 validation fold), and then started moving the window forward in time (1 fold at a time) to obtain the following results (Figure~\ref{fig:slidingwindow}): for  $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$, we obtain \gls{auroc} values of 89\%, 88\% and 73\%, respectively.
-These results come to reaffirm our argument that we can reduce the size of the training set, without losing any significant score (2\% for each scenario).
-
-\begin{figure}[!h]
-	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/slidingwindow.png}
-	\caption[Single layer results for static features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
-	\label{fig:slidingwindow}
-\end{figure}
-
-Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from 91\% to 89\% for $\mathcal{S}_{\strict}$, from 90\% to 88\% for $\mathcal{S}_{\loose}$ and from 75\% to 73\% for $\mathcal{S}_{\real}$.
-We should highlight that the results that use temporal consistency should better reflect the reality than standard cross-validation, since we are requiring temporally ordered samples.
-Another important idea that should be stressed is that for cross validation we used a fairly reasonable amount of data for training purposes, whereas in this last case we used a restricted amount of data.
-This might be a relevant issue in a few year's time.
-The results obtained are summarized in Table \ref{tab:singlelayer_results}.
-
-\begin{table}[!htb]
-	\renewcommand{\arraystretch}{1.2} % more space between rows
-	\centering
-	\begin{tabular}{ccccc}
-	\hline AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
-	\hline Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
-	\hline Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
-	\hline Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
-	\hline Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
-	\hline
-\end{tabular}
-\caption{Single layer results summary.}
-\label{tab:singlelayer_results}
-\end{table}
-
-\medskip
-
-With a better understanding of how the model behaves under different methodologies, we now diverge to how we improved not only the overall results, but also the information provided by the model.
+%\todo[inline]{}
+%\medskip 
+%
+%Our first and simplest model $\mathcal{LR}$, takes as features a sample's static imports, as a binary vector, and outputs the likelihood of a sample being malware.
+As we are dealing with a variety of malware samples, in the sense they have different behaviour and characteristics (\eg, trojan, virus), a generic classifier like this one may not perform well, as not only goodware may be diverse, but also is malware.
+%
+%\medskip
+%
+%To mitigate this shortcoming we provide a more complex model $\mathcal{E}$, which comprises a simple ensemble stacking approach.
+This model also takes as features the static imports and outputs the likelihood of a sample being malware, but it is layered in two steps.
+%
+%The first step (layer $\mathcal{E}_{\mathcal{L}_{0}}$) is composed of 6 \gls{lr} models, where each model is trained to output the likelihood of a sample belonging to one of six classes: trojan $\mathcal{C}_{t}$, worm $\mathcal{C}_{w}$, virus $\mathcal{C}_{v}$, spyware $\mathcal{C}_{s}$, ransom $\mathcal{C}_{r}$, or other $\mathcal{C}_{o}$, in a \textit{one-vs-all} methodology, \ie, a sample either belongs to $\mathcal{C}_{n}$ or not, with a sample's static imports as features.
+%
+%%\todo[inline]{maybe talk about AVClass here. PA: concordo}
+%%\todo[inline]{\color{blue}moved to III-B.}
+%
+%The second step (layer $\mathcal{E}_{\mathcal{L}_{1}}$) is identical to $\mathcal{LR}$, but now takes as features the output of each classifier from the previous layer, outputting the likelihood of a sample being malware.
+%
+%In summary, we make use of two models $\mathcal{LR}$ and $\mathcal{E}$, both using a linear logistic regression as base, but where $\mathcal{LR}$ uses a single classifier, and $\mathcal{E}$ uses a 2 layer ensemble stacking with 6 classifiers on the first layer providing input to a single classifier in the second layer, with the hope of facilitating class separation.

### 2018-08-12 extended_abstract/improvements.tex
latest

--- 2018-05-10 thesis_report/Thesis_Model_Improvements.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                                                                      %
-%     File: Thesis_Model_Improvements.tex                              %
-%     Tex Master: Thesis.tex                                           %
-%                                                                      %
-%     Author: João C. Godinho                                          %
-%     Last modified : Apr 2018                                         %
-%                                                                      %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%!TEX root = main.tex
 
-\chapter{Model Improvements}
-\label{chapter:model_improvements}
+\section{Model Improvements}
+\label{sec:improvements}
 
-Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this chapter to present the improvements made in order to obtain a more robust model to detect malware.
+Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this section to present the improvements made in order to obtain a more robust model to detect malware.
 We start by describing our first improvement, applying a multi layer model to extract more information regarding a sample.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Multi Layer Model}
+\subsection{Multi Layer Model}
 \label{section:improvements_multi_layer}
@@
 On the previous chapter we ended up with a simple \gls{lr} model $\LR$ that given a set of static imports from a sample, would give the probability of it being malware.
-Although that was the end purpose of our work, it does not give any deeper understanding of the sample, other than being or not malware.
-With that in mind, we propose a new model that not only provides the probability of a sample being malware, but also the probability of being from a specific malware type.
 
-This new model $\mathcal{E}$ comprises a simple ensemble stacking approach, which instead of using a single \gls{lr} classifier, multiple ones are used, layered into two steps.
+In this section we provide a new model $\mathcal{E}$ comprises a simple ensemble stacking approach, which instead of using a single \gls{lr} classifier, multiple ones are used, layered into two steps.
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}[h]
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/dia_multilayer.png}
+	\includegraphics[width=\columnwidth]{dia_multilayer}
 	\caption{Multi layer model representation.}
@@
 
-With this new model defined, we now present our approach on selecting the $n$ classes of interest.
+We now present our approach on selecting the $n$ classes of interest.
 This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware as belonging to some subclass.
 
-To choose the malware classes for our datasets, we take into account our description in Chapter \ref{chapter:related_work} regarding propagation methods and malware names based on purpose.
 With this in mind, we chose 6 malware classes: \textit{virus}, \textit{trojan}, \textit{worm}, \textit{ransom}, \textit{spyware} and \textit{other}.
-The first three classes, \textit{virus}, \textit{trojan} and \textit{worm}, were chosen from the propagation methods, whereas \textit{ransom} and \textit{spyware}
-where chosen due to their popularity in recent years. The last class, \textit{other}, serves for any malware that does not fit the previous five.
 
-As mentioned in Chapter \ref{chapter:related_work}, there is no agreed upon naming convention for malware, which translates into different names for the same malware sample.
-To minimize this problem, we referenced a tool by Sebastián, M.
-et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, using the names provided by VirusTotal~\cite{tool:virustotal}.
-
-We took advantage of this tool and modified it such that instead of providing a family name, it would provide one (or more) of the 6 previously defined classes.
+To help label our malware samples into the aforementioned classes, we take a tool by Sebastián, M.
+et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, and modify it such that instead of providing a family name, it would provide one (or more) of the 6 previously defined classes.
 Specifically, we changed it in a way that given a set of malware names, the output would be a distribution over the 6 malware classes.
@@
 For example, if a given set of names contain the name \textit{trojan} 3 times and the name \textit{virus} one time, then the weights would be
+
 \begin{eqnarray*}
-	\mathcal{W}_{trojan}=\dfrac{3}{4}=0.75,~\mathcal{W}_{virus}=\dfrac{1}{4}=0.25,~ \mathcal{W}_{c}=0, c \in \{worm, spyware, other, ransom\}
+	\mathcal{W}_{trojan}=\dfrac{3}{4}=0.75,~\mathcal{W}_{virus}=\dfrac{1}{4}=0.25,\\\mathcal{W}_{c}=0, c \in \{worm, spyware, other, ransom\}
 \end{eqnarray*}
@@
 It is worth mentioning that if a sample belongs to the \textit{goodware} class, it cannot belong to any other, likewise, if it belongs to any malware class, it cannot belong to the \textit{goodware} class.
-
-The following Table \ref{tab:dataset_sizes_new} discriminates the amount of each malware class.
-Note that the classes are not exclusive, as such the represented numbers show how many samples, out of each dataset $\CC_{strict}$, $\CC_{loose}$ and $\CC_{real}$, are classified by at least one vendor as belonging to any of the defined classes.
+Table \ref{tab:dataset_sizes_new} discriminates the amount of each malware class.
 
@@
 	\begin{tabular}{lccc}
-		\toprule
 		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
-		\midrule
+		\hline
 		Trojan			& 97,054 & 44,329 & 24,176\\
@@
 		Ransom			& 7,761 & 1,924 & 1,160\\
-		\bottomrule
+		\hline
+		\hline
+		Malware Total	& 98,582 & 45,306 & 24,658\\
+		\hline
 	\end{tabular}
+	\medskip
 	\caption{Samples belonging to each of the 6 malware classes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
@@
 
-One can see that the \textit{trojan} class is present in almost all samples, whereas \textit{ransom} is the least common class.
-More detailed statistics regarding each dataset is presented in Appendix \ref{chapter:appendixClasses}, containing information like the average weight for each class and the minimum and maximum values.
-
 \medskip
 
-In sum, we added a layer of labeling to our dataset, where a sample can either belong to the \textit{goodware} class, or to a set of the other 6 malware classes.
-This concludes the description of our first improvement to the model, what follows are our improvements regarding features.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Dynamic Features}
-\label{section:improvements_dynamic_features}
-
-Having described our first model improvement, we now proceed to our second enhancement, which takes advantage of the amount of information available in Cuckoo's~\cite{tool:cuckoo} reports.
-
-In the previous chapter, we used static imports as features for our malware detection model.
+In the base model, we used static imports as features for our malware detection model.
 Although the results are reasonable, the information which can be retrieved from static imports alone is limited.
@@
 To overcome these limitations, we resort to more dynamic information provided by Cuckoo.
-
-Given that Cuckoo runs the provided samples inside a virtual machine, it monitors the sample from a dynamic point of view.
-This monitoring includes information like the sequence of library calls and their categories and network and file activity.
-With this in mind, we now detail what dynamic information we extracted from the reports.
 
@@
 
-Figure \ref{fig:boxplot_category_calls} shows a boxplot for the category calls, where one can see the categories \textit{filesystem}, \textit{registry} and \textit{system} are the ones with higher mean, hence more used than the other categories.
-Although not depicted, some outliers for the number of category calls go up to 1,400,000, which do not provide useful information other than the number of category calls is very high.
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=\textwidth]{Figures/boxplot_category_calls.png}
-	\caption{Category calls boxplot.}
-	\label{fig:boxplot_category_calls}
-\end{figure}
-
-With that in mind, we decide to transform the values to follow a normal distribution.
+To normalize the category calls value, we decide to transform the values to follow a normal distribution.
 We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
@@
 
-Our second type of dynamic information are the number of library calls, which closely relate to the previous feature.
+Our second type of dynamic information are the number of library calls.
 While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call, hence being a subset of the previous.
@@
 \subsection{Cuckoo Signatures}
-\label{section:improvements_categories}
+\label{section:improvements_signatures}
 
@@
 For example, if a sample allocates memory and then makes it executable, it might suggest some sort of packing or obfuscation.
-
-Although some signatures are defined by a sequence of library calls, which could also be attained by creating \textit{n-grams} of our previously defined feature, using the already defined Cuckoo signatures saves training time.
-As an example, creating bi-grams of library calls, given there are 144 unique calls, would yield over 20,000 possible features, which becomes highly unpractical when considering a higher degree of n-grams.
 
@@
 	\begin{tabular}{lccc}
-		\toprule
 		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
-		\midrule
+		\hline
 		Malware			& 94,248 & 42,911 & 23,437\\
 		Goodware		& 28,385 & 1,741 & 1,741\\
-		\midrule
+		\hline
 		Total			& 122,633 & 44,652 & 25,178\\
-		\bottomrule
+		\hline
 	\end{tabular}
+	\medskip
 	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
@@
 
-In the previous sections we described all our modifications to the original model $\LR$: a two layer approach that splits malware into subclasses and the addition of three dynamic features.
-We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{section:evaluation}.
+We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{sec:eval_results}.
 
@@
 We test each methodology using the three different scenarios: $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
-
-We again use the resources provided by INOV - Inesc Inovação, together with scikit-learn~\cite{tool:sklearn}, Pandas~\cite{tool:pandas} and Jupyter Interactive Notebooks~\cite{tool:jupyter} to train and analyze the data.
-
 \medskip
@@
 For this scenario, we obtain an \gls{auroc} of 98\%, as presented by the red curve in Figure \ref{fig:improved_xval}.
-These provide the highest results, provided by factors like a small and reliable dataset, more features and the use of cross-validation.
 
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/xval_improved.png}
+	\includegraphics[width=\columnwidth]{xval_improved}
 	\caption{Multi layer results for dynamic features in laboratory conditions.}
@@
 
-As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are affected at all, when the change between the scenarios is merely in the number of malware samples.
+As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are not affected at all, when the change between the scenarios is merely in the number of malware samples.
 Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the most realistic dataset.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/pastpresent_improved.png}
+	\includegraphics[width=\columnwidth]{pastpresent_improved}
 	\caption[Multi layer results for dynamic features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/presentpast_improved.png}
+	\includegraphics[width=\columnwidth]{presentpast_improved}
 	\caption[Multi layer results for dynamic features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation.
@@
 Finally, we retest how a reduced training set behaves in our scenarios by using our \textit{Temporal Window} methodology.
-As previously mentioned, the first 3 folds seem to provide enough information to obtain good results, hence we apply the same sliding window size of $w=3$ as in Section \ref{section:single_layer_results}.
-In summary, we selected 30\% of each dataset for training and 10\% for validation.
+As previously mentioned, the first 3 folds seem to provide enough information to obtain good results, hence we apply the same sliding window size of $n=3$ as in Section \ref{section:single_layer_results}.
 Starting at the oldest fold, we apply this window and slide it by one fold at each iteration.
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{Figures/slidingwindow_improved.png}
+	\includegraphics[width=\columnwidth]{Figures/slidingwindow_improved.png}
 	\caption[Multi layer results for dynamic features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
@@
 
-As a closing remark, we reaffirm on how it is possible to yield reasonable results when considering temporal consistent samples, specifically in this improved model, which not only uses more features, but has the ability to provide more information regarding a malware sample, as it can discriminate malware classes due to its multi layer design.
+%As a closing remark, we reaffirm on how it is possible to yield reasonable results when considering temporal consistent samples, specifically in this improved model, which not only uses more features, but has the ability to provide more information regarding a malware sample, as it can discriminate malware classes due to its multi layer design.
 The improved layer results are summarized in Table \ref{tab:multilayer_results}.
@@
 	\begin{tabular}{ccccc}
-		\toprule
 		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
-		\midrule
+		\hline
 		Cross-Validation & 0.98 & 0.98 & 0.95 & 90 / 10\\
@@
 		Sliding-Window & 0.94 & 0.94 & 0.94 & 30 / 10\\
-		\bottomrule
+		\hline
 	\end{tabular}
+	\medskip
 	\caption{Multi layer results summary.}
@@
 \end{table}
-
-Having studied how our modifications improved the proposed model, the following chapter goes on to provide practical applications for our work.

### 2018-08-12 extended_abstract/introduction.tex
latest

%!TEX root = main.tex

\section{Introduction}\label{sec:intro}

The number of reported security breaches due to \emph{virus}, \emph{worms}, \emph{trojans}, etc, has been growing considerably in recent years~\cite{av-test:report} with reports of infections due to \emph{malware}  making the headlines, now more than ever.
Almost every week one such security vulnerability is reported which may be seen as a failure by the security community on the control and detection of malicious content.

The very first challenge that we face is the definition of \emph{malware}.
Malware is considered to be \say{a program with malicious intent}~\cite{christodorescu:semantics} which in itself is a dubious definition. 
Not only the same programs are classified differently as \emph{malware} and \emph{goodware} depending on the vendor, but also some programs fall within a gray area for which no clear classification can be deemed correct.
An example is what is called \emph{adware}, advertising-supported software, that although not performing directly malicious actions, perform arguably non-requested actions.
The non-existence of concrete metrics and properties that uniquely distinguish malware from goodware requires an extra effort in the preparation of datasets for evaluation of malware detectors.

Due to the significant number of malware attacks, and taking into consideration the increasing popularity and huge success of \gls{ml} methodologies in classification in different domains~\cite{lee2003learning,joachims2002learning,li2010object,ding2001multi}, it is only natural to see these techniques applied to complement classical methodologies for malicious-content detection~\cite{arp2014drebin,christodorescu:semantics,kolosnjaji2016deep,kolter:learning,miller:rev_int,nissim:al_pdf,perdisci:behavior,rieck:dynamic,santos2013opcode,schultz:data_mining,schwenk2012autonomous,vsrndic2013detection,deo2016prescience,gandotra2014malware,jordaney2017transcend,rossow:practices}, in particular, supervised learning techniques.
However, several issues have arised regarding the usage of these ML techniques in the scope of malware analysis~\cite{deo2016prescience,gandotra2014malware,jordaney2017transcend,rossow:practices}.

On the sometimes ambiguous scenario that is the task of distinguishing malware from goodware, some of these results largely depend on the datasets used for evaluation, often not representing the real-world. 
In this paper we address this pertinent questions and make a comparative analysis of a supervised learning approach in three different scenarios (depicted in Figure~\ref{fig:scenarios}): a \emph{strict scenario} where only very well-characterized samples are considered, a \emph{loose scenario} where a wider set of still well-studied samples is considered, and finally a \emph{realistic scenario} where we get closer to the reality faced by vendors of malware detection solutions.

\begin{figure}[!h]
	\centering
	\includegraphics[width=2.5in]{dataset_sizes}
	\caption{Representation of our $\mathcal{S}_{strict}$, $\mathcal{S}_{loose}$ and $\mathcal{S}_{real}$ scenarios.}
	\label{fig:scenarios}
\end{figure}

Throughout this work we will perform a comparative analysis of the three above scenarios under two distinct environments:

\begin{itemize}[noitemsep]
	\item \emph{laboratory conditions} where traditional cross-validation methodologies can be applied;
	\item \emph{real-world conditions} where time is relevant and we analyze the behavior of the classifier with temporal-based methodologies.
\end{itemize}

We show that without much tweaking, and using simple features, a \gls{lr} model is able to achieve an \gls{auroc} of 0.91 when ideal laboratory conditions are met, but these results go down to an \gls{auroc} of 0.75 when we change into a real-world scenario.

With the purpose of actually evaluating the usage of \gls{ml} methodologies for malware detection as a complement to traditional malware detection techniques, we also analyze how the size of the training set can influence the performance of the classifier under the three studied scenarios.
For this, we use the aforementioned temporal-based methodologies to come up with conclusions on how large the training set need to be to ensure optimal results from the classifier.

We finish off by improving our base model to include more dynamic features, as well as providing a multi layer approach which adds the ability of classifying malware samples from different classes, in contrast with simply providing if a sample is malware or not.
These improvements boost our \gls{auroc} to 0.98 in ideal laboratory conditions and 0.95 in a real-world scenario.

As the main contributions of this paper
\begin{enumerate*}[label={\alph*)},font={\color{red!80!black}\bfseries}]
	\item we propose three different scenarios to train and validate the model which range from a more simulated scenario to a more realistic one,
	\item temporal-based methodologies to train and validate the classifier, 
	\item we study how much can we reduce the training dataset without compromising the optimal results,
	\item we improve our base model by 7\% to 27\%.

\end{enumerate*}

This paper is outlined as follows: in Section~\ref{sec:related} we present
the related work and justify how it motivated our work; in Section~\ref{sec:data} we describe and analyze the available dataset and how it was labeled; in Section~\ref{sec:feature_model} we propose a feature selection and describe the used model; Section~\ref{sec:eval_results} encloses our main contributions, describing our three scenarios and how they are to be evaluated, using regular cross-validation and our defined temporal-based methodology, together with the results of said evaluation;
in Section~\ref{sec:improvements} we provide improvements to our base model; in Section~\ref{sec:discussion} we discuss our main achievements; Section~\ref{sec:conclusion} concludes the paper and discusses avenues for further research.

### 2018-08-12 extended_abstract/relatedwork.tex
latest

%!TEX root = main.tex

\section{Related Work}\label{sec:related}

We use this section to present prior work that closely relates to the topic of our research and our areas of contribution. Specifically, detecting malicious software by training supervised models on static information, and validation methodologies that resemble real-world conditions.

Shabtai et al.\cite{shabtai:survey} provide a survey directed
at the application of \gls{ml} classifiers to detect malware from static
features. Their work concerns the design and evaluation of such systems. Our contributions are inspired by the concern of how to correctly evaluate \gls{ml} classifiers, regarding size, reliability of labeling metrics and chronological evaluation.

In the topic of methodologies that resemble real-world conditions, Srndi\'c et al.\cite{vsrndic2013detection}
train and validate their malicious PDF detector under laboratory and real-world conditions. Laboratory conditions consist on applying regular cross-validation, whereas real-world conditions validate a newer dataset with a model created from outdated data (i.e.\ older then the validation), and also validate the model when the validation set spans one week and the training is gathered in the previous 4 weeks. They show that laboratory conditions inflate the results, when compared to real-world conditions.

Miller et al.\cite{miller:rev_int} also introduce sample temporal consistency. They show the impact of performance measurement technique on a dataset containing 1.1 million samples, when using cross-validation and temporally consistent samples. As noted by others, regular cross-validation showed inflated results when compared to temporally consistent samples.

Our work enhances these methodologies by analyzing the performance variation when the distance between the training and validation set increases and decreases, as well as analysis on how reducing the size of the training without compromising the results.

Kolter et al.\cite{kolter:learning} learn to detect malicious executables in a dataset with under 4,000 samples, obtained from reliable sources, and evaluate their model under standard cross-validation and by gathering newer malware samples to validate for new and unseen samples. They show how the model provides optimal results under cross-validation, but for the unseen samples lower scores are obtained. Our work considers these results to compare how reliability affects performance.

Sebastián, M. et al.\cite{sebastian2016avclass} develop \textit{AVClass}, a tool that given a set of antivirus vendors, outputs the most likely family name. They test their tool under 10 datasets, totaling 8.9 M samples, with results showing an F1 measure up to 93.9 on labeled datasets. The tool takes as input the labels as seen in VirusTotal, tokenizes the labels, replaces known aliases and general names (e.g.\ win32, trojan, generic), ending with possible family names. These remaining names are counted and the most frequent one is given as family name. Our work takes advantage of this tool, not to label malware families, but minimal modifications, to label more general malware classes (e.g.\ trojan, virus).

Deo, A. et al.~\cite{deo2016prescience} focus over the problem of \gls{ml} models becoming antiquated over time, given how malware evolves. This problem is seen as concept drift, where the performance of a model over time diminishes, as the statistical properties of malware (i.e.\ features) change over time. They study how probabilistic predictors can help minimize the aforementioned problem, by indicating when retraining of a given model is necessary.

Jordaney, R. et al.~\cite{jordaney2017transcend} also study the problem of concept drift in malware classification models, focusing on providing metrics, based on statistical comparison of samples, to detect when should a model be retrained.

Both ~\cite{deo2016prescience} and ~\cite{jordaney2017transcend} are on the subject of our work, regarding the problem of concept drift, but differ on the study-case. In our work we acknowledge the concept but focus on its relative effects and how one can balance the size of needed training data vs. validation data, when maintaining a temporal consistent dataset. Whereas their work provide indicators for when should a model be retrained when the problem of concept drift becomes significant.

### 2018-08-12 thesis_report/Thesis_Acknowledgements.tex
latest

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Acknowledgments.tex                                 %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified :  May 2018                                        %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\acknowledgments}

% Add entry in the table of contents as section
\addcontentsline{toc}{section}{\acknowledgments}

Firstly I would like to thank my supervisor, Prof. Pedro Adão, for accepting this challenge, as well as his support during this period of time.
I would also like to thank INOV - Inesc Inovação, specifically its cyber unit, who believed in my work and provided means for it to be developed.

Lastly, I would like to thank my family, significant other, friends and colleagues for their support and feedback throughout this time.

### 2018-08-12 thesis_report/Thesis_Background.tex
latest

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO mais umas linhas sobre estatico e dinamico
 \section{Cuckoo Sandbox and Malwr}
@@
 For example, only binary executables (\eg\ \gls{dll}, \gls{pe}) provide static analysis information.
+Using the defined sections provided by Cuckoo's reports, one can obtain static and dynamic features from the samples.
 
@@
 
+\subsection{Best Practices for \gls{ml} in Malware Detection}
+
 Before analyzing the work done with \gls{ml} in malware detection, it is worth noting how such work is to be conducted.
@@
 
-\medskip
+\subsection{Malware Detection using Static and Dynamic Features}
 
@@
 
-\medskip
+\subsection{Pitfalls of Using Classical Validation Methodologies}
 

### 2018-08-12 thesis_report/Thesis_Conclusions.tex
latest

@@
 
+% TODO mudar pratical para aqui
+
+\section{Practical Applications}
+\label{section:practical_applications}
+
+We take this section to briefly describe how we took the previously defined model to the real world, providing practical implementations for it.
+The following work is still on-going as part of a scholarship provided by INOV - Inesc Inovação, as such the end products are still being developed and test.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Malware Detection Service}
+\label{section:malware_service}
+
+Our first practical application for the obtained model comes as a malware detection service.
+The idea is to take advantage of Cuckoo~\cite{tool:cuckoo} to implement a web service identical to Malwr~\cite{tool:malwr}, but also adding the malware detection by applying our model.
+
+To implement this service, we developed a simple web application that receives files from users and then sends them to a Cuckoo instance running on a server.
+We created a custom signature for Cuckoo that takes the analysis results and converts it into features suitable for our model, which is then used to classify the sample.
+The output of the model is then appended as a signature to the final report.
+After analysis our web application extracts basic information regarding the file, as well as the output of the classifier.
+
+We implemented our web application using Django~\cite{tool:django}.
+The implemented application provides a simple upload form for users to submit their own files.
+Upon submission, Django sends the sample to a Cuckoo instance running locally.
+While Cuckoo is analyzing the sample, users are redirected to the report page, which shows the current state of analysis.
+Upon finishing analysis, the page changes to yield the classifier results.
+
+Figure \ref{fig:malware_detetor} shows the service being tested with a \texttt{petya}\footnote{Symantec - Petya Ransomware, October 27, 2017 [https://www.symantec.com/blogs/threat-intelligence/petya-ransomware-wiper]} sample, using early versions of our $\LR$ model.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/malware_detector.png}
+	\caption{Malware Detector in early stages.}
+	\label{fig:malware_detetor}
+\end{figure}
+
+We are currently testing and optimizing the service, as it is not currently using our latest and best model $\mathcal{E}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Email Attachment Scanner}
+\label{section:email_scanner}
+
+Our second practical application uses the previous service, but in an automated way, to scan email attachments arriving and departing INOV's mail server.
+The idea is to insert the malware detector in the mail server pipeline as a scanner.
+
+INOV's mail server uses MailScanner~\cite{tool:mailscanner} to filter and scan emails.
+In Figure \ref{fig:mailscanner} we provide a diagram on the MailScanner pipeline.
+
+Our implementation works by creating a custom \textit{Virus Scanner}, which takes email attachments and sends them to the previous described malware detection service.
+Upon receiving the results, our custom scanner modifies the mail object to include the classification results.
+
+Further down in the \textit{Message Processing} step, we added a custom step to modify an unused \texttt{Header} in the email to include the classification results.
+
+Currently we are in the process of moving the developed code into the mail server to start obtaining real world results.
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.8\textwidth]{Figures/mailscanner.png}
+	\caption{MailScanner process diagram.}
+	\label{fig:mailscanner}
+\end{figure}
+
 % ----------------------------------------------------------------------
@@
 	\item We provided means to achieve metrics as to facilitate the malware labeling task.
-	\item We developed methodologies that test real-world scenarios which take into account time consistency.
-	\item We provided a simple model to detect malware and how it behaves under different methodologies.
-	\item We improved the previous model to boost the results in all of our tested methodologies.
+	\item We developed methodologies that test real-world scenarios which take into account time consistency, rather than laboratory created scenarios.
+	\item We provided a simple model to detect malware and investigated how it behaves under different methodologies.
+	\item We improved the previous model with a two layer approach and more dynamic features to boost the results in all of our tested methodologies.
 	\item We started developing practical applications of our model, to be inserted in real world scenarios.
@@
 
-As future work, the developed model could be improved by adding more features, and by doing better research regarding the impacts of each feature.
+As future work, the developed model could be improved by adding more features, and by doing more in depth research regarding the impacts of each feature.
 Other models could be tested, as well as used in conjunction with our logistic regression, to improve the overall results.

### 2018-08-12 thesis_report/Thesis_Introduction.tex
latest

@@
 The amount of malware has shown an exponential increase since the early 2000s, from 1 million in 2006 to over 500 million by the end of 2016~\cite{av-test:report,mcafee:report}.
-According to a 2007 study by McAfee\footnote{McAfee Study Finds 4 Percent of Search Results Malicious, by Frederick Lane, June 4, 2007 [http:\/\/www.newsfactor.com\/story.xhtml?story\_id=010000CEUEQO]} about 4\% of the query results by major search engines lead to potentially dangerous websites.
+According to a 2007 study by McAfee\footnote{McAfee Study Finds 4 Percent of Search Results Malicious, by Frederick Lane, June 4, 2007 [http://www.newsfactor.com/story.xhtml?story\_id=010000CEUEQO]} about 4\% of the query results by major search engines lead to potentially dangerous websites.
 A more recent study by Norton in 2010\footnote{Symantec Press Release, July 29, 2010 [https://www.symantec.com/en/aa/about/newsroom/press-releases/2010/symantec\_0729\_02]} showed that at least 10\% of top trending search terms returned malicious results.
@@
 
+By designing these three scenarios, we can better understand how having a smaller and very distinctive dataset compares to a bigger and more uneven dataset, providing a greater understanding of the effects of \textit{ground-truth}.
+
 \begin{figure}[!h]
@@
 
+This will help us understand if one should take precautions when using classical evaluation methodologies, as these do not take into account time dependency between samples.
+
 We also provide several improvements to the model.
 One of these come under the form of a multi-layer approach, which instead of classifying a sample as malware or goodware, also provides information regarding the type of malware.
-Our other improvement regards providing more information about a sample to the model, in order to obtain higher scores.
+Our other improvement regards providing more information about a sample to the model, by using dynamic information extract during runtime, in order to obtain higher scores.
 
@@
 
-Chapter \ref{chapter:static_features} and \ref{chapter:model_improvements} takes the created datasets to build a malware detection model based on both static and dynamic features, as well as how these perform under different validation methodologies.
+Chapter \ref{chapter:static_features} and \ref{chapter:model_improvements} take the created datasets to build a malware detection model based on both static and dynamic features, as well as how these perform under different validation methodologies.
 
-Chapter \ref{chapter:practical_applications} describes how our model is practically implemented both as an online malware detection service and how it was inserted into a corporate email server to analyze attachments.
+% Chapter \ref{chapter:practical_applications} describes how our model is practically implemented both as an online malware detection service and how it was inserted into a email server to analyze attachments.
 
-Chapter \ref{chapter:conclusions} provides some final remarks on our work and contributions, together with possible ideas that may improve the presented work.
+Chapter \ref{chapter:conclusions} provides some final remarks on our work, how our model is practically implemented, contributions and possible ideas that may improve the presented work.

### 2018-08-12 thesis_report/Thesis_Model_Improvements.tex
latest

@@
 Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this chapter to present the improvements made in order to obtain a more robust model to detect malware.
-We start by describing our first improvement, applying a multi layer model to extract more information regarding a sample.
+We start by describing our first improvement, the application of a multi layer model to extract more information regarding a sample.
 We then take this enhanced model and increase the number of features to include dynamic content and how it impacted the model's results.
@@
 On the previous chapter we ended up with a simple \gls{lr} model $\LR$ that given a set of static imports from a sample, would give the probability of it being malware.
-Although that was the end purpose of our work, it does not give any deeper understanding of the sample, other than being or not malware.
+Although that was the goal of our work, it does not give any deeper understanding of the sample, other than being or not malware.
 With that in mind, we propose a new model that not only provides the probability of a sample being malware, but also the probability of being from a specific malware type.
 
-This new model $\mathcal{E}$ comprises a simple ensemble stacking approach, which instead of using a single \gls{lr} classifier, multiple ones are used, layered into two steps.
+This new model $\mathcal{E}$ comprises a simple ensemble stacking approach, that uses multiple \gls{lr} classifiers instead of one, layered into two steps.
 
 The first step (layer $\mathcal{E}_{\mathcal{L}_{0}}$) is composed of $n$ \gls{lr} models, where $n$ is the number of possible classes.
-Each model is trained to output the likelihood of sample belonging to one of the $n$ classes, in a \textit{one-vs-all} methodology (\ie\ a sample either belongs to $\mathcal{C}_{n}$ or not), having as input the raw features (\eg\ static imports).
+Each model is trained to output the likelihood of a sample belonging to one of the $n$ classes, in a \textit{one-vs-all} methodology (\ie\ a sample either belongs to $\mathcal{C}_{n}$ or not), having as input the raw features (\eg\ static imports).
 
@@
 With this new model defined, we now present our approach on selecting the $n$ classes of interest.
-This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware as belonging to some subclass.
+This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware according to some subclass.
 
@@
 As mentioned in Chapter \ref{chapter:related_work}, there is no agreed upon naming convention for malware, which translates into different names for the same malware sample.
-To minimize this problem, we referenced a tool by Sebastián, M.
+To minimize this problem, we used a tool by Sebastián, M.
 et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, using the names provided by VirusTotal~\cite{tool:virustotal}.
@@
 Specifically, we changed it in a way that given a set of malware names, the output would be a distribution over the 6 malware classes.
+
+Our modifications affected steps \textit{(d)} and \textit{(e)} of the AVClass process, depicted in Figure \ref{fig:avclass}.
+In step \textit{(d)}, instead of filtering family name tokens, we filtered based on our malware classes.
+Doing this affects step \textit{(e)}, where the alias detection is now for our classes.
+The remaining step is identical, but the names are now malware classes instead of family names.
 
@@
 		Ransom			& 7,761 & 1,924 & 1,160\\
+		\midrule
+		\midrule
+		Malware Total	& 98,582 & 45,306 & 24,658\\
 		\bottomrule
@@
 In the previous chapter, we used static imports as features for our malware detection model.
-Although the results are reasonable, the information which can be retrieved from static imports alone is limited.
+Although the results were interesting, the information which can be retrieved from static imports alone is limited.
 As an example, if a sample is compressed, encrypted or packed, its behavior cannot be inferred from static imports only.
@@
 Given that Cuckoo runs the provided samples inside a virtual machine, it monitors the sample from a dynamic point of view.
-This monitoring includes information like the sequence of library calls and their categories and network and file activity.
+This monitoring includes information like the \emph{sequence of library calls} and their \emph{categories} and \emph{network} and \emph{file activity}.
 With this in mind, we now detail what dynamic information we extracted from the reports.
@@
 We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
-This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large values fall into the same bin.
+This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large but uncommon values fall into the same bin.
 
@@
 Our second type of dynamic information are the number of library calls, which closely relate to the previous feature.
-While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call, hence being a subset of the previous.
+While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call.
+Each category from the previous feature contains a set of library calls, hence each library call is contained in some category.
 
@@
 
-We joined these features $\FF_{dynamic}$ to the labeled samples $\CC_{real}$, obtaining a total of 122,633 labeled samples with the new features $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$.
+Having these three new features $\FF_{dynamic}$, we made sure our labeled samples $\CC_{real}$ all contained the features, hence we joined the datasets into a slightly smaller dataset $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$, containing 122,633 samples.
 
@@
 
-As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are affected at all, when the change between the scenarios is merely in the number of malware samples.
-Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the most realistic dataset.
+As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are not affected at all, when the change between the scenarios is merely in the number of malware samples.
+Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the more realistic dataset.
 
@@
 
-Having applied the same cross-validation to our modified model, obtaining interesting results, we now go over to test how our temporal based methodologies are affected.
+Having applied the same cross-validation to our modified model, and obtaining interesting results, we now go over to test how our temporal based methodologies are affected.
 
@@
 
-The results are consistent with was previously seen in Section \ref{section:single_layer_results}, with the added factor that the absolute values are higher, and the tendency to increase is more present as we move forward in time, close to the validation set.
+The results are consistent with what was previously seen in Section \ref{section:single_layer_results}, with the added factor that the absolute values are higher, and the tendency to increase is more present as we move forward in time, close to the validation set.
 
@@
 Following the previous evaluation order, we now present the results using our \textit{Present-to-Past} validation methodology.
-In Figure \ref{fig:presentpast_improved} we present the \gls{auroc} for each iteration, where higher folds represent older samples.
+In Figure \ref{fig:presentpast_improved} we present the \gls{auroc} for each iteration, where higher folds represent the inclusion of older samples.
 Here we see values of 97\% for $\mathcal{S}_{\strict}$, 98\% for $\mathcal{S}_{\loose}$ and 96\% $\mathcal{S}_{\real}$.
@@
 For $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ the score is equally and negatively affected by 4\%.
-The jagged curve on both scenarios indicate that slight changes on the \gls{fpr} threshold have significant impact on the \gls{tpr}, this may be caused by the smaller dataset size, which in turn creates uneven folds for malware and goodware.
+The jagged curve on both scenarios indicates that slight changes on the \gls{fpr} threshold have significant impact on the \gls{tpr}.
+This may be caused by the smaller dataset size, which in turn creates uneven folds for malware and goodware.
 

### 2018-08-12 thesis_report/Thesis_Practical_Applications.tex
latest

@@
 We created a custom signature for Cuckoo that takes the analysis results and converts it into features suitable for our model, which is then used to classify the sample.
-The output of the model is then appended as a signature the final report.
+The output of the model is then appended as a signature to the final report.
 After analysis our web application extracts basic information regarding the file, as well as the output of the classifier.
+
+We implemented our web application using Django~\cite{tool:django}.
+The implemented application provides a simple upload form for users to submit their own files.
+Upon submission, Django sends the sample to a Cuckoo instance running locally.
+While Cuckoo is analyzing the sample, users are redirected to the report page, which shows the current state of analysis.
+Upon finishing analysis, the page changes to yield the classifier results.
 

### 2018-08-12 thesis_report/Thesis_Resumo.tex
latest

% !TeX spellcheck = pt_PT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Resumo.tex                                          %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified :  May 2018                                        %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Resumo}

% Add entry in the table of contents as section
\addcontentsline{toc}{section}{Resumo}

O uso de técnicas de aprendizagem supervisionada para a detecção de programas maliciosos tem sido cada vez mais utilizada para melhorar métodos clássicos de detecção.
Este trabalho desenvolveu um modelo para a detecção de programas maliciosos, analisou o impacto da confiabilidade dos dados de treino no resultado final do classificador, e definiu métricas para \textit{verdade absoluta}.
Para tal, desenvolvemos três cenários para conjuntos de dados, cujo conteúdo varia entre amostras inequivocamente maliciosas e legítimas para amostras mais ambíguas e reais.
Analisámos cada cenário em condições laboratoriais, onde metodologias padrão de validação cruzada são aplicadas, descartando a importância \textit{temporal} na detecção de programas maliciosos, e também em condições de mundo real, onde a dependência temporal é proposta e aplicada.
Além disso, modificámos o nosso modelo original de modo a possibilitar a extração de mais informação sobre uma amostra, ao implementar um modelo com multiplas camadas, e de maneira a melhorar os resultados finais, ao usar informação dinâmica sobre a amostras.
Usámos depois a nossa metodologia baseada na ordem temporal das amostras para reduzir o tamanho dos dados de treino sem comprometer resultados ótimos, concluindo que existe um número ideal de dados de treino, temporalmente consistentes com os dados de validação, tal que o resultado final é ótimo.
Finalmente, fornecemos aplicações práticas do nosso modelo ao implementar o serviço de detecção de programas maliciosos, que é também inserido num servidor de correio eletrónico para validar anexos.

\vfill

\textbf{\Large Palavras-chave:} Segurança, Aprendizagem Automática, Detecção de Programas Maliciosos, Consistência Temporal, Verdade Absoluta

### 2018-08-12 thesis_report/Thesis_Static_Features.tex
Slight fixes.
latest

@@
 Although the cross-validation methodology enables to measure the generalization capabilities of a model, it does not account for temporal ordering of the samples.
-Since we want to measure the when training samples predate validation, we now define a couple of temporal based validations, based on the concept of cross-validation:
+Since we want to measure when training samples predate validation, we now define a couple of temporal based validations, based on the concept of cross-validation:
 
@@
 
-\paragraph{\textit{Temporal Window}:} This validation, depicted in Figure \ref{fig:sliding_window}, also splits the dataset into $k$ folds, but takes $w$ temporal consistent and contiguous folds (\ie each fold immediately precedes the next one), using the first $w-1$ folds for training (older samples) and the remaining fold (more recent samples) for validation.
+\paragraph{\textit{Temporal Window}:} This validation, depicted in Figure \ref{fig:sliding_window}, also splits the dataset into $k$ folds, but takes $w$ temporal consistent and contiguous folds (\ie\ each fold immediately precedes the next one), using the first $w-1$ folds for training (older samples) and the remaining fold (more recent samples) for validation.
 By starting with the $w$ first folds of the dataset and sliding one fold on each iteration, we apply a sliding window of size $w$ over the dataset.
@@
 \medskip
+
+We measure the \gls{auroc} during each validation's iteration and use the average measurement to discuss the results.
 
@@
 
-We define this our single layer model $\LR$ results, as we are directly taking a set of features (static imports in this case) and outputting the probability of a sample being malware.
+We define this as our single layer model $\LR$ results, as we are directly taking a set of features (static imports in this case) and outputting the probability of a sample being malware.
 
@@
 
-When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 0.75\% to 0.68\%.
+When directly comparing the average \gls{auroc} for cross-validation and our \textit{Past-to-Present} validation, we note that for both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ the \gls{auroc} remains identical, while for $\mathcal{S}_{\real}$ the score decreases from 75\% to 67\%.
 
@@
 
+\medskip
+
 The comparison between scenarios is identical to what was observed in cross-validation and \textit{Past-to-Present}: scenarios $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ display very similar results, with $\mathcal{S}_{\real}$ dropping behind due to its less reliable labeling metric.
@@
 
+\medskip
+
 Comparing these results with the baseline cross-validation, we note a decrease for each scenario, specifically a decrease from 91\% to 89\% for $\mathcal{S}_{\strict}$, from 90\% to 88\% for $\mathcal{S}_{\loose}$ and from 75\% to 73\% for $\mathcal{S}_{\real}$.
@@
 \begin{table}[!htb]
-	\renewcommand{\arraystretch}{1.2} % more space between rows
+	\renewcommand{\arraystretch}{1.0} % more space between rows
 	\centering
 	\begin{tabular}{ccccc}
-	\hline AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
-	\hline Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
-	\hline Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
-	\hline Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
-	\hline Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
-	\hline
+		\toprule
+		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
+		\midrule
+		Cross-Validation & 0.91 & 0.90 & 0.75 & 90 / 10\\
+		Past-to-Present & 0.90 & 0.90 & 0.67 & 10 to 90 / 10\\
+		Present-to-Past & 0.90 & 0.90 & 0.69 & 10 to 90 / 10\\
+		Sliding-Window & 0.89 & 0.88 & 0.73 & 30 / 10\\
+		\bottomrule
 \end{tabular}
@@
 
-\medskip
-
 With a better understanding of how the model behaves under different methodologies, we now diverge to how we improved not only the overall results, but also the information provided by the model.

### 2018-08-12 learn/Real Dataset/tasks.py

@@ -1,0 +1,15 @@
+import celery
+import pickle
+from sklearn.linear_model import LogisticRegression
+
+app = celery.Celery('tasks', backend='rpc://', broker='pyamqp://jcfg:jcfg@localhost/thesis')
+app.conf.accept_content = ['pickle', 'json']
+print(app.conf.accept_content)
+
+@app.task
+def lr_train(x_train, y_train):
+    return LogisticRegression(max_iter=1).fit(x_train, y_train)
+
+@app.task
+def lr_predict(lr, x_test):
+    return picker.loads(lr).predict_proba(x_test)

### 2018-08-12 learn/helpers.py

@@ -1,0 +1,73 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from sklearn.linear_model import LogisticRegression
+from sklearn.metrics import roc_curve, auc
+from sklearn.model_selection import StratifiedKFold
+from scipy import interp
+
+
+def train_plot_roc(X, y, n_splits=10):
+    # Trying ROC
+    skf = StratifiedKFold(n_splits=n_splits)
+    classifier = LogisticRegression(C=1)
+
+    tprs = []
+    aucs = []
+    mean_fpr = np.linspace(0, 1, 100)
+    plt.figure(figsize=(10, 4), dpi=100)
+
+    i = 0
+    for train, test in skf.split(X, y):
+        probas_ = classifier.fit(X[train], y[train]).predict_proba(X[test])
+        # Compute ROC curve and area the curve
+        fpr, tpr, thresholds = roc_curve(y[test], probas_[:, 1])
+        tprs.append(interp(mean_fpr, fpr, tpr))
+        tprs[-1][0] = 0.0
+        roc_auc = auc(fpr, tpr)
+        aucs.append(roc_auc)
+        plt.plot(fpr, tpr, lw=1, alpha=0.3,
+                 label='ROC fold %d (AUC = %0.2f)' % (i, roc_auc))
+
+        i += 1
+    plt.plot([0, 1], [0, 1], linestyle='--', lw=2, color='r',
+             label='Luck', alpha=.8)
+
+    mean_tpr = np.mean(tprs, axis=0)
+    mean_tpr[-1] = 1.0
+    mean_auc = auc(mean_fpr, mean_tpr)
+    std_auc = np.std(aucs)
+    plt.plot(mean_fpr, mean_tpr, color='b',
+             label=r'Mean ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc),
+             lw=2, alpha=.8)
+
+    std_tpr = np.std(tprs, axis=0)
+    tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
+    tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
+    plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='grey', alpha=.2,
+                     label=r'$\pm$ 1 std. dev.')
+
+    plt.xlim([-0.05, 1.05])
+    plt.ylim([-0.05, 1.05])
[…]
